### 2021-08-07 analysis/paper.Rmd
Updating the flow of the paper.
Shuffled some sections around, copied text from the presentation notes from the caa conference talk, moved some unused stuff to scrap section at the end.

---
title: "openarchaeoCollaboration"
author:
  - name: Zack Batist
    url: https://zackbatist.github.io
    affiliation: University of Toronto
  - name: Joe Roe
    url: https://joeroe.io
    affiliation: University of Copenhagen
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	layout = "l-body-outset"
)
```

```{r dependencies, include=FALSE}
library("tidyverse")
library("lubridate")
library("openarchaeoCollaboration")
library("khroma")
library("gt")
```

```{r plot-theme}
theme_oarch <- function() theme(
  text = element_text(family = "Lato")
)
```

```{r data-oarch}
read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
  mutate(category = na_if(category, "?")) ->
  oarch

# Extract clean repos from GitHub URLs
oarch %>%
  mutate(
    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
  ) ->
  oarch

# Amalgamate categories into fewer rough categories
# TODO: redo
oarch <- mutate(oarch, rough_category = recode(category,
                                 "R" = "Packages (R, Python, etc.)",
                                 "Web apps" = "Standalone apps",
                                 "Datasets" = "Analyses & datasets",
                                 "Tutorials" = "Lists & tutorials",
                                 "Python" = "Packages (R, Python, etc.)",
                                 "Desktop apps" = "Standalone apps",
                                 "QGIS" = "Plugins",
                                 "Lists" = "Lists & tutorials",
                                 "Mobile apps" = "Standalone apps",
                                 "Analyses" = "Analyses & datasets",
                                 "Spreadsheets" = "Forms & spreadsheets",
                                 "Command line tools" = "Standalone apps",
                                 "Plugins" = "Plugins",
                                 "ArcGis" = "Plugins",
                                 "Filemaker" = "Plugins",
                                 "ODK" = "Forms & spreadsheets",
                                 "Lisp" = "Packages (R, Python, etc.)",
                                 .default = "Other",
                                 .missing = "Other"
                                 ))
```

```{r data-github}
# Warning: slow
oarch <- mutate(oarch,
                gh_langs = map(oarch$gh_repo, gh_lang),
                gh_contribs = map(oarch$gh_repo, gh_contrib),
                gh_issues = map(oarch$gh_repo, gh_issue),
                gh_comments = map(oarch$gh_repo, gh_comment),
                gh_commits = map(oarch$gh_repo, gh_commit)
)
```

## Introduction

In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools, and leveraging them in ways to serve our own unique needs that are typically unanticipated by their original designers.

A lot of this is surely due to resourcing and budgeting constraints, but holding prerequisite knowledge to create new tools, specifically designed for archaeology, may be a significant factor as well.

And while most archaeologists are probably content to outsource the metallurgical skills needed to forge their trowels, an increasingly large number of us are able to code our own programs and libraries, which serve functions that warrant archaeological input and initiative.

In fact, archaeologists are prolific programmers, as this paper will demonstrate.

Here, we present a survey of archaeological software development with a couple goals in mind.

First, we want to see what kinds of software archaeologists are making, and for what general purposes.

And second, we are interested in seeing how archaeologists collaborate in their programming work, and experience open source development processes, which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.

In this sense, we want to see how archaeological software is benefitting from, or missing out on, the affordances that open source development models -- particularly, the value added by a broader community of invested stakeholders, the pragmatic implementation of iterative updates, and increased transparency -- and consider how our own unique software development culture emerged and continues to grow.

## Background

## Data and evidence

Our findings here are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources. As of right now it includes 408 items, which are nicely rendered at open-archaeo.info.

The main criteria for inclusion is whether a project is identified by its maintainers as relating to archaeological work.

The dataset was compiled by browsing collaborative software development platforms, essentially manually crawling through the profiles of users who identify as archaeologists or who have contributed to tools that pertain to archaeological work.

This quasi-systematic collection strategy is also supplemented by contributions made by interested individuals who identified relevant work that was initially overlooked.

Open-archaeo is a relatively comprehensive list, however it generally lacks code written before the growth and general use of collaborative software development platforms, such as GitHub, among archaeologists, and also lacks references to code that is not available on the web.

It is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in your field, to contribute.

Please get in touch if you’re interested in adding to the list or doing a deep dive relating to your area of expertise!

### Tags and categories
Each record is also annotated with tags that describe the aspects of archaeological work that each tool contributes to.

Items are tagged based on how they have been generally identified or presented by their maintainers.

The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.

Right now there are 63 tags.

Records are also categorized based on how each tool or resource is accessed or used.

This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.

In other words, categories hint at some of the decisions that developers make regarding how they want their tools to be used and accessed.

|Category|Kind|Scope|
|:-------|:---|:----|
|Packages and libraries|Software|Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.|
|Standalone software|Software|Software that may be operated without needing to first access an underlying platform.|
|Scripts|Software|Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.|
|Specifications, protocols and schemas|Documents|A formal data structure or framework intended to be used as a model.|
|Products|Documents|Stable outcomes of creative work.|
|Guides|Documents|An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.|
|Lists and datasets|Documents|A series of consistently organized observations assembled with purpose.|

Two kinds of categories have been defined: software and documents.

Software comprise collections of instructions that dictate how a computer should operate, and documents are bounded sets of recorded information.

Three categories of software and four categories of documents are described in the table.

Packages and libraries are sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.

Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.

Standalone software are tools that may be operated without needing to first access an underlying platform.

And the third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation. Things like code generalized sets of code snippets.

As for documents, specifications, protocols and schemas are formal data structures or frameworks intended to be used as models.

Products are stable outcomes of creative work, loosely defined.

Guides are educational resources or documented protocols meant to instruct readers how to apply relevant tools or techniques.

And finally, lists or datasets, including open-archaeo itself, are series of consistently organized observations assembled with purpose.

## Findings
### What archaeologists are making
Our first set of analyses were focused on these questions:

What kinds of software are archaeologists making?

And what are the general purposes of their software?

---

Most projects included in open-archaeo are designed to be used as a package or library atop am existing ‘platform’ – for example a package for a programming language like R, or a plugin for an application like QGIS.

The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.

Others create standalone software that can be run independently of such platforms, for example desktop or web apps.

A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.

```{r plot-categories}
oarch <- mutate(oarch,
                category = recode(category,
                                  "Specifications, protocols and schemas" = "Specifications etc.",
                                  .default = category))
oarch %>%
  drop_na(category) %>%
  group_by(category) %>%
  summarise(n = n()) %>%
  ggplot(aes(fct_reorder(category, n), n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Most popular categories",
    subtitle = "By number of projects",
    caption = "Data: open-archaeo.info",
    x = NULL, y = NULL
  ) +
  theme_oarch()

# Export for slides (half slide)
# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
```

---

R is overwhelmingly dominant as a platform, closely followed by Python.

Plugins for QGIS and mobile apps are also relatively popular.

Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g. AutoCAD, ArcGIS); a number of lesser used programming languages; and a genre consisting of custom forms and spreadsheet templates.

In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms—for example Excel templates—are targeted by only one or two developers. The larger platforms tend to be more diverse.

```{r plot-platforms}
oarch %>%
  drop_na(platform) %>%
  group_by(platform, author1_name) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(total = sum(n)) %>%
  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
  geom_col(colour = "#333333") +
  coord_flip() +
  labs(
    title = "Most popular platforms",
    subtitle = "By number of projects, split by primary author",
    caption = "Data: open-archaeo.info",
    x = NULL, y = NULL
  ) +
  theme_oarch()

# Export for slides (half slide)
# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
```

---

If we compare categories with tags, we can see the general domains that each kind of resource is designed to serve.

We see that packages are fairly common across the board.

Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.

These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.

```{r plot-tags-cats}
oarch %>%
  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
               values_drop_na = TRUE) %>%
  drop_na(tag) %>%
  filter(!tag %in% detail_tags) %>%
  mutate(category = recode(category,
                                  "Specifications etc." = "Documents",
                                  "Products" = "Documents",
                                  "Guides" = "Documents",
                                  "Lists and datasets" = "Documents",
                                  .default = category)) %>%
  group_by(tag, category) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(total = sum(n)) %>%
  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
  # facet_wrap(vars(category)) +
  geom_col() +
  scale_fill_bright() +
  coord_flip() +
  labs(
    title = "Tags by category",
    caption = "Data: open-archaeo.info",
    x = NULL, y = "Number of projects",
    fill = NULL
  ) +
  theme_oarch() +
  theme(
    axis.text.y = element_text(size = 7)
  )

# Export for slides (full slide)
# ggsave("figures/tags-cats.png", width = 254, height = 127, units = "mm")
```

**ZB NOTE:** Do a temporal analysis to get a sense of when platforms, tags and categories have been gaining popularity. There may not be a need for extreme detail here, but simply noting the dates of first commits might be helpful. I'm imagining a stacked bar chart, with each bar representing a month or a year, and with the stacked portions of each bar representing a platform or category.

### Development processes
While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about _how_ archaeologists develop software.

The remainder of the paper is driven by a concern about how archaeologists collaborate in their programming work.

How is their work sustained, and how does it leverage distributed work processes that are a key aspect of open source software development in general?

To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub’s public API.

That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API.

This does reduce our sample size slightly; about an eighth of projects don’t use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.

```{r plot-repo-hosts}
oarch <- mutate(oarch, none = ifelse(is.na(github) &
                                       is.na(gitlab) &
                                       is.na(bitbucket) &
                                       is.na(launchpad) &
                                       is.na(codeberg),
                                     "", NA))

oarch %>%
  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
               names_to = "repo_host", values_to = "repo_url") %>%
  drop_na(repo_url) %>%
  group_by(repo_host, author1_name) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(total = sum(n)) %>%
  ungroup() %>%
  mutate(repo_host = fct_reorder(repo_host, total),
         repo_host = fct_relevel(repo_host, "none")) %>%
  ggplot(aes(repo_host, n, group = author1_name)) +
  geom_col(colour = "#333333") +
  coord_flip() +
  labs(title = "Most popular repository hosts",
       subtitle = "By number of projects, split by primary author",
       caption = "Data: open-archaeo.info",
       x = NULL, y = NULL)

# Export for slides (half slide)
# ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
```

We obtained this data in R using the {gh} package; those data, and the analysis code, are available in our research compendium on GitHub.

---

Starting with a relatively simple plot, this is a chart of the cumulative growth of open archaeology over the years, measured by code contributions committed and pushed to GitHub repositories.

As we can see, archaeologists have been using git from even before GitHub was launched in 2008. But they really began to get into it around 2014–2015, when we see an uptick in the rate of growth that has continued ever since.

Around this time we also see that GitHub starts being used to host documents and scripts. This may represent a recognition of GitHub’s ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.

```{r plot-gh_commits-cum}
oarch %>%
  unnest(gh_commits) %>%
  drop_na(sha) %>%
  mutate(date = floor_date(datetime, "week")) %>%
  mutate(lumped_category = recode(category,
                                  "Specifications etc." = "Documents",
                                  "Products" = "Documents",
                                  "Guides" = "Documents",
                                  "Lists and datasets" = "Documents",
                                  .default = category)) %>%
  group_by(lumped_category, date) %>%
  summarise(n_commits = n(), .groups = "drop_last") %>%
  mutate(cat_total = sum(n_commits)) %>%
  mutate(cum_commits = cumsum(n_commits)) %>%
  ggplot(aes(date, cum_commits,
             fill = fct_reorder(lumped_category, cat_total),
             colour = fct_reorder(lumped_category, cat_total))) +
  geom_area() +
  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
                              ymd("2021-05-31", tz = "UTC")),
                   date_breaks = "2 years",
                   date_labels = "%Y") +
  scale_y_continuous(labels = scales::comma) +
  scale_colour_bright(guide = NULL) +
  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
  labs(
    title = "Growth of open archaeology",
    subtitle = "Measured by cumulative commits in GitHub repositories",
    caption = "Data: open-archaeo.info and GitHub API",
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_oarch()

# Export for slides (full slide)
# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
```

---

But how is this activity distributed across repositories?

This visualizes the ‘pulse’ of active projects.

Each vertical line represents the lifespan of a repository; the points on the line are individual commits.

This is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.

```{r plot-lifespan}
oarch %>%
  unnest(gh_commits) %>%
  drop_na(datetime) %>%
  group_by(item_name) %>%
  add_tally() %>%
  filter(n > 1) %>%
  mutate(first_commit = min(datetime),
         mid_commit = median(datetime),
         last_commit = max(datetime)) %>%
  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
             y = datetime,
             ymin = first_commit,
             ymax = last_commit)) +
  geom_linerange(size = 0.05) +
  geom_point(size = 0.1) +
  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
                              ymd("2021-05-31", tz = "UTC")),
                   date_breaks = "2 years", date_labels = "%Y") +
  # coord_flip() +
  labs(
    title = "Project lifespan",
    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
    x = NULL, y = NULL,
    caption = "Data: open-archaeo.info and GitHub API"
  ) +
  theme_oarch() +
  theme(
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank()
  )

# Export for slides (full slide)
# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
```

---

The span of time between the first and latest commits is visualized more clearly in summary here.

Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.

Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.

Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects.

The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e. the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.

But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.

This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.

**ZB NOTE:** Use patchwork package to combine these plots into a unified figure.

```{r plot-lifespan-hist}
oarch_lifespan %>%
  ggplot(aes(x = lifespan)) +
  geom_histogram(binwidth = 100) +
  scale_x_continuous(breaks = scales::breaks_width(1000)) +
  labs(
    title = "Project lifespan",
    subtitle = "Time between first and latest commit, by repository",
    x = "Days", y = NULL,
    caption = "Data: open-archaeo.info and GitHub API"
  ) +
  theme_oarch()

# Export for slides (half slide)
# ggsave("figures/lifespan-hist.png", width = 127, height = 127, units = "mm")
```

```{r plot-lifespan-vs-rate}
oarch_lifespan %>%
  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
  geom_point(shape = 21, fill = NA) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = "Project lifespan vs. average commit rate",
    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
    caption = NULL
  ) +
  theme_oarch()
# Export for slides (half slide)
# ggsave("figures/lifespan-vs-rate-total.png", width = 127, height = 60, units = "mm")

oarch_lifespan %>%
  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
  geom_point(shape = 21, fill = NA) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    title = NULL,
    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
    caption = "Data: open-archaeo.info and GitHub API"
  ) +
  theme_oarch()
# Export for slides (half slide)
# ggsave("figures/lifespan-vs-rate-contrib.png", width = 127, height = 60, units = "mm")

```

---

Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed [<3%].

A “contribution”, in this case, is a commit, issue, comment applied by a user to a repository.

```{r plot-gh_contributors-histogram}
oarch %>%
  drop_na(gh_repo) %>%
  unnest(gh_contribs) %>%
  group_by(gh_repo) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = n)) +
  geom_histogram() +
  scale_x_log10() +
  labs(
    title = "Repository contributors",
    x = "Number of contributors", y = "Repositories",
    caption = "Data: open-archaeo.info and GitHub API"
  ) +
  theme_oarch()

# Export for slides (half slide)
# ggsave("figures/contrib-hist.png", width = 127, height = 127, units = "mm")
```

---

Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.

Of multi-contributor repositories, 92% have a third of contributions made by a single person;

65% have half of all contributions made by a single person;

and 36% have three quarters of all contributions made a single person.

Just to clarify, the two rows near the bottom with what seem like very long yellow bars are actually projects that have equal number of contributions made by the top contributors.

```{r plot-contrib-dist}
oarch %>%
  unnest(gh_contribs) %>%
  drop_na(contributor) %>%
  group_by(item_name) %>%
  mutate(n_contributors = n()) %>%
  filter(n_contributors > 1) %>%
  mutate(big_boy = ifelse(contributions == max(contributions),
                          "Main contributor", "Other contributors"),
         big_boy_pct = max(contributions) / sum(contributions)) ->x %>%
  ggplot(aes(x = fct_reorder(item_name, big_boy_pct, .desc = FALSE),
             y = contributions,
             # group = contributor,
             fill = fct_relevel(big_boy, c("Other contributors",
                                           "Main contributor")))) +
  geom_col(position = "fill", colour = "white", size = 0.1) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_contrast(guide = guide_legend(reverse = TRUE)) +
  labs(title = "Repository contributors by proportion of contributions",
       x = NULL, y = "Contributions",
       fill = NULL,
       caption = "Data: open-archaeo.info and GitHub API") +
  coord_flip() +
  theme_oarch() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

# Export for slides (half slide)
# ggsave("figures/contrib-dist.png", width = 127, height = 127, units = "mm")
```

Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).

Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.

However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.

This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.

```{r plot-issues-hist}
oarch %>%
  unnest(gh_issues) %>%
  drop_na(title) %>%
  group_by(repo) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = n)) +
  geom_histogram() +
  scale_x_log10() +
  labs(
    title = "Use of issues",
    caption = "Data: open-archaeo.info and GitHub API",
    x = "Number of issues", y = "Repositories"
  ) +
  theme_oarch()

# Export for slides (half slide)
# ggsave("figures/issues-hist.png", width = 127, height = 127, units = "mm")
```

```{r plot-comments-hist}
oarch %>%
  unnest(gh_comments) %>%
  drop_na(issue_num) %>%
  group_by(repo) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = n)) +
  geom_histogram() +
  scale_x_log10() +
  labs(
    title = "Use of comments",
    caption = "Data: open-archaeo.info and GitHub API",
    x = "Number of comments", y = "Repositories"
  ) +
  theme_oarch()

# Export for slides (half slide)
# ggsave("figures/comments-hist.png", width = 127, height = 127, units = "mm")
```

### Formation of collaborative development communities
If we construct a network linking repositories by their common contributors, can we identify emerging networks of collaboration, or are individual projects still largely isolated? If there are networks, what characterises their links?

**ZB NOTE:** I'm speculating that there are going to be various users who have their fingers in many projects, and that those projects tend to be the most active. If this is the case, I would further speculate that these people hold a different attitude regarding authorship and the dividuality of projects that is more in sync with the collaborative capabilities of open source, which are at odds with traditional academic norms regarding how contributions are credited and how products are produced are disseminated.

## Conclusions

Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.

In terms of collaboration, our analysis shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host.

The vast majority of repositories have 1–3 contributors, with only a few distinguished by an active and diverse developer base.

Generally speaking, we believe that this is because people do not want to step on other people’s shoes by raising issues or intruding on other people’s projects.

This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are ‘owned’ by individuals or communities, and how work should be iteratively improved upon.

We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.

# Scrap

The GitHub API provides data on the languages used in a repository based on an automated machine learning algorithm.

Since most repositories use more than one language—if only one for the actual software, and another (e.g. HTML) for documentation—this data is presented as a table of languages and the number of bytes of code in that language detected in the repository.

We can therefore measure the popularity of languages across repositories in a couple of ways: based on how many repositories use a language at all; or based on how many bytes of that language are present.

Looking at number of bytes puts C++ and the web languages (HTML, PHP, and Javascript) on top, but this is probably affected by the relative verboseness of these languages and the popularity of web-based documentation across implementation languages. The number of repositories is probably more reliable and shows the popularity of scientific programming languages—R and Python—followed by web languages and then compiled C and C++.

This largely confirms the pattern seen in our manual classification of the 'platform' of each project.

```{r plot-gh_langs, fig.show="hold", out.width="50%"}
oarch %>%
  unnest(gh_langs) %>%
  drop_na(lang, bytes) %>%
  group_by(lang) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(-n) %>%
  slice(1:10) %>%
  ggplot(aes(fct_reorder(lang, n, .desc = TRUE), n)) +
  geom_col() +
  labs(title = "Repository languages",
       subtitle = "Number of repositories using top 10 languages",
       caption = "Data: open-archaeo.info and GitHub API",
       x = NULL, y = "Number of repositories")

oarch %>%
  unnest(gh_langs) %>%
  drop_na(lang, bytes) %>%
  group_by(lang) %>%
  summarise(bytes = sum(bytes), .groups = "drop") %>%
  arrange(-bytes) %>%
  slice(1:10) %>%
  ggplot(aes(fct_reorder(lang, bytes, .desc = TRUE), bytes)) +
  geom_col() +
  labs(title = "Repository languages",
       subtitle = "Total bytes of code in top 10 languages",
       caption = "Data: open-archaeo.info and GitHub API",
       x = NULL, y = "Bytes")
```

---

About one in eight projects do not use one of these repository host.

Some may use a different host, or not make their repositories publicly available, but it is probably safe to assume that the majority simply do not use version control.

Many of these are standalone documents or code fragments (scripts) hosted, for example, as 'gists', but there are also some pieces of standalone software that apparently do not include version control.

```{r table-no-repo}
# TODO: prettify with gt()
filter(oarch, none == "") %>%
  count(category)
```

---

The measures of activity may also be compared with the tags and categories.

```{r plot-lifespan-activity-tags}
left_join(oarch_lifespan, oarch, by = c("item_name")) %>%
  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
               values_drop_na = TRUE) %>%
  drop_na(tag) %>%
  group_by(tag) %>%
  summarise(avg_lifespan = mean(lifespan),
            avg_commit_rate = mean(commit_rate)) ->
  oarch_lifespan_by_tag

oarch_lifespan_by_tag %>%
  arrange(-avg_lifespan) %>%
  slice(1:10) %>%
  ggplot(aes(x = fct_reorder(tag, avg_lifespan, .desc = TRUE),
             y = avg_lifespan)) +
  geom_col() +
  labs(
    title = "Project lifespan by tag",
    subtitle = "Top 10 longest-lived tags",
    x = NULL, y = "Average lifespan (days)",
    caption = "Data: open-archaeo.info and GitHub API"
  )

oarch_lifespan_by_tag %>%
  arrange(-avg_commit_rate) %>%
  slice(1:10) %>%
  ggplot(aes(x = fct_reorder(tag, avg_commit_rate, .desc = TRUE),
             y = avg_commit_rate)) +
  geom_col() +
  labs(
    title = "Project activity by tag",
    subtitle = "Top 10 most activity tags",
    x = NULL, y = "Average activity (commits per day)",
    caption = "Data: open-archaeo.info and GitHub API"
  )
```

```{r plot-lifespan-activity-cats}
#TODO: Eliminate Infs

left_join(oarch_lifespan, oarch, by = c("item_name")) %>%
  drop_na(category) %>%
  group_by(category) %>%
  summarise(avg_lifespan = mean(lifespan),
            avg_commit_rate = mean(commit_rate)) ->
  oarch_lifespan_by_cat

oarch_lifespan_by_cat %>%
  arrange(-avg_lifespan) %>%
  slice(1:10) %>%
  ggplot(aes(x = fct_reorder(category, avg_lifespan, .desc = TRUE),
             y = avg_lifespan)) +
  geom_col() +
  labs(
    title = "Project lifespan by category",
    subtitle = "Top 10 longest-lived categories",
    x = NULL, y = "Average lifespan (days)",
    caption = "Data: open-archaeo.info and GitHub API"
  )

oarch_lifespan_by_cat %>%
  arrange(-avg_commit_rate) %>%
  slice(1:10) %>%
  ggplot(aes(x = fct_reorder(category, avg_commit_rate, .desc = TRUE),
             y = avg_commit_rate)) +
  geom_col() +
  labs(
    title = "Project activity by categories",
    subtitle = "Top 10 most activity categories",
    x = NULL, y = "Average activity (commits per day)",
    caption = "Data: open-archaeo.info and GitHub API"
  )
```

### 2021-11-17 analysis/paper.Rmd
Outlining background section, updating intro.

@@
 ## Introduction
+In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools to serve our own unique purposes, which are not typically unanticipated by their original designers.
+In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways.
+For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase.
+However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems.
+In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes.
 
-In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools, and leveraging them in ways to serve our own unique needs that are typically unanticipated by their original designers.
+This article presents a survey of archaeological software development with two goals in mind:
 
-A lot of this is surely due to resourcing and budgeting constraints, but holding prerequisite knowledge to create new tools, specifically designed for archaeology, may be a significant factor as well.
+(1) we identify _what_ kinds of software archaeologists are making;
+(2) and evaluate _how_ archaeologists create these tools, with particular emphasis on practices of collaboration.
 
-And while most archaeologists are probably content to outsource the metallurgical skills needed to forge their trowels, an increasingly large number of us are able to code our own programs and libraries, which serve functions that warrant archaeological input and initiative.
+We examine archaeological software development as reflections of a unique set of cultural practices.
 
-In fact, archaeologists are prolific programmers, as this paper will demonstrate.
-
-Here, we present a survey of archaeological software development with a couple goals in mind.
-
-First, we want to see what kinds of software archaeologists are making, and for what general purposes.
-
-And second, we are interested in seeing how archaeologists collaborate in their programming work, and experience open source development processes, which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.
-
-In this sense, we want to see how archaeological software is benefitting from, or missing out on, the affordances that open source development models -- particularly, the value added by a broader community of invested stakeholders, the pragmatic implementation of iterative updates, and increased transparency -- and consider how our own unique software development culture emerged and continues to grow.
+Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
 
 ## Background
+- The notion of the archaeological toolbox, and its inclusion of digital tools and skillsets
+  - The use of off-the-shelf tools like MS Access, Excel, Agisoft PhotoScan, Dropbox, etc
+  - Link to related factors:
+    - Lack of experienced programmers among archaeologists
+    - The built-in flexibility of these tools
+    - Ease of administrative purchasing through university departments
+  - What changed?
+    - Software is expensive, shift to subscription models, locking into a dependency on a commercial service provider
+    - Archaeologists are better at computer programming, digital literacy on the rise, digital archaeologists grows into its own legitimate sub-discipline
+    - Open source is cool, everyone has a github profile (even if not really)
+
+- About open source practices and their influence on archaeology
+  - What is open source software and open source software development?
+    - Free cost as an end-user
+    - Free to manipulate and extend
+    - Community-based support and development
+  - Software as product vs software as collective process
+    - Basically what distinguishes abandonware from thriving software ecosystems
+    - Technical, legal and social notions of open
+      - Technical: source code is open, use of a platform and systems that support software integration and implementations
+      - Legal: open licenses, backing by the courts
+      - Social: inviting community, meaningful contributions provided by users, common vision and values directing the project into future directions
+    - Consider how these aspects of open have been implemented or prioritized by digital archaeologists in different ways
+      - Technical: reproducible research, archiving and publishing
+      - Legal: archiving and publishing, budgeting for projects
+      - Social: ensuring that software is actually useful for general use and not just one-off implementation, coordinating to reduce redundant effort and not reinvent the wheel, writing documentation, making it fit within existing systems or infrastructures (i.e.
+allocation of credit, citing software directly versus articles that present software, sustainable development after funding runs out, etc)
+    - This paper largely focuses on the social aspect of open, which is arguably the most important for ensuring a positive future for archaeological software development
+
+The 'archaeological toolbox' is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely upon in their work.
+
+collaborate in their programming work, and how they experience open source development processes which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.
 

### 2022-04-25 analysis/paper.Rmd
Update cached GitHub data and put caching logic in the Rmd

@@
 library("gt")
+library("fs")
+library("here")
 ```
@@
 ```{r data-github}
-# Warning: slow
+# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
+# chunk will prefer to load it rather than make requests to the GitHub API.
+# Delete this file to fetch fresh data from the GitHub or modify 
+# `cached_gh_data` to read a different cached version.
+#
+# Note that fetching data from the GitHub API takes some time and is subject to
+# a rate limit:
+# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
+#
+# You will need to authenticate authenticate with GitHub credentials to get a
+# rate limit sufficient to obtain the required data, see 
+# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
+# for options on how to do this.
+cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
+if (!file_exists(cached_gh_data)) {
   oarch <- mutate(oarch,
@@
   )
+  saveRDS(oarch, cached_gh_data)
+}
+oarch <- readRDS(cached_gh_data)
 ```

### 2022-04-25 analysis/paper.Rmd
Note or fix some errors preventing knit

@@
 ```{r plot-lifespan-hist}
+# TODO: where does oarch_lifespan come from?
 oarch_lifespan %>%
@@
                           "Main contributor", "Other contributors"),
-         big_boy_pct = max(contributions) / sum(contributions)) ->x %>%
+         big_boy_pct = max(contributions) / sum(contributions)) %>%
   ggplot(aes(x = fct_reorder(item_name, big_boy_pct, .desc = FALSE),
@@
   scale_y_continuous(labels = scales::percent) +
-  scale_fill_contrast(guide = guide_legend(reverse = TRUE)) +
+  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
   labs(title = "Repository contributors by proportion of contributions",
@@
 ```{r plot-lifespan-activity-tags}
+# TODO: where does oarch_lifespan come from?
 left_join(oarch_lifespan, oarch, by = c("item_name")) %>%

### 2022-04-25 analysis/paper.Rmd
Comment on GitHub hegemony

@@
     - Open source is cool, everyone has a github profile (even if not really)
+      - (JR:) People are presumably guided towards GitHub (or at least git) via other routes, because it's seen as the standard platform for 'open' digital work.
+For example, the literature on reproducible compendiums recommends GitHub, and almost all R development work is now done there.
+And once there, you're sort of forced into engaging in a (particular set of) open source development.
+I know there's controversy about how dominant this has become and people who say it's not 'true' (free and) open source, and perhaps we could add to that the question of whether it's the right model for archaeology.
 

### 2022-04-25 analysis/paper.Rmd
Initial network analysis
Including repo-user and repo-repo graphs

@@
 library("here")
+library("tidygraph")
+library("ggraph")
 ```
@@
 
+To construct our graphs we need a table linking users to repos by collaborative activity.
+For this purpose commits, issues, and comments are all considered contributions.
+
+```{r graph-data}
+oarch |>
+  drop_na(gh_repo) |>
+  nest(tags = c(tag1, tag2, tag3, tag4, tag5)) |>
+  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
+  oarch_graph
+
+oarch_graph <- bind_rows(
+  oarch_graph |>
+    unnest(gh_comments) |>
+    group_by(repo, author) |>
+    summarise(category, tags,
+              contrib = "comment", n_contrib = n(), .groups = "drop"),
+  oarch_graph |>
+    unnest(gh_issues) |>
+    group_by(repo, author) |>
+    summarise(category, tags,
+              contrib = "issue", n_contrib = n(), .groups = "drop"),
+  oarch_graph |>
+    unnest(gh_commits) |>
+    group_by(repo, author) |>
+    summarise(category, tags,
+              contrib = "commit", n_contrib = n(), .groups = "drop")
+)
+
+oarch_graph <- rename(oarch_graph, user = author)
+```
+
+With this, we can construct our three graphs: repo–user, repo–repo (by common users), and user–user (by common repos).
+
+### Repo–user
+
+```{r plot-graph-repo-user}
+oarch_graph |>
+  pivot_longer(c(repo, user), names_to = "node_type", values_to = "node") |>
+  select(node, node_type) |>
+  distinct() ->
+  nodes
+
+oarch_graph |>
+  rename(to = repo, from = user) ->
+  edges
+
+tbl_graph(nodes, edges) |>
+  ggraph(layout = "graphopt") +
+  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
+  geom_node_point(aes(colour = node_type)) +
+  scale_edge_alpha(trans = "log")
+```
+
+### Repo–repo
+
+```{r plot-graph-repo-repo}
+oarch_graph |>
+  select(repo, category) |>
+  distinct() ->
+  nodes
+
+oarch_graph |>
+  left_join(oarch_graph, by = "user") |>
+  select(repo.x, user, repo.y) |>
+  filter(repo.x != repo.y) |>
+  filter(!str_detect(user, coll("[bot]"))) |>
+  distinct() |>
+  count(repo.x, repo.y) |>
+  rename(to = repo.x, from = repo.y) ->
+  edges
+
+tbl_graph(nodes, edges) |>
+  ggraph(layout = "graphopt") +
+  geom_edge_fan(aes(alpha = n)) +
+  geom_node_point(aes(colour = category)) +
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common contributors")
+```
+
+### User–user
+
+```{r plot-graph-repo-user}
+# TODO
+```
+
 ## Conclusions

### 2022-04-26 analysis/paper.Rmd
Refine graph data and refactor

--- 2022-04-25 analysis/paper.Rmd
@@
 
+To construct our graphs we need a table linking users to repos by collaborative activity.
+For this purpose commits, issues, and comments are all considered contributions.
+
+```{r graph-data}
+oarch |>
+  drop_na(gh_repo) |>
+  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
+  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
+  oarch_graph
+
+oarch_graph <- bind_rows(
+  oarch_graph |>
+    unnest(gh_commits) |>
+    drop_na(author) |> # Apparently real commits, but we can't connect them
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "commit", n_contrib = n(), .groups = "drop"),
+  oarch_graph |>
+    unnest(gh_comments) |>
+    drop_na(author) |> # Repos with no comments
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "comment", n_contrib = n(), .groups = "drop"),
+  oarch_graph |>
+    unnest(gh_issues) |>
+    drop_na(author) |> # Repos with no issues
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "issue", n_contrib = n(), .groups = "drop")
+)
+
+# Cleaning
+oarch_graph |>
+  rename(user = author) |>
+  filter(!str_detect(user, coll("[bot]"))) ->
+  oarch_graph
+```
+
+With this, we can construct our three graphs: repo–user, repo–repo (by common users), and user–user (by common repos).
+
+#### Repo–user
+
+```{r plot-graph-repo-user}
+oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
+                                   edges_user_to_repo(oarch_graph))
+
+oarch_graph_repo_user |>
+  ggraph(layout = "graphopt") +
+  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
+  geom_node_point(aes(colour = node_type)) +
+  scale_edge_alpha(trans = "log")
+```
+
+#### Repo–repo
+
+```{r plot-graph-repo-repo}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
+                                   edges_common_users(oarch_graph))
+
+oarch_graph_repo_repo |>
+  ggraph(layout = "graphopt") +
+  geom_edge_link(aes(alpha = n)) +
+  geom_node_point(aes(colour = category)) +
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common contributors")
+```
+
+#### User–user
+
+```{r plot-graph-repo-user}
+# TODO
+```
+
+#### Over time
+
+Could animate the graphs??
+
 ## Conclusions

### 2022-04-26 R/graph.R

@@ -1,0 +1,45 @@
+#' @export
+nodes_repos_users <- function(oarch_graph) {
+  oarch_graph |>
+    pivot_longer(c(repo, user), names_to = "node_type", values_to = "node") |>
+    select(node, node_type) |>
+    distinct()
+}
+
+#' @export
+nodes_repos <- function(oarch_graph) {
+  oarch_graph |>
+    select(repo, category) |>
+    distinct()
+}
+
+#' @export
+nodes_users <- function(oarch_graph) {
+
+}
+
+#' @export
+edges_user_to_repo <- function(oarch_graph) {
+  oarch_graph |>
+    rename(to = repo, from = user)
+}
+
+#' @export
+edges_repo_to_user <- function(oarch_graph) {
+  oarch_graph |>
+    rename(to = user, from = repo)
+}
+
+#' @export
+edges_common_users <- function(oarch_graph) {
+  oarch_graph |>
+    left_join(oarch_graph, by = "user") |>
+    select(repo.x, user, repo.y) |>
+    filter(repo.x != repo.y) |>
+    mutate(repo.xy = map2_chr(repo.x, repo.y,
+                              ~paste0(sort(c(.x, .y)), collapse = ""))) |>
+    distinct(user, repo.xy, .keep_all = TRUE) |>
+    select(-repo.xy) |>
+    count(repo.x, repo.y) |>
+    rename(to = repo.x, from = repo.y)
+}

### 2022-04-27 analysis/paper.Rmd
Add interactive graph vis.
Cite {gh}

@@
 
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API.
+That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh].
 
@@
 
+```{r iplot-graph-repo-repo}
+iplot <- oarch_graph_repo_repo |>
+    ggraph(layout = "graphopt") +
+    geom_edge_link(aes(alpha = n)) +
+    geom_point_interactive(aes(x, y, colour = category, tooltip = repo)) + 
+    scale_edge_alpha(trans = "log") +
+    labs(edge_alpha = "Common contributors") +
+    theme_graph()
+
+girafe(ggobj = iplot)
+```
+
 #### User–user

### 2023-01-30 analysis/paper.Rmd
Minor formatting changes

@@
 ## Introduction
+
 In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools to serve our own unique purposes, which are not typically unanticipated by their original designers.
@@
 
-(1) we identify _what_ kinds of software archaeologists are making;
-(2) and evaluate _how_ archaeologists create these tools, with particular emphasis on practices of collaboration.
+(1) we identify *what* kinds of software archaeologists are making;
+(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
 
@@
 ## Background
+
 -   The notion of the archaeological toolbox, and its inclusion of digital tools and skillsets
@@
 I know there's controversy about how dominant this has become and people who say it's not 'true' (free and) open source, and perhaps we could add to that the question of whether it's the right model for archaeology.
-
 -   About open source practices and their influence on archaeology
@@
 
-Please get in touch if you’re interested in adding to the list or doing a deep dive relating to your area of expertise!
+Please get in touch if you're interested in adding to the list or doing a deep dive relating to your area of expertise!
 
 ### Tags and categories
+
 Each record is also annotated with tags that describe the aspects of archaeological work that each tool contributes to.
@@
 
-|Category|Kind|Scope|
-|:-------|:---|:----|
-|Packages and libraries|Software|Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.|
-|Standalone software|Software|Software that may be operated without needing to first access an underlying platform.|
-|Scripts|Software|Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.|
-|Specifications, protocols and schemas|Documents|A formal data structure or framework intended to be used as a model.|
-|Products|Documents|Stable outcomes of creative work.|
-|Guides|Documents|An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.|
-|Lists and datasets|Documents|A series of consistently organized observations assembled with purpose.|
+| Category                              | Kind      | Scope                                                                                                                                                                                                        |
+|:--------------------------------------|:----------|:-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
+| Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
+|
+| Standalone software                   | Software  | Software that may be operated without needing to first access an underlying platform.
+|
+| Scripts                               | Software  | Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
+|
+| Specifications, protocols and schemas | Documents | A formal data structure or framework intended to be used as a model.
+|
+| Products                              | Documents | Stable outcomes of creative work.
+|
+| Guides                                | Documents | An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.
+|
+| Lists and datasets                    | Documents | A series of consistently organized observations assembled with purpose.
+|
 
@@
 ## Findings
+
 ### What archaeologists are making
+
 Our first set of analyses were focused on these questions:
@@
 
----
+------------------------------------------------------------------------
 
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing ‘platform’ – for example a package for a programming language like R, or a plugin for an application like QGIS.
+Most projects included in open-archaeo are designed to be used as a package or library atop am existing 'platform' -- for example a package for a programming language like R, or a plugin for an application like QGIS.
 
@@
 
----
+------------------------------------------------------------------------
 
@@
 
-In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms—for example Excel templates—are targeted by only one or two developers.
+In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
 The larger platforms tend to be more diverse.
@@
 ### Development processes
-While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about _how_ archaeologists develop software.
+
+While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about *how* archaeologists develop software.
 
@@
 
-To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub’s public API.
+To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
 
@@
 
-This does reduce our sample size slightly; about an eighth of projects don’t use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.
+This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.
 
@@
 As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
-But they really began to get into it around 2014–2015, when we see an uptick in the rate of growth that has continued ever since.
+But they really began to get into it around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
 
 Around this time we also see that GitHub starts being used to host documents and scripts.
-This may represent a recognition of GitHub’s ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.
+This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.
 
@@
 
-This visualizes the ‘pulse’ of active projects.
+This visualizes the 'pulse' of active projects.
 
@@
 
----
+------------------------------------------------------------------------
 
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed [<3%].
+Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed [\<3%].
 
-A “contribution”, in this case, is a commit, issue, comment applied by a user to a repository.
+A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.
 
@@
 ### Formation of collaborative development communities
+
 If we construct a network linking repositories by their common contributors, can we identify emerging networks of collaboration, or are individual projects still largely isolated?
@@
 
-To construct our graphs we need a table linking users to repos by collaborative activity.
-For this purpose commits, issues, and comments are all considered contributions.
+To construct our graphs we need a table linking users to repos by collaborative activity. For this purpose commits, issues, and comments are all considered contributions.
 
@@
 
-With this, we can construct our three graphs: repo–user, repo–repo (by common users), and user–user (by common repos).
+With this, we can construct our three graphs: repo--user, repo--repo (by common users), and user--user (by common repos).
 
-#### Repo–user
+#### Repo--user
 
@@
 
-#### Repo–repo
+#### Repo--repo
 
@@
 
-#### User–user
+#### User--user
 
@@
 
-The vast majority of repositories have 1–3 contributors, with only a few distinguished by an active and diverse developer base.
+The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base.
 
-Generally speaking, we believe that this is because people do not want to step on other people’s shoes by raising issues or intruding on other people’s projects.
+Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects.
 
-This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are ‘owned’ by individuals or communities, and how work should be iteratively improved upon.
+This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
 
@@
 
-Since most repositories use more than one language—if only one for the actual software, and another (e.g.
-HTML) for documentation—this data is presented as a table of languages and the number of bytes of code in that language detected in the repository.
+Since most repositories use more than one language---if only one for the actual software, and another (e.g.
+HTML) for documentation---this data is presented as a table of languages and the number of bytes of code in that language detected in the repository.
 
@@
 Looking at number of bytes puts C++ and the web languages (HTML, PHP, and Javascript) on top, but this is probably affected by the relative verboseness of these languages and the popularity of web-based documentation across implementation languages.
-The number of repositories is probably more reliable and shows the popularity of scientific programming languages—R and Python—followed by web languages and then compiled C and C++.
+The number of repositories is probably more reliable and shows the popularity of scientific programming languages---R and Python---followed by web languages and then compiled C and C++.
 

### 2023-01-30 analysis/paper.qmd
Switch to quarto

--- 2023-01-30 analysis/paper.Rmd
@@
 date: "`r Sys.Date()`"
-output: distill::distill_article
 ---
@@
 | Category                              | Kind      | Scope                                                                                                                                                                                                        |
-|:--------------------------------------|:----------|:-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
+|:-----------|:-----------|:------------------------------------------------|
 | Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
@@
 
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed [\<3%].
+Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
 
@@
 
-```{r plot-graph-repo-user}
+```{r plot-graph-user-user}
 # TODO

### 2023-01-30 analysis/paper.qmd
Restore variable lost in 4f60afd1fffcfef7a7713493de469bd1b59f6e8a

@@
 ```{r plot-tags-cats}
+detail_tags <- c("Instrumental Neutron activation analysis",
+                 "Harris Matrix",
+                 "aDNA Simulators",
+                 "Damage profilers",
+                 "Metagenomic classifers",
+                 "aDNA simulators",
+                 "aDNA damage-aware genotypers",
+                 "Low-coverage contamination estimation",
+                 "Low-coverage adapted population-genetics tools",
+                 "aDNA preprocessing",
+                 "aDNA aligner",
+                 "aDNA assembler",
+                 "aDNA toolkit")
+
 oarch %>%

### 2023-01-30 analysis/paper.qmd
Remove custom plot font
Restore variable lost in e4fb22d1b0f66a9eb0b3d91bcb5aee4c761eb3bc

@@
 theme_oarch <- function() theme(
-  text = element_text(family = "Lato")
 )
@@
 
+```{r lifespan-stats}
+oarch %>% 
+  unnest(gh_commits) %>% 
+  drop_na(datetime) %>% 
+  group_by(item_name) %>%
+  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
+  summarise(
+    category = first(category),
+    n_contribs = first(n_contribs),
+    n = n(),
+    first_commit = min(datetime),
+    last_commit = max(datetime),
+    lifespan = as.integer(last_commit - first_commit),
+    commit_rate = n / lifespan
+  ) ->
+  oarch_lifespan
+```
+
 ```{r plot-lifespan-hist}

### 2023-01-30 analysis/paper.qmd
Fix order of plot-gh_commits-cum

@@
   mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Main contributor", "Other contributors"),
-         big_boy_pct = max(contributions) / sum(contributions)) %>%
-  ggplot(aes(x = fct_reorder(item_name, big_boy_pct, .desc = FALSE),
-             y = contributions,
-             # group = contributor,
+                          "Leading contributor", "Other contributors")) %>% 
+  group_by(item_name, big_boy) %>%
+  summarise(contributions = sum(contributions)) %>%
+  mutate(
+    p_contributions = contributions / sum(contributions),
+    p_big_boy = first(p_contributions)
+  ) %>%
+  ungroup() %>%
+  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
+  ggplot(aes(x = item_name, y = contributions,
              fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Main contributor")))) +
+                                           "Leading contributor")))) +
   geom_col(position = "fill", colour = "white", size = 0.1) +
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
   scale_y_continuous(labels = scales::percent) +

### 2023-01-30 analysis/paper.qmd
Fix plot-lifespan-activity-cats

@@
 #TODO: Eliminate Infs
-
-left_join(oarch_lifespan, oarch, by = c("item_name")) %>%
+oarch_lifespan %>% 
+  select(-category) %>% 
+  left_join(oarch, by = c("item_name")) %>%
   drop_na(category) %>%

### 2023-01-30 analysis/paper.qmd
Add missing dependency

@@
 library("ggraph")
+library("ggiraph")
 ```

### 2023-01-30 analysis/paper.qmd
Remove scrapped figures and reknit

@@
 	message = FALSE,
-	warning = FALSE,
-	layout = "l-body-outset"
+	warning = FALSE
 )
@@
 ```{r plot-repo-hosts}
+# TODO: turn into table or just text
 oarch <- mutate(oarch, none = ifelse(is.na(github) &
@@
 # ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
+```
+
+```{r table-no-repo}
+# TODO: prettify with gt()
+filter(oarch, none == "") %>%
+  count(category)
 ```
@@
 We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.
-
-# Scrap
-
-The GitHub API provides data on the languages used in a repository based on an automated machine learning algorithm.
-
-Since most repositories use more than one language---if only one for the actual software, and another (e.g.
-HTML) for documentation---this data is presented as a table of languages and the number of bytes of code in that language detected in the repository.
-
-We can therefore measure the popularity of languages across repositories in a couple of ways: based on how many repositories use a language at all; or based on how many bytes of that language are present.
-
-Looking at number of bytes puts C++ and the web languages (HTML, PHP, and Javascript) on top, but this is probably affected by the relative verboseness of these languages and the popularity of web-based documentation across implementation languages.
-The number of repositories is probably more reliable and shows the popularity of scientific programming languages---R and Python---followed by web languages and then compiled C and C++.
-
-This largely confirms the pattern seen in our manual classification of the 'platform' of each project.
-
-```{r plot-gh_langs, fig.show="hold", out.width="50%"}
-oarch %>%
-  unnest(gh_langs) %>%
-  drop_na(lang, bytes) %>%
-  group_by(lang) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  arrange(-n) %>%
-  slice(1:10) %>%
-  ggplot(aes(fct_reorder(lang, n, .desc = TRUE), n)) +
-  geom_col() +
-  labs(title = "Repository languages",
-       subtitle = "Number of repositories using top 10 languages",
-       caption = "Data: open-archaeo.info and GitHub API",
-       x = NULL, y = "Number of repositories")
-
-oarch %>%
-  unnest(gh_langs) %>%
-  drop_na(lang, bytes) %>%
-  group_by(lang) %>%
-  summarise(bytes = sum(bytes), .groups = "drop") %>%
-  arrange(-bytes) %>%
-  slice(1:10) %>%
-  ggplot(aes(fct_reorder(lang, bytes, .desc = TRUE), bytes)) +
-  geom_col() +
-  labs(title = "Repository languages",
-       subtitle = "Total bytes of code in top 10 languages",
-       caption = "Data: open-archaeo.info and GitHub API",
-       x = NULL, y = "Bytes")
-```
-
-------------------------------------------------------------------------
-
-About one in eight projects do not use one of these repository host.
-
-Some may use a different host, or not make their repositories publicly available, but it is probably safe to assume that the majority simply do not use version control.
-
-Many of these are standalone documents or code fragments (scripts) hosted, for example, as 'gists', but there are also some pieces of standalone software that apparently do not include version control.
-
-```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
-```
-
-------------------------------------------------------------------------
-
-The measures of activity may also be compared with the tags and categories.
-
-```{r plot-lifespan-activity-tags}
-# TODO: where does oarch_lifespan come from?
-left_join(oarch_lifespan, oarch, by = c("item_name")) %>%
-  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  group_by(tag) %>%
-  summarise(avg_lifespan = mean(lifespan),
-            avg_commit_rate = mean(commit_rate)) ->
-  oarch_lifespan_by_tag
-
-oarch_lifespan_by_tag %>%
-  arrange(-avg_lifespan) %>%
-  slice(1:10) %>%
-  ggplot(aes(x = fct_reorder(tag, avg_lifespan, .desc = TRUE),
-             y = avg_lifespan)) +
-  geom_col() +
-  labs(
-    title = "Project lifespan by tag",
-    subtitle = "Top 10 longest-lived tags",
-    x = NULL, y = "Average lifespan (days)",
-    caption = "Data: open-archaeo.info and GitHub API"
-  )
-
-oarch_lifespan_by_tag %>%
-  arrange(-avg_commit_rate) %>%
-  slice(1:10) %>%
-  ggplot(aes(x = fct_reorder(tag, avg_commit_rate, .desc = TRUE),
-             y = avg_commit_rate)) +
-  geom_col() +
-  labs(
-    title = "Project activity by tag",
-    subtitle = "Top 10 most activity tags",
-    x = NULL, y = "Average activity (commits per day)",
-    caption = "Data: open-archaeo.info and GitHub API"
-  )
-```
-
-```{r plot-lifespan-activity-cats}
-#TODO: Eliminate Infs
-oarch_lifespan %>% 
-  select(-category) %>% 
-  left_join(oarch, by = c("item_name")) %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(avg_lifespan = mean(lifespan),
-            avg_commit_rate = mean(commit_rate)) ->
-  oarch_lifespan_by_cat
-
-oarch_lifespan_by_cat %>%
-  arrange(-avg_lifespan) %>%
-  slice(1:10) %>%
-  ggplot(aes(x = fct_reorder(category, avg_lifespan, .desc = TRUE),
-             y = avg_lifespan)) +
-  geom_col() +
-  labs(
-    title = "Project lifespan by category",
-    subtitle = "Top 10 longest-lived categories",
-    x = NULL, y = "Average lifespan (days)",
-    caption = "Data: open-archaeo.info and GitHub API"
-  )
-
-oarch_lifespan_by_cat %>%
-  arrange(-avg_commit_rate) %>%
-  slice(1:10) %>%
-  ggplot(aes(x = fct_reorder(category, avg_commit_rate, .desc = TRUE),
-             y = avg_commit_rate)) +
-  geom_col() +
-  labs(
-    title = "Project activity by categories",
-    subtitle = "Top 10 most activity categories",
-    x = NULL, y = "Average activity (commits per day)",
-    caption = "Data: open-archaeo.info and GitHub API"
-  )
-```

### 2023-01-30 analysis/paper.qmd
Update my affiliation

@@
     url: https://joeroe.io
-    affiliation: University of Copenhagen
+    affiliation: University of Bern
 date: "`r Sys.Date()`"

### 2023-01-30 analysis/paper.qmd
Add user-user graph

@@
 oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
+                                   edges_user_to_repo(oarch_graph),
+                                   directed = FALSE)
 
@@
 ```{r plot-graph-user-user}
-# TODO
+oarch_graph_user_user <- tbl_graph(nodes_users(oarch_graph), 
+                                   edges_common_repos(oarch_graph))
+
+oarch_graph_user_user |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 0) |>
+  ggraph(layout = "graphopt") +
+  geom_edge_link(aes(alpha = n)) +
+  geom_node_point(aes(colour = category)) +
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common repositories")
 ```

### 2023-01-30 R/graph.R

@@ -15,7 +15,9 @@ nodes_repos <- function(oarch_graph) {
 
 #' @export
 nodes_users <- function(oarch_graph) {
-
+  oarch_graph |>
+    select(user, category) |>
+    distinct()
 }
 
 #' @export
@@ -42,4 +44,18 @@ edges_common_users <- function(oarch_graph) {
     select(-repo.xy) |>
     count(repo.x, repo.y) |>
     rename(to = repo.x, from = repo.y)
+}
+
+#' @export
+edges_common_repos <- function(oarch_graph) {
+  oarch_graph |>
+    left_join(oarch_graph, by = "repo") |>
+    select(user.x, repo, user.y) |>
+    filter(user.x != user.y) |>
+    mutate(user.xy = map2_chr(user.x, user.y,
+                              ~paste0(sort(c(.x, .y)), collapse = ""))) |>
+    distinct(repo, user.xy, .keep_all = TRUE) |>
+    select(-user.xy) |>
+    count(user.x, user.y) |>
+    rename(to = user.x, from = user.y)
 }

### 2023-01-30 analysis/paper.qmd
Add dots to nodes_repos and nodes_users

@@
 ```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
                                    edges_common_users(oarch_graph))
@@
 ```{r plot-graph-user-user}
-oarch_graph_user_user <- tbl_graph(nodes_users(oarch_graph), 
+oarch_graph_user_user <- tbl_graph(nodes_users(oarch_graph, category), 
                                    edges_common_repos(oarch_graph))

### 2023-01-30 R/graph.R

@@ -7,16 +7,16 @@ nodes_repos_users <- function(oarch_graph) {
 }
 
 #' @export
-nodes_repos <- function(oarch_graph) {
+nodes_repos <- function(oarch_graph, ...) {
   oarch_graph |>
-    select(repo, category) |>
+    select(repo, ...) |>
     distinct()
 }
 
 #' @export
-nodes_users <- function(oarch_graph) {
+nodes_users <- function(oarch_graph, ...) {
   oarch_graph |>
-    select(user, category) |>
+    select(user, ...) |>
     distinct()
 }
 

### 2023-01-30 analysis/paper.qmd
Community detection visuals

@@
 ```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category, tags),
                                    edges_common_users(oarch_graph))
@@
 
-```{r iplot-graph-repo-repo}
+```{r plot-graph-repo-repo-edge-betweenness}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
+                                   edges_common_users(oarch_graph),
+                                   directed = FALSE)
+
+oarch_graph_repo_repo |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 1) |>
+  mutate(group = as.character(group_edge_betweenness())) ->
+  oarch_graph_repo_repo
+
 iplot <- oarch_graph_repo_repo |>
@@
   geom_edge_link(aes(alpha = n)) +
-    geom_point_interactive(aes(x, y, colour = category, tooltip = repo)) + 
+  #geom_node_point(aes(colour = group)) +
+  #geom_node_text(aes(label = group)) +
+  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
   scale_edge_alpha(trans = "log") +
-    labs(edge_alpha = "Common contributors") +
-    theme_graph()
+  labs(edge_alpha = "Common contributors")
 
 girafe(ggobj = iplot)
+```
+
+```{r plot-graph-repo-repo-edge-betweenness-dendrogram}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
+                                   edges_common_users(oarch_graph),
+                                   directed = TRUE)
+
+oarch_graph_repo_repo |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 1) |>
+  mutate(group = as.character(group_edge_betweenness())) ->
+  oarch_graph_repo_repo
+
+xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
+xx <- as.dendrogram(xx)
+plot(xx)
+
+ggraph(xx, 'dendrogram', height = height)
+
+```
+
+```{r plot-graph-repo-repo-louvain}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
+                                   edges_common_users(oarch_graph),
+                                   directed = FALSE)
+
+oarch_graph_repo_repo |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 1) |>
+  mutate(group = as.character(group_louvain())) ->
+  oarch_graph_repo_repo
+
+oarch_graph_repo_repo |>
+  ggraph(layout = "graphopt") +
+  geom_edge_link(aes(alpha = n)) +
+  geom_node_point(aes(colour = group)) +
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common contributors")
 ```

### 2023-01-30 analysis/paper.qmd
added notes from end of day

@@
 We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.
+
+---
+
+#### Notes from January 30, end of day
+In cleaning the data we made a few significant decisions that revealed insights into our dataset. 
+We constructed the user-user graph and filtered out several bots. 
+We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community. 
+We filtered isolates. 
+We calculated the degree and betweenness to assess whether this was realistic representation of what we know. 
+We wanted to figure out who are the key nodes, what they had in common. 
+The people who are high betweenness are people who primarily produce computational archaeology code as a job. 
+We can investgate further by comparing with some other measures of activity, like volume of commits or something else. 
+We also looked at the same thing, but with only commits, or without commits. 
+For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.

### 2023-01-30 analysis/paper.qmd
Fix conflicts

@@
 
+Q's:
+
+1. Are there significant networks? If so, how big and how many?
+  * Only one, it seems!
+2. What proportion of users and repos are connected to this network?
+3. What predicts membership in this network?
+4. Does centrality in the network correlate with general rate of activity?
+5. Who are the most central nodes in this network? Is there anything they have in common?
+   * High activity?
+   * Certain platforms/categories/languages?
+
 ```{r graph-data}
@@
   rename(user = author) |>
-  filter(!str_detect(user, coll("[bot]"))) ->
+  # Remove non-human users
+  filter(!str_detect(user, coll("[bot]"))) |>
+  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
+                      "gitter-badger", "dependabot-support")) |>
+  # Remove big, non-archaeology projects that flood the network
+  filter(!repo %in% c("Stellarium/stellarium", 
+                     "opengisch/QField",
+                     "lutraconsulting/input")) ->
   oarch_graph
@@
 oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph),
-                                   directed = FALSE)
+                                   edges_user_to_repo(oarch_graph))
 
-oarch_graph_repo_user |>
+g_repo_user <- oarch_graph_repo_user |>
   ggraph(layout = "graphopt") +
@@
   scale_edge_alpha(trans = "log")
+
+#g_repo_user
+girafe(g_repo_user)
 ```
@@
 ```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category, tags),
-                                   edges_common_users(oarch_graph))
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
+                                   edges_common_users(oarch_graph),
+                                   directed = FALSE)
 
@@
 
+```{r graph-user-user}
+oarch_graph_user_user <- tbl_graph(
+  nodes = nodes_users(oarch_graph), 
+  edges = edges_common_repos(oarch_graph),
+  directed = FALSE
+)
+```
+
+```{r summary-graph-user-user}
+oarch_graph_user_user |>
+  mutate(degree = centrality_degree(), 
+         betweenness = centrality_betweenness(weights = n), 
+         strength = centrality_degree(weights = n)) |>
+  filter(degree > 0) |> 
+  data.frame() |> 
+  View()
+```
+
 ```{r plot-graph-user-user}
-oarch_graph_user_user <- tbl_graph(nodes_users(oarch_graph, category), 
-                                   edges_common_repos(oarch_graph))
-
 oarch_graph_user_user |>
-  mutate(degree = centrality_degree()) |>
+  mutate(degree = centrality_degree(), strength = centrality_degree(weights = n)) |>
   filter(degree > 0) |>
@@
   geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
+  # geom_node_point(aes(colour = category)) +
   scale_edge_alpha(trans = "log") +

### 2023-01-31 analysis/paper.qmd
Add analysis of betweenness vs. activity

@@
 ```{r dependencies, include=FALSE}
-library("tidyverse")
+library("fs")
+library("geomtextpath")
+library("ggiraph")
+library("ggraph")
+library("gt")
+library("here")
+library("khroma")
 library("lubridate")
 library("openarchaeoCollaboration")
-library("khroma")
-library("gt")
-library("fs")
-library("here")
+library("patchwork")
 library("tidygraph")
-library("ggraph")
-library("ggiraph")
+library("tidyverse")
 ```
@@
 
+### Who's doing what?
+
+```{r user-contributors}
+oarch |>
+  unnest(gh_contribs) |>
+  drop_na(contributor) |>
+  group_by(contributor) |>
+  mutate(
+    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
+  ) |>
+  group_by(contributor, repo_owner) |>
+  summarise(repos = n(), contribs = sum(contributions)) |>
+  pivot_wider(c(contributor), names_from = repo_owner, 
+              values_from = c(repos, contribs), values_fill = 0) |>
+  mutate(
+    repos = repos_other + repos_own,
+    contribs = contribs_other + contribs_own,
+    .after = contributor
+  ) ->
+  oarch_contributors
+```
+
 ### Development processes
@@
 
-```{r plot-gh_commits-cum}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  mutate(lumped_category = recode(category,
+```{r plot-gh-cum}
+oarch %>% 
+  mutate(
+    lumped_category = recode(category,
                              "Specifications etc." = "Documents",
@@
                              "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
+                             .default = category),
+    lumped_category = factor(lumped_category, c("Scripts",
+                                                "Documents",
+                                                "Packages and libraries",
+                                                "Standalone software")) ) ->
+  oarch_lumped
+
+p_gh_commits_cum <- oarch_lumped %>%
+  unnest(gh_commits) %>%
+  drop_na(sha) %>%
+  mutate(date = floor_date(datetime, "week")) %>%
+  group_by(lumped_category, date) %>%
+  summarise(n_commits = n(), .groups = "drop_last") %>%
+  mutate(cum_commits = cumsum(n_commits)) %>%
+  ggplot(aes(date, cum_commits,
+             fill = lumped_category,
+             colour = lumped_category,
+             label = lumped_category)) +
+  geom_area() +
+  annotate("segment", 
+           x = ymd("2008-04-01", tz = "UTC"), 
+           xend = ymd("2008-04-01", tz = "UTC"),
+           y = 16000, yend = 4000,
+           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
+  annotate("text", 
+           x = ymd("2008-04-01", tz = "UTC"), y = 19000,
+           label = "GitHub\nlaunched",
+           vjust = 0) +
+  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
+  scale_y_continuous(labels = scales::comma) +
+  scale_colour_bright(guide = guide_none()) +
+  scale_fill_bright() +
+  labs(
+    y = "cumulative commits",
+    x = NULL, fill = NULL
+  ) +
+  theme_oarch() +
+  theme(legend.position = "top")
+
+# TODO: extend to end
+p_gh_repos_cum <- oarch_lumped %>%
+  unnest(gh_commits) %>%
+  drop_na(sha) %>%
+  group_by(repo) %>% 
+  filter(datetime == min(datetime)) %>% # Only first commits
+  mutate(date = floor_date(datetime, "week")) %>%
   group_by(lumped_category, date) %>%
@@
   ggplot(aes(date, cum_commits,
-             fill = fct_reorder(lumped_category, cat_total),
-             colour = fct_reorder(lumped_category, cat_total))) +
+             fill = lumped_category,
+             colour = lumped_category)) +
   geom_area() +
-  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years",
-                   date_labels = "%Y") +
+  annotate("segment",
+           x = ymd("2008-04-01", tz = "UTC"),
+           xend = ymd("2008-04-01", tz = "UTC"),
+           y = 40, yend = 10,
+           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
+  annotate("text",
+           x = ymd("2008-04-01", tz = "UTC"), y = 50,
+           label = "GitHub\nlaunched",
+           vjust = 0) +
+  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
   scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = NULL) +
-  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
-  labs(
-    title = "Growth of open archaeology",
-    subtitle = "Measured by cumulative commits in GitHub repositories",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = NULL, y = NULL, fill = NULL
+  scale_colour_bright(guide = guide_none()) +
+  scale_fill_bright(guide = guide_none()) +
+  labs(
+    y = "cumulative repositories",
+    x = NULL, fill = NULL
   ) +
   theme_oarch()
+
+p_gh_commits_cum / p_gh_repos_cum
 
@@
   filter(degree > 0) |> 
-  data.frame() |> 
-  View()
+  data.frame()
 ```
@@
 oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |>
+  mutate(betweenness = centrality_betweenness(weights = n)) ->
+  oarch_graph_user_user
+
+oarch_graph_user_user |> 
   ggraph(layout = "graphopt") +
   geom_edge_link(aes(alpha = n)) +
-  # geom_node_point(aes(colour = category)) +
+  # geom_node_point() +
   scale_edge_alpha(trans = "log") +
   labs(edge_alpha = "Common repositories")
+```
+
+```{r plot-user-betweenness-dist}
+oarch_graph_user_user |>
+  mutate(
+    betweenness = centrality_betweenness(weights = n)
+  ) |>
+  data.frame() |>
+  ggplot(aes(x = betweenness)) +
+  geom_density(bw = "nrd") +
+  scale_x_log10(labels = scales::number)
+```
+
+```{r plot-user-betweenness-activity}
+oarch_graph_user_user |>
+  mutate(
+    degree = centrality_degree(), 
+    betweenness = centrality_betweenness(weights = n), 
+    strength = centrality_degree(weights = n)
+  ) |>
+  as_tibble() |> 
+  left_join(oarch_contributors, by = c("user" = "contributor")) |>
+  ggplot(aes(betweenness, contribs)) +
+  geom_point() +
+  scale_x_log10() +
+  scale_y_log10() +
+  geom_smooth(method = "lm")
 ```

### 2023-01-31 analysis/paper.qmd
Edits to the primary text

--- 2023-01-30 analysis/paper.qmd
@@
 (2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
-
-We examine archaeological software development as reflections of a unique set of cultural practices.
 
@@
 
-Just to clarify, the two rows near the bottom with what seem like very long yellow bars are actually projects that have equal number of contributions made by the top contributors.
+**ZB NOTE:** Re-calculate these numbers to ensure they are up to date.
+Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic.
 
@@
 Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
+
+**ZB NOTE:** Recalcute these numbers.
 
@@
 
-If we construct a network linking repositories by their common contributors, can we identify emerging networks of collaboration, or are individual projects still largely isolated?
-If there are networks, what characterises their links?
+We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
 
-**ZB NOTE:** I'm speculating that there are going to be various users who have their fingers in many projects, and that those projects tend to be the most active.
-If this is the case, I would further speculate that these people hold a different attitude regarding authorship and the dividuality of projects that is more in sync with the collaborative capabilities of open source, which are at odds with traditional academic norms regarding how contributions are credited and how products are produced are disseminated.
-
-To construct our graphs we need a table linking users to repos by collaborative activity. For this purpose commits, issues, and comments are all considered contributions.
-
-Q's:
-
-1. Are there significant networks? If so, how big and how many?
-  * Only one, it seems!
-2. What proportion of users and repos are connected to this network?
-3. What predicts membership in this network?
-4. Does centrality in the network correlate with general rate of activity?
-5. Who are the most central nodes in this network? Is there anything they have in common?
-   * High activity?
-   * Certain platforms/categories/languages?
+We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
 
@@
 
-With this, we can construct our three graphs: repo--user, repo--repo (by common users), and user--user (by common repos).
+We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations.
+We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
 
@@
 #### Repo--repo
+
+We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
+Aside from isolate notes, which represent repositories with only a single contributor, we detected 21 clusters.
+While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
+
+Are the clusters associated with particular topics or tags?
+Are there some sorts of associations between clusters that break down from a parent in the community detection dendrogram?
+How does all of this differ when we differentiate commits from issues and comments?
+
+**ZB NOTE:** Remove the cells titled `plot-graph-repo-repo` and `plot-graph-repo-repo-louvain`.
+Use patchwork to compare sets of graphs and dendrograms with links based on all contributions, only commits, and only issues and comments.
 
@@
 
+We assembled graphs linking users based on common contributions to the same repositories.
+We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
+
+We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as a job.
+
+**ZB NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits or something else.
+
+We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
+For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the all-contributions graph also appear here.
+These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
+
+**ZB NOTE:** Look at the histogram that maps the distribution of betweenness scores across all users.
+Interesting double hump of low-moderate betweenness scores, which may correspond with distinct or protruding components.
+
 ```{r graph-user-user}

### 2023-01-31 analysis/paper.qmd

--- 2023-01-31 analysis/paper.qmd
@@
 
+**ZB NOTE:** Recalcute these numbers.
+
 However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
@@
 
-If we construct a network linking repositories by their common contributors, can we identify emerging networks of collaboration, or are individual projects still largely isolated?
-If there are networks, what characterises their links?
-
-**ZB NOTE:** I'm speculating that there are going to be various users who have their fingers in many projects, and that those projects tend to be the most active.
-If this is the case, I would further speculate that these people hold a different attitude regarding authorship and the dividuality of projects that is more in sync with the collaborative capabilities of open source, which are at odds with traditional academic norms regarding how contributions are credited and how products are produced are disseminated.
-
-To construct our graphs we need a table linking users to repos by collaborative activity. For this purpose commits, issues, and comments are all considered contributions.
-
-Q's:
-
-1. Are there significant networks? If so, how big and how many?
-  * Only one, it seems!
-2. What proportion of users and repos are connected to this network?
-3. What predicts membership in this network?
-4. Does centrality in the network correlate with general rate of activity?
-5. Who are the most central nodes in this network? Is there anything they have in common?
-   * High activity?
-   * Certain platforms/categories/languages?
+We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
+
+We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
 
@@
 #### User--user
+
+We assembled graphs linking users based on common contributions to the same repositories.
+We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
+
+We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as a job.
+
+**ZB NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits or something else.
+
+We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
+For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the all-contributions graph also appear here.
+These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
+
+**ZB NOTE:** Look at the histogram that maps the distribution of betweenness scores across all users.
+Interesting double hump of low-moderate betweenness scores, which may correspond with distinct or protruding components.
 

### 2023-01-31 analysis/paper.qmd
Move frontmatter

@@
----
-title: "openarchaeoCollaboration"
-author:
-  - name: Zack Batist
-    url: https://zackbatist.github.io
-    affiliation: University of Toronto
-  - name: Joe Roe
-    url: https://joeroe.io
-    affiliation: University of Bern
-date: "`r Sys.Date()`"
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	message = FALSE,
-	warning = FALSE
-)
-```
-
-```{r dependencies, include=FALSE}
-library("fs")
-library("geomtextpath")
-library("ggiraph")
-library("ggraph")
-library("gt")
-library("here")
-library("khroma")
-library("lubridate")
-library("openarchaeoCollaboration")
-library("patchwork")
-library("tidygraph")
-library("tidyverse")
-```
-
-```{r plot-theme}
-theme_oarch <- function() theme(
-)
-```
-
 ```{r data-oarch}

### 2023-03-27 analysis/98-scrap.qmd
Created a file for abandoned words.
Initially containts work-in-progress from earlier un-committed iterations of the introduction, and I may draw from this as I continue work on the background section.

The ability to write code is of course fundamental to effective use of digital tools and media. However, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are underappreciated aspects of software development. This is particularly true pertaining to open source projects, which entail publishing source code on the web for free, maintaining transparent programming practices, and inviting community participation and feedback. Open source projects, which have been heralded as revolutionary, community-oriented, and anti-capitalist means of production, literally run the web (and, by extension, the modern global economy), but they are not the asocial, meritocratic systems that many proponents claim they are. In fact, open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants. They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer  maintainers involved, and generally ensure that the project can be sustained over the long-term. Open source is more than just putting your code online; when done right, it is a social experience.

The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings. Open science advocates have put a lot of effort into developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale, but, at least among archaeologists, their transformative value has not yet been realized. For instance, numerous studies have highlighted how archaeologists are not reusing data published on open data platforms, and instead opt to communicate directly with those who created the dataset to obtain a relationship that more closely resembles a more traditional form of collaboration. In other words, these pltforms do little to alter the professional norms and expectations that frame archaeological research. They provide the capacity and the potential to enable change, but do not in themselves change the ways in which archaeologists create and value research outcomes. This must come from community buy-in, and the extension and establishing of novel forms of collaborative commitments that reflect the ideal state of affairs. In other words, it is a social problem, not a technical one.

---

Archaeologists have been developing specialized software since the 1960s, when computers were first introduced in the social science (cf. Cowgill). The popular adoption of personal computers has contributed to the growth of digital archaeology as a distinct field of practice. Although computer programming is not generally taught in archaeological curricula, archaeologists are generally eager to apply novel technologies to address their research questions.

Software development entails designing, programming, distributing, and maintaining software, and is a key aspect of contemporary digital archaeology.

* There is open archaeology
The open science movement has emerged as a series of professional practices that are aimed at making research more accessible and transparent.

The term 'open' is somewhat nebulous, but a generaly agreed upon definition defines open as "A piece of content or data is open if anyone is free to use, reuse, and redistribute it – subject only, at most, to the requirement to attribute and/or share-alike."

This definition of open does not state what open is for, or what motivates the open science movement. However, a brief reading of open science manifestos leads us to believe that open is about a few things: facilitating novel research opportunities, making participation in scienctific research more equitable, reclaiming science as a public good, and enhancing validation of findings.

A lot of work has gone in to developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale. Open data repositories, open peer-review platforms, and online social spaces have been set up to foster participation in new forms of scholarly communication.

- By making research materials and processes more accessible and transparent, it is believed that scientific knowledge production can be ehanhced.
- In part, this is through the increased capacity to use these materials, i.e. it is an efficiency concern. Simply having these materials available is a fundamental requirement to facilitate their use. It is about reducing barriers, in terms of information transmission and legal accomodations.
- As such, a lot of attention goes into developing technical protocols and standards to facilitate data sharing and reuse
- But, as some studies have noted, these infrastructures often fail to deliver on these promises, and casual observation indicates how people value connections that these platforms fail to support
- These issues centre on the inability to account for social and professional expectations
- Either the professional culture has not cought up with the technical infrastructures (which impose novel sociotechnical arrangeemnts), or the infrastructures serve needs that archaeologists do not feel they actually need
- What is the point of establishing these infrastructures, and in making research materials available, if no one is going to use them, or use them in a way that actually serves their intended purpose?

---

For instance, the Linux kernel is generally dependent on a coterie of senior developers, who are financially supported by the Linux Foundation, who delegate and vet contributions made by unpaid volunteers. Simularly, the Firefox web browser depends on funding from Alphabet, the corporate entity which dominates the space in which they operate, to stay afloat.

Open source software development models are crucial for maintaining key software that runs the internet (and, by extension, the modern global economy), and have been generally posed as revolutionary, community-oriented, anti-capitalist means of production. But even the most prominent open source projects contain bottlenecks that disrupt these idealistic imaginaries of heterarchical organizational structures, meritocratic decision-making, and potential for multivocal and community-led growth. For instance, the Linux kernel is generally dependent on a coterie of senior developers who delegate and vet contributions (cf. Cyberchiefs), the Firefox web browser is dependent on funding from the corporate entity which dominates the space in which they operate, and the massive distributed web of libraries and extensions that make up various software development ecosystems (such as JavaScript, Python and R) struggles to keep maintainers interested in continuing to provide support on a volunteer basis (as illustrated by the left-pad incident). Open source is not a panacea for institutional inertia and inequity, and actually depends on social relations and governance strategies to remain sustainable (Cyber chiefs, two bits, Matt's dissertation).

What does this mean for the adoption of open source ideas and practices among archaeologists? Archaeologists work in a very different institutional space than these very prominent projects, and are influenced by different profesional pressures, such as the warrant to publish, inadequate access to long-term research funding, the obligation to train students, the need to work as part of university systems with intricate institutional budgeting practices (which pit likeminded teams of researchers against each other), and the need to move on from projects after producing stable research outputs. At first glance, many of these aspects of being an archaeologist are in opposition with open source strategies. These tensions lead us to wonder how archaeologists actually implement open source. Beyond simply making their code available on the web, do they also implement the social strategies that make open source productive? Does archaeological open source actually help achieve greater transparency, sustainability, and community participation? And if not, what does it actually achieve?

### 2023-03-28 analysis/03-data_methodology.qmd
Initial text for the data and methodology section

* open-archaeo

Our findings are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources. As of right now it includes 408 items, which are nicely rendered at open-archaeo.info. The dataset was compiled by browsing collaborative software development platforms. This entailed manually crawling through the profiles of users who identify as archaeologists or who have contributed to tools that pertain to archaeological work. This quasi-systematic collection strategy is also supplemented by contributions made by interested individuals who identified relevant work that was initially overlooked. The main criteria for inclusion is whether a project is identified by its maintainers as relating to archaeological work.

Open-archaeo is a relatively comprehensive list, however it generally lacks code written before the growth and general use of collaborative software development platforms, such as GitHub, among archaeologists. It also lacks references to code that is not available on the web. Open-archaeo is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their field, to contribute.

* GitHub API
* Indicate how much is on GitHub or elsewhere

For the purpose of this study, we augmented the core open-archaeo dataset with data obtained from GitHub's public API. That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh]. This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but amongst those that do, GitHub is by far the most popular host.

{r plot-repo-hosts}
{r table-no-repo}

The data and analysis code are available in our research compendium on GitHub, which is also indexed on Zenodo.

### 2023-03-28 analysis/04-what_is_open_archaeology.qmd
Copying info from paper.qmd and organizing it for more productive work

--- 2023-01-31 analysis/paper.qmd
@@
----
-title: "openarchaeoCollaboration"
-author:
-  - name: Zack Batist
-    url: https://zackbatist.github.io
-    affiliation: University of Toronto
-  - name: Joe Roe
-    url: https://joeroe.io
-    affiliation: University of Bern
-date: "`r Sys.Date()`"
----
+## Breakdown of tags and categories
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	message = FALSE,
-	warning = FALSE
-)
-```
-
-```{r dependencies, include=FALSE}
-library("tidyverse")
-library("lubridate")
-library("openarchaeoCollaboration")
-library("khroma")
-library("gt")
-library("fs")
-library("here")
-library("tidygraph")
-library("ggraph")
-library("ggiraph")
-```
-
-```{r plot-theme}
-theme_oarch <- function() theme(
-)
-```
-
-```{r data-oarch}
-read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
-  mutate(category = na_if(category, "?")) ->
-  oarch
-
-# Extract clean repos from GitHub URLs
-oarch %>%
-  mutate(
-    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
-
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
-```
-
-```{r data-github}
-# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
-# chunk will prefer to load it rather than make requests to the GitHub API.
-# Delete this file to fetch fresh data from the GitHub or modify 
-# `cached_gh_data` to read a different cached version.
-#
-# Note that fetching data from the GitHub API takes some time and is subject to
-# a rate limit:
-# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
-#
-# You will need to authenticate authenticate with GitHub credentials to get a
-# rate limit sufficient to obtain the required data, see 
-# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
-# for options on how to do this.
-cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
-if (!file_exists(cached_gh_data)) {
-  oarch <- mutate(oarch,
-                  gh_langs = map(oarch$gh_repo, gh_lang),
-                  gh_contribs = map(oarch$gh_repo, gh_contrib),
-                  gh_issues = map(oarch$gh_repo, gh_issue),
-                  gh_comments = map(oarch$gh_repo, gh_comment),
-                  gh_commits = map(oarch$gh_repo, gh_commit)
-  )
-  saveRDS(oarch, cached_gh_data)
-}
-oarch <- readRDS(cached_gh_data)
-```
-
-## Introduction
-
-In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools to serve our own unique purposes, which are not typically unanticipated by their original designers.
-In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways.
-For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase.
-However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems.
-In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes.
-
-This article presents a survey of archaeological software development with two goals in mind:
-
-(1) we identify *what* kinds of software archaeologists are making;
-(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
-
-Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
-
-## Background
-
--   The notion of the archaeological toolbox, and its inclusion of digital tools and skillsets
-    -   The use of off-the-shelf tools like MS Access, Excel, Agisoft PhotoScan, Dropbox, etc
-    -   Link to related factors:
-        -   Lack of experienced programmers among archaeologists
-        -   The built-in flexibility of these tools
-        -   Ease of administrative purchasing through university departments
-    -   What changed?
-        -   Software is expensive, shift to subscription models, locking into a dependency on a commercial service provider
-        -   Archaeologists are better at computer programming, digital literacy on the rise, digital archaeologists grows into its own legitimate sub-discipline
-        -   Open source is cool, everyone has a github profile (even if not really)
-            -   (JR:) People are presumably guided towards GitHub (or at least git) via other routes, because it's seen as the standard platform for 'open' digital work.
-For example, the literature on reproducible compendiums recommends GitHub, and almost all R development work is now done there.
-And once there, you're sort of forced into engaging in a (particular set of) open source development.
-I know there's controversy about how dominant this has become and people who say it's not 'true' (free and) open source, and perhaps we could add to that the question of whether it's the right model for archaeology.
--   About open source practices and their influence on archaeology
-    -   What is open source software and open source software development?
-        -   Free cost as an end-user
-        -   Free to manipulate and extend
-        -   Community-based support and development
-    -   Software as product vs software as collective process
-        -   Basically what distinguishes abandonware from thriving software ecosystems
-        -   Technical, legal and social notions of open
-            -   Technical: source code is open, use of a platform and systems that support software integration and implementations
-            -   Legal: open licenses, backing by the courts
-            -   Social: inviting community, meaningful contributions provided by users, common vision and values directing the project into future directions
-        -   Consider how these aspects of open have been implemented or prioritized by digital archaeologists in different ways
-            -   Technical: reproducible research, archiving and publishing
-            -   Legal: archiving and publishing, budgeting for projects
-            -   Social: ensuring that software is actually useful for general use and not just one-off implementation, coordinating to reduce redundant effort and not reinvent the wheel, writing documentation, making it fit within existing systems or infrastructures (i.e.
-allocation of credit, citing software directly versus articles that present software, sustainable development after funding runs out, etc)
-        -   This paper largely focuses on the social aspect of open, which is arguably the most important for ensuring a positive future for archaeological software development
-
-The 'archaeological toolbox' is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely upon in their work.
-
-collaborate in their programming work, and how they experience open source development processes which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.
-
-## Data and evidence
-
-Our findings here are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
-As of right now it includes 408 items, which are nicely rendered at open-archaeo.info.
-
-The main criteria for inclusion is whether a project is identified by its maintainers as relating to archaeological work.
-
-The dataset was compiled by browsing collaborative software development platforms, essentially manually crawling through the profiles of users who identify as archaeologists or who have contributed to tools that pertain to archaeological work.
-
-This quasi-systematic collection strategy is also supplemented by contributions made by interested individuals who identified relevant work that was initially overlooked.
-
-Open-archaeo is a relatively comprehensive list, however it generally lacks code written before the growth and general use of collaborative software development platforms, such as GitHub, among archaeologists, and also lacks references to code that is not available on the web.
-
-It is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in your field, to contribute.
-
-Please get in touch if you're interested in adding to the list or doing a deep dive relating to your area of expertise!
-
-### Tags and categories
-
-Each record is also annotated with tags that describe the aspects of archaeological work that each tool contributes to.
-
+Each record is annotated with tags that describe the aspects of archaeological work that each tool contributes to.
 Items are tagged based on how they have been generally identified or presented by their maintainers.
-
 The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.
-
 Right now there are 63 tags.
@@
 Records are also categorized based on how each tool or resource is accessed or used.
-
 This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-
 In other words, categories hint at some of the decisions that developers make regarding how they want their tools to be used and accessed.
@@
 
-Two kinds of categories have been defined: software and documents.
-
+We defined two kinds of categories: software and documents.
 Software comprise collections of instructions that dictate how a computer should operate, and documents are bounded sets of recorded information.
-
-Three categories of software and four categories of documents are described in the table.
-
+We define three categories of software and four categories of documents.
 Packages and libraries are sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-
 Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.
-
 Standalone software are tools that may be operated without needing to first access an underlying platform.
-
-And the third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
+The third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
 Things like code generalized sets of code snippets.
@@
 As for documents, specifications, protocols and schemas are formal data structures or frameworks intended to be used as models.
-
 Products are stable outcomes of creative work, loosely defined.
-
 Guides are educational resources or documented protocols meant to instruct readers how to apply relevant tools or techniques.
-
 And finally, lists or datasets, including open-archaeo itself, are series of consistently organized observations assembled with purpose.
 
-## Findings
-
-### What archaeologists are making
+## Breakdown of platforms and languages
 
@@
 
-------------------------------------------------------------------------
-
 Most projects included in open-archaeo are designed to be used as a package or library atop am existing 'platform' -- for example a package for a programming language like R, or a plugin for an application like QGIS.
-
 The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
-
 Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-
 A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
 
-```{r plot-categories}
-oarch <- mutate(oarch,
-                category = recode(category,
-                                  "Specifications, protocols and schemas" = "Specifications etc.",
-                                  .default = category))
-oarch %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(n = n()) %>%
-  ggplot(aes(fct_reorder(category, n), n)) +
-  geom_col() +
-  coord_flip() +
-  labs(
-    title = "Most popular categories",
-    subtitle = "By number of projects",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
+{r plot-categories}
 
 R is overwhelmingly dominant as a platform, closely followed by Python.
-
 Plugins for QGIS and mobile apps are also relatively popular.
-
 Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
 AutoCAD, ArcGIS); a number of lesser used programming languages; and a genre consisting of custom forms and spreadsheet templates.
-
 In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
@@
 
-```{r plot-platforms}
-oarch %>%
-  drop_na(platform) %>%
-  group_by(platform, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(
-    title = "Most popular platforms",
-    subtitle = "By number of projects, split by primary author",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
+{r plot-platforms}
 
 If we compare categories with tags, we can see the general domains that each kind of resource is designed to serve.
-
 We see that packages are fairly common across the board.
-
 Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
-
 These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.
 
-```{r plot-tags-cats}
-detail_tags <- c("Instrumental Neutron activation analysis",
-                 "Harris Matrix",
-                 "aDNA Simulators",
-                 "Damage profilers",
-                 "Metagenomic classifers",
-                 "aDNA simulators",
-                 "aDNA damage-aware genotypers",
-                 "Low-coverage contamination estimation",
-                 "Low-coverage adapted population-genetics tools",
-                 "aDNA preprocessing",
-                 "aDNA aligner",
-                 "aDNA assembler",
-                 "aDNA toolkit")
-
-oarch %>%
-  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  filter(!tag %in% detail_tags) %>%
-  mutate(category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(tag, category) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  # facet_wrap(vars(category)) +
-  geom_col() +
-  scale_fill_bright() +
-  coord_flip() +
-  labs(
-    title = "Tags by category",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = "Number of projects",
-    fill = NULL
-  ) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_text(size = 7)
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/tags-cats.png", width = 254, height = 127, units = "mm")
-```
-
-**ZB NOTE:** Do a temporal analysis to get a sense of when platforms, tags and categories have been gaining popularity.
-There may not be a need for extreme detail here, but simply noting the dates of first commits might be helpful.
-I'm imagining a stacked bar chart, with each bar representing a month or a year, and with the stacked portions of each bar representing a platform or category.
-
-### Development processes
-
-While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about *how* archaeologists develop software.
-
-The remainder of the paper is driven by a concern about how archaeologists collaborate in their programming work.
-
-How is their work sustained, and how does it leverage distributed work processes that are a key aspect of open source software development in general?
-
-To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
-
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh].
-
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.
-
-```{r plot-repo-hosts}
-# TODO: turn into table or just text
-oarch <- mutate(oarch, none = ifelse(is.na(github) &
-                                       is.na(gitlab) &
-                                       is.na(bitbucket) &
-                                       is.na(launchpad) &
-                                       is.na(codeberg),
-                                     "", NA))
-
-oarch %>%
-  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
-               names_to = "repo_host", values_to = "repo_url") %>%
-  drop_na(repo_url) %>%
-  group_by(repo_host, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ungroup() %>%
-  mutate(repo_host = fct_reorder(repo_host, total),
-         repo_host = fct_relevel(repo_host, "none")) %>%
-  ggplot(aes(repo_host, n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(title = "Most popular repository hosts",
-       subtitle = "By number of projects, split by primary author",
-       caption = "Data: open-archaeo.info",
-       x = NULL, y = NULL)
-
-# Export for slides (half slide)
-# ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
-```
-
-```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
-```
-
-We obtained this data in R using the {gh} package; those data, and the analysis code, are available in our research compendium on GitHub.
-
-------------------------------------------------------------------------
-
-Starting with a relatively simple plot, this is a chart of the cumulative growth of open archaeology over the years, measured by code contributions committed and pushed to GitHub repositories.
-
-As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
-But they really began to get into it around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
-
-Around this time we also see that GitHub starts being used to host documents and scripts.
-This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.
-
-```{r plot-gh_commits-cum}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  mutate(lumped_category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cat_total = sum(n_commits)) %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = fct_reorder(lumped_category, cat_total),
-             colour = fct_reorder(lumped_category, cat_total))) +
-  geom_area() +
-  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years",
-                   date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = NULL) +
-  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
-  labs(
-    title = "Growth of open archaeology",
-    subtitle = "Measured by cumulative commits in GitHub repositories",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = NULL, y = NULL, fill = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (full slide)
-# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-But how is this activity distributed across repositories?
-
-This visualizes the 'pulse' of active projects.
-
-Each vertical line represents the lifespan of a repository; the points on the line are individual commits.
-
-This is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
-
-```{r plot-lifespan}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(datetime) %>%
-  group_by(item_name) %>%
-  add_tally() %>%
-  filter(n > 1) %>%
-  mutate(first_commit = min(datetime),
-         mid_commit = median(datetime),
-         last_commit = max(datetime)) %>%
-  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
-             y = datetime,
-             ymin = first_commit,
-             ymax = last_commit)) +
-  geom_linerange(size = 0.05) +
-  geom_point(size = 0.1) +
-  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years", date_labels = "%Y") +
-  # coord_flip() +
-  labs(
-    title = "Project lifespan",
-    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
-    x = NULL, y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch() +
-  theme(
-    axis.line.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    panel.grid.major.x = element_blank()
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-The span of time between the first and latest commits is visualized more clearly in summary here.
-
-Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
-
-Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.
-
-Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects.
-
-The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e.
-the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.
-
-But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
-
-This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.
-
-**ZB NOTE:** Use patchwork package to combine these plots into a unified figure.
-
-```{r lifespan-stats}
-oarch %>% 
-  unnest(gh_commits) %>% 
-  drop_na(datetime) %>% 
-  group_by(item_name) %>%
-  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
-  summarise(
-    category = first(category),
-    n_contribs = first(n_contribs),
-    n = n(),
-    first_commit = min(datetime),
-    last_commit = max(datetime),
-    lifespan = as.integer(last_commit - first_commit),
-    commit_rate = n / lifespan
-  ) ->
-  oarch_lifespan
-```
-
-```{r plot-lifespan-hist}
-# TODO: where does oarch_lifespan come from?
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan)) +
-  geom_histogram(binwidth = 100) +
-  scale_x_continuous(breaks = scales::breaks_width(1000)) +
-  labs(
-    title = "Project lifespan",
-    subtitle = "Time between first and latest commit, by repository",
-    x = "Days", y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/lifespan-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-lifespan-vs-rate}
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = "Project lifespan vs. average commit rate",
-    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
-    caption = NULL
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-total.png", width = 127, height = 60, units = "mm")
-
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = NULL,
-    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-contrib.png", width = 127, height = 60, units = "mm")
-
-```
-
-------------------------------------------------------------------------
-
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
-
-A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.
-
-```{r plot-gh_contributors-histogram}
-oarch %>%
-  drop_na(gh_repo) %>%
-  unnest(gh_contribs) %>%
-  group_by(gh_repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Repository contributors",
-    x = "Number of contributors", y = "Repositories",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-hist.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
-
-Of multi-contributor repositories, 92% have a third of contributions made by a single person;
-
-65% have half of all contributions made by a single person;
-
-and 36% have three quarters of all contributions made a single person.
-
-**ZB NOTE:** Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic.
-
-```{r plot-contrib-dist}
-oarch %>%
-  unnest(gh_contribs) %>%
-  drop_na(contributor) %>%
-  group_by(item_name) %>%
-  mutate(n_contributors = n()) %>%
-  filter(n_contributors > 1) %>%
-  mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Leading contributor", "Other contributors")) %>% 
-  group_by(item_name, big_boy) %>%
-  summarise(contributions = sum(contributions)) %>%
-  mutate(
-    p_contributions = contributions / sum(contributions),
-    p_big_boy = first(p_contributions)
-  ) %>%
-  ungroup() %>%
-  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
-  ggplot(aes(x = item_name, y = contributions,
-             fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Leading contributor")))) +
-  geom_col(position = "fill", colour = "white", size = 0.1) +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
-  labs(title = "Repository contributors by proportion of contributions",
-       x = NULL, y = "Contributions",
-       fill = NULL,
-       caption = "Data: open-archaeo.info and GitHub API") +
-  coord_flip() +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-dist.png", width = 127, height = 127, units = "mm")
-```
-
-Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-
-Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-
-**ZB NOTE:** Recalcute these numbers.
-
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
-
-```{r plot-issues-hist}
-oarch %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of issues",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of issues", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/issues-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-comments-hist}
-oarch %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of comments",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of comments", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/comments-hist.png", width = 127, height = 127, units = "mm")
-```
-
-### Formation of collaborative development communities
-
-We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
-
-We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
-
-```{r graph-data}
-oarch |>
-  drop_na(gh_repo) |>
-  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
-  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
-  oarch_graph
-
-oarch_graph <- bind_rows(
-  oarch_graph |>
-    unnest(gh_commits) |>
-    drop_na(author) |> # Apparently real commits, but we can't connect them
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "commit", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_comments) |>
-    drop_na(author) |> # Repos with no comments
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "comment", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_issues) |>
-    drop_na(author) |> # Repos with no issues
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "issue", n_contrib = n(), .groups = "drop")
-)
-
-# Cleaning
-oarch_graph |>
-  rename(user = author) |>
-  # Remove non-human users
-  filter(!str_detect(user, coll("[bot]"))) |>
-  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
-                      "gitter-badger", "dependabot-support")) |>
-  # Remove big, non-archaeology projects that flood the network
-  filter(!repo %in% c("Stellarium/stellarium", 
-                     "opengisch/QField",
-                     "lutraconsulting/input")) ->
-  oarch_graph
-```
-
-We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations.
-We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
-
-#### Repo--user
-
-```{r plot-graph-repo-user}
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
-
-g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt") +
-  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
-  geom_node_point(aes(colour = node_type)) +
-  scale_edge_alpha(trans = "log")
-
-#g_repo_user
-girafe(g_repo_user)
-```
-
-#### Repo--repo
-
-We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate notes, which represent repositories with only a single contributor, we detected 21 clusters.
-While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
-
-Are the clusters associated with particular topics or tags?
-Are there some sorts of associations between clusters that break down from a parent in the community detection dendrogram?
-How does all of this differ when we differentiate commits from issues and comments?
-
-**ZB NOTE:** Remove the cells titled `plot-graph-repo-repo` and `plot-graph-repo-repo-louvain`.
-Use patchwork to compare sets of graphs and dendrograms with links based on all contributions, only commits, and only issues and comments.
-
-```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-```{r plot-graph-repo-repo-edge-betweenness}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  #geom_node_point(aes(colour = group)) +
-  #geom_node_text(aes(label = group)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-
-girafe(ggobj = iplot)
-```
-
-```{r plot-graph-repo-repo-edge-betweenness-dendrogram}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = TRUE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
-xx <- as.dendrogram(xx)
-plot(xx)
-
-ggraph(xx, 'dendrogram', height = height)
-
-```
-
-```{r plot-graph-repo-repo-louvain}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_louvain())) ->
-  oarch_graph_repo_repo
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = group)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-#### User--user
-
-We assembled graphs linking users based on common contributions to the same repositories.
-We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
-
-We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as a job.
-
-**ZB NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits or something else.
-
-We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
-For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the all-contributions graph also appear here.
-These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
-
-**ZB NOTE:** Look at the histogram that maps the distribution of betweenness scores across all users.
-Interesting double hump of low-moderate betweenness scores, which may correspond with distinct or protruding components.
-
-```{r graph-user-user}
-oarch_graph_user_user <- tbl_graph(
-  nodes = nodes_users(oarch_graph), 
-  edges = edges_common_repos(oarch_graph),
-  directed = FALSE
-)
-```
-
-```{r summary-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), 
-         betweenness = centrality_betweenness(weights = n), 
-         strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |> 
-  data.frame() |> 
-  View()
-```
-
-```{r plot-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  # geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common repositories")
-```
-
-#### Over time
-
-Could animate the graphs??
-
-## Conclusions
-
-Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.
-
-In terms of collaboration, our analysis shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host.
-
-The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base.
-
-Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects.
-
-This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
-
-We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.
+{r plot-tags-cats}
 
@@
 
-#### Notes from January 30, end of day
-In cleaning the data we made a few significant decisions that revealed insights into our dataset. 
-We constructed the user-user graph and filtered out several bots. 
-We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community. 
-We filtered isolates. 
-We calculated the degree and betweenness to assess whether this was realistic representation of what we know. 
-We wanted to figure out who are the key nodes, what they had in common. 
-The people who are high betweenness are people who primarily produce computational archaeology code as a job. 
-We can investgate further by comparing with some other measures of activity, like volume of commits or something else. 
-We also looked at the same thing, but with only commits, or without commits. 
-For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.
+- Add in a breakdown on licensing practices here too

### 2023-03-28 analysis/05-history_of_open_archaeology.qmd
Copying info from paper.qmd and organizing it for more productive work

--- 2023-01-31 analysis/paper.qmd
@@
----
-title: "openarchaeoCollaboration"
-author:
-  - name: Zack Batist
-    url: https://zackbatist.github.io
-    affiliation: University of Toronto
-  - name: Joe Roe
-    url: https://joeroe.io
-    affiliation: University of Bern
-date: "`r Sys.Date()`"
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	message = FALSE,
-	warning = FALSE
-)
-```
-
-```{r dependencies, include=FALSE}
-library("tidyverse")
-library("lubridate")
-library("openarchaeoCollaboration")
-library("khroma")
-library("gt")
-library("fs")
-library("here")
-library("tidygraph")
-library("ggraph")
-library("ggiraph")
-```
-
-```{r plot-theme}
-theme_oarch <- function() theme(
-)
-```
-
-```{r data-oarch}
-read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
-  mutate(category = na_if(category, "?")) ->
-  oarch
-
-# Extract clean repos from GitHub URLs
-oarch %>%
-  mutate(
-    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
-
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
-```
-
-```{r data-github}
-# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
-# chunk will prefer to load it rather than make requests to the GitHub API.
-# Delete this file to fetch fresh data from the GitHub or modify 
-# `cached_gh_data` to read a different cached version.
-#
-# Note that fetching data from the GitHub API takes some time and is subject to
-# a rate limit:
-# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
-#
-# You will need to authenticate authenticate with GitHub credentials to get a
-# rate limit sufficient to obtain the required data, see 
-# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
-# for options on how to do this.
-cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
-if (!file_exists(cached_gh_data)) {
-  oarch <- mutate(oarch,
-                  gh_langs = map(oarch$gh_repo, gh_lang),
-                  gh_contribs = map(oarch$gh_repo, gh_contrib),
-                  gh_issues = map(oarch$gh_repo, gh_issue),
-                  gh_comments = map(oarch$gh_repo, gh_comment),
-                  gh_commits = map(oarch$gh_repo, gh_commit)
-  )
-  saveRDS(oarch, cached_gh_data)
-}
-oarch <- readRDS(cached_gh_data)
-```
-
-## Introduction
-
-In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools to serve our own unique purposes, which are not typically unanticipated by their original designers.
-In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways.
-For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase.
-However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems.
-In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes.
-
-This article presents a survey of archaeological software development with two goals in mind:
-
-(1) we identify *what* kinds of software archaeologists are making;
-(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
-
-Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
-
-## Background
-
--   The notion of the archaeological toolbox, and its inclusion of digital tools and skillsets
-    -   The use of off-the-shelf tools like MS Access, Excel, Agisoft PhotoScan, Dropbox, etc
-    -   Link to related factors:
-        -   Lack of experienced programmers among archaeologists
-        -   The built-in flexibility of these tools
-        -   Ease of administrative purchasing through university departments
-    -   What changed?
-        -   Software is expensive, shift to subscription models, locking into a dependency on a commercial service provider
-        -   Archaeologists are better at computer programming, digital literacy on the rise, digital archaeologists grows into its own legitimate sub-discipline
-        -   Open source is cool, everyone has a github profile (even if not really)
-            -   (JR:) People are presumably guided towards GitHub (or at least git) via other routes, because it's seen as the standard platform for 'open' digital work.
-For example, the literature on reproducible compendiums recommends GitHub, and almost all R development work is now done there.
-And once there, you're sort of forced into engaging in a (particular set of) open source development.
-I know there's controversy about how dominant this has become and people who say it's not 'true' (free and) open source, and perhaps we could add to that the question of whether it's the right model for archaeology.
--   About open source practices and their influence on archaeology
-    -   What is open source software and open source software development?
-        -   Free cost as an end-user
-        -   Free to manipulate and extend
-        -   Community-based support and development
-    -   Software as product vs software as collective process
-        -   Basically what distinguishes abandonware from thriving software ecosystems
-        -   Technical, legal and social notions of open
-            -   Technical: source code is open, use of a platform and systems that support software integration and implementations
-            -   Legal: open licenses, backing by the courts
-            -   Social: inviting community, meaningful contributions provided by users, common vision and values directing the project into future directions
-        -   Consider how these aspects of open have been implemented or prioritized by digital archaeologists in different ways
-            -   Technical: reproducible research, archiving and publishing
-            -   Legal: archiving and publishing, budgeting for projects
-            -   Social: ensuring that software is actually useful for general use and not just one-off implementation, coordinating to reduce redundant effort and not reinvent the wheel, writing documentation, making it fit within existing systems or infrastructures (i.e.
-allocation of credit, citing software directly versus articles that present software, sustainable development after funding runs out, etc)
-        -   This paper largely focuses on the social aspect of open, which is arguably the most important for ensuring a positive future for archaeological software development
-
-The 'archaeological toolbox' is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely upon in their work.
-
-collaborate in their programming work, and how they experience open source development processes which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.
-
-## Data and evidence
-
-Our findings here are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
-As of right now it includes 408 items, which are nicely rendered at open-archaeo.info.
-
-The main criteria for inclusion is whether a project is identified by its maintainers as relating to archaeological work.
-
-The dataset was compiled by browsing collaborative software development platforms, essentially manually crawling through the profiles of users who identify as archaeologists or who have contributed to tools that pertain to archaeological work.
-
-This quasi-systematic collection strategy is also supplemented by contributions made by interested individuals who identified relevant work that was initially overlooked.
-
-Open-archaeo is a relatively comprehensive list, however it generally lacks code written before the growth and general use of collaborative software development platforms, such as GitHub, among archaeologists, and also lacks references to code that is not available on the web.
-
-It is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in your field, to contribute.
-
-Please get in touch if you're interested in adding to the list or doing a deep dive relating to your area of expertise!
-
-### Tags and categories
-
-Each record is also annotated with tags that describe the aspects of archaeological work that each tool contributes to.
-
-Items are tagged based on how they have been generally identified or presented by their maintainers.
-
-The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.
-
-Right now there are 63 tags.
-
-Records are also categorized based on how each tool or resource is accessed or used.
-
-This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-
-In other words, categories hint at some of the decisions that developers make regarding how they want their tools to be used and accessed.
-
-| Category                              | Kind      | Scope                                                                                                                                                                                                        |
-|:-----------|:-----------|:------------------------------------------------|
-| Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-|
-| Standalone software                   | Software  | Software that may be operated without needing to first access an underlying platform.
-|
-| Scripts                               | Software  | Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-|
-| Specifications, protocols and schemas | Documents | A formal data structure or framework intended to be used as a model.
-|
-| Products                              | Documents | Stable outcomes of creative work.
-|
-| Guides                                | Documents | An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.
-|
-| Lists and datasets                    | Documents | A series of consistently organized observations assembled with purpose.
-|
-
-Two kinds of categories have been defined: software and documents.
-
-Software comprise collections of instructions that dictate how a computer should operate, and documents are bounded sets of recorded information.
-
-Three categories of software and four categories of documents are described in the table.
-
-Packages and libraries are sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-
-Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.
-
-Standalone software are tools that may be operated without needing to first access an underlying platform.
-
-And the third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-Things like code generalized sets of code snippets.
-
-As for documents, specifications, protocols and schemas are formal data structures or frameworks intended to be used as models.
-
-Products are stable outcomes of creative work, loosely defined.
-
-Guides are educational resources or documented protocols meant to instruct readers how to apply relevant tools or techniques.
-
-And finally, lists or datasets, including open-archaeo itself, are series of consistently organized observations assembled with purpose.
-
-## Findings
-
-### What archaeologists are making
-
-Our first set of analyses were focused on these questions:
-
-What kinds of software are archaeologists making?
-
-And what are the general purposes of their software?
-
-------------------------------------------------------------------------
-
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing 'platform' -- for example a package for a programming language like R, or a plugin for an application like QGIS.
-
-The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
-
-Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-
-A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
-
-```{r plot-categories}
-oarch <- mutate(oarch,
-                category = recode(category,
-                                  "Specifications, protocols and schemas" = "Specifications etc.",
-                                  .default = category))
-oarch %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(n = n()) %>%
-  ggplot(aes(fct_reorder(category, n), n)) +
-  geom_col() +
-  coord_flip() +
-  labs(
-    title = "Most popular categories",
-    subtitle = "By number of projects",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-R is overwhelmingly dominant as a platform, closely followed by Python.
-
-Plugins for QGIS and mobile apps are also relatively popular.
-
-Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
-AutoCAD, ArcGIS); a number of lesser used programming languages; and a genre consisting of custom forms and spreadsheet templates.
-
-In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
-The larger platforms tend to be more diverse.
-
-```{r plot-platforms}
-oarch %>%
-  drop_na(platform) %>%
-  group_by(platform, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(
-    title = "Most popular platforms",
-    subtitle = "By number of projects, split by primary author",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-If we compare categories with tags, we can see the general domains that each kind of resource is designed to serve.
-
-We see that packages are fairly common across the board.
-
-Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
-
-These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.
-
-```{r plot-tags-cats}
-detail_tags <- c("Instrumental Neutron activation analysis",
-                 "Harris Matrix",
-                 "aDNA Simulators",
-                 "Damage profilers",
-                 "Metagenomic classifers",
-                 "aDNA simulators",
-                 "aDNA damage-aware genotypers",
-                 "Low-coverage contamination estimation",
-                 "Low-coverage adapted population-genetics tools",
-                 "aDNA preprocessing",
-                 "aDNA aligner",
-                 "aDNA assembler",
-                 "aDNA toolkit")
-
-oarch %>%
-  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  filter(!tag %in% detail_tags) %>%
-  mutate(category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(tag, category) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  # facet_wrap(vars(category)) +
-  geom_col() +
-  scale_fill_bright() +
-  coord_flip() +
-  labs(
-    title = "Tags by category",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = "Number of projects",
-    fill = NULL
-  ) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_text(size = 7)
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/tags-cats.png", width = 254, height = 127, units = "mm")
-```
-
-**ZB NOTE:** Do a temporal analysis to get a sense of when platforms, tags and categories have been gaining popularity.
-There may not be a need for extreme detail here, but simply noting the dates of first commits might be helpful.
-I'm imagining a stacked bar chart, with each bar representing a month or a year, and with the stacked portions of each bar representing a platform or category.
-
-### Development processes
-
-While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about *how* archaeologists develop software.
-
-The remainder of the paper is driven by a concern about how archaeologists collaborate in their programming work.
-
-How is their work sustained, and how does it leverage distributed work processes that are a key aspect of open source software development in general?
-
-To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
-
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh].
-
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.
-
-```{r plot-repo-hosts}
-# TODO: turn into table or just text
-oarch <- mutate(oarch, none = ifelse(is.na(github) &
-                                       is.na(gitlab) &
-                                       is.na(bitbucket) &
-                                       is.na(launchpad) &
-                                       is.na(codeberg),
-                                     "", NA))
-
-oarch %>%
-  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
-               names_to = "repo_host", values_to = "repo_url") %>%
-  drop_na(repo_url) %>%
-  group_by(repo_host, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ungroup() %>%
-  mutate(repo_host = fct_reorder(repo_host, total),
-         repo_host = fct_relevel(repo_host, "none")) %>%
-  ggplot(aes(repo_host, n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(title = "Most popular repository hosts",
-       subtitle = "By number of projects, split by primary author",
-       caption = "Data: open-archaeo.info",
-       x = NULL, y = NULL)
-
-# Export for slides (half slide)
-# ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
-```
-
-```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
-```
-
-We obtained this data in R using the {gh} package; those data, and the analysis code, are available in our research compendium on GitHub.
-
-------------------------------------------------------------------------
-
 Starting with a relatively simple plot, this is a chart of the cumulative growth of open archaeology over the years, measured by code contributions committed and pushed to GitHub repositories.
@@
 
-```{r plot-gh_commits-cum}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  mutate(lumped_category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cat_total = sum(n_commits)) %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = fct_reorder(lumped_category, cat_total),
-             colour = fct_reorder(lumped_category, cat_total))) +
-  geom_area() +
-  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years",
-                   date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = NULL) +
-  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
-  labs(
-    title = "Growth of open archaeology",
-    subtitle = "Measured by cumulative commits in GitHub repositories",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = NULL, y = NULL, fill = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (full slide)
-# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-But how is this activity distributed across repositories?
-
-This visualizes the 'pulse' of active projects.
-
-Each vertical line represents the lifespan of a repository; the points on the line are individual commits.
-
-This is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
-
-```{r plot-lifespan}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(datetime) %>%
-  group_by(item_name) %>%
-  add_tally() %>%
-  filter(n > 1) %>%
-  mutate(first_commit = min(datetime),
-         mid_commit = median(datetime),
-         last_commit = max(datetime)) %>%
-  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
-             y = datetime,
-             ymin = first_commit,
-             ymax = last_commit)) +
-  geom_linerange(size = 0.05) +
-  geom_point(size = 0.1) +
-  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years", date_labels = "%Y") +
-  # coord_flip() +
-  labs(
-    title = "Project lifespan",
-    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
-    x = NULL, y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch() +
-  theme(
-    axis.line.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    panel.grid.major.x = element_blank()
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-The span of time between the first and latest commits is visualized more clearly in summary here.
-
-Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
-
-Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.
-
-Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects.
-
-The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e.
-the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.
-
-But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
-
-This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.
-
-**ZB NOTE:** Use patchwork package to combine these plots into a unified figure.
-
-```{r lifespan-stats}
-oarch %>% 
-  unnest(gh_commits) %>% 
-  drop_na(datetime) %>% 
-  group_by(item_name) %>%
-  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
-  summarise(
-    category = first(category),
-    n_contribs = first(n_contribs),
-    n = n(),
-    first_commit = min(datetime),
-    last_commit = max(datetime),
-    lifespan = as.integer(last_commit - first_commit),
-    commit_rate = n / lifespan
-  ) ->
-  oarch_lifespan
-```
-
-```{r plot-lifespan-hist}
-# TODO: where does oarch_lifespan come from?
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan)) +
-  geom_histogram(binwidth = 100) +
-  scale_x_continuous(breaks = scales::breaks_width(1000)) +
-  labs(
-    title = "Project lifespan",
-    subtitle = "Time between first and latest commit, by repository",
-    x = "Days", y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/lifespan-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-lifespan-vs-rate}
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = "Project lifespan vs. average commit rate",
-    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
-    caption = NULL
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-total.png", width = 127, height = 60, units = "mm")
-
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = NULL,
-    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-contrib.png", width = 127, height = 60, units = "mm")
-
-```
-
-------------------------------------------------------------------------
-
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
-
-A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.
-
-```{r plot-gh_contributors-histogram}
-oarch %>%
-  drop_na(gh_repo) %>%
-  unnest(gh_contribs) %>%
-  group_by(gh_repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Repository contributors",
-    x = "Number of contributors", y = "Repositories",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-hist.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
-
-Of multi-contributor repositories, 92% have a third of contributions made by a single person;
-
-65% have half of all contributions made by a single person;
-
-and 36% have three quarters of all contributions made a single person.
-
-**ZB NOTE:** Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic.
-
-```{r plot-contrib-dist}
-oarch %>%
-  unnest(gh_contribs) %>%
-  drop_na(contributor) %>%
-  group_by(item_name) %>%
-  mutate(n_contributors = n()) %>%
-  filter(n_contributors > 1) %>%
-  mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Leading contributor", "Other contributors")) %>% 
-  group_by(item_name, big_boy) %>%
-  summarise(contributions = sum(contributions)) %>%
-  mutate(
-    p_contributions = contributions / sum(contributions),
-    p_big_boy = first(p_contributions)
-  ) %>%
-  ungroup() %>%
-  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
-  ggplot(aes(x = item_name, y = contributions,
-             fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Leading contributor")))) +
-  geom_col(position = "fill", colour = "white", size = 0.1) +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
-  labs(title = "Repository contributors by proportion of contributions",
-       x = NULL, y = "Contributions",
-       fill = NULL,
-       caption = "Data: open-archaeo.info and GitHub API") +
-  coord_flip() +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-dist.png", width = 127, height = 127, units = "mm")
-```
-
-Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-
-Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-
-**ZB NOTE:** Recalcute these numbers.
-
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
-
-```{r plot-issues-hist}
-oarch %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of issues",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of issues", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/issues-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-comments-hist}
-oarch %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of comments",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of comments", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/comments-hist.png", width = 127, height = 127, units = "mm")
-```
-
-### Formation of collaborative development communities
-
-We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
-
-We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
-
-```{r graph-data}
-oarch |>
-  drop_na(gh_repo) |>
-  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
-  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
-  oarch_graph
-
-oarch_graph <- bind_rows(
-  oarch_graph |>
-    unnest(gh_commits) |>
-    drop_na(author) |> # Apparently real commits, but we can't connect them
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "commit", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_comments) |>
-    drop_na(author) |> # Repos with no comments
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "comment", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_issues) |>
-    drop_na(author) |> # Repos with no issues
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "issue", n_contrib = n(), .groups = "drop")
-)
-
-# Cleaning
-oarch_graph |>
-  rename(user = author) |>
-  # Remove non-human users
-  filter(!str_detect(user, coll("[bot]"))) |>
-  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
-                      "gitter-badger", "dependabot-support")) |>
-  # Remove big, non-archaeology projects that flood the network
-  filter(!repo %in% c("Stellarium/stellarium", 
-                     "opengisch/QField",
-                     "lutraconsulting/input")) ->
-  oarch_graph
-```
-
-We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations.
-We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
-
-#### Repo--user
-
-```{r plot-graph-repo-user}
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
-
-g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt") +
-  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
-  geom_node_point(aes(colour = node_type)) +
-  scale_edge_alpha(trans = "log")
-
-#g_repo_user
-girafe(g_repo_user)
-```
-
-#### Repo--repo
-
-We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate notes, which represent repositories with only a single contributor, we detected 21 clusters.
-While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
-
-Are the clusters associated with particular topics or tags?
-Are there some sorts of associations between clusters that break down from a parent in the community detection dendrogram?
-How does all of this differ when we differentiate commits from issues and comments?
-
-**ZB NOTE:** Remove the cells titled `plot-graph-repo-repo` and `plot-graph-repo-repo-louvain`.
-Use patchwork to compare sets of graphs and dendrograms with links based on all contributions, only commits, and only issues and comments.
-
-```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-```{r plot-graph-repo-repo-edge-betweenness}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  #geom_node_point(aes(colour = group)) +
-  #geom_node_text(aes(label = group)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-
-girafe(ggobj = iplot)
-```
-
-```{r plot-graph-repo-repo-edge-betweenness-dendrogram}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = TRUE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
-xx <- as.dendrogram(xx)
-plot(xx)
-
-ggraph(xx, 'dendrogram', height = height)
-
-```
-
-```{r plot-graph-repo-repo-louvain}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_louvain())) ->
-  oarch_graph_repo_repo
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = group)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-#### User--user
-
-We assembled graphs linking users based on common contributions to the same repositories.
-We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
-
-We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as a job.
-
-**ZB NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits or something else.
-
-We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
-For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the all-contributions graph also appear here.
-These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
-
-**ZB NOTE:** Look at the histogram that maps the distribution of betweenness scores across all users.
-Interesting double hump of low-moderate betweenness scores, which may correspond with distinct or protruding components.
-
-```{r graph-user-user}
-oarch_graph_user_user <- tbl_graph(
-  nodes = nodes_users(oarch_graph), 
-  edges = edges_common_repos(oarch_graph),
-  directed = FALSE
-)
-```
-
-```{r summary-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), 
-         betweenness = centrality_betweenness(weights = n), 
-         strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |> 
-  data.frame() |> 
-  View()
-```
-
-```{r plot-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  # geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common repositories")
-```
-
-#### Over time
-
-Could animate the graphs??
-
-## Conclusions
-
-Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.
-
-In terms of collaboration, our analysis shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host.
-
-The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base.
-
-Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects.
-
-This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
-
-We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.
+{r plot-gh-cum}
 
@@
 
-#### Notes from January 30, end of day
-In cleaning the data we made a few significant decisions that revealed insights into our dataset. 
-We constructed the user-user graph and filtered out several bots. 
-We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community. 
-We filtered isolates. 
-We calculated the degree and betweenness to assess whether this was realistic representation of what we know. 
-We wanted to figure out who are the key nodes, what they had in common. 
-The people who are high betweenness are people who primarily produce computational archaeology code as a job. 
-We can investgate further by comparing with some other measures of activity, like volume of commits or something else. 
-We also looked at the same thing, but with only commits, or without commits. 
-For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.
+- Growth of open archaeology by number of commits
+- Growth of open archaeology by number of repos
+- Compare changing proportions of each category on a year by year basis
+- Identify periods associated with different rates of growth, and relate these with different general attitudes in the history of digital archaeology

### 2023-03-28 analysis/06-project_activity.qmd
Copying info from paper.qmd and organizing it for more productive work

--- 2023-01-31 analysis/paper.qmd
@@
----
-title: "openarchaeoCollaboration"
-author:
-  - name: Zack Batist
-    url: https://zackbatist.github.io
-    affiliation: University of Toronto
-  - name: Joe Roe
-    url: https://joeroe.io
-    affiliation: University of Bern
-date: "`r Sys.Date()`"
----
+In this section we assess projects' pulse, or the amount of activity pertaining to a repository.
+In the following figure, each vertical line represents the lifespan of a repository, and the points on the line are individual commits.
+The graph is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	message = FALSE,
-	warning = FALSE
-)
-```
-
-```{r dependencies, include=FALSE}
-library("tidyverse")
-library("lubridate")
-library("openarchaeoCollaboration")
-library("khroma")
-library("gt")
-library("fs")
-library("here")
-library("tidygraph")
-library("ggraph")
-library("ggiraph")
-```
-
-```{r plot-theme}
-theme_oarch <- function() theme(
-)
-```
-
-```{r data-oarch}
-read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
-  mutate(category = na_if(category, "?")) ->
-  oarch
-
-# Extract clean repos from GitHub URLs
-oarch %>%
-  mutate(
-    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
-
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
-```
-
-```{r data-github}
-# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
-# chunk will prefer to load it rather than make requests to the GitHub API.
-# Delete this file to fetch fresh data from the GitHub or modify 
-# `cached_gh_data` to read a different cached version.
-#
-# Note that fetching data from the GitHub API takes some time and is subject to
-# a rate limit:
-# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
-#
-# You will need to authenticate authenticate with GitHub credentials to get a
-# rate limit sufficient to obtain the required data, see 
-# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
-# for options on how to do this.
-cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
-if (!file_exists(cached_gh_data)) {
-  oarch <- mutate(oarch,
-                  gh_langs = map(oarch$gh_repo, gh_lang),
-                  gh_contribs = map(oarch$gh_repo, gh_contrib),
-                  gh_issues = map(oarch$gh_repo, gh_issue),
-                  gh_comments = map(oarch$gh_repo, gh_comment),
-                  gh_commits = map(oarch$gh_repo, gh_commit)
-  )
-  saveRDS(oarch, cached_gh_data)
-}
-oarch <- readRDS(cached_gh_data)
-```
-
-## Introduction
-
-In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools to serve our own unique purposes, which are not typically unanticipated by their original designers.
-In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways.
-For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase.
-However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems.
-In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes.
-
-This article presents a survey of archaeological software development with two goals in mind:
-
-(1) we identify *what* kinds of software archaeologists are making;
-(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
-
-Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
-
-## Background
-
--   The notion of the archaeological toolbox, and its inclusion of digital tools and skillsets
-    -   The use of off-the-shelf tools like MS Access, Excel, Agisoft PhotoScan, Dropbox, etc
-    -   Link to related factors:
-        -   Lack of experienced programmers among archaeologists
-        -   The built-in flexibility of these tools
-        -   Ease of administrative purchasing through university departments
-    -   What changed?
-        -   Software is expensive, shift to subscription models, locking into a dependency on a commercial service provider
-        -   Archaeologists are better at computer programming, digital literacy on the rise, digital archaeologists grows into its own legitimate sub-discipline
-        -   Open source is cool, everyone has a github profile (even if not really)
-            -   (JR:) People are presumably guided towards GitHub (or at least git) via other routes, because it's seen as the standard platform for 'open' digital work.
-For example, the literature on reproducible compendiums recommends GitHub, and almost all R development work is now done there.
-And once there, you're sort of forced into engaging in a (particular set of) open source development.
-I know there's controversy about how dominant this has become and people who say it's not 'true' (free and) open source, and perhaps we could add to that the question of whether it's the right model for archaeology.
--   About open source practices and their influence on archaeology
-    -   What is open source software and open source software development?
-        -   Free cost as an end-user
-        -   Free to manipulate and extend
-        -   Community-based support and development
-    -   Software as product vs software as collective process
-        -   Basically what distinguishes abandonware from thriving software ecosystems
-        -   Technical, legal and social notions of open
-            -   Technical: source code is open, use of a platform and systems that support software integration and implementations
-            -   Legal: open licenses, backing by the courts
-            -   Social: inviting community, meaningful contributions provided by users, common vision and values directing the project into future directions
-        -   Consider how these aspects of open have been implemented or prioritized by digital archaeologists in different ways
-            -   Technical: reproducible research, archiving and publishing
-            -   Legal: archiving and publishing, budgeting for projects
-            -   Social: ensuring that software is actually useful for general use and not just one-off implementation, coordinating to reduce redundant effort and not reinvent the wheel, writing documentation, making it fit within existing systems or infrastructures (i.e.
-allocation of credit, citing software directly versus articles that present software, sustainable development after funding runs out, etc)
-        -   This paper largely focuses on the social aspect of open, which is arguably the most important for ensuring a positive future for archaeological software development
-
-The 'archaeological toolbox' is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely upon in their work.
-
-collaborate in their programming work, and how they experience open source development processes which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.
-
-## Data and evidence
-
-Our findings here are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
-As of right now it includes 408 items, which are nicely rendered at open-archaeo.info.
-
-The main criteria for inclusion is whether a project is identified by its maintainers as relating to archaeological work.
-
-The dataset was compiled by browsing collaborative software development platforms, essentially manually crawling through the profiles of users who identify as archaeologists or who have contributed to tools that pertain to archaeological work.
-
-This quasi-systematic collection strategy is also supplemented by contributions made by interested individuals who identified relevant work that was initially overlooked.
-
-Open-archaeo is a relatively comprehensive list, however it generally lacks code written before the growth and general use of collaborative software development platforms, such as GitHub, among archaeologists, and also lacks references to code that is not available on the web.
-
-It is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in your field, to contribute.
-
-Please get in touch if you're interested in adding to the list or doing a deep dive relating to your area of expertise!
-
-### Tags and categories
-
-Each record is also annotated with tags that describe the aspects of archaeological work that each tool contributes to.
-
-Items are tagged based on how they have been generally identified or presented by their maintainers.
-
-The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.
-
-Right now there are 63 tags.
-
-Records are also categorized based on how each tool or resource is accessed or used.
-
-This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-
-In other words, categories hint at some of the decisions that developers make regarding how they want their tools to be used and accessed.
-
-| Category                              | Kind      | Scope                                                                                                                                                                                                        |
-|:-----------|:-----------|:------------------------------------------------|
-| Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-|
-| Standalone software                   | Software  | Software that may be operated without needing to first access an underlying platform.
-|
-| Scripts                               | Software  | Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-|
-| Specifications, protocols and schemas | Documents | A formal data structure or framework intended to be used as a model.
-|
-| Products                              | Documents | Stable outcomes of creative work.
-|
-| Guides                                | Documents | An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.
-|
-| Lists and datasets                    | Documents | A series of consistently organized observations assembled with purpose.
-|
-
-Two kinds of categories have been defined: software and documents.
-
-Software comprise collections of instructions that dictate how a computer should operate, and documents are bounded sets of recorded information.
-
-Three categories of software and four categories of documents are described in the table.
-
-Packages and libraries are sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-
-Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.
-
-Standalone software are tools that may be operated without needing to first access an underlying platform.
-
-And the third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-Things like code generalized sets of code snippets.
-
-As for documents, specifications, protocols and schemas are formal data structures or frameworks intended to be used as models.
-
-Products are stable outcomes of creative work, loosely defined.
-
-Guides are educational resources or documented protocols meant to instruct readers how to apply relevant tools or techniques.
-
-And finally, lists or datasets, including open-archaeo itself, are series of consistently organized observations assembled with purpose.
-
-## Findings
-
-### What archaeologists are making
-
-Our first set of analyses were focused on these questions:
-
-What kinds of software are archaeologists making?
-
-And what are the general purposes of their software?
-
-------------------------------------------------------------------------
-
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing 'platform' -- for example a package for a programming language like R, or a plugin for an application like QGIS.
-
-The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
-
-Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-
-A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
-
-```{r plot-categories}
-oarch <- mutate(oarch,
-                category = recode(category,
-                                  "Specifications, protocols and schemas" = "Specifications etc.",
-                                  .default = category))
-oarch %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(n = n()) %>%
-  ggplot(aes(fct_reorder(category, n), n)) +
-  geom_col() +
-  coord_flip() +
-  labs(
-    title = "Most popular categories",
-    subtitle = "By number of projects",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-R is overwhelmingly dominant as a platform, closely followed by Python.
-
-Plugins for QGIS and mobile apps are also relatively popular.
-
-Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
-AutoCAD, ArcGIS); a number of lesser used programming languages; and a genre consisting of custom forms and spreadsheet templates.
-
-In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
-The larger platforms tend to be more diverse.
-
-```{r plot-platforms}
-oarch %>%
-  drop_na(platform) %>%
-  group_by(platform, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(
-    title = "Most popular platforms",
-    subtitle = "By number of projects, split by primary author",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-If we compare categories with tags, we can see the general domains that each kind of resource is designed to serve.
-
-We see that packages are fairly common across the board.
-
-Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
-
-These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.
-
-```{r plot-tags-cats}
-detail_tags <- c("Instrumental Neutron activation analysis",
-                 "Harris Matrix",
-                 "aDNA Simulators",
-                 "Damage profilers",
-                 "Metagenomic classifers",
-                 "aDNA simulators",
-                 "aDNA damage-aware genotypers",
-                 "Low-coverage contamination estimation",
-                 "Low-coverage adapted population-genetics tools",
-                 "aDNA preprocessing",
-                 "aDNA aligner",
-                 "aDNA assembler",
-                 "aDNA toolkit")
-
-oarch %>%
-  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  filter(!tag %in% detail_tags) %>%
-  mutate(category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(tag, category) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  # facet_wrap(vars(category)) +
-  geom_col() +
-  scale_fill_bright() +
-  coord_flip() +
-  labs(
-    title = "Tags by category",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = "Number of projects",
-    fill = NULL
-  ) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_text(size = 7)
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/tags-cats.png", width = 254, height = 127, units = "mm")
-```
-
-**ZB NOTE:** Do a temporal analysis to get a sense of when platforms, tags and categories have been gaining popularity.
-There may not be a need for extreme detail here, but simply noting the dates of first commits might be helpful.
-I'm imagining a stacked bar chart, with each bar representing a month or a year, and with the stacked portions of each bar representing a platform or category.
-
-### Development processes
-
-While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about *how* archaeologists develop software.
-
-The remainder of the paper is driven by a concern about how archaeologists collaborate in their programming work.
-
-How is their work sustained, and how does it leverage distributed work processes that are a key aspect of open source software development in general?
-
-To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
-
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh].
-
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.
-
-```{r plot-repo-hosts}
-# TODO: turn into table or just text
-oarch <- mutate(oarch, none = ifelse(is.na(github) &
-                                       is.na(gitlab) &
-                                       is.na(bitbucket) &
-                                       is.na(launchpad) &
-                                       is.na(codeberg),
-                                     "", NA))
-
-oarch %>%
-  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
-               names_to = "repo_host", values_to = "repo_url") %>%
-  drop_na(repo_url) %>%
-  group_by(repo_host, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ungroup() %>%
-  mutate(repo_host = fct_reorder(repo_host, total),
-         repo_host = fct_relevel(repo_host, "none")) %>%
-  ggplot(aes(repo_host, n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(title = "Most popular repository hosts",
-       subtitle = "By number of projects, split by primary author",
-       caption = "Data: open-archaeo.info",
-       x = NULL, y = NULL)
-
-# Export for slides (half slide)
-# ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
-```
-
-```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
-```
-
-We obtained this data in R using the {gh} package; those data, and the analysis code, are available in our research compendium on GitHub.
-
-------------------------------------------------------------------------
-
-Starting with a relatively simple plot, this is a chart of the cumulative growth of open archaeology over the years, measured by code contributions committed and pushed to GitHub repositories.
-
-As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
-But they really began to get into it around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
-
-Around this time we also see that GitHub starts being used to host documents and scripts.
-This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.
-
-```{r plot-gh_commits-cum}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  mutate(lumped_category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cat_total = sum(n_commits)) %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = fct_reorder(lumped_category, cat_total),
-             colour = fct_reorder(lumped_category, cat_total))) +
-  geom_area() +
-  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years",
-                   date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = NULL) +
-  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
-  labs(
-    title = "Growth of open archaeology",
-    subtitle = "Measured by cumulative commits in GitHub repositories",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = NULL, y = NULL, fill = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (full slide)
-# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-But how is this activity distributed across repositories?
-
-This visualizes the 'pulse' of active projects.
-
-Each vertical line represents the lifespan of a repository; the points on the line are individual commits.
-
-This is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
-
-```{r plot-lifespan}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(datetime) %>%
-  group_by(item_name) %>%
-  add_tally() %>%
-  filter(n > 1) %>%
-  mutate(first_commit = min(datetime),
-         mid_commit = median(datetime),
-         last_commit = max(datetime)) %>%
-  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
-             y = datetime,
-             ymin = first_commit,
-             ymax = last_commit)) +
-  geom_linerange(size = 0.05) +
-  geom_point(size = 0.1) +
-  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years", date_labels = "%Y") +
-  # coord_flip() +
-  labs(
-    title = "Project lifespan",
-    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
-    x = NULL, y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch() +
-  theme(
-    axis.line.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    panel.grid.major.x = element_blank()
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
+{r plot-lifespan}
 
 The span of time between the first and latest commits is visualized more clearly in summary here.
-
 Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
-
 Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.
-
 Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects.
-
 The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e.
 the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.
-
 But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
-
 This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.
 
-**ZB NOTE:** Use patchwork package to combine these plots into a unified figure.
+{r lifespan-stats}
 
-```{r lifespan-stats}
-oarch %>% 
-  unnest(gh_commits) %>% 
-  drop_na(datetime) %>% 
-  group_by(item_name) %>%
-  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
-  summarise(
-    category = first(category),
-    n_contribs = first(n_contribs),
-    n = n(),
-    first_commit = min(datetime),
-    last_commit = max(datetime),
-    lifespan = as.integer(last_commit - first_commit),
-    commit_rate = n / lifespan
-  ) ->
-  oarch_lifespan
-```
+{r plot-lifespan-hist}
 
-```{r plot-lifespan-hist}
-# TODO: where does oarch_lifespan come from?
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan)) +
-  geom_histogram(binwidth = 100) +
-  scale_x_continuous(breaks = scales::breaks_width(1000)) +
-  labs(
-    title = "Project lifespan",
-    subtitle = "Time between first and latest commit, by repository",
-    x = "Days", y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/lifespan-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-lifespan-vs-rate}
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = "Project lifespan vs. average commit rate",
-    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
-    caption = NULL
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-total.png", width = 127, height = 60, units = "mm")
-
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = NULL,
-    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-contrib.png", width = 127, height = 60, units = "mm")
-
-```
-
-------------------------------------------------------------------------
+{r plot-lifespan-vs-rate}
 
 Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
-
 A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.
 
-```{r plot-gh_contributors-histogram}
-oarch %>%
-  drop_na(gh_repo) %>%
-  unnest(gh_contribs) %>%
-  group_by(gh_repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Repository contributors",
-    x = "Number of contributors", y = "Repositories",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-hist.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
+{r plot-gh_contributors-histogram}
 
 Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
+Of multi-contributor repositories, 92% have a third of contributions made by a single person; 65% have half of all contributions made by a single person; and 36% have three quarters of all contributions made a single person.
 
-Of multi-contributor repositories, 92% have a third of contributions made by a single person;
-
-65% have half of all contributions made by a single person;
-
-and 36% have three quarters of all contributions made a single person.
-
-**ZB NOTE:** Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic.
-
-```{r plot-contrib-dist}
-oarch %>%
-  unnest(gh_contribs) %>%
-  drop_na(contributor) %>%
-  group_by(item_name) %>%
-  mutate(n_contributors = n()) %>%
-  filter(n_contributors > 1) %>%
-  mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Leading contributor", "Other contributors")) %>% 
-  group_by(item_name, big_boy) %>%
-  summarise(contributions = sum(contributions)) %>%
-  mutate(
-    p_contributions = contributions / sum(contributions),
-    p_big_boy = first(p_contributions)
-  ) %>%
-  ungroup() %>%
-  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
-  ggplot(aes(x = item_name, y = contributions,
-             fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Leading contributor")))) +
-  geom_col(position = "fill", colour = "white", size = 0.1) +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
-  labs(title = "Repository contributors by proportion of contributions",
-       x = NULL, y = "Contributions",
-       fill = NULL,
-       caption = "Data: open-archaeo.info and GitHub API") +
-  coord_flip() +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-dist.png", width = 127, height = 127, units = "mm")
-```
+{r plot-contrib-dist}
 
 Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-
-Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-
-**ZB NOTE:** Recalcute these numbers.
-
+Only 39% of repositories make use of these features.
+Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
 However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-
 This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
 
-```{r plot-issues-hist}
-oarch %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of issues",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of issues", y = "Repositories"
-  ) +
-  theme_oarch()
+{r plot-issues-hist}
 
-# Export for slides (half slide)
-# ggsave("figures/issues-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-comments-hist}
-oarch %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of comments",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of comments", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/comments-hist.png", width = 127, height = 127, units = "mm")
-```
-
-### Formation of collaborative development communities
-
-We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
-
-We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
-
-```{r graph-data}
-oarch |>
-  drop_na(gh_repo) |>
-  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
-  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
-  oarch_graph
-
-oarch_graph <- bind_rows(
-  oarch_graph |>
-    unnest(gh_commits) |>
-    drop_na(author) |> # Apparently real commits, but we can't connect them
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "commit", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_comments) |>
-    drop_na(author) |> # Repos with no comments
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "comment", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_issues) |>
-    drop_na(author) |> # Repos with no issues
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "issue", n_contrib = n(), .groups = "drop")
-)
-
-# Cleaning
-oarch_graph |>
-  rename(user = author) |>
-  # Remove non-human users
-  filter(!str_detect(user, coll("[bot]"))) |>
-  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
-                      "gitter-badger", "dependabot-support")) |>
-  # Remove big, non-archaeology projects that flood the network
-  filter(!repo %in% c("Stellarium/stellarium", 
-                     "opengisch/QField",
-                     "lutraconsulting/input")) ->
-  oarch_graph
-```
-
-We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations.
-We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
-
-#### Repo--user
-
-```{r plot-graph-repo-user}
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
-
-g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt") +
-  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
-  geom_node_point(aes(colour = node_type)) +
-  scale_edge_alpha(trans = "log")
-
-#g_repo_user
-girafe(g_repo_user)
-```
-
-#### Repo--repo
-
-We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate notes, which represent repositories with only a single contributor, we detected 21 clusters.
-While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
-
-Are the clusters associated with particular topics or tags?
-Are there some sorts of associations between clusters that break down from a parent in the community detection dendrogram?
-How does all of this differ when we differentiate commits from issues and comments?
-
-**ZB NOTE:** Remove the cells titled `plot-graph-repo-repo` and `plot-graph-repo-repo-louvain`.
-Use patchwork to compare sets of graphs and dendrograms with links based on all contributions, only commits, and only issues and comments.
-
-```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-```{r plot-graph-repo-repo-edge-betweenness}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  #geom_node_point(aes(colour = group)) +
-  #geom_node_text(aes(label = group)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-
-girafe(ggobj = iplot)
-```
-
-```{r plot-graph-repo-repo-edge-betweenness-dendrogram}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = TRUE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
-xx <- as.dendrogram(xx)
-plot(xx)
-
-ggraph(xx, 'dendrogram', height = height)
-
-```
-
-```{r plot-graph-repo-repo-louvain}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_louvain())) ->
-  oarch_graph_repo_repo
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = group)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-#### User--user
-
-We assembled graphs linking users based on common contributions to the same repositories.
-We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
-
-We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as a job.
-
-**ZB NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits or something else.
-
-We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
-For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the all-contributions graph also appear here.
-These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
-
-**ZB NOTE:** Look at the histogram that maps the distribution of betweenness scores across all users.
-Interesting double hump of low-moderate betweenness scores, which may correspond with distinct or protruding components.
-
-```{r graph-user-user}
-oarch_graph_user_user <- tbl_graph(
-  nodes = nodes_users(oarch_graph), 
-  edges = edges_common_repos(oarch_graph),
-  directed = FALSE
-)
-```
-
-```{r summary-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), 
-         betweenness = centrality_betweenness(weights = n), 
-         strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |> 
-  data.frame() |> 
-  View()
-```
-
-```{r plot-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  # geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common repositories")
-```
-
-#### Over time
-
-Could animate the graphs??
-
-## Conclusions
-
-Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.
-
-In terms of collaboration, our analysis shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host.
-
-The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base.
-
-Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects.
-
-This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
-
-We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.
-
----
-
-#### Notes from January 30, end of day
-In cleaning the data we made a few significant decisions that revealed insights into our dataset. 
-We constructed the user-user graph and filtered out several bots. 
-We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community. 
-We filtered isolates. 
-We calculated the degree and betweenness to assess whether this was realistic representation of what we know. 
-We wanted to figure out who are the key nodes, what they had in common. 
-The people who are high betweenness are people who primarily produce computational archaeology code as a job. 
-We can investgate further by comparing with some other measures of activity, like volume of commits or something else. 
-We also looked at the same thing, but with only commits, or without commits. 
-For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.
+{r plot-comments-hist}

### 2023-03-28 analysis/07-network.qmd
Copying info from paper.qmd and organizing it for more productive work

--- 2023-01-31 analysis/paper.qmd
@@
----
-title: "openarchaeoCollaboration"
-author:
-  - name: Zack Batist
-    url: https://zackbatist.github.io
-    affiliation: University of Toronto
-  - name: Joe Roe
-    url: https://joeroe.io
-    affiliation: University of Bern
-date: "`r Sys.Date()`"
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	message = FALSE,
-	warning = FALSE
-)
-```
-
-```{r dependencies, include=FALSE}
-library("tidyverse")
-library("lubridate")
-library("openarchaeoCollaboration")
-library("khroma")
-library("gt")
-library("fs")
-library("here")
-library("tidygraph")
-library("ggraph")
-library("ggiraph")
-```
-
-```{r plot-theme}
-theme_oarch <- function() theme(
-)
-```
-
-```{r data-oarch}
-read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
-  mutate(category = na_if(category, "?")) ->
-  oarch
-
-# Extract clean repos from GitHub URLs
-oarch %>%
-  mutate(
-    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
-
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
-```
-
-```{r data-github}
-# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
-# chunk will prefer to load it rather than make requests to the GitHub API.
-# Delete this file to fetch fresh data from the GitHub or modify 
-# `cached_gh_data` to read a different cached version.
-#
-# Note that fetching data from the GitHub API takes some time and is subject to
-# a rate limit:
-# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
-#
-# You will need to authenticate authenticate with GitHub credentials to get a
-# rate limit sufficient to obtain the required data, see 
-# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
-# for options on how to do this.
-cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
-if (!file_exists(cached_gh_data)) {
-  oarch <- mutate(oarch,
-                  gh_langs = map(oarch$gh_repo, gh_lang),
-                  gh_contribs = map(oarch$gh_repo, gh_contrib),
-                  gh_issues = map(oarch$gh_repo, gh_issue),
-                  gh_comments = map(oarch$gh_repo, gh_comment),
-                  gh_commits = map(oarch$gh_repo, gh_commit)
-  )
-  saveRDS(oarch, cached_gh_data)
-}
-oarch <- readRDS(cached_gh_data)
-```
-
-## Introduction
-
-In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools to serve our own unique purposes, which are not typically unanticipated by their original designers.
-In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways.
-For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase.
-However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems.
-In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes.
-
-This article presents a survey of archaeological software development with two goals in mind:
-
-(1) we identify *what* kinds of software archaeologists are making;
-(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
-
-Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
-
-## Background
-
--   The notion of the archaeological toolbox, and its inclusion of digital tools and skillsets
-    -   The use of off-the-shelf tools like MS Access, Excel, Agisoft PhotoScan, Dropbox, etc
-    -   Link to related factors:
-        -   Lack of experienced programmers among archaeologists
-        -   The built-in flexibility of these tools
-        -   Ease of administrative purchasing through university departments
-    -   What changed?
-        -   Software is expensive, shift to subscription models, locking into a dependency on a commercial service provider
-        -   Archaeologists are better at computer programming, digital literacy on the rise, digital archaeologists grows into its own legitimate sub-discipline
-        -   Open source is cool, everyone has a github profile (even if not really)
-            -   (JR:) People are presumably guided towards GitHub (or at least git) via other routes, because it's seen as the standard platform for 'open' digital work.
-For example, the literature on reproducible compendiums recommends GitHub, and almost all R development work is now done there.
-And once there, you're sort of forced into engaging in a (particular set of) open source development.
-I know there's controversy about how dominant this has become and people who say it's not 'true' (free and) open source, and perhaps we could add to that the question of whether it's the right model for archaeology.
--   About open source practices and their influence on archaeology
-    -   What is open source software and open source software development?
-        -   Free cost as an end-user
-        -   Free to manipulate and extend
-        -   Community-based support and development
-    -   Software as product vs software as collective process
-        -   Basically what distinguishes abandonware from thriving software ecosystems
-        -   Technical, legal and social notions of open
-            -   Technical: source code is open, use of a platform and systems that support software integration and implementations
-            -   Legal: open licenses, backing by the courts
-            -   Social: inviting community, meaningful contributions provided by users, common vision and values directing the project into future directions
-        -   Consider how these aspects of open have been implemented or prioritized by digital archaeologists in different ways
-            -   Technical: reproducible research, archiving and publishing
-            -   Legal: archiving and publishing, budgeting for projects
-            -   Social: ensuring that software is actually useful for general use and not just one-off implementation, coordinating to reduce redundant effort and not reinvent the wheel, writing documentation, making it fit within existing systems or infrastructures (i.e.
-allocation of credit, citing software directly versus articles that present software, sustainable development after funding runs out, etc)
-        -   This paper largely focuses on the social aspect of open, which is arguably the most important for ensuring a positive future for archaeological software development
-
-The 'archaeological toolbox' is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely upon in their work.
-
-collaborate in their programming work, and how they experience open source development processes which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.
-
-## Data and evidence
-
-Our findings here are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
-As of right now it includes 408 items, which are nicely rendered at open-archaeo.info.
-
-The main criteria for inclusion is whether a project is identified by its maintainers as relating to archaeological work.
-
-The dataset was compiled by browsing collaborative software development platforms, essentially manually crawling through the profiles of users who identify as archaeologists or who have contributed to tools that pertain to archaeological work.
-
-This quasi-systematic collection strategy is also supplemented by contributions made by interested individuals who identified relevant work that was initially overlooked.
-
-Open-archaeo is a relatively comprehensive list, however it generally lacks code written before the growth and general use of collaborative software development platforms, such as GitHub, among archaeologists, and also lacks references to code that is not available on the web.
-
-It is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in your field, to contribute.
-
-Please get in touch if you're interested in adding to the list or doing a deep dive relating to your area of expertise!
-
-### Tags and categories
-
-Each record is also annotated with tags that describe the aspects of archaeological work that each tool contributes to.
-
-Items are tagged based on how they have been generally identified or presented by their maintainers.
-
-The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.
-
-Right now there are 63 tags.
-
-Records are also categorized based on how each tool or resource is accessed or used.
-
-This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-
-In other words, categories hint at some of the decisions that developers make regarding how they want their tools to be used and accessed.
-
-| Category                              | Kind      | Scope                                                                                                                                                                                                        |
-|:-----------|:-----------|:------------------------------------------------|
-| Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-|
-| Standalone software                   | Software  | Software that may be operated without needing to first access an underlying platform.
-|
-| Scripts                               | Software  | Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-|
-| Specifications, protocols and schemas | Documents | A formal data structure or framework intended to be used as a model.
-|
-| Products                              | Documents | Stable outcomes of creative work.
-|
-| Guides                                | Documents | An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.
-|
-| Lists and datasets                    | Documents | A series of consistently organized observations assembled with purpose.
-|
-
-Two kinds of categories have been defined: software and documents.
-
-Software comprise collections of instructions that dictate how a computer should operate, and documents are bounded sets of recorded information.
-
-Three categories of software and four categories of documents are described in the table.
-
-Packages and libraries are sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-
-Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.
-
-Standalone software are tools that may be operated without needing to first access an underlying platform.
-
-And the third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-Things like code generalized sets of code snippets.
-
-As for documents, specifications, protocols and schemas are formal data structures or frameworks intended to be used as models.
-
-Products are stable outcomes of creative work, loosely defined.
-
-Guides are educational resources or documented protocols meant to instruct readers how to apply relevant tools or techniques.
-
-And finally, lists or datasets, including open-archaeo itself, are series of consistently organized observations assembled with purpose.
-
-## Findings
-
-### What archaeologists are making
-
-Our first set of analyses were focused on these questions:
-
-What kinds of software are archaeologists making?
-
-And what are the general purposes of their software?
-
-------------------------------------------------------------------------
-
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing 'platform' -- for example a package for a programming language like R, or a plugin for an application like QGIS.
-
-The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
-
-Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-
-A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
-
-```{r plot-categories}
-oarch <- mutate(oarch,
-                category = recode(category,
-                                  "Specifications, protocols and schemas" = "Specifications etc.",
-                                  .default = category))
-oarch %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(n = n()) %>%
-  ggplot(aes(fct_reorder(category, n), n)) +
-  geom_col() +
-  coord_flip() +
-  labs(
-    title = "Most popular categories",
-    subtitle = "By number of projects",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-R is overwhelmingly dominant as a platform, closely followed by Python.
-
-Plugins for QGIS and mobile apps are also relatively popular.
-
-Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
-AutoCAD, ArcGIS); a number of lesser used programming languages; and a genre consisting of custom forms and spreadsheet templates.
-
-In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
-The larger platforms tend to be more diverse.
-
-```{r plot-platforms}
-oarch %>%
-  drop_na(platform) %>%
-  group_by(platform, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(
-    title = "Most popular platforms",
-    subtitle = "By number of projects, split by primary author",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-If we compare categories with tags, we can see the general domains that each kind of resource is designed to serve.
-
-We see that packages are fairly common across the board.
-
-Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
-
-These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.
-
-```{r plot-tags-cats}
-detail_tags <- c("Instrumental Neutron activation analysis",
-                 "Harris Matrix",
-                 "aDNA Simulators",
-                 "Damage profilers",
-                 "Metagenomic classifers",
-                 "aDNA simulators",
-                 "aDNA damage-aware genotypers",
-                 "Low-coverage contamination estimation",
-                 "Low-coverage adapted population-genetics tools",
-                 "aDNA preprocessing",
-                 "aDNA aligner",
-                 "aDNA assembler",
-                 "aDNA toolkit")
-
-oarch %>%
-  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  filter(!tag %in% detail_tags) %>%
-  mutate(category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(tag, category) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  # facet_wrap(vars(category)) +
-  geom_col() +
-  scale_fill_bright() +
-  coord_flip() +
-  labs(
-    title = "Tags by category",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = "Number of projects",
-    fill = NULL
-  ) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_text(size = 7)
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/tags-cats.png", width = 254, height = 127, units = "mm")
-```
-
-**ZB NOTE:** Do a temporal analysis to get a sense of when platforms, tags and categories have been gaining popularity.
-There may not be a need for extreme detail here, but simply noting the dates of first commits might be helpful.
-I'm imagining a stacked bar chart, with each bar representing a month or a year, and with the stacked portions of each bar representing a platform or category.
-
-### Development processes
-
-While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about *how* archaeologists develop software.
-
-The remainder of the paper is driven by a concern about how archaeologists collaborate in their programming work.
-
-How is their work sustained, and how does it leverage distributed work processes that are a key aspect of open source software development in general?
-
-To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
-
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh].
-
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.
-
-```{r plot-repo-hosts}
-# TODO: turn into table or just text
-oarch <- mutate(oarch, none = ifelse(is.na(github) &
-                                       is.na(gitlab) &
-                                       is.na(bitbucket) &
-                                       is.na(launchpad) &
-                                       is.na(codeberg),
-                                     "", NA))
-
-oarch %>%
-  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
-               names_to = "repo_host", values_to = "repo_url") %>%
-  drop_na(repo_url) %>%
-  group_by(repo_host, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ungroup() %>%
-  mutate(repo_host = fct_reorder(repo_host, total),
-         repo_host = fct_relevel(repo_host, "none")) %>%
-  ggplot(aes(repo_host, n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(title = "Most popular repository hosts",
-       subtitle = "By number of projects, split by primary author",
-       caption = "Data: open-archaeo.info",
-       x = NULL, y = NULL)
-
-# Export for slides (half slide)
-# ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
-```
-
-```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
-```
-
-We obtained this data in R using the {gh} package; those data, and the analysis code, are available in our research compendium on GitHub.
-
-------------------------------------------------------------------------
-
-Starting with a relatively simple plot, this is a chart of the cumulative growth of open archaeology over the years, measured by code contributions committed and pushed to GitHub repositories.
-
-As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
-But they really began to get into it around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
-
-Around this time we also see that GitHub starts being used to host documents and scripts.
-This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.
-
-```{r plot-gh_commits-cum}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  mutate(lumped_category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cat_total = sum(n_commits)) %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = fct_reorder(lumped_category, cat_total),
-             colour = fct_reorder(lumped_category, cat_total))) +
-  geom_area() +
-  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years",
-                   date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = NULL) +
-  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
-  labs(
-    title = "Growth of open archaeology",
-    subtitle = "Measured by cumulative commits in GitHub repositories",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = NULL, y = NULL, fill = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (full slide)
-# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-But how is this activity distributed across repositories?
-
-This visualizes the 'pulse' of active projects.
-
-Each vertical line represents the lifespan of a repository; the points on the line are individual commits.
-
-This is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
-
-```{r plot-lifespan}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(datetime) %>%
-  group_by(item_name) %>%
-  add_tally() %>%
-  filter(n > 1) %>%
-  mutate(first_commit = min(datetime),
-         mid_commit = median(datetime),
-         last_commit = max(datetime)) %>%
-  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
-             y = datetime,
-             ymin = first_commit,
-             ymax = last_commit)) +
-  geom_linerange(size = 0.05) +
-  geom_point(size = 0.1) +
-  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years", date_labels = "%Y") +
-  # coord_flip() +
-  labs(
-    title = "Project lifespan",
-    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
-    x = NULL, y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch() +
-  theme(
-    axis.line.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    panel.grid.major.x = element_blank()
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-The span of time between the first and latest commits is visualized more clearly in summary here.
-
-Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
-
-Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.
-
-Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects.
-
-The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e.
-the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.
-
-But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
-
-This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.
-
-**ZB NOTE:** Use patchwork package to combine these plots into a unified figure.
-
-```{r lifespan-stats}
-oarch %>% 
-  unnest(gh_commits) %>% 
-  drop_na(datetime) %>% 
-  group_by(item_name) %>%
-  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
-  summarise(
-    category = first(category),
-    n_contribs = first(n_contribs),
-    n = n(),
-    first_commit = min(datetime),
-    last_commit = max(datetime),
-    lifespan = as.integer(last_commit - first_commit),
-    commit_rate = n / lifespan
-  ) ->
-  oarch_lifespan
-```
-
-```{r plot-lifespan-hist}
-# TODO: where does oarch_lifespan come from?
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan)) +
-  geom_histogram(binwidth = 100) +
-  scale_x_continuous(breaks = scales::breaks_width(1000)) +
-  labs(
-    title = "Project lifespan",
-    subtitle = "Time between first and latest commit, by repository",
-    x = "Days", y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/lifespan-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-lifespan-vs-rate}
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = "Project lifespan vs. average commit rate",
-    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
-    caption = NULL
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-total.png", width = 127, height = 60, units = "mm")
-
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = NULL,
-    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-contrib.png", width = 127, height = 60, units = "mm")
-
-```
-
-------------------------------------------------------------------------
-
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
-
-A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.
-
-```{r plot-gh_contributors-histogram}
-oarch %>%
-  drop_na(gh_repo) %>%
-  unnest(gh_contribs) %>%
-  group_by(gh_repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Repository contributors",
-    x = "Number of contributors", y = "Repositories",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-hist.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
-
-Of multi-contributor repositories, 92% have a third of contributions made by a single person;
-
-65% have half of all contributions made by a single person;
-
-and 36% have three quarters of all contributions made a single person.
-
-**ZB NOTE:** Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic.
-
-```{r plot-contrib-dist}
-oarch %>%
-  unnest(gh_contribs) %>%
-  drop_na(contributor) %>%
-  group_by(item_name) %>%
-  mutate(n_contributors = n()) %>%
-  filter(n_contributors > 1) %>%
-  mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Leading contributor", "Other contributors")) %>% 
-  group_by(item_name, big_boy) %>%
-  summarise(contributions = sum(contributions)) %>%
-  mutate(
-    p_contributions = contributions / sum(contributions),
-    p_big_boy = first(p_contributions)
-  ) %>%
-  ungroup() %>%
-  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
-  ggplot(aes(x = item_name, y = contributions,
-             fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Leading contributor")))) +
-  geom_col(position = "fill", colour = "white", size = 0.1) +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
-  labs(title = "Repository contributors by proportion of contributions",
-       x = NULL, y = "Contributions",
-       fill = NULL,
-       caption = "Data: open-archaeo.info and GitHub API") +
-  coord_flip() +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-dist.png", width = 127, height = 127, units = "mm")
-```
-
-Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-
-Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-
-**ZB NOTE:** Recalcute these numbers.
-
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
-
-```{r plot-issues-hist}
-oarch %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of issues",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of issues", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/issues-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-comments-hist}
-oarch %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of comments",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of comments", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/comments-hist.png", width = 127, height = 127, units = "mm")
-```
-
-### Formation of collaborative development communities
-
 We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
-
 We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
 
-```{r graph-data}
-oarch |>
-  drop_na(gh_repo) |>
-  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
-  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
-  oarch_graph
-
-oarch_graph <- bind_rows(
-  oarch_graph |>
-    unnest(gh_commits) |>
-    drop_na(author) |> # Apparently real commits, but we can't connect them
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "commit", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_comments) |>
-    drop_na(author) |> # Repos with no comments
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "comment", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_issues) |>
-    drop_na(author) |> # Repos with no issues
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "issue", n_contrib = n(), .groups = "drop")
-)
-
-# Cleaning
-oarch_graph |>
-  rename(user = author) |>
-  # Remove non-human users
-  filter(!str_detect(user, coll("[bot]"))) |>
-  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
-                      "gitter-badger", "dependabot-support")) |>
-  # Remove big, non-archaeology projects that flood the network
-  filter(!repo %in% c("Stellarium/stellarium", 
-                     "opengisch/QField",
-                     "lutraconsulting/input")) ->
-  oarch_graph
-```
+{r graph-data}
 
@@
 
-#### Repo--user
+{r plot-graph-repo-user}
 
-```{r plot-graph-repo-user}
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
-
-g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt") +
-  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
-  geom_node_point(aes(colour = node_type)) +
-  scale_edge_alpha(trans = "log")
-
-#g_repo_user
-girafe(g_repo_user)
-```
-
-#### Repo--repo
-
+### repo-repo
 We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
@@
 
-**ZB NOTE:** Remove the cells titled `plot-graph-repo-repo` and `plot-graph-repo-repo-louvain`.
-Use patchwork to compare sets of graphs and dendrograms with links based on all contributions, only commits, and only issues and comments.
+{r plot-graph-repo-repo-edge-betweenness}
 
-```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
+{r plot-graph-repo-repo-edge-betweenness-dendrogram}
 
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
+* Identify the repos/organisations that are the focus of the most intense collaboration?
+  * Are they thematically or topically linked?
+  * Do they map onto real-world research centres?
+* Is the network a new thing, or just a translation of existing offline networks (e.g. fields, institutions) onto GitHub?
+  * cf. is open source really open?
 
-```{r plot-graph-repo-repo-edge-betweenness}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  #geom_node_point(aes(colour = group)) +
-  #geom_node_text(aes(label = group)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-
-girafe(ggobj = iplot)
-```
-
-```{r plot-graph-repo-repo-edge-betweenness-dendrogram}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = TRUE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
-xx <- as.dendrogram(xx)
-plot(xx)
-
-ggraph(xx, 'dendrogram', height = height)
-
-```
-
-```{r plot-graph-repo-repo-louvain}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_louvain())) ->
-  oarch_graph_repo_repo
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = group)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-#### User--user
+### user-user
 
@@
 
-**ZB NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits or something else.
-
 We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
@@
 
-**ZB NOTE:** Look at the histogram that maps the distribution of betweenness scores across all users.
-Interesting double hump of low-moderate betweenness scores, which may correspond with distinct or protruding components.
+{r graph-user-user}
 
-```{r graph-user-user}
-oarch_graph_user_user <- tbl_graph(
-  nodes = nodes_users(oarch_graph), 
-  edges = edges_common_repos(oarch_graph),
-  directed = FALSE
-)
-```
+{r summary-graph-user-user}
 
-```{r summary-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), 
-         betweenness = centrality_betweenness(weights = n), 
-         strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |> 
-  data.frame() |> 
-  View()
-```
+{r plot-graph-user-user}
 
-```{r plot-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  # geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common repositories")
-```
+{r plot-user-betweenness-dist}
 
-#### Over time
+{r plot-user-betweenness-activity}
 
-Could animate the graphs??
-
-## Conclusions
-
-Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.
-
-In terms of collaboration, our analysis shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host.
-
-The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base.
-
-Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects.
-
-This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
-
-We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.
-
----
-
-#### Notes from January 30, end of day
-In cleaning the data we made a few significant decisions that revealed insights into our dataset. 
-We constructed the user-user graph and filtered out several bots. 
-We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community. 
-We filtered isolates. 
-We calculated the degree and betweenness to assess whether this was realistic representation of what we know. 
-We wanted to figure out who are the key nodes, what they had in common. 
-The people who are high betweenness are people who primarily produce computational archaeology code as a job. 
-We can investgate further by comparing with some other measures of activity, like volume of commits or something else. 
-We also looked at the same thing, but with only commits, or without commits. 
-For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.
+* What proportion of users and repos are connected to this network?
+* What predicts membership in this network?
+* Does centrality in the network correlate with general rate of activity?
+*  Who are the most central nodes in this network? Is there anything they have in common?
+   * High activity?
+   * Certain platforms/categories/languages?
+* Qualitative analysis of top 50? 100? most connected users
+  * Career stage? Precarity? Employer? Location?

### 2023-03-28 analysis/99-conclusion.qmd
Copying info from paper.qmd and organizing it for more productive work

--- 2023-01-31 analysis/paper.qmd
@@
----
-title: "openarchaeoCollaboration"
-author:
-  - name: Zack Batist
-    url: https://zackbatist.github.io
-    affiliation: University of Toronto
-  - name: Joe Roe
-    url: https://joeroe.io
-    affiliation: University of Bern
-date: "`r Sys.Date()`"
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	message = FALSE,
-	warning = FALSE
-)
-```
-
-```{r dependencies, include=FALSE}
-library("tidyverse")
-library("lubridate")
-library("openarchaeoCollaboration")
-library("khroma")
-library("gt")
-library("fs")
-library("here")
-library("tidygraph")
-library("ggraph")
-library("ggiraph")
-```
-
-```{r plot-theme}
-theme_oarch <- function() theme(
-)
-```
-
-```{r data-oarch}
-read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
-  mutate(category = na_if(category, "?")) ->
-  oarch
-
-# Extract clean repos from GitHub URLs
-oarch %>%
-  mutate(
-    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
-
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
-```
-
-```{r data-github}
-# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
-# chunk will prefer to load it rather than make requests to the GitHub API.
-# Delete this file to fetch fresh data from the GitHub or modify 
-# `cached_gh_data` to read a different cached version.
-#
-# Note that fetching data from the GitHub API takes some time and is subject to
-# a rate limit:
-# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
-#
-# You will need to authenticate authenticate with GitHub credentials to get a
-# rate limit sufficient to obtain the required data, see 
-# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
-# for options on how to do this.
-cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
-if (!file_exists(cached_gh_data)) {
-  oarch <- mutate(oarch,
-                  gh_langs = map(oarch$gh_repo, gh_lang),
-                  gh_contribs = map(oarch$gh_repo, gh_contrib),
-                  gh_issues = map(oarch$gh_repo, gh_issue),
-                  gh_comments = map(oarch$gh_repo, gh_comment),
-                  gh_commits = map(oarch$gh_repo, gh_commit)
-  )
-  saveRDS(oarch, cached_gh_data)
-}
-oarch <- readRDS(cached_gh_data)
-```
-
-## Introduction
-
-In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools to serve our own unique purposes, which are not typically unanticipated by their original designers.
-In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways.
-For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase.
-However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems.
-In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes.
-
-This article presents a survey of archaeological software development with two goals in mind:
-
-(1) we identify *what* kinds of software archaeologists are making;
-(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
-
-Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
-
-## Background
-
--   The notion of the archaeological toolbox, and its inclusion of digital tools and skillsets
-    -   The use of off-the-shelf tools like MS Access, Excel, Agisoft PhotoScan, Dropbox, etc
-    -   Link to related factors:
-        -   Lack of experienced programmers among archaeologists
-        -   The built-in flexibility of these tools
-        -   Ease of administrative purchasing through university departments
-    -   What changed?
-        -   Software is expensive, shift to subscription models, locking into a dependency on a commercial service provider
-        -   Archaeologists are better at computer programming, digital literacy on the rise, digital archaeologists grows into its own legitimate sub-discipline
-        -   Open source is cool, everyone has a github profile (even if not really)
-            -   (JR:) People are presumably guided towards GitHub (or at least git) via other routes, because it's seen as the standard platform for 'open' digital work.
-For example, the literature on reproducible compendiums recommends GitHub, and almost all R development work is now done there.
-And once there, you're sort of forced into engaging in a (particular set of) open source development.
-I know there's controversy about how dominant this has become and people who say it's not 'true' (free and) open source, and perhaps we could add to that the question of whether it's the right model for archaeology.
--   About open source practices and their influence on archaeology
-    -   What is open source software and open source software development?
-        -   Free cost as an end-user
-        -   Free to manipulate and extend
-        -   Community-based support and development
-    -   Software as product vs software as collective process
-        -   Basically what distinguishes abandonware from thriving software ecosystems
-        -   Technical, legal and social notions of open
-            -   Technical: source code is open, use of a platform and systems that support software integration and implementations
-            -   Legal: open licenses, backing by the courts
-            -   Social: inviting community, meaningful contributions provided by users, common vision and values directing the project into future directions
-        -   Consider how these aspects of open have been implemented or prioritized by digital archaeologists in different ways
-            -   Technical: reproducible research, archiving and publishing
-            -   Legal: archiving and publishing, budgeting for projects
-            -   Social: ensuring that software is actually useful for general use and not just one-off implementation, coordinating to reduce redundant effort and not reinvent the wheel, writing documentation, making it fit within existing systems or infrastructures (i.e.
-allocation of credit, citing software directly versus articles that present software, sustainable development after funding runs out, etc)
-        -   This paper largely focuses on the social aspect of open, which is arguably the most important for ensuring a positive future for archaeological software development
-
-The 'archaeological toolbox' is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely upon in their work.
-
-collaborate in their programming work, and how they experience open source development processes which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.
-
-## Data and evidence
-
-Our findings here are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
-As of right now it includes 408 items, which are nicely rendered at open-archaeo.info.
-
-The main criteria for inclusion is whether a project is identified by its maintainers as relating to archaeological work.
-
-The dataset was compiled by browsing collaborative software development platforms, essentially manually crawling through the profiles of users who identify as archaeologists or who have contributed to tools that pertain to archaeological work.
-
-This quasi-systematic collection strategy is also supplemented by contributions made by interested individuals who identified relevant work that was initially overlooked.
-
-Open-archaeo is a relatively comprehensive list, however it generally lacks code written before the growth and general use of collaborative software development platforms, such as GitHub, among archaeologists, and also lacks references to code that is not available on the web.
-
-It is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in your field, to contribute.
-
-Please get in touch if you're interested in adding to the list or doing a deep dive relating to your area of expertise!
-
-### Tags and categories
-
-Each record is also annotated with tags that describe the aspects of archaeological work that each tool contributes to.
-
-Items are tagged based on how they have been generally identified or presented by their maintainers.
-
-The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.
-
-Right now there are 63 tags.
-
-Records are also categorized based on how each tool or resource is accessed or used.
-
-This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-
-In other words, categories hint at some of the decisions that developers make regarding how they want their tools to be used and accessed.
-
-| Category                              | Kind      | Scope                                                                                                                                                                                                        |
-|:-----------|:-----------|:------------------------------------------------|
-| Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-|
-| Standalone software                   | Software  | Software that may be operated without needing to first access an underlying platform.
-|
-| Scripts                               | Software  | Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-|
-| Specifications, protocols and schemas | Documents | A formal data structure or framework intended to be used as a model.
-|
-| Products                              | Documents | Stable outcomes of creative work.
-|
-| Guides                                | Documents | An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.
-|
-| Lists and datasets                    | Documents | A series of consistently organized observations assembled with purpose.
-|
-
-Two kinds of categories have been defined: software and documents.
-
-Software comprise collections of instructions that dictate how a computer should operate, and documents are bounded sets of recorded information.
-
-Three categories of software and four categories of documents are described in the table.
-
-Packages and libraries are sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-
-Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.
-
-Standalone software are tools that may be operated without needing to first access an underlying platform.
-
-And the third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-Things like code generalized sets of code snippets.
-
-As for documents, specifications, protocols and schemas are formal data structures or frameworks intended to be used as models.
-
-Products are stable outcomes of creative work, loosely defined.
-
-Guides are educational resources or documented protocols meant to instruct readers how to apply relevant tools or techniques.
-
-And finally, lists or datasets, including open-archaeo itself, are series of consistently organized observations assembled with purpose.
-
-## Findings
-
-### What archaeologists are making
-
-Our first set of analyses were focused on these questions:
-
-What kinds of software are archaeologists making?
-
-And what are the general purposes of their software?
-
-------------------------------------------------------------------------
-
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing 'platform' -- for example a package for a programming language like R, or a plugin for an application like QGIS.
-
-The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
-
-Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-
-A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
-
-```{r plot-categories}
-oarch <- mutate(oarch,
-                category = recode(category,
-                                  "Specifications, protocols and schemas" = "Specifications etc.",
-                                  .default = category))
-oarch %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(n = n()) %>%
-  ggplot(aes(fct_reorder(category, n), n)) +
-  geom_col() +
-  coord_flip() +
-  labs(
-    title = "Most popular categories",
-    subtitle = "By number of projects",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-R is overwhelmingly dominant as a platform, closely followed by Python.
-
-Plugins for QGIS and mobile apps are also relatively popular.
-
-Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
-AutoCAD, ArcGIS); a number of lesser used programming languages; and a genre consisting of custom forms and spreadsheet templates.
-
-In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
-The larger platforms tend to be more diverse.
-
-```{r plot-platforms}
-oarch %>%
-  drop_na(platform) %>%
-  group_by(platform, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(
-    title = "Most popular platforms",
-    subtitle = "By number of projects, split by primary author",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-If we compare categories with tags, we can see the general domains that each kind of resource is designed to serve.
-
-We see that packages are fairly common across the board.
-
-Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
-
-These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.
-
-```{r plot-tags-cats}
-detail_tags <- c("Instrumental Neutron activation analysis",
-                 "Harris Matrix",
-                 "aDNA Simulators",
-                 "Damage profilers",
-                 "Metagenomic classifers",
-                 "aDNA simulators",
-                 "aDNA damage-aware genotypers",
-                 "Low-coverage contamination estimation",
-                 "Low-coverage adapted population-genetics tools",
-                 "aDNA preprocessing",
-                 "aDNA aligner",
-                 "aDNA assembler",
-                 "aDNA toolkit")
-
-oarch %>%
-  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  filter(!tag %in% detail_tags) %>%
-  mutate(category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(tag, category) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  # facet_wrap(vars(category)) +
-  geom_col() +
-  scale_fill_bright() +
-  coord_flip() +
-  labs(
-    title = "Tags by category",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = "Number of projects",
-    fill = NULL
-  ) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_text(size = 7)
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/tags-cats.png", width = 254, height = 127, units = "mm")
-```
-
-**ZB NOTE:** Do a temporal analysis to get a sense of when platforms, tags and categories have been gaining popularity.
-There may not be a need for extreme detail here, but simply noting the dates of first commits might be helpful.
-I'm imagining a stacked bar chart, with each bar representing a month or a year, and with the stacked portions of each bar representing a platform or category.
-
-### Development processes
-
-While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about *how* archaeologists develop software.
-
-The remainder of the paper is driven by a concern about how archaeologists collaborate in their programming work.
-
-How is their work sustained, and how does it leverage distributed work processes that are a key aspect of open source software development in general?
-
-To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
-
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh].
-
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.
-
-```{r plot-repo-hosts}
-# TODO: turn into table or just text
-oarch <- mutate(oarch, none = ifelse(is.na(github) &
-                                       is.na(gitlab) &
-                                       is.na(bitbucket) &
-                                       is.na(launchpad) &
-                                       is.na(codeberg),
-                                     "", NA))
-
-oarch %>%
-  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
-               names_to = "repo_host", values_to = "repo_url") %>%
-  drop_na(repo_url) %>%
-  group_by(repo_host, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ungroup() %>%
-  mutate(repo_host = fct_reorder(repo_host, total),
-         repo_host = fct_relevel(repo_host, "none")) %>%
-  ggplot(aes(repo_host, n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(title = "Most popular repository hosts",
-       subtitle = "By number of projects, split by primary author",
-       caption = "Data: open-archaeo.info",
-       x = NULL, y = NULL)
-
-# Export for slides (half slide)
-# ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
-```
-
-```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
-```
-
-We obtained this data in R using the {gh} package; those data, and the analysis code, are available in our research compendium on GitHub.
-
-------------------------------------------------------------------------
-
-Starting with a relatively simple plot, this is a chart of the cumulative growth of open archaeology over the years, measured by code contributions committed and pushed to GitHub repositories.
-
-As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
-But they really began to get into it around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
-
-Around this time we also see that GitHub starts being used to host documents and scripts.
-This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.
-
-```{r plot-gh_commits-cum}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  mutate(lumped_category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cat_total = sum(n_commits)) %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = fct_reorder(lumped_category, cat_total),
-             colour = fct_reorder(lumped_category, cat_total))) +
-  geom_area() +
-  scale_x_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years",
-                   date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = NULL) +
-  scale_fill_bright(guide = guide_legend(reverse = TRUE)) +
-  labs(
-    title = "Growth of open archaeology",
-    subtitle = "Measured by cumulative commits in GitHub repositories",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = NULL, y = NULL, fill = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (full slide)
-# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-But how is this activity distributed across repositories?
-
-This visualizes the 'pulse' of active projects.
-
-Each vertical line represents the lifespan of a repository; the points on the line are individual commits.
-
-This is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
-
-```{r plot-lifespan}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(datetime) %>%
-  group_by(item_name) %>%
-  add_tally() %>%
-  filter(n > 1) %>%
-  mutate(first_commit = min(datetime),
-         mid_commit = median(datetime),
-         last_commit = max(datetime)) %>%
-  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
-             y = datetime,
-             ymin = first_commit,
-             ymax = last_commit)) +
-  geom_linerange(size = 0.05) +
-  geom_point(size = 0.1) +
-  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years", date_labels = "%Y") +
-  # coord_flip() +
-  labs(
-    title = "Project lifespan",
-    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
-    x = NULL, y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch() +
-  theme(
-    axis.line.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    panel.grid.major.x = element_blank()
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-The span of time between the first and latest commits is visualized more clearly in summary here.
-
-Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
-
-Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.
-
-Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects.
-
-The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e.
-the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.
-
-But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
-
-This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.
-
-**ZB NOTE:** Use patchwork package to combine these plots into a unified figure.
-
-```{r lifespan-stats}
-oarch %>% 
-  unnest(gh_commits) %>% 
-  drop_na(datetime) %>% 
-  group_by(item_name) %>%
-  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
-  summarise(
-    category = first(category),
-    n_contribs = first(n_contribs),
-    n = n(),
-    first_commit = min(datetime),
-    last_commit = max(datetime),
-    lifespan = as.integer(last_commit - first_commit),
-    commit_rate = n / lifespan
-  ) ->
-  oarch_lifespan
-```
-
-```{r plot-lifespan-hist}
-# TODO: where does oarch_lifespan come from?
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan)) +
-  geom_histogram(binwidth = 100) +
-  scale_x_continuous(breaks = scales::breaks_width(1000)) +
-  labs(
-    title = "Project lifespan",
-    subtitle = "Time between first and latest commit, by repository",
-    x = "Days", y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/lifespan-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-lifespan-vs-rate}
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = "Project lifespan vs. average commit rate",
-    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
-    caption = NULL
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-total.png", width = 127, height = 60, units = "mm")
-
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = NULL,
-    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-contrib.png", width = 127, height = 60, units = "mm")
-
-```
-
-------------------------------------------------------------------------
-
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
-
-A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.
-
-```{r plot-gh_contributors-histogram}
-oarch %>%
-  drop_na(gh_repo) %>%
-  unnest(gh_contribs) %>%
-  group_by(gh_repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Repository contributors",
-    x = "Number of contributors", y = "Repositories",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-hist.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
-
-Of multi-contributor repositories, 92% have a third of contributions made by a single person;
-
-65% have half of all contributions made by a single person;
-
-and 36% have three quarters of all contributions made a single person.
-
-**ZB NOTE:** Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic.
-
-```{r plot-contrib-dist}
-oarch %>%
-  unnest(gh_contribs) %>%
-  drop_na(contributor) %>%
-  group_by(item_name) %>%
-  mutate(n_contributors = n()) %>%
-  filter(n_contributors > 1) %>%
-  mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Leading contributor", "Other contributors")) %>% 
-  group_by(item_name, big_boy) %>%
-  summarise(contributions = sum(contributions)) %>%
-  mutate(
-    p_contributions = contributions / sum(contributions),
-    p_big_boy = first(p_contributions)
-  ) %>%
-  ungroup() %>%
-  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
-  ggplot(aes(x = item_name, y = contributions,
-             fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Leading contributor")))) +
-  geom_col(position = "fill", colour = "white", size = 0.1) +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
-  labs(title = "Repository contributors by proportion of contributions",
-       x = NULL, y = "Contributions",
-       fill = NULL,
-       caption = "Data: open-archaeo.info and GitHub API") +
-  coord_flip() +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-dist.png", width = 127, height = 127, units = "mm")
-```
-
-Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-
-Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-
-**ZB NOTE:** Recalcute these numbers.
-
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
-
-```{r plot-issues-hist}
-oarch %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of issues",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of issues", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/issues-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-comments-hist}
-oarch %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of comments",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of comments", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/comments-hist.png", width = 127, height = 127, units = "mm")
-```
-
-### Formation of collaborative development communities
-
-We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
-
-We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
-
-```{r graph-data}
-oarch |>
-  drop_na(gh_repo) |>
-  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
-  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
-  oarch_graph
-
-oarch_graph <- bind_rows(
-  oarch_graph |>
-    unnest(gh_commits) |>
-    drop_na(author) |> # Apparently real commits, but we can't connect them
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "commit", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_comments) |>
-    drop_na(author) |> # Repos with no comments
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "comment", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_issues) |>
-    drop_na(author) |> # Repos with no issues
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "issue", n_contrib = n(), .groups = "drop")
-)
-
-# Cleaning
-oarch_graph |>
-  rename(user = author) |>
-  # Remove non-human users
-  filter(!str_detect(user, coll("[bot]"))) |>
-  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
-                      "gitter-badger", "dependabot-support")) |>
-  # Remove big, non-archaeology projects that flood the network
-  filter(!repo %in% c("Stellarium/stellarium", 
-                     "opengisch/QField",
-                     "lutraconsulting/input")) ->
-  oarch_graph
-```
-
-We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations.
-We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
-
-#### Repo--user
-
-```{r plot-graph-repo-user}
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
-
-g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt") +
-  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
-  geom_node_point(aes(colour = node_type)) +
-  scale_edge_alpha(trans = "log")
-
-#g_repo_user
-girafe(g_repo_user)
-```
-
-#### Repo--repo
-
-We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate notes, which represent repositories with only a single contributor, we detected 21 clusters.
-While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
-
-Are the clusters associated with particular topics or tags?
-Are there some sorts of associations between clusters that break down from a parent in the community detection dendrogram?
-How does all of this differ when we differentiate commits from issues and comments?
-
-**ZB NOTE:** Remove the cells titled `plot-graph-repo-repo` and `plot-graph-repo-repo-louvain`.
-Use patchwork to compare sets of graphs and dendrograms with links based on all contributions, only commits, and only issues and comments.
-
-```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-```{r plot-graph-repo-repo-edge-betweenness}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  #geom_node_point(aes(colour = group)) +
-  #geom_node_text(aes(label = group)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-
-girafe(ggobj = iplot)
-```
-
-```{r plot-graph-repo-repo-edge-betweenness-dendrogram}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = TRUE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
-xx <- as.dendrogram(xx)
-plot(xx)
-
-ggraph(xx, 'dendrogram', height = height)
-
-```
-
-```{r plot-graph-repo-repo-louvain}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_louvain())) ->
-  oarch_graph_repo_repo
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = group)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-#### User--user
-
-We assembled graphs linking users based on common contributions to the same repositories.
-We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
-
-We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as a job.
-
-**ZB NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits or something else.
-
-We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
-For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the all-contributions graph also appear here.
-These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
-
-**ZB NOTE:** Look at the histogram that maps the distribution of betweenness scores across all users.
-Interesting double hump of low-moderate betweenness scores, which may correspond with distinct or protruding components.
-
-```{r graph-user-user}
-oarch_graph_user_user <- tbl_graph(
-  nodes = nodes_users(oarch_graph), 
-  edges = edges_common_repos(oarch_graph),
-  directed = FALSE
-)
-```
-
-```{r summary-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), 
-         betweenness = centrality_betweenness(weights = n), 
-         strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |> 
-  data.frame() |> 
-  View()
-```
-
-```{r plot-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  # geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common repositories")
-```
-
-#### Over time
-
-Could animate the graphs??
-
-## Conclusions
-
 Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.
@@
 
-#### Notes from January 30, end of day
-In cleaning the data we made a few significant decisions that revealed insights into our dataset. 
-We constructed the user-user graph and filtered out several bots. 
-We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community. 
-We filtered isolates. 
-We calculated the degree and betweenness to assess whether this was realistic representation of what we know. 
-We wanted to figure out who are the key nodes, what they had in common. 
-The people who are high betweenness are people who primarily produce computational archaeology code as a job. 
-We can investgate further by comparing with some other measures of activity, like volume of commits or something else. 
-We also looked at the same thing, but with only commits, or without commits. 
-For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.
+There is an emerging community of practice around open source in archaeology.
+
+But:
+
+* It is limited and fragmented
+  * A small set of subfields are represented
+  * Probably most people still haven't heard it?
+* Collaboration is limited
+  * Most work remains solo & short-lived
+  * Perhaps because open source norms conflict with academic norms?
+* The members of this community face structural challenges in academia

### 2023-03-29 analysis/02-background.qmd
Initial and probably overly ranty draft of the background section.

* What is open source?

Open source is a software development model that prioritizes transparent work processes. Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development (Kelty 2008: chapter 3, especially starting at page 99). By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base. Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision. An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.

Open source has traditionally been referred to as being based on meritocratic principles. A good test of whether a contribution should be included in a published software release is whether it is functional. Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues. This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion. Personal ego is minimized in favour of co-creating stable and functional outcomes. (Coleman 2012)

However, this is not the same as saying that open source is completely anarchic or based on the "wisdom of crowds". In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants. They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term. Open source is more than just putting your code online; to be successful, it requires participation in a social experience. (Kelty 2008; O'Neill 2009; Ratto dissertation and derived works)

In other words, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are grossly underappreciated factors that contribute to an open source project's success.

We therefore consider open source to be a means of collaboration more than a means of transmitting information. It involves developing software as part of a group, developing consensus, and working with common purpose. Crucially, it also involves having a welcoming attitude, a sense of humility, and an understanding that one's work may be appropriated and used in unanticipated ways.

* How does open source relate to open science?

The open science movement comprises a series of practices and principles intended to make research more accessible, transparent, and efficient. Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" (https://opendefinition.org/). This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement. However, most open science advocates claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.

The idea that scientistis should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production. However, while open science does have the _potential_ to effect radical change, this is not a given. The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals. Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails. Namely, it considers science as the accumulation and assembly of a species-level understanding of the world, which is not held by any one individual but is stored in seemingly value-neutral and disembodied media, facts and observations. This is manifested by digital telecommunications systems that host files, document processes, facilitate co-working opportunities, and perform automated processes. However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.

Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain. But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce. Whereas open source emerged from concern of consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible. Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (inlcuding: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions. Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience. Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of tenure and promotion. Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out. Once a project is completed and credit has been allocated, it is common for scientists to mark their projects as completed and move onto new endeavours. Open source projects, on the other hand, are motivated by a more practical need for the software to function properly, and contributors may remain actively or sporadically involved to satisfy users' needs.

* Situating open archaeology in open science

The "archaeological toolbox" is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely on to achieve their goals. In applying either physical or digital methods, archaeology has a long tradition of adopting common or "off the shelf" tools to serve their own unique purposes, which are not typically unanticipated by their original designers. In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways. For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase. However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems. In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes. This software is typically made to support tasks that archaeologists find uniquely meaningful, and which are not supported by mainstream technology companies. This differs from the use of purpose-specific scripts written to address specific research questions. Such work often falls under the purview of open science, since making one's scripts openly available is said to enhance the potential for reproducibility (Marwick and Schmidt).

The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings. Open science advocates have put a lot of effort into developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale, but, at least among archaeologists, their transformative value has not yet been realized. For instance, numerous studies have highlighted how archaeologists are not reusing data published on open data platforms, and instead opt to communicate directly with those who created the dataset to obtain a relationship that more closely resembles a more traditional form of collaboration (Huggett 2018; Atici 2013; Faniel 2013; Sobotkova 2018; Opitz 2021). In other words, these pltforms do little to alter the professional norms and expectations that frame archaeological research, and which seem to be fundamental factors that influence the nature of archaeological research. Open data platforms may provide the capacity and the potential to enable change, but do not in themselves change the ways in which archaeologists create and value research outcomes. This must come from community buy-in, and organic extension and establishing of novel forms of collaborative commitments that reflect the ideal state of affairs. In other words, it is a social problem, not a technical one.

Similarly, archaeological software development is framed by the circumstances under which research that supports and/or motivates the effort occurs. This means that software development processes are shaped by the pressures and circumstances of being an archaeologist, such as the need to publish and to generate stable research outputs, archaeologists' tendency to move on from projects after they are 'completed' (i.e. when money runs out and people lose interest), the need to be recognized as a citable author for purposes of hiring and promotion, and the general lack of formal digital literacy among archaeologists (which either produces a digital archaeology subdiscipline, a kind of ad-hoc cobbled together system (made by the most digitally literate person in the room), or cross-disciplinary connections with non-archaeologists (spouses, childhood friends, maybe a computer scientists, rarely someone in private industry)). (Morgan 2015; Morgan and Eve 2012; Batist et al 2021; Caraher 2022; Graham 2020; maybe also Thorpe 2012 and Zorzin 2015 on economic models that frame (non-digital) archaeological research)

A lot of work has gone in to developing information infrastructures, or systems designed to facilitate the transmission of open materials on a global scale, including open data repositories, open peer-review platforms, and online social spaces meant to foster participation in new forms of scholarly communication. However, the general failure of open science to deliver on its lofty promises has led archaeologists to reflect on what openness really means and entails. In the past, open archaeology was very much wrapped up in discourse on the best tools to use, optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards (e.g., establishing proper and secure web servers, open api practices, using orcid, proper licencing, using markdown, git, R package development practices, etc) (cf. Kansa and Kansa 2014; Kintigh 2014; Roosevelt 2015; others), and this continues to be a prominent part of discourse on the development of open infrastructures (cf. Kansa on open identifiers; others). However, more recent reflections have attempted to re-frame research practices, particularly those corresponding with "data work", as situated experiences and associated with a continuum of decisions made within pragmatic contexts of action. Archaeological research is thus framed as inherently collaborative, and the materials and media that archaeologists rely on to do their research are framed as discursive entities, or communicatve media that bridge a gap of experience or understanding. Bridging this gap entails an act of reconciliation between different pragmatic experiences, and the technical infrastructures we have set up to handle these acts of communication do not adequately account for the range of factors that must be considered to successfully pass this reconciliation.

### 2023-03-30 analysis/02-background.qmd
Re-wrote the third section with a more positive tone

@@
-* What is open source?
+## What is open source?
 
@@
 
-* How does open source relate to open science?
+## How does open source relate to open science?
 
@@
 
-* Situating open archaeology in open science
-
-The "archaeological toolbox" is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely on to achieve their goals.
-In applying either physical or digital methods, archaeology has a long tradition of adopting common or "off the shelf" tools to serve their own unique purposes, which are not typically unanticipated by their original designers.
-In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways.
-For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase.
-However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems.
-In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes.
-This software is typically made to support tasks that archaeologists find uniquely meaningful, and which are not supported by mainstream technology companies.
-This differs from the use of purpose-specific scripts written to address specific research questions.
-Such work often falls under the purview of open science, since making one's scripts openly available is said to enhance the potential for reproducibility (Marwick and Schmidt).
+## Open archaeology at a crossroads
 
 The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings.
-Open science advocates have put a lot of effort into developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale, but, at least among archaeologists, their transformative value has not yet been realized.
-For instance, numerous studies have highlighted how archaeologists are not reusing data published on open data platforms, and instead opt to communicate directly with those who created the dataset to obtain a relationship that more closely resembles a more traditional form of collaboration (Huggett 2018; Atici 2013; Faniel 2013; Sobotkova 2018; Opitz 2021).
-In other words, these pltforms do little to alter the professional norms and expectations that frame archaeological research, and which seem to be fundamental factors that influence the nature of archaeological research.
-Open data platforms may provide the capacity and the potential to enable change, but do not in themselves change the ways in which archaeologists create and value research outcomes.
-This must come from community buy-in, and organic extension and establishing of novel forms of collaborative commitments that reflect the ideal state of affairs.
-In other words, it is a social problem, not a technical one.
+Open science advocates, including archaeologists, have put a lot of effort into developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale.
+And as with open science in general, open archaeology has been wrapped up in discourse on the best tools to use, optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards (cf.
+Kansa and Kansa 2014; Kintigh 2014; Roosevelt 2015).
+This continues to be a prominent part of discourse on the development of open infrastructures (cf.
+Kansa 2022 on unique  identifiers; something about aDNA pipelines; Hacıgüzeller 2021 on CIDOC-CRM (though she is actually very critical); Huvila and his team on formal paradata), but more recent reflections on the use of these infrastructures recognize underlying and systemtic commitments and assumptions that inform how these systems are used and valued by various stakeholders.
+For instance, the pressures and circumstances of being an archaeologist, such as the need to publish stable research outputs, and archaeologists' tendency to move on from projects after they are completed, contribute towards shaping data sharing platforms as delegated archival services, whereupon the obligation to care for data over the long-term is passed on to others.
+In response to criticisms that highlight how these systems add little practical downstream value (Huggett 2018; Sobotkova 2018), open archaeology advocates have begun to prioritize the role that information infrastructures may play in contexts of data reuse, namely by implementing the FAIR principles in order to ensure that data are findable, accessible, interoperable, and reusable (Kansa 2023 in that recent special issue in Advances in Archaeological Practice; one or two additional sources).
+This example is matched by similar efforts to raise the value proposition of using open science tools and systems, such as through an emphasis on enhancing R package development and documentation skills so that code may be run and evaluated in alternative computing environments (cite Marwick and/or Roe), and [another example].
 
-Similarly, archaeological software development is framed by the circumstances under which research that supports and/or motivates the effort occurs.
-This means that software development processes are shaped by the pressures and circumstances of being an archaeologist, such as the need to publish and to generate stable research outputs, archaeologists' tendency to move on from projects after they are 'completed' (i.e.
-when money runs out and people lose interest), the need to be recognized as a citable author for purposes of hiring and promotion, and the general lack of formal digital literacy among archaeologists (which either produces a digital archaeology subdiscipline, a kind of ad-hoc cobbled together system (made by the most digitally literate person in the room), or cross-disciplinary connections with non-archaeologists (spouses, childhood friends, maybe a computer scientists, rarely someone in private industry)).
-(Morgan 2015; Morgan and Eve 2012; Batist et al 2021; Caraher 2022; Graham 2020; maybe also Thorpe 2012 and Zorzin 2015 on economic models that frame (non-digital) archaeological research)
-
-A lot of work has gone in to developing information infrastructures, or systems designed to facilitate the transmission of open materials on a global scale, including open data repositories, open peer-review platforms, and online social spaces meant to foster participation in new forms of scholarly communication.
-However, the general failure of open science to deliver on its lofty promises has led archaeologists to reflect on what openness really means and entails.
-In the past, open archaeology was very much wrapped up in discourse on the best tools to use, optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards (e.g., establishing proper and secure web servers, open api practices, using orcid, proper licencing, using markdown, git, R package development practices, etc) (cf.
-Kansa and Kansa 2014; Kintigh 2014; Roosevelt 2015; others), and this continues to be a prominent part of discourse on the development of open infrastructures (cf.
-Kansa on open identifiers; others).
-However, more recent reflections have attempted to re-frame research practices, particularly those corresponding with "data work", as situated experiences and associated with a continuum of decisions made within pragmatic contexts of action.
-Archaeological research is thus framed as inherently collaborative, and the materials and media that archaeologists rely on to do their research are framed as discursive entities, or communicatve media that bridge a gap of experience or understanding.
-Bridging this gap entails an act of reconciliation between different pragmatic experiences, and the technical infrastructures we have set up to handle these acts of communication do not adequately account for the range of factors that must be considered to successfully pass this reconciliation.
+Additionally, archaeologists are becoming more comfortable showcasing and actively discussing undervalued research ouputs such as software and datasets, which are reviewed and published by venues like Peer Community In, the Journal of Archaeological Data, and the Journal of Open Source Software.
+Professional blogs, social media and podcasts are also gaining traction as proper venues for disseminating one's work.
+This is consistent with archaeology's historical willingness to experiment with new methods and media and its casual attitude towards scholarly communication.
+Archaeologists' embrace of methodological plurality and their positions as interdisciplinary scholars scattered across institutional boundaries has prepared them for this kind of pragmatic, improvisatory approach to scholarly communication via the web.
+There is enormous potential for archaeologists to reinvent novel forms of scholarly communication and extend a new set of collaborative commitments that go beyond those typically recognized by scholarly institutions, and which correspond with open science principles.
+But this depends on archaeologists' willingness to continue experimenting and on the community's willingness to buy into novel value regimes.

### 2023-03-31 analysis/06-project_activity.qmd
Updating the text of analysis and interpretation.
Re-wording for more active voice, adding some interpretation (specifically to section 07), and updates the notes for further work to be done.

@@
-In this section we assess projects' pulse, or the amount of activity pertaining to a repository.
-In the following figure, each vertical line represents the lifespan of a repository, and the points on the line are individual commits.
+But how is this activity distributed across repositories? In this section we assess projects' pulse, or the amount of activity pertaining to a repository.
+
+In the following figure, each vertical line represents a repository's lifespan, and the points on each line are individual commits.
 The graph is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
@@
 {r plot-lifespan}
+{r lifespan-stats}
 
-The span of time between the first and latest commits is visualized more clearly in summary here.
-Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
+The span of time between the first and latest commits is visualized more clearly in the following charts.
+Notably, many projects are active for only a short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
 Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.
@@
 
-{r lifespan-stats}
-
+**NOTE:** Combine these figures using patchwork.
 {r plot-lifespan-hist}
-
 {r plot-lifespan-vs-rate}
@@
 Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
-A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.
+A "contribution", in this case, is a commit, issue, or comment applied by a user to a repository.
 
@@
 Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
-Of multi-contributor repositories, 92% have a third of contributions made by a single person; 65% have half of all contributions made by a single person; and 36% have three quarters of all contributions made a single person.
+Of multi-contributor repositories, 92% have a third of contributions made by a single person, 65% have half of all contributions made by a single person, and 36% have three quarters of all contributions made a single person.
+
+**NOTE:** Re-calculate these numbers to ensure they are up to date.
+Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic, which are now divided into quarters rather than thirds.
 
@@
 
+**NOTE:** Combine these figures using patchwork.
 {r plot-issues-hist}
-
 {r plot-comments-hist}

### 2023-03-31 analysis/07-network.qmd
Updating the text of analysis and interpretation.
Re-wording for more active voice, adding some interpretation (specifically to section 07), and updates the notes for further work to be done.

@@
-We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
+We constructed and analyzed networks representing users' contributions to GitHub repositories to examine the emergence of collaborative tendencies.
 We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
@@
 We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations.
-We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
 
@@
 
-### repo-repo
+We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
+
+**NOTE:** Combine these figures using patchwork.
+**NOTE:** Integrate {r graph-user-user} and {r plot-graph-user-user} into a single cell, similar to how {r plot-graph-repo-repo} integrates both components.
+
+{r plot-graph-repo-repo}
+{r graph-user-user}
+{r plot-graph-user-user}
+
+## repo-repo
 We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate notes, which represent repositories with only a single contributor, we detected 21 clusters.
+Aside from isolate nodes, which represent repositories with only a single contributor, we detected 21 clusters.
 While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
 
-Are the clusters associated with particular topics or tags?
-Are there some sorts of associations between clusters that break down from a parent in the community detection dendrogram?
-How does all of this differ when we differentiate commits from issues and comments?
-
 {r plot-graph-repo-repo-edge-betweenness}
-
 {r plot-graph-repo-repo-edge-betweenness-dendrogram}
 
-* Identify the repos/organisations that are the focus of the most intense collaboration?
-  * Are they thematically or topically linked?
-  * Do they map onto real-world research centres?
-* Is the network a new thing, or just a translation of existing offline networks (e.g. fields, institutions) onto GitHub?
-  * cf. is open source really open?
+This graph identifies three core clusters characterized by repositories whose contributors commit to projects other than their own.
+The clusters that appear along the peiphery largely correspond with the work of single individuals, and sometimes also their colleagues.
+Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (and perhaps initial) integrations of lone developers into a broader community.^[**NOTE:** I wonder what these repositories that prompt greater integration are?
+They should correspond with the nodes of highest betweenness centrality among those in the core cluster.]^[**NOTE:** I also wonder how recently these connections were forged, or how they are situated within users' history of commits.]
 
-### user-user
+The three core clusters also have distinct character.
+One is characterized by its focus on archaeogenetics, which consists of a very well established collaborative network and general reliance on data modelling and processing tools.
+A second cluster is mostly centred on fieldwork-oriented data collection tools, and particularly the emergence of dominant platforms that attract more attention than other independent projects scattered across the network.
+The third and most significant cluster includes a schmorgasborg of projects whose contributors share varied interests.
+The emphasis in this latter cluster is on the formation of a central software development community rather than on any specific topic of work.^[**NOTE:** How does this map onto real-world research centres or partnerships?
+I'm wary of getting too into the weeds on specific people or relationships, but we might highlight the fact that projects led by people who are more active in the SSLA are well represented in this cluster, as well as members of ISAA-Kiel.
+Note that this observation is not backed by statistics, it is just a casual remark based on me browsing the network visualization.]
+
+**NOTE:** What happens to this interpretation when we differentiate commits from issues and comments?
+
+## user-user
 
@@
 We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
+We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as their job.
 
-We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as a job.
+**NOTE:** Create a graph visualization showing either edge-betweenness clustering or rendering nodes' relative betweenness by size.
+**NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits.
 
 We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
-For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the all-contributions graph also appear here.
+For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the graph representing all contributions also appear here.
 These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
-
-{r graph-user-user}
 
@@
 
-{r plot-graph-user-user}
+**NOTE:** I have no idea what these next two charts depict.
 
 {r plot-user-betweenness-dist}
+{r plot-user-betweenness-activity}
 
-{r plot-user-betweenness-activity}
+---
 
@@
   * Career stage? Precarity? Employer? Location?
+
+---
+
+In cleaning the data we made a few significant decisions that revealed insights into our dataset.
+We constructed the user-user graph and filtered out several bots.
+We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community.
+We filtered isolates.
+We calculated the degree and betweenness to assess whether this was a realistic representation of what we know.
+We wanted to figure out who are the key nodes, what they had in common.
+
+The people who are high betweenness are people who primarily produce computational archaeology code as a job.
+We can investgate further by comparing with some other measures of activity, like volume of commits or something else.
+We also looked at the same thing, but with only commits, or without commits.
+For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.

### 2023-03-31 analysis/99-conclusion.qmd
Updating the text of analysis and interpretation.
Re-wording for more active voice, adding some interpretation (specifically to section 07), and updates the notes for further work to be done.

@@
 Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.
-
-In terms of collaboration, our analysis shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host.
+This is marked by a variety of applications and use cases.
+While there is an emerging community of practice around open source in archaeology, collaboration remains limited.
+Most work is performed solo and is short-lived.^[**NOTE:** I wonder if we can say something about change over time, for instance, if things are getting better.]
 
 The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base.
-
+Our analysis also shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host.
 Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects.
-
 This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
-
-We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.
-
----
-
-There is an emerging community of practice around open source in archaeology.
-
-But:
-
-* It is limited and fragmented
-  * A small set of subfields are represented
-  * Probably most people still haven't heard it?
-* Collaboration is limited
-  * Most work remains solo & short-lived
-  * Perhaps because open source norms conflict with academic norms?
-* The members of this community face structural challenges in academia

### 2023-06-12 analysis/paper.qmd
Added this top matter so that I could run the code

@@
+---
+title: "openarchaeoCollaboration"
+author:
+  - name: Zack Batist
+    url: https://zackbatist.info
+    affiliation: Faculty of Information, University of Toronto
+    orcid: 0000-0003-0435-508X 
+  - name: Joe Roe
+    url: https://joeroe.io
+    affiliation: Institute of Archaeological Sciences, University of Bern
+    orcid: 0000-0002-1011-1244
+date: "`r Sys.Date()`"
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(
+	echo = FALSE,
+	message = FALSE,
+	warning = FALSE
+)
+```
+
+```{r dependencies, include=FALSE}
+library("fs")
+library("geomtextpath")
+library("ggiraph")
+library("ggraph")
+library("gt")
+library("here")
+library("khroma")
+library("lubridate")
+library("openarchaeoCollaboration")
+library("patchwork")
+library("tidygraph")
+library("tidyverse")
+library("here")
+```
+
+```{r plot-theme}
+theme_oarch <- function() theme(
+)
+```
+
 ```{r data-oarch}
@@
 | Category                              | Kind      | Scope                                                                                                                                                                                                        |
-|:-----------|:-----------|:------------------------------------------------|
+|:------------------|:------------------|:-----------------------------------|
 | Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
@@
 
----
+------------------------------------------------------------------------
 
 #### Notes from January 30, end of day
+
 In cleaning the data we made a few significant decisions that revealed insights into our dataset.

### 2023-06-14 analysis/_03-data_methodology.qmd
Add main document openarchaeo_collaboration.qmd

Our findings are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources. As of right now it includes 408 items, which are rendered for the web at [open-archaeo.info](open-archaeo.info). We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, and by tracing connections to other personal, professional, and institutional websites that describe and host additional archaeological software. This entailed manually crawling through users' profiles, particularly those who identify as archaeologists or who have contributed to tools that pertain to archaeological work. We supplemented this quasi-systematic collection strategy with word-of-mouth contributions made by interested individuals who identified relevant work that we initially overlooked.

Open-archaeo is a relatively comprehensive list, however it generally overlooks records pertaining to code written before archaeologists started using collaborative software development platforms such as GitHub. It also lacks references to code that is not available on the web. Open-archaeo is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to contribute.

For the purpose of this study, we augmented the core open-archaeo dataset with data obtained from GitHub's public API. That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh]. This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but amongst those that do, GitHub is by far the most popular host.

{r plot-repo-hosts}

**NOTE:** Indicate the numerical value at the end of each bar in the plot-repo-hosts figure.

{r table-no-repo}

While our initial intention was to only list open source software, open-archaeo's scope has expanded to include all software created by and for archaeologists. We found that many repositories do not explicitly denote an open source license, or any license at all.

**NOTE:** Add a breakdown on licensing practices.

The data and analysis code are available in our research compendium on GitHub, which is also indexed on Zenodo.

### 2023-06-14 analysis/_04-what_is_open_archaeology.qmd
Add main document openarchaeo_collaboration.qmd

--- 2023-03-28 analysis/04-what_is_open_archaeology.qmd
@@
 
-Each record is annotated with tags that describe the aspects of archaeological work that each tool contributes to.
-Items are tagged based on how they have been generally identified or presented by their maintainers.
+We annotated each record with tags that describe aspects of archaeological work that each tool contributes to.
+We assigned tags based on how their maintainers identified their projects.
 The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.
@@
 
-Records are also categorized based on how each tool or resource is accessed or used.
+We also categorized records based on how each tool or resource is meant to be accessed or used.
 This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
@@
 Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.
-Standalone software are tools that may be operated without needing to first access an underlying platform.
+Standalone software are tools that users may operate without needing to first access an underlying platform.
 The third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-Things like code generalized sets of code snippets.
 
-As for documents, specifications, protocols and schemas are formal data structures or frameworks intended to be used as models.
+The first kind of document, which we refer to as specifications, protocols and schemas, are formal data structures or frameworks intended to be used as models.
 Products are stable outcomes of creative work, loosely defined.
@@
 
-Our first set of analyses were focused on these questions:
+Our first set of analyses focused on these questions:
 
-What kinds of software are archaeologists making?
+* What kinds of software are archaeologists making?
+* What are the general purposes of their software?
 
-And what are the general purposes of their software?
-
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing 'platform' -- for example a package for a programming language like R, or a plugin for an application like QGIS.
+Most projects included in open-archaeo are designed to be used as a package or library atop am existing "platform" -- for example a package for a programming language like R, or a plugin for an application like QGIS.
 The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
@@
 Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
-AutoCAD, ArcGIS); a number of lesser used programming languages; and a genre consisting of custom forms and spreadsheet templates.
+AutoCAD, ArcGIS), a number of lesser used programming languages, and a genre consisting of custom forms and spreadsheet templates.
 In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
@@
 
-If we compare categories with tags, we can see the general domains that each kind of resource is designed to serve.
+When we compare categories with tags, we see the general domains that each kind of resource is designed to serve.
 We see that packages are fairly common across the board.
@@
 {r plot-tags-cats}
-
----
-
-- Add in a breakdown on licensing practices here too

### 2023-06-14 analysis/_05-history_of_open_archaeology.qmd
Add main document openarchaeo_collaboration.qmd

Archaeological software development activity has increased significantly over the years. The following charts portray the cumulative growth of code contributions committed and pushed to GitHub repositories, and the number of GitHub repositories that host archaeological software and resources.

{r plot-gh-cum}

As we can see, archaeologists have been using git from even before GitHub was launched in 2008. But use of git really began to take off around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.

Around this time we also see that GitHub starts being used to host documents and scripts. This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work in an open and somewhat nerdy way.

---

- Compare changing proportions of each category on a year by year basis
- Identify temporal trends in the use of licenses

### 2023-06-14 analysis/_07-network.qmd
Add main document openarchaeo_collaboration.qmd

--- 2023-03-31 analysis/07-network.qmd
@@
 The clusters that appear along the peiphery largely correspond with the work of single individuals, and sometimes also their colleagues.
-Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (and perhaps initial) integrations of lone developers into a broader community.^[**NOTE:** I wonder what these repositories that prompt greater integration are?
-They should correspond with the nodes of highest betweenness centrality among those in the core cluster.]^[**NOTE:** I also wonder how recently these connections were forged, or how they are situated within users' history of commits.]
+Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (and perhaps initial) integrations of lone developers into a broader community.
 
-The three core clusters also have distinct character.
-One is characterized by its focus on archaeogenetics, which consists of a very well established collaborative network and general reliance on data modelling and processing tools.
-A second cluster is mostly centred on fieldwork-oriented data collection tools, and particularly the emergence of dominant platforms that attract more attention than other independent projects scattered across the network.
+**NOTE:** Create a table containing the nodes with highest betweenness centrality from the repo-repo graph.
+Columns should specify the name of the repository, the number of commits, the category, and tags.
+If possible also include a summary of the pulse as a line chart with x axis as time and y axis as activity.
+This will allow us to identify the characteristics of these repositories that prompt greater integration.
+
+The three core clusters have their own distinct character.
+One has a focus on archaeogenetics, which consists of a very well established collaborative network and general reliance on data modelling and processing tools.
+A second cluster is mostly centred on fieldwork-oriented data collection tools, and particularly the emergence of well-funded and well-supported dominant platforms that attract more attention than other independent projects scattered across the network.
 The third and most significant cluster includes a schmorgasborg of projects whose contributors share varied interests.
-The emphasis in this latter cluster is on the formation of a central software development community rather than on any specific topic of work.^[**NOTE:** How does this map onto real-world research centres or partnerships?
-I'm wary of getting too into the weeds on specific people or relationships, but we might highlight the fact that projects led by people who are more active in the SSLA are well represented in this cluster, as well as members of ISAA-Kiel.
-Note that this observation is not backed by statistics, it is just a casual remark based on me browsing the network visualization.]
+The emphasis in this latter cluster is on the formation of a central software development community rather than on any specific topic of work.
+Many of the projects represented in this third cluster emerge from underlying professional partnerships, namely research labs (e.g.
+ISAA-Kiel) and special interest groups (SSLA).
 
@@
 We assembled graphs linking users based on common contributions to the same repositories.
+
+**NOTE:** Plot a graph that renders nodes' sizes based on their betweenness centrality.
+
+{r summary-graph-user-user}
+
 We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
 We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as their job.
+Moreover, we found that these people tend to be employed under precarious circumstances.
+Although precarious employment is part of our sad reality, in the context of developing and maintaining open source software, this presents a serious source of risk.
+The people who occupy central positions in these networks are crucial community members that make the network whole, and if they are either unable to continue on in their contributions or decide to leave archaeology entirely, then the overall network would fragment.
 
-**NOTE:** Create a graph visualization showing either edge-betweenness clustering or rendering nodes' relative betweenness by size.
-**NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits.
+**NOTE:** Create a table containing the top 10 users based on betweenness centrality values, with columns specifying theur username, real name, job title/affiliation, location, career stage (student, early/mid/late career, tenured/non-tenured, independent, etc).
 
-We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
-For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the graph representing all contributions also appear here.
+We also applied the same betweenness centrality algorithm on a subnetwork whose links are based only on issues and comments, and not code contributions.
+In this subnetwork, people who commit less code have higher betweenness scores.
+However, many of the people with high betweenness from the graph representing all contributions also appear here.
 These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
+This list also includes a series of contributors who never or rarely commit code.
+Although it is out of this study's scope a qualitative analysis of issues and comments may yield more insight on the kinds of contributions that each of these participants make.
 
-{r summary-graph-user-user}
+**NOTE:** Create a table containing the top 10 users based on betweenness centrality values derived from the issues and comments only subnetwork, with columns specifying theur username, real name, job title/affiliation.
+
+---
 
@@
 {r plot-user-betweenness-activity}
-
----
-
-* What proportion of users and repos are connected to this network?
-* What predicts membership in this network?
-* Does centrality in the network correlate with general rate of activity?
-*  Who are the most central nodes in this network? Is there anything they have in common?
-   * High activity?
-   * Certain platforms/categories/languages?
-* Qualitative analysis of top 50? 100? most connected users
-  * Career stage? Precarity? Employer? Location?
-
----
-
-In cleaning the data we made a few significant decisions that revealed insights into our dataset.
-We constructed the user-user graph and filtered out several bots.
-We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community.
-We filtered isolates.
-We calculated the degree and betweenness to assess whether this was a realistic representation of what we know.
-We wanted to figure out who are the key nodes, what they had in common.
-
-The people who are high betweenness are people who primarily produce computational archaeology code as a job.
-We can investgate further by comparing with some other measures of activity, like volume of commits or something else.
-We also looked at the same thing, but with only commits, or without commits.
-For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.

### 2023-06-14 analysis/openarchaeo_collaboration.qmd
Add main document openarchaeo_collaboration.qmd

--- 2023-06-12 analysis/paper.qmd
@@
 library("tidyverse")
-library("here")
 ```
@@
 
-```{r data-oarch}
-read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
-  mutate(category = na_if(category, "?")) ->
-  oarch
-
-# Extract clean repos from GitHub URLs
-oarch %>%
-  mutate(
-    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
-
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
-```
-
-```{r data-github}
-# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
-# chunk will prefer to load it rather than make requests to the GitHub API.
-# Delete this file to fetch fresh data from the GitHub or modify 
-# `cached_gh_data` to read a different cached version.
-#
-# Note that fetching data from the GitHub API takes some time and is subject to
-# a rate limit:
-# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
-#
-# You will need to authenticate authenticate with GitHub credentials to get a
-# rate limit sufficient to obtain the required data, see 
-# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
-# for options on how to do this.
-cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
-if (!file_exists(cached_gh_data)) {
-  oarch <- mutate(oarch,
-                  gh_langs = map(oarch$gh_repo, gh_lang),
-                  gh_contribs = map(oarch$gh_repo, gh_contrib),
-                  gh_issues = map(oarch$gh_repo, gh_issue),
-                  gh_comments = map(oarch$gh_repo, gh_comment),
-                  gh_commits = map(oarch$gh_repo, gh_commit)
-  )
-  saveRDS(oarch, cached_gh_data)
-}
-oarch <- readRDS(cached_gh_data)
-```
-
-## Introduction
-
-In applying either physical or digital methods, archaeology has a long tradition of adopting common or 'off the shelf' tools to serve our own unique purposes, which are not typically unanticipated by their original designers.
-In other words, the apparatus of archaeology is an open system that relies on products and skillsets from outside the discipline, but which exhibit enough flexibility to be used in novel ways.
-For example, archaeologists are content to outsource the metallurgical skills needed to forge their trowels, and delegate the manufacture and acquisition of key electronic tools, like laptop computers and cameras, to the global supply chains and labour networks that produce consumer-grade products for purchase.
-However, an increasingly large number of archaeologists are taking it upon themselves to program their own software, rather than rely on generic and commercially-produced information processing systems.
-In fact, archaeologists are prolific computer programmers and have been coding a variety of functions and libraries that serve uniquely archaeological problems and purposes.
-
-This article presents a survey of archaeological software development with two goals in mind:
-
-(1) we identify *what* kinds of software archaeologists are making;
-(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
-
-Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
-
-## Background
-
--   The notion of the archaeological toolbox, and its inclusion of digital tools and skillsets
-    -   The use of off-the-shelf tools like MS Access, Excel, Agisoft PhotoScan, Dropbox, etc
-    -   Link to related factors:
-        -   Lack of experienced programmers among archaeologists
-        -   The built-in flexibility of these tools
-        -   Ease of administrative purchasing through university departments
-    -   What changed?
-        -   Software is expensive, shift to subscription models, locking into a dependency on a commercial service provider
-        -   Archaeologists are better at computer programming, digital literacy on the rise, digital archaeologists grows into its own legitimate sub-discipline
-        -   Open source is cool, everyone has a github profile (even if not really)
-            -   (JR:) People are presumably guided towards GitHub (or at least git) via other routes, because it's seen as the standard platform for 'open' digital work.
-For example, the literature on reproducible compendiums recommends GitHub, and almost all R development work is now done there.
-And once there, you're sort of forced into engaging in a (particular set of) open source development.
-I know there's controversy about how dominant this has become and people who say it's not 'true' (free and) open source, and perhaps we could add to that the question of whether it's the right model for archaeology.
--   About open source practices and their influence on archaeology
-    -   What is open source software and open source software development?
-        -   Free cost as an end-user
-        -   Free to manipulate and extend
-        -   Community-based support and development
-    -   Software as product vs software as collective process
-        -   Basically what distinguishes abandonware from thriving software ecosystems
-        -   Technical, legal and social notions of open
-            -   Technical: source code is open, use of a platform and systems that support software integration and implementations
-            -   Legal: open licenses, backing by the courts
-            -   Social: inviting community, meaningful contributions provided by users, common vision and values directing the project into future directions
-        -   Consider how these aspects of open have been implemented or prioritized by digital archaeologists in different ways
-            -   Technical: reproducible research, archiving and publishing
-            -   Legal: archiving and publishing, budgeting for projects
-            -   Social: ensuring that software is actually useful for general use and not just one-off implementation, coordinating to reduce redundant effort and not reinvent the wheel, writing documentation, making it fit within existing systems or infrastructures (i.e.
-allocation of credit, citing software directly versus articles that present software, sustainable development after funding runs out, etc)
-        -   This paper largely focuses on the social aspect of open, which is arguably the most important for ensuring a positive future for archaeological software development
-
-The 'archaeological toolbox' is a common metaphor that describes the set of physical and informational tools and processes that archaeologists rely upon in their work.
-
-collaborate in their programming work, and how they experience open source development processes which have their own benefits, challenges and organizing principles that may conflict with or relate to means of collaboration and sharing that archaeologists are already generally accustomed to.
-
-## Data and evidence
-
-Our findings here are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
-As of right now it includes 408 items, which are nicely rendered at open-archaeo.info.
-
-The main criteria for inclusion is whether a project is identified by its maintainers as relating to archaeological work.
-
-The dataset was compiled by browsing collaborative software development platforms, essentially manually crawling through the profiles of users who identify as archaeologists or who have contributed to tools that pertain to archaeological work.
-
-This quasi-systematic collection strategy is also supplemented by contributions made by interested individuals who identified relevant work that was initially overlooked.
-
-Open-archaeo is a relatively comprehensive list, however it generally lacks code written before the growth and general use of collaborative software development platforms, such as GitHub, among archaeologists, and also lacks references to code that is not available on the web.
-
-It is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in your field, to contribute.
-
-Please get in touch if you're interested in adding to the list or doing a deep dive relating to your area of expertise!
-
-### Tags and categories
-
-Each record is also annotated with tags that describe the aspects of archaeological work that each tool contributes to.
-
-Items are tagged based on how they have been generally identified or presented by their maintainers.
-
-The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.
-
-Right now there are 63 tags.
-
-Records are also categorized based on how each tool or resource is accessed or used.
-
-This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-
-In other words, categories hint at some of the decisions that developers make regarding how they want their tools to be used and accessed.
-
-| Category                              | Kind      | Scope                                                                                                                                                                                                        |
-|:------------------|:------------------|:-----------------------------------|
-| Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-|
-| Standalone software                   | Software  | Software that may be operated without needing to first access an underlying platform.
-|
-| Scripts                               | Software  | Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-|
-| Specifications, protocols and schemas | Documents | A formal data structure or framework intended to be used as a model.
-|
-| Products                              | Documents | Stable outcomes of creative work.
-|
-| Guides                                | Documents | An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.
-|
-| Lists and datasets                    | Documents | A series of consistently organized observations assembled with purpose.
-|
-
-Two kinds of categories have been defined: software and documents.
-
-Software comprise collections of instructions that dictate how a computer should operate, and documents are bounded sets of recorded information.
-
-Three categories of software and four categories of documents are described in the table.
-
-Packages and libraries are sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-
-Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.
-
-Standalone software are tools that may be operated without needing to first access an underlying platform.
-
-And the third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-Things like code generalized sets of code snippets.
-
-As for documents, specifications, protocols and schemas are formal data structures or frameworks intended to be used as models.
-
-Products are stable outcomes of creative work, loosely defined.
-
-Guides are educational resources or documented protocols meant to instruct readers how to apply relevant tools or techniques.
-
-And finally, lists or datasets, including open-archaeo itself, are series of consistently organized observations assembled with purpose.
-
-## Findings
-
-### What archaeologists are making
-
-Our first set of analyses were focused on these questions:
-
-What kinds of software are archaeologists making?
-
-And what are the general purposes of their software?
-
-------------------------------------------------------------------------
-
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing 'platform' -- for example a package for a programming language like R, or a plugin for an application like QGIS.
-
-The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
-
-Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-
-A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
-
-```{r plot-categories}
-oarch <- mutate(oarch,
-                category = recode(category,
-                                  "Specifications, protocols and schemas" = "Specifications etc.",
-                                  .default = category))
-oarch %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(n = n()) %>%
-  ggplot(aes(fct_reorder(category, n), n)) +
-  geom_col() +
-  coord_flip() +
-  labs(
-    title = "Most popular categories",
-    subtitle = "By number of projects",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-R is overwhelmingly dominant as a platform, closely followed by Python.
-
-Plugins for QGIS and mobile apps are also relatively popular.
-
-Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
-AutoCAD, ArcGIS); a number of lesser used programming languages; and a genre consisting of custom forms and spreadsheet templates.
-
-In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
-The larger platforms tend to be more diverse.
-
-```{r plot-platforms}
-oarch %>%
-  drop_na(platform) %>%
-  group_by(platform, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(
-    title = "Most popular platforms",
-    subtitle = "By number of projects, split by primary author",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-If we compare categories with tags, we can see the general domains that each kind of resource is designed to serve.
-
-We see that packages are fairly common across the board.
-
-Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
-
-These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.
-
-```{r plot-tags-cats}
-detail_tags <- c("Instrumental Neutron activation analysis",
-                 "Harris Matrix",
-                 "aDNA Simulators",
-                 "Damage profilers",
-                 "Metagenomic classifers",
-                 "aDNA simulators",
-                 "aDNA damage-aware genotypers",
-                 "Low-coverage contamination estimation",
-                 "Low-coverage adapted population-genetics tools",
-                 "aDNA preprocessing",
-                 "aDNA aligner",
-                 "aDNA assembler",
-                 "aDNA toolkit")
-
-oarch %>%
-  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  filter(!tag %in% detail_tags) %>%
-  mutate(category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(tag, category) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  # facet_wrap(vars(category)) +
-  geom_col() +
-  scale_fill_bright() +
-  coord_flip() +
-  labs(
-    title = "Tags by category",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = "Number of projects",
-    fill = NULL
-  ) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_text(size = 7)
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/tags-cats.png", width = 254, height = 127, units = "mm")
-```
-
-**ZB NOTE:** Do a temporal analysis to get a sense of when platforms, tags and categories have been gaining popularity.
-There may not be a need for extreme detail here, but simply noting the dates of first commits might be helpful.
-I'm imagining a stacked bar chart, with each bar representing a month or a year, and with the stacked portions of each bar representing a platform or category.
-
-### Who's doing what?
-
-```{r user-contributors}
-oarch |>
-  unnest(gh_contribs) |>
-  drop_na(contributor) |>
-  group_by(contributor) |>
-  mutate(
-    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
-  ) |>
-  group_by(contributor, repo_owner) |>
-  summarise(repos = n(), contribs = sum(contributions)) |>
-  pivot_wider(c(contributor), names_from = repo_owner, 
-              values_from = c(repos, contribs), values_fill = 0) |>
-  mutate(
-    repos = repos_other + repos_own,
-    contribs = contribs_other + contribs_own,
-    .after = contributor
-  ) ->
-  oarch_contributors
-```
-
-### Development processes
-
-While somewhat informative, breaking down the distribution of categories and tags does not really tell us that much about *how* archaeologists develop software.
-
-The remainder of the paper is driven by a concern about how archaeologists collaborate in their programming work.
-
-How is their work sustained, and how does it leverage distributed work processes that are a key aspect of open source software development in general?
-
-To address these questions, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
-
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@gh].
-
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but as yu can see in the figure below, amongst those that do, GitHub is by far the most popular host.
-
-```{r plot-repo-hosts}
-# TODO: turn into table or just text
-oarch <- mutate(oarch, none = ifelse(is.na(github) &
-                                       is.na(gitlab) &
-                                       is.na(bitbucket) &
-                                       is.na(launchpad) &
-                                       is.na(codeberg),
-                                     "", NA))
-
-oarch %>%
-  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
-               names_to = "repo_host", values_to = "repo_url") %>%
-  drop_na(repo_url) %>%
-  group_by(repo_host, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ungroup() %>%
-  mutate(repo_host = fct_reorder(repo_host, total),
-         repo_host = fct_relevel(repo_host, "none")) %>%
-  ggplot(aes(repo_host, n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(title = "Most popular repository hosts",
-       subtitle = "By number of projects, split by primary author",
-       caption = "Data: open-archaeo.info",
-       x = NULL, y = NULL)
-
-# Export for slides (half slide)
-# ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
-```
-
-```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
-```
-
-We obtained this data in R using the {gh} package; those data, and the analysis code, are available in our research compendium on GitHub.
-
-------------------------------------------------------------------------
-
-Starting with a relatively simple plot, this is a chart of the cumulative growth of open archaeology over the years, measured by code contributions committed and pushed to GitHub repositories.
-
-As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
-But they really began to get into it around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
-
-Around this time we also see that GitHub starts being used to host documents and scripts.
-This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work, especially work or manifestations of undervalued skills that may not necessarily be easily represented using traditional publication venues, in an open and somewhat nerdy way.
-
-```{r plot-gh-cum}
-oarch %>% 
-  mutate(
-    lumped_category = recode(category,
-                             "Specifications etc." = "Documents",
-                             "Products" = "Documents",
-                             "Guides" = "Documents",
-                             "Lists and datasets" = "Documents",
-                             .default = category),
-    lumped_category = factor(lumped_category, c("Scripts",
-                                                "Documents",
-                                                "Packages and libraries",
-                                                "Standalone software")) ) ->
-  oarch_lumped
-
-p_gh_commits_cum <- oarch_lumped %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = lumped_category,
-             colour = lumped_category,
-             label = lumped_category)) +
-  geom_area() +
-  annotate("segment", 
-           x = ymd("2008-04-01", tz = "UTC"), 
-           xend = ymd("2008-04-01", tz = "UTC"),
-           y = 16000, yend = 4000,
-           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
-  annotate("text", 
-           x = ymd("2008-04-01", tz = "UTC"), y = 19000,
-           label = "GitHub\nlaunched",
-           vjust = 0) +
-  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = guide_none()) +
-  scale_fill_bright() +
-  labs(
-    y = "cumulative commits",
-    x = NULL, fill = NULL
-  ) +
-  theme_oarch() +
-  theme(legend.position = "top")
-
-# TODO: extend to end
-p_gh_repos_cum <- oarch_lumped %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  group_by(repo) %>% 
-  filter(datetime == min(datetime)) %>% # Only first commits
-  mutate(date = floor_date(datetime, "week")) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cat_total = sum(n_commits)) %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = lumped_category,
-             colour = lumped_category)) +
-  geom_area() +
-  annotate("segment",
-           x = ymd("2008-04-01", tz = "UTC"),
-           xend = ymd("2008-04-01", tz = "UTC"),
-           y = 40, yend = 10,
-           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
-  annotate("text",
-           x = ymd("2008-04-01", tz = "UTC"), y = 50,
-           label = "GitHub\nlaunched",
-           vjust = 0) +
-  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = guide_none()) +
-  scale_fill_bright(guide = guide_none()) +
-  labs(
-    y = "cumulative repositories",
-    x = NULL, fill = NULL
-  ) +
-  theme_oarch()
-
-p_gh_commits_cum / p_gh_repos_cum
-
-# Export for slides (full slide)
-# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-But how is this activity distributed across repositories?
-
-This visualizes the 'pulse' of active projects.
-
-Each vertical line represents the lifespan of a repository; the points on the line are individual commits.
-
-This is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
-
-```{r plot-lifespan}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(datetime) %>%
-  group_by(item_name) %>%
-  add_tally() %>%
-  filter(n > 1) %>%
-  mutate(first_commit = min(datetime),
-         mid_commit = median(datetime),
-         last_commit = max(datetime)) %>%
-  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
-             y = datetime,
-             ymin = first_commit,
-             ymax = last_commit)) +
-  geom_linerange(size = 0.05) +
-  geom_point(size = 0.1) +
-  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years", date_labels = "%Y") +
-  # coord_flip() +
-  labs(
-    title = "Project lifespan",
-    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
-    x = NULL, y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch() +
-  theme(
-    axis.line.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    panel.grid.major.x = element_blank()
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-The span of time between the first and latest commits is visualized more clearly in summary here.
-
-Notably, many projects are active for only a few short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
-
-Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.
-
-Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects.
-
-The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e.
-the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.
-
-But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
-
-This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.
-
-**ZB NOTE:** Use patchwork package to combine these plots into a unified figure.
-
-```{r lifespan-stats}
-oarch %>% 
-  unnest(gh_commits) %>% 
-  drop_na(datetime) %>% 
-  group_by(item_name) %>%
-  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
-  summarise(
-    category = first(category),
-    n_contribs = first(n_contribs),
-    n = n(),
-    first_commit = min(datetime),
-    last_commit = max(datetime),
-    lifespan = as.integer(last_commit - first_commit),
-    commit_rate = n / lifespan
-  ) ->
-  oarch_lifespan
-```
-
-```{r plot-lifespan-hist}
-# TODO: where does oarch_lifespan come from?
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan)) +
-  geom_histogram(binwidth = 100) +
-  scale_x_continuous(breaks = scales::breaks_width(1000)) +
-  labs(
-    title = "Project lifespan",
-    subtitle = "Time between first and latest commit, by repository",
-    x = "Days", y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/lifespan-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-lifespan-vs-rate}
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = "Project lifespan vs. average commit rate",
-    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
-    caption = NULL
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-total.png", width = 127, height = 60, units = "mm")
-
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = NULL,
-    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-contrib.png", width = 127, height = 60, units = "mm")
-
-```
-
-------------------------------------------------------------------------
-
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
-
-A "contribution", in this case, is a commit, issue, comment applied by a user to a repository.
-
-```{r plot-gh_contributors-histogram}
-oarch %>%
-  drop_na(gh_repo) %>%
-  unnest(gh_contribs) %>%
-  group_by(gh_repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Repository contributors",
-    x = "Number of contributors", y = "Repositories",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-hist.png", width = 127, height = 127, units = "mm")
-```
-
-------------------------------------------------------------------------
-
-Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
-
-Of multi-contributor repositories, 92% have a third of contributions made by a single person;
-
-65% have half of all contributions made by a single person;
-
-and 36% have three quarters of all contributions made a single person.
-
-**ZB NOTE:** Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic.
-
-```{r plot-contrib-dist}
-oarch %>%
-  unnest(gh_contribs) %>%
-  drop_na(contributor) %>%
-  group_by(item_name) %>%
-  mutate(n_contributors = n()) %>%
-  filter(n_contributors > 1) %>%
-  mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Leading contributor", "Other contributors")) %>% 
-  group_by(item_name, big_boy) %>%
-  summarise(contributions = sum(contributions)) %>%
-  mutate(
-    p_contributions = contributions / sum(contributions),
-    p_big_boy = first(p_contributions)
-  ) %>%
-  ungroup() %>%
-  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
-  ggplot(aes(x = item_name, y = contributions,
-             fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Leading contributor")))) +
-  geom_col(position = "fill", colour = "white", size = 0.1) +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
-  labs(title = "Repository contributors by proportion of contributions",
-       x = NULL, y = "Contributions",
-       fill = NULL,
-       caption = "Data: open-archaeo.info and GitHub API") +
-  coord_flip() +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-dist.png", width = 127, height = 127, units = "mm")
-```
-
-Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-
-Only 39% of repositories make use of these features. Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-
-**ZB NOTE:** Recalcute these numbers.
-
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
-
-```{r plot-issues-hist}
-oarch %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of issues",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of issues", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/issues-hist.png", width = 127, height = 127, units = "mm")
-```
-
-```{r plot-comments-hist}
-oarch %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of comments",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of comments", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/comments-hist.png", width = 127, height = 127, units = "mm")
-```
-
-### Formation of collaborative development communities
-
-We constructed and analyzed networks representing users' contributions to GitHub repositories in order to examine the emergence of collaborative tendencies.
-
-We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
-
-```{r graph-data}
-oarch |>
-  drop_na(gh_repo) |>
-  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
-  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
-  oarch_graph
-
-oarch_graph <- bind_rows(
-  oarch_graph |>
-    unnest(gh_commits) |>
-    drop_na(author) |> # Apparently real commits, but we can't connect them
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "commit", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_comments) |>
-    drop_na(author) |> # Repos with no comments
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "comment", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_issues) |>
-    drop_na(author) |> # Repos with no issues
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "issue", n_contrib = n(), .groups = "drop")
-)
-
-# Cleaning
-oarch_graph |>
-  rename(user = author) |>
-  # Remove non-human users
-  filter(!str_detect(user, coll("[bot]"))) |>
-  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
-                      "gitter-badger", "dependabot-support")) |>
-  # Remove big, non-archaeology projects that flood the network
-  filter(!repo %in% c("Stellarium/stellarium", 
-                     "opengisch/QField",
-                     "lutraconsulting/input")) ->
-  oarch_graph
-```
-
-We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations.
-We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
-
-#### Repo--user
-
-```{r plot-graph-repo-user}
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
-
-g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt") +
-  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
-  geom_node_point(aes(colour = node_type)) +
-  scale_edge_alpha(trans = "log")
-
-#g_repo_user
-girafe(g_repo_user)
-```
-
-#### Repo--repo
-
-We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate notes, which represent repositories with only a single contributor, we detected 21 clusters.
-While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
-
-Are the clusters associated with particular topics or tags?
-Are there some sorts of associations between clusters that break down from a parent in the community detection dendrogram?
-How does all of this differ when we differentiate commits from issues and comments?
-
-**ZB NOTE:** Remove the cells titled `plot-graph-repo-repo` and `plot-graph-repo-repo-louvain`.
-Use patchwork to compare sets of graphs and dendrograms with links based on all contributions, only commits, and only issues and comments.
-
-```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-```{r plot-graph-repo-repo-edge-betweenness}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  #geom_node_point(aes(colour = group)) +
-  #geom_node_text(aes(label = group)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-
-girafe(ggobj = iplot)
-```
-
-```{r plot-graph-repo-repo-edge-betweenness-dendrogram}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = TRUE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
-xx <- as.dendrogram(xx)
-plot(xx)
-
-ggraph(xx, 'dendrogram', height = height)
-
-```
-
-```{r plot-graph-repo-repo-louvain}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_louvain())) ->
-  oarch_graph_repo_repo
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = group)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-```
-
-#### User--user
-
-We assembled graphs linking users based on common contributions to the same repositories.
-We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
-
-We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as a job.
-
-**ZB NOTE:** We can further investigate this point made above by comparing with some other measures of activity, like volume of commits or something else.
-
-We applied the same betweenness centrality algorithm on user-user networks whose links are based on only commits and on only issues and comments.
-For the latter, people who commit less code are more represented with higher betweenness scores, but many of the people with high betweenness from the all-contributions graph also appear here.
-These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
-
-**ZB NOTE:** Look at the histogram that maps the distribution of betweenness scores across all users.
-Interesting double hump of low-moderate betweenness scores, which may correspond with distinct or protruding components.
-
-```{r graph-user-user}
-oarch_graph_user_user <- tbl_graph(
-  nodes = nodes_users(oarch_graph), 
-  edges = edges_common_repos(oarch_graph),
-  directed = FALSE
-)
-```
-
-```{r summary-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), 
-         betweenness = centrality_betweenness(weights = n), 
-         strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |> 
-  data.frame()
-```
-
-```{r plot-graph-user-user}
-oarch_graph_user_user |>
-  mutate(betweenness = centrality_betweenness(weights = n)) ->
-  oarch_graph_user_user
-
-oarch_graph_user_user |> 
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  # geom_node_point() +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common repositories")
-```
-
-```{r plot-user-betweenness-dist}
-oarch_graph_user_user |>
-  mutate(
-    betweenness = centrality_betweenness(weights = n)
-  ) |>
-  data.frame() |>
-  ggplot(aes(x = betweenness)) +
-  geom_density(bw = "nrd") +
-  scale_x_log10(labels = scales::number)
-```
-
-```{r plot-user-betweenness-activity}
-oarch_graph_user_user |>
-  mutate(
-    degree = centrality_degree(), 
-    betweenness = centrality_betweenness(weights = n), 
-    strength = centrality_degree(weights = n)
-  ) |>
-  as_tibble() |> 
-  left_join(oarch_contributors, by = c("user" = "contributor")) |>
-  ggplot(aes(betweenness, contribs)) +
-  geom_point() +
-  scale_x_log10() +
-  scale_y_log10() +
-  geom_smooth(method = "lm")
-```
-
-#### Over time
-
-Could animate the graphs??
-
-## Conclusions
-
-Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.
-
-In terms of collaboration, our analysis shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host.
-
-The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base.
-
-Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects.
-
-This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
-
-We intend to look deeper at the social aspects of collaborative software engineering among archaeologists with a network analysis, and perhaps by leveraging more data from documentation included in each repository.
-
-------------------------------------------------------------------------
-
-#### Notes from January 30, end of day
-
-In cleaning the data we made a few significant decisions that revealed insights into our dataset.
-We constructed the user-user graph and filtered out several bots.
-We filtered out several large non-archaeology projects that effected the network in ways that distracted us from what we wanted to say about the archaeological community.
-We filtered isolates.
-We calculated the degree and betweenness to assess whether this was realistic representation of what we know.
-We wanted to figure out who are the key nodes, what they had in common.
-The people who are high betweenness are people who primarily produce computational archaeology code as a job.
-We can investgate further by comparing with some other measures of activity, like volume of commits or something else.
-We also looked at the same thing, but with only commits, or without commits.
-For the subgraph of non-commits, it includes many people from the commits-only or commits-included graphs, but also more people who do not commit as much.
+{{< include _01-introduction.qmd >}}
+{{< include _02-background.qmd >}}
+{{< include _03-data_methodology.qmd >}}
+{{< include _04-what_is_open_archaeology.qmd >}}
+{{< include _05-history_of_open_archaeology.qmd >}}
+{{< include _06-project_activity.qmd >}}
+{{< include _07-network.qmd >}}
+{{< include _99-conclusion.qmd >}}

### 2023-06-14 analysis/openarchaeo_collaboration.qmd
Update usage instructions

@@
 ```{r dependencies, include=FALSE}
+library("openarchaeoCollaboration") # this package, install with devtools::build()
+
 library("fs")
@@
 library("lubridate")
-library("openarchaeoCollaboration")
 library("patchwork")

### 2023-06-14 analysis/_03-data_methodology.qmd
Merge code chunks into section qmds
Leaving the network analysis unevaluated for now.

@@
 
+```{r data-oarch}
+read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
+  mutate(category = na_if(category, "?")) ->
+  oarch
+
+# Extract clean repos from GitHub URLs
+oarch %>%
+  mutate(
+    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
+  ) ->
+  oarch
+
+# Amalgamate categories into fewer rough categories
+# TODO: redo
+oarch <- mutate(oarch, rough_category = recode(category,
+                                 "R" = "Packages (R, Python, etc.)",
+                                 "Web apps" = "Standalone apps",
+                                 "Datasets" = "Analyses & datasets",
+                                 "Tutorials" = "Lists & tutorials",
+                                 "Python" = "Packages (R, Python, etc.)",
+                                 "Desktop apps" = "Standalone apps",
+                                 "QGIS" = "Plugins",
+                                 "Lists" = "Lists & tutorials",
+                                 "Mobile apps" = "Standalone apps",
+                                 "Analyses" = "Analyses & datasets",
+                                 "Spreadsheets" = "Forms & spreadsheets",
+                                 "Command line tools" = "Standalone apps",
+                                 "Plugins" = "Plugins",
+                                 "ArcGis" = "Plugins",
+                                 "Filemaker" = "Plugins",
+                                 "ODK" = "Forms & spreadsheets",
+                                 "Lisp" = "Packages (R, Python, etc.)",
+                                 .default = "Other",
+                                 .missing = "Other"
+                                 ))
+```
+
 For the purpose of this study, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
@@
 
-{r plot-repo-hosts}
+```{r data-github}
+# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
+# chunk will prefer to load it rather than make requests to the GitHub API.
+# Delete this file to fetch fresh data from the GitHub or modify 
+# `cached_gh_data` to read a different cached version.
+#
+# Note that fetching data from the GitHub API takes some time and is subject to
+# a rate limit:
+# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
+#
+# You will need to authenticate authenticate with GitHub credentials to get a
+# rate limit sufficient to obtain the required data, see 
+# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
+# for options on how to do this.
+cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
+if (!file_exists(cached_gh_data)) {
+  oarch <- mutate(oarch,
+                  gh_langs = map(oarch$gh_repo, gh_lang),
+                  gh_contribs = map(oarch$gh_repo, gh_contrib),
+                  gh_issues = map(oarch$gh_repo, gh_issue),
+                  gh_comments = map(oarch$gh_repo, gh_comment),
+                  gh_commits = map(oarch$gh_repo, gh_commit)
+  )
+  saveRDS(oarch, cached_gh_data)
+}
+oarch <- readRDS(cached_gh_data)
+```
 
-**NOTE:** Indicate the numerical value at the end of each bar in the plot-repo-hosts figure.
+```{r plot-repo-hosts}
+# TODO: turn into table or just text
+# TODO: Indicate the numerical value at the end of each bar in the plot-repo-hosts figure.
+oarch <- mutate(oarch, none = ifelse(is.na(github) &
+                                       is.na(gitlab) &
+                                       is.na(bitbucket) &
+                                       is.na(launchpad) &
+                                       is.na(codeberg),
+                                     "", NA))
 
-{r table-no-repo}
+oarch %>%
+  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
+               names_to = "repo_host", values_to = "repo_url") %>%
+  drop_na(repo_url) %>%
+  group_by(repo_host, author1_name) %>%
+  summarise(n = n(), .groups = "drop_last") %>%
+  mutate(total = sum(n)) %>%
+  ungroup() %>%
+  mutate(repo_host = fct_reorder(repo_host, total),
+         repo_host = fct_relevel(repo_host, "none")) %>%
+  ggplot(aes(repo_host, n, group = author1_name)) +
+  geom_col(colour = "#333333") +
+  coord_flip() +
+  labs(title = "Most popular repository hosts",
+       subtitle = "By number of projects, split by primary author",
+       caption = "Data: open-archaeo.info",
+       x = NULL, y = NULL)
+```
+
+```{r table-no-repo}
+# TODO: prettify with gt()
+filter(oarch, none == "") %>%
+  count(category)
+```
 

### 2023-06-14 analysis/_04-what_is_open_archaeology.qmd
Merge code chunks into section qmds
Leaving the network analysis unevaluated for now.

@@
 
-{r plot-categories}
+```{r plot-categories}
+oarch <- mutate(oarch,
+                category = recode(category,
+                                  "Specifications, protocols and schemas" = "Specifications etc.",
+                                  .default = category))
+oarch %>%
+  drop_na(category) %>%
+  group_by(category) %>%
+  summarise(n = n()) %>%
+  ggplot(aes(fct_reorder(category, n), n)) +
+  geom_col() +
+  coord_flip() +
+  labs(
+    title = "Most popular categories",
+    subtitle = "By number of projects",
+    caption = "Data: open-archaeo.info",
+    x = NULL, y = NULL
+  ) +
+  theme_oarch()
+```
 
@@
 
-{r plot-platforms}
+```{r plot-platforms}
+oarch %>%
+  drop_na(platform) %>%
+  group_by(platform, author1_name) %>%
+  summarise(n = n(), .groups = "drop_last") %>%
+  mutate(total = sum(n)) %>%
+  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
+  geom_col(colour = "#333333") +
+  coord_flip() +
+  labs(
+    title = "Most popular platforms",
+    subtitle = "By number of projects, split by primary author",
+    caption = "Data: open-archaeo.info",
+    x = NULL, y = NULL
+  ) +
+  theme_oarch()
+```
 
@@
 
-{r plot-tags-cats}
+```{r plot-tags-cats}
+detail_tags <- c("Instrumental Neutron activation analysis",
+                 "Harris Matrix",
+                 "aDNA Simulators",
+                 "Damage profilers",
+                 "Metagenomic classifers",
+                 "aDNA simulators",
+                 "aDNA damage-aware genotypers",
+                 "Low-coverage contamination estimation",
+                 "Low-coverage adapted population-genetics tools",
+                 "aDNA preprocessing",
+                 "aDNA aligner",
+                 "aDNA assembler",
+                 "aDNA toolkit")
+
+oarch %>%
+  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
+               values_drop_na = TRUE) %>%
+  drop_na(tag) %>%
+  filter(!tag %in% detail_tags) %>%
+  mutate(category = recode(category,
+                                  "Specifications etc." = "Documents",
+                                  "Products" = "Documents",
+                                  "Guides" = "Documents",
+                                  "Lists and datasets" = "Documents",
+                                  .default = category)) %>%
+  group_by(tag, category) %>%
+  summarise(n = n(), .groups = "drop_last") %>%
+  mutate(total = sum(n)) %>%
+  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
+  # facet_wrap(vars(category)) +
+  geom_col() +
+  scale_fill_bright() +
+  coord_flip() +
+  labs(
+    title = "Tags by category",
+    caption = "Data: open-archaeo.info",
+    x = NULL, y = "Number of projects",
+    fill = NULL
+  ) +
+  theme_oarch() +
+  theme(
+    axis.text.y = element_text(size = 7)
+  )
+```

### 2023-06-14 analysis/_05-history_of_open_archaeology.qmd
Merge code chunks into section qmds
Leaving the network analysis unevaluated for now.

@@
 
-{r plot-gh-cum}
+```{r plot-gh-cum}
+oarch %>% 
+  mutate(
+    lumped_category = recode(category,
+                             "Specifications etc." = "Documents",
+                             "Products" = "Documents",
+                             "Guides" = "Documents",
+                             "Lists and datasets" = "Documents",
+                             .default = category),
+    lumped_category = factor(lumped_category, c("Scripts",
+                                                "Documents",
+                                                "Packages and libraries",
+                                                "Standalone software")) ) ->
+  oarch_lumped
+
+p_gh_commits_cum <- oarch_lumped %>%
+  unnest(gh_commits) %>%
+  drop_na(sha) %>%
+  mutate(date = floor_date(datetime, "week")) %>%
+  group_by(lumped_category, date) %>%
+  summarise(n_commits = n(), .groups = "drop_last") %>%
+  mutate(cum_commits = cumsum(n_commits)) %>%
+  ggplot(aes(date, cum_commits,
+             fill = lumped_category,
+             colour = lumped_category,
+             label = lumped_category)) +
+  geom_area() +
+  annotate("segment", 
+           x = ymd("2008-04-01", tz = "UTC"), 
+           xend = ymd("2008-04-01", tz = "UTC"),
+           y = 16000, yend = 4000,
+           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
+  annotate("text", 
+           x = ymd("2008-04-01", tz = "UTC"), y = 19000,
+           label = "GitHub\nlaunched",
+           vjust = 0) +
+  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
+  scale_y_continuous(labels = scales::comma) +
+  scale_colour_bright(guide = guide_none()) +
+  scale_fill_bright() +
+  labs(
+    y = "cumulative commits",
+    x = NULL, fill = NULL
+  ) +
+  theme_oarch() +
+  theme(legend.position = "top")
+
+# TODO: extend to end
+p_gh_repos_cum <- oarch_lumped %>%
+  unnest(gh_commits) %>%
+  drop_na(sha) %>%
+  group_by(repo) %>% 
+  filter(datetime == min(datetime)) %>% # Only first commits
+  mutate(date = floor_date(datetime, "week")) %>%
+  group_by(lumped_category, date) %>%
+  summarise(n_commits = n(), .groups = "drop_last") %>%
+  mutate(cat_total = sum(n_commits)) %>%
+  mutate(cum_commits = cumsum(n_commits)) %>%
+  ggplot(aes(date, cum_commits,
+             fill = lumped_category,
+             colour = lumped_category)) +
+  geom_area() +
+  annotate("segment",
+           x = ymd("2008-04-01", tz = "UTC"),
+           xend = ymd("2008-04-01", tz = "UTC"),
+           y = 40, yend = 10,
+           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
+  annotate("text",
+           x = ymd("2008-04-01", tz = "UTC"), y = 50,
+           label = "GitHub\nlaunched",
+           vjust = 0) +
+  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
+  scale_y_continuous(labels = scales::comma) +
+  scale_colour_bright(guide = guide_none()) +
+  scale_fill_bright(guide = guide_none()) +
+  labs(
+    y = "cumulative repositories",
+    x = NULL, fill = NULL
+  ) +
+  theme_oarch()
+
+p_gh_commits_cum / p_gh_repos_cum
+```
 

### 2023-06-14 analysis/_06-project_activity.qmd
Merge code chunks into section qmds
Leaving the network analysis unevaluated for now.

--- 2023-03-31 analysis/06-project_activity.qmd
@@
 
-{r plot-lifespan}
-{r lifespan-stats}
+```{r plot-lifespan}
+oarch %>%
+  unnest(gh_commits) %>%
+  drop_na(datetime) %>%
+  group_by(item_name) %>%
+  add_tally() %>%
+  filter(n > 1) %>%
+  mutate(first_commit = min(datetime),
+         mid_commit = median(datetime),
+         last_commit = max(datetime)) %>%
+  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
+             y = datetime,
+             ymin = first_commit,
+             ymax = last_commit)) +
+  geom_linerange(size = 0.05) +
+  geom_point(size = 0.1) +
+  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
+                              ymd("2021-05-31", tz = "UTC")),
+                   date_breaks = "2 years", date_labels = "%Y") +
+  # coord_flip() +
+  labs(
+    title = "Project lifespan",
+    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
+    x = NULL, y = NULL,
+    caption = "Data: open-archaeo.info and GitHub API"
+  ) +
+  theme_oarch() +
+  theme(
+    axis.line.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.x = element_blank()
+  )
+```
+
+```{r lifespan-stats}
+oarch %>% 
+  unnest(gh_commits) %>% 
+  drop_na(datetime) %>% 
+  group_by(item_name) %>%
+  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
+  summarise(
+    category = first(category),
+    n_contribs = first(n_contribs),
+    n = n(),
+    first_commit = min(datetime),
+    last_commit = max(datetime),
+    lifespan = as.integer(last_commit - first_commit),
+    commit_rate = n / lifespan
+  ) ->
+  oarch_lifespan
+```
 
@@
 
-**NOTE:** Combine these figures using patchwork.
-{r plot-lifespan-hist}
-{r plot-lifespan-vs-rate}
+```{r plot-lifespan-hist}
+# TODO: combine with patchwork
+# TODO: where does oarch_lifespan come from?
+oarch_lifespan %>%
+  ggplot(aes(x = lifespan)) +
+  geom_histogram(binwidth = 100) +
+  scale_x_continuous(breaks = scales::breaks_width(1000)) +
+  labs(
+    title = "Project lifespan",
+    subtitle = "Time between first and latest commit, by repository",
+    x = "Days", y = NULL,
+    caption = "Data: open-archaeo.info and GitHub API"
+  ) +
+  theme_oarch()
+```
+
+```{r plot-lifespan-vs-rate}
+oarch_lifespan %>%
+  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
+  geom_point(shape = 21, fill = NA) +
+  scale_x_log10() +
+  scale_y_log10() +
+  labs(
+    title = "Project lifespan vs. average commit rate",
+    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
+    caption = NULL
+  ) +
+  theme_oarch()
+
+oarch_lifespan %>%
+  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
+  geom_point(shape = 21, fill = NA) +
+  scale_x_log10() +
+  scale_y_log10() +
+  labs(
+    title = NULL,
+    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
+    caption = "Data: open-archaeo.info and GitHub API"
+  ) +
+  theme_oarch()
+```
 
@@
 
-{r plot-gh_contributors-histogram}
+```{r plot-gh_contributors-histogram}
+oarch %>%
+  drop_na(gh_repo) %>%
+  unnest(gh_contribs) %>%
+  group_by(gh_repo) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  ggplot(aes(x = n)) +
+  geom_histogram() +
+  scale_x_log10() +
+  labs(
+    title = "Repository contributors",
+    x = "Number of contributors", y = "Repositories",
+    caption = "Data: open-archaeo.info and GitHub API"
+  ) +
+  theme_oarch()
+```
 
@@
 
-{r plot-contrib-dist}
+```{r plot-contrib-dist}
+oarch %>%
+  unnest(gh_contribs) %>%
+  drop_na(contributor) %>%
+  group_by(item_name) %>%
+  mutate(n_contributors = n()) %>%
+  filter(n_contributors > 1) %>%
+  mutate(big_boy = ifelse(contributions == max(contributions),
+                          "Leading contributor", "Other contributors")) %>% 
+  group_by(item_name, big_boy) %>%
+  summarise(contributions = sum(contributions)) %>%
+  mutate(
+    p_contributions = contributions / sum(contributions),
+    p_big_boy = first(p_contributions)
+  ) %>%
+  ungroup() %>%
+  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
+  ggplot(aes(x = item_name, y = contributions,
+             fill = fct_relevel(big_boy, c("Other contributors",
+                                           "Leading contributor")))) +
+  geom_col(position = "fill", colour = "white", size = 0.1) +
+  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
+  scale_y_continuous(labels = scales::percent) +
+  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
+  labs(title = "Repository contributors by proportion of contributions",
+       x = NULL, y = "Contributions",
+       fill = NULL,
+       caption = "Data: open-archaeo.info and GitHub API") +
+  coord_flip() +
+  theme_oarch() +
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank()
+  )
+```
 
@@
 
-**NOTE:** Combine these figures using patchwork.
-{r plot-issues-hist}
-{r plot-comments-hist}
+```{r plot-issues-hist}
+# TODO: combine with comments-hist
+oarch %>%
+  unnest(gh_issues) %>%
+  drop_na(title) %>%
+  group_by(repo) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  ggplot(aes(x = n)) +
+  geom_histogram() +
+  scale_x_log10() +
+  labs(
+    title = "Use of issues",
+    caption = "Data: open-archaeo.info and GitHub API",
+    x = "Number of issues", y = "Repositories"
+  ) +
+  theme_oarch()
+```
+
+```{r plot-comments-hist}
+oarch %>%
+  unnest(gh_comments) %>%
+  drop_na(issue_num) %>%
+  group_by(repo) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  ggplot(aes(x = n)) +
+  geom_histogram() +
+  scale_x_log10() +
+  labs(
+    title = "Use of comments",
+    caption = "Data: open-archaeo.info and GitHub API",
+    x = "Number of comments", y = "Repositories"
+  ) +
+  theme_oarch()
+```

### 2023-06-14 analysis/_07-network.qmd
Merge code chunks into section qmds
Leaving the network analysis unevaluated for now.

@@
 
-{r graph-data}
+```{r graph-data, eval=FALSE}
+# TODO: refactor some of this as functions?
+oarch |>
+  drop_na(gh_repo) |>
+  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
+  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
+  oarch_graph
+
+oarch_graph <- bind_rows(
+  oarch_graph |>
+    unnest(gh_commits) |>
+    drop_na(author) |> # Apparently real commits, but we can't connect them
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "commit", n_contrib = n(), .groups = "drop"),
+  oarch_graph |>
+    unnest(gh_comments) |>
+    drop_na(author) |> # Repos with no comments
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "comment", n_contrib = n(), .groups = "drop"),
+  oarch_graph |>
+    unnest(gh_issues) |>
+    drop_na(author) |> # Repos with no issues
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "issue", n_contrib = n(), .groups = "drop")
+)
+
+# Cleaning
+oarch_graph |>
+  rename(user = author) |>
+  # Remove non-human users
+  filter(!str_detect(user, coll("[bot]"))) |>
+  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
+                      "gitter-badger", "dependabot-support")) |>
+  # Remove big, non-archaeology projects that flood the network
+  filter(!repo %in% c("Stellarium/stellarium", 
+                     "opengisch/QField",
+                     "lutraconsulting/input")) ->
+  oarch_graph
+```
 
@@
 
-{r plot-graph-repo-user}
+```{r plot-graph-repo-user, eval=FALSE}
+oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
+                                   edges_user_to_repo(oarch_graph))
+
+g_repo_user <- oarch_graph_repo_user |>
+  ggraph(layout = "graphopt") +
+  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
+  geom_node_point(aes(colour = node_type)) +
+  scale_edge_alpha(trans = "log")
+
+#g_repo_user
+girafe(g_repo_user)
+```
 
@@
 
-{r plot-graph-repo-repo}
-{r graph-user-user}
-{r plot-graph-user-user}
+```{r plot-graph-repo-repo, eval=FALSE}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
+                                   edges_common_users(oarch_graph),
+                                   directed = FALSE)
+
+oarch_graph_repo_repo |>
+  ggraph(layout = "graphopt") +
+  geom_edge_link(aes(alpha = n)) +
+  geom_node_point(aes(colour = category)) +
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common contributors")
+```
+
+```{r graph-user-user, eval=FALSE}
+oarch_graph_user_user <- tbl_graph(
+  nodes = nodes_users(oarch_graph), 
+  edges = edges_common_repos(oarch_graph),
+  directed = FALSE
+)
+```
+```{r plot-graph-user-user, eval=FALSE}
+oarch_graph_user_user |>
+  mutate(betweenness = centrality_betweenness(weights = n)) ->
+  oarch_graph_user_user
+
+oarch_graph_user_user |> 
+  ggraph(layout = "graphopt") +
+  geom_edge_link(aes(alpha = n)) +
+  # geom_node_point() +
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common repositories")
+```
 
@@
 
-{r plot-graph-repo-repo-edge-betweenness}
-{r plot-graph-repo-repo-edge-betweenness-dendrogram}
+```{r plot-graph-repo-repo-edge-betweenness, eval=FALSE}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
+                                   edges_common_users(oarch_graph),
+                                   directed = FALSE)
+
+oarch_graph_repo_repo |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 1) |>
+  mutate(group = as.character(group_edge_betweenness())) ->
+  oarch_graph_repo_repo
+
+iplot <- oarch_graph_repo_repo |>
+  ggraph(layout = "graphopt") +
+  geom_edge_link(aes(alpha = n)) +
+  #geom_node_point(aes(colour = group)) +
+  #geom_node_text(aes(label = group)) +
+  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common contributors")
+
+girafe(ggobj = iplot)
+```
+
+```{r plot-graph-repo-repo-edge-betweenness-dendrogram, eval=FALSE}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
+                                   edges_common_users(oarch_graph),
+                                   directed = TRUE)
+
+oarch_graph_repo_repo |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 1) |>
+  mutate(group = as.character(group_edge_betweenness())) ->
+  oarch_graph_repo_repo
+
+xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
+xx <- as.dendrogram(xx)
+plot(xx)
+
+ggraph(xx, 'dendrogram', height = height)
+```
 
@@
 
-{r summary-graph-user-user}
+```{r summary-graph-user-user, eval=FALSE}
+oarch_graph_user_user |>
+  mutate(degree = centrality_degree(), 
+         betweenness = centrality_betweenness(weights = n), 
+         strength = centrality_degree(weights = n)) |>
+  filter(degree > 0) |> 
+  data.frame()
+```
 
@@
 
-{r plot-user-betweenness-dist}
-{r plot-user-betweenness-activity}
+```{r plot-user-betweenness-dist, eval=FALSE}
+oarch_graph_user_user |>
+  mutate(
+    betweenness = centrality_betweenness(weights = n)
+  ) |>
+  data.frame() |>
+  ggplot(aes(x = betweenness)) +
+  geom_density(bw = "nrd") +
+  scale_x_log10(labels = scales::number)
+```
+
+```{r plot-user-betweenness-activity, eval=FALSE}
+oarch_graph_user_user |>
+  mutate(
+    degree = centrality_degree(), 
+    betweenness = centrality_betweenness(weights = n), 
+    strength = centrality_degree(weights = n)
+  ) |>
+  as_tibble() |> 
+  left_join(oarch_contributors, by = c("user" = "contributor")) |>
+  ggplot(aes(betweenness, contribs)) +
+  geom_point() +
+  scale_x_log10() +
+  scale_y_log10() +
+  geom_smooth(method = "lm")
+```

### 2023-06-14 analysis/_paper.qmd
Exclude old files from quarto renders

--- 2023-06-12 analysis/paper.qmd
@@
 ```{r data-oarch}
-read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
-  mutate(category = na_if(category, "?")) ->
-  oarch
-
-# Extract clean repos from GitHub URLs
-oarch %>%
-  mutate(
-    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
-
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
+#merged
 ```
@@
 ```{r data-github}
-# If cached data at `analysis/data/derived_data/oarch.RData` is present, this
-# chunk will prefer to load it rather than make requests to the GitHub API.
-# Delete this file to fetch fresh data from the GitHub or modify 
-# `cached_gh_data` to read a different cached version.
-#
-# Note that fetching data from the GitHub API takes some time and is subject to
-# a rate limit:
-# <https://docs.github.com/en/rest/overview/resources-in-the-rest-api#rate-limiting>
-#
-# You will need to authenticate authenticate with GitHub credentials to get a
-# rate limit sufficient to obtain the required data, see 
-# <https://gh.r-lib.org/articles/managing-personal-access-tokens.html>
-# for options on how to do this.
-cached_gh_data <- here("analysis/data/derived_data/oarch.RData")
-if (!file_exists(cached_gh_data)) {
-  oarch <- mutate(oarch,
-                  gh_langs = map(oarch$gh_repo, gh_lang),
-                  gh_contribs = map(oarch$gh_repo, gh_contrib),
-                  gh_issues = map(oarch$gh_repo, gh_issue),
-                  gh_comments = map(oarch$gh_repo, gh_comment),
-                  gh_commits = map(oarch$gh_repo, gh_commit)
-  )
-  saveRDS(oarch, cached_gh_data)
-}
-oarch <- readRDS(cached_gh_data)
+#merged
 ```
@@
 ```{r plot-categories}
-oarch <- mutate(oarch,
-                category = recode(category,
-                                  "Specifications, protocols and schemas" = "Specifications etc.",
-                                  .default = category))
-oarch %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(n = n()) %>%
-  ggplot(aes(fct_reorder(category, n), n)) +
-  geom_col() +
-  coord_flip() +
-  labs(
-    title = "Most popular categories",
-    subtitle = "By number of projects",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/categories.png", width = 127, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r plot-platforms}
-oarch %>%
-  drop_na(platform) %>%
-  group_by(platform, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(
-    title = "Most popular platforms",
-    subtitle = "By number of projects, split by primary author",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/platforms.png", width = 127, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r plot-tags-cats}
-detail_tags <- c("Instrumental Neutron activation analysis",
-                 "Harris Matrix",
-                 "aDNA Simulators",
-                 "Damage profilers",
-                 "Metagenomic classifers",
-                 "aDNA simulators",
-                 "aDNA damage-aware genotypers",
-                 "Low-coverage contamination estimation",
-                 "Low-coverage adapted population-genetics tools",
-                 "aDNA preprocessing",
-                 "aDNA aligner",
-                 "aDNA assembler",
-                 "aDNA toolkit")
-
-oarch %>%
-  pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  filter(!tag %in% detail_tags) %>%
-  mutate(category = recode(category,
-                                  "Specifications etc." = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(tag, category) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  # facet_wrap(vars(category)) +
-  geom_col() +
-  scale_fill_bright() +
-  coord_flip() +
-  labs(
-    title = "Tags by category",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = "Number of projects",
-    fill = NULL
-  ) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_text(size = 7)
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/tags-cats.png", width = 254, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r plot-repo-hosts}
-# TODO: turn into table or just text
-oarch <- mutate(oarch, none = ifelse(is.na(github) &
-                                       is.na(gitlab) &
-                                       is.na(bitbucket) &
-                                       is.na(launchpad) &
-                                       is.na(codeberg),
-                                     "", NA))
-
-oarch %>%
-  pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
-               names_to = "repo_host", values_to = "repo_url") %>%
-  drop_na(repo_url) %>%
-  group_by(repo_host, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ungroup() %>%
-  mutate(repo_host = fct_reorder(repo_host, total),
-         repo_host = fct_relevel(repo_host, "none")) %>%
-  ggplot(aes(repo_host, n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(title = "Most popular repository hosts",
-       subtitle = "By number of projects, split by primary author",
-       caption = "Data: open-archaeo.info",
-       x = NULL, y = NULL)
-
-# Export for slides (half slide)
-# ggsave("figures/repo-hosts.png", width = 127, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
+#merged
 ```
@@
 ```{r plot-gh-cum}
-oarch %>% 
-  mutate(
-    lumped_category = recode(category,
-                             "Specifications etc." = "Documents",
-                             "Products" = "Documents",
-                             "Guides" = "Documents",
-                             "Lists and datasets" = "Documents",
-                             .default = category),
-    lumped_category = factor(lumped_category, c("Scripts",
-                                                "Documents",
-                                                "Packages and libraries",
-                                                "Standalone software")) ) ->
-  oarch_lumped
-
-p_gh_commits_cum <- oarch_lumped %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = lumped_category,
-             colour = lumped_category,
-             label = lumped_category)) +
-  geom_area() +
-  annotate("segment", 
-           x = ymd("2008-04-01", tz = "UTC"), 
-           xend = ymd("2008-04-01", tz = "UTC"),
-           y = 16000, yend = 4000,
-           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
-  annotate("text", 
-           x = ymd("2008-04-01", tz = "UTC"), y = 19000,
-           label = "GitHub\nlaunched",
-           vjust = 0) +
-  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = guide_none()) +
-  scale_fill_bright() +
-  labs(
-    y = "cumulative commits",
-    x = NULL, fill = NULL
-  ) +
-  theme_oarch() +
-  theme(legend.position = "top")
-
-# TODO: extend to end
-p_gh_repos_cum <- oarch_lumped %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  group_by(repo) %>% 
-  filter(datetime == min(datetime)) %>% # Only first commits
-  mutate(date = floor_date(datetime, "week")) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cat_total = sum(n_commits)) %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
-  ggplot(aes(date, cum_commits,
-             fill = lumped_category,
-             colour = lumped_category)) +
-  geom_area() +
-  annotate("segment",
-           x = ymd("2008-04-01", tz = "UTC"),
-           xend = ymd("2008-04-01", tz = "UTC"),
-           y = 40, yend = 10,
-           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
-  annotate("text",
-           x = ymd("2008-04-01", tz = "UTC"), y = 50,
-           label = "GitHub\nlaunched",
-           vjust = 0) +
-  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
-  scale_y_continuous(labels = scales::comma) +
-  scale_colour_bright(guide = guide_none()) +
-  scale_fill_bright(guide = guide_none()) +
-  labs(
-    y = "cumulative repositories",
-    x = NULL, fill = NULL
-  ) +
-  theme_oarch()
-
-p_gh_commits_cum / p_gh_repos_cum
-
-# Export for slides (full slide)
-# ggsave("figures/commits-cum.png", width = 254, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r plot-lifespan}
-oarch %>%
-  unnest(gh_commits) %>%
-  drop_na(datetime) %>%
-  group_by(item_name) %>%
-  add_tally() %>%
-  filter(n > 1) %>%
-  mutate(first_commit = min(datetime),
-         mid_commit = median(datetime),
-         last_commit = max(datetime)) %>%
-  ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
-             y = datetime,
-             ymin = first_commit,
-             ymax = last_commit)) +
-  geom_linerange(size = 0.05) +
-  geom_point(size = 0.1) +
-  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years", date_labels = "%Y") +
-  # coord_flip() +
-  labs(
-    title = "Project lifespan",
-    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
-    x = NULL, y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch() +
-  theme(
-    axis.line.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    axis.text.x = element_blank(),
-    panel.grid.minor.x = element_blank(),
-    panel.grid.major.x = element_blank()
-  )
-
-# Export for slides (full slide)
-# ggsave("figures/lifespan.png", width = 254, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r lifespan-stats}
-oarch %>% 
-  unnest(gh_commits) %>% 
-  drop_na(datetime) %>% 
-  group_by(item_name) %>%
-  mutate(n_contribs = map_int(gh_contribs, vctrs::vec_size)) %>% 
-  summarise(
-    category = first(category),
-    n_contribs = first(n_contribs),
-    n = n(),
-    first_commit = min(datetime),
-    last_commit = max(datetime),
-    lifespan = as.integer(last_commit - first_commit),
-    commit_rate = n / lifespan
-  ) ->
-  oarch_lifespan
+#merged
 ```
@@
 ```{r plot-lifespan-hist}
-# TODO: where does oarch_lifespan come from?
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan)) +
-  geom_histogram(binwidth = 100) +
-  scale_x_continuous(breaks = scales::breaks_width(1000)) +
-  labs(
-    title = "Project lifespan",
-    subtitle = "Time between first and latest commit, by repository",
-    x = "Days", y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/lifespan-hist.png", width = 127, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r plot-lifespan-vs-rate}
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = "Project lifespan vs. average commit rate",
-    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
-    caption = NULL
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-total.png", width = 127, height = 60, units = "mm")
-
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = NULL,
-    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-# Export for slides (half slide)
-# ggsave("figures/lifespan-vs-rate-contrib.png", width = 127, height = 60, units = "mm")
-
+#merged
 ```
@@
 ```{r plot-gh_contributors-histogram}
-oarch %>%
-  drop_na(gh_repo) %>%
-  unnest(gh_contribs) %>%
-  group_by(gh_repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Repository contributors",
-    x = "Number of contributors", y = "Repositories",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-hist.png", width = 127, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r plot-contrib-dist}
-oarch %>%
-  unnest(gh_contribs) %>%
-  drop_na(contributor) %>%
-  group_by(item_name) %>%
-  mutate(n_contributors = n()) %>%
-  filter(n_contributors > 1) %>%
-  mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Leading contributor", "Other contributors")) %>% 
-  group_by(item_name, big_boy) %>%
-  summarise(contributions = sum(contributions)) %>%
-  mutate(
-    p_contributions = contributions / sum(contributions),
-    p_big_boy = first(p_contributions)
-  ) %>%
-  ungroup() %>%
-  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
-  ggplot(aes(x = item_name, y = contributions,
-             fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Leading contributor")))) +
-  geom_col(position = "fill", colour = "white", size = 0.1) +
-  geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
-  scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
-  labs(title = "Repository contributors by proportion of contributions",
-       x = NULL, y = "Contributions",
-       fill = NULL,
-       caption = "Data: open-archaeo.info and GitHub API") +
-  coord_flip() +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
-  )
-
-# Export for slides (half slide)
-# ggsave("figures/contrib-dist.png", width = 127, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r plot-issues-hist}
-oarch %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of issues",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of issues", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/issues-hist.png", width = 127, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r plot-comments-hist}
-oarch %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Use of comments",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of comments", y = "Repositories"
-  ) +
-  theme_oarch()
-
-# Export for slides (half slide)
-# ggsave("figures/comments-hist.png", width = 127, height = 127, units = "mm")
+#merged
 ```
@@
 ```{r graph-data}
-oarch |>
-  drop_na(gh_repo) |>
-  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
-  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
-  oarch_graph
-
-oarch_graph <- bind_rows(
-  oarch_graph |>
-    unnest(gh_commits) |>
-    drop_na(author) |> # Apparently real commits, but we can't connect them
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "commit", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_comments) |>
-    drop_na(author) |> # Repos with no comments
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "comment", n_contrib = n(), .groups = "drop"),
-  oarch_graph |>
-    unnest(gh_issues) |>
-    drop_na(author) |> # Repos with no issues
-    group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "issue", n_contrib = n(), .groups = "drop")
-)
-
-# Cleaning
-oarch_graph |>
-  rename(user = author) |>
-  # Remove non-human users
-  filter(!str_detect(user, coll("[bot]"))) |>
-  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
-                      "gitter-badger", "dependabot-support")) |>
-  # Remove big, non-archaeology projects that flood the network
-  filter(!repo %in% c("Stellarium/stellarium", 
-                     "opengisch/QField",
-                     "lutraconsulting/input")) ->
-  oarch_graph
+#merged
 ```
@@
 ```{r plot-graph-repo-user}
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
-
-g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt") +
-  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
-  geom_node_point(aes(colour = node_type)) +
-  scale_edge_alpha(trans = "log")
-
-#g_repo_user
-girafe(g_repo_user)
+#merged
 ```
@@
 ```{r plot-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph, category),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
+#merged
 ```
@@
 ```{r plot-graph-repo-repo-edge-betweenness}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  #geom_node_point(aes(colour = group)) +
-  #geom_node_text(aes(label = group)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
-
-girafe(ggobj = iplot)
+#merged
 ```
@@
 ```{r plot-graph-repo-repo-edge-betweenness-dendrogram}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = TRUE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
-xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
-xx <- as.dendrogram(xx)
-plot(xx)
-
-ggraph(xx, 'dendrogram', height = height)
-
+#merged
 ```
@@
 ```{r graph-user-user}
-oarch_graph_user_user <- tbl_graph(
-  nodes = nodes_users(oarch_graph), 
-  edges = edges_common_repos(oarch_graph),
-  directed = FALSE
-)
+#merged
 ```
@@
 ```{r summary-graph-user-user}
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree(), 
-         betweenness = centrality_betweenness(weights = n), 
-         strength = centrality_degree(weights = n)) |>
-  filter(degree > 0) |> 
-  data.frame()
+#merged
 ```
@@
 ```{r plot-graph-user-user}
-oarch_graph_user_user |>
-  mutate(betweenness = centrality_betweenness(weights = n)) ->
-  oarch_graph_user_user
-
-oarch_graph_user_user |> 
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  # geom_node_point() +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common repositories")
+#merged
 ```
@@
 ```{r plot-user-betweenness-dist}
-oarch_graph_user_user |>
-  mutate(
-    betweenness = centrality_betweenness(weights = n)
-  ) |>
-  data.frame() |>
-  ggplot(aes(x = betweenness)) +
-  geom_density(bw = "nrd") +
-  scale_x_log10(labels = scales::number)
+#merged
 ```
@@
 ```{r plot-user-betweenness-activity}
-oarch_graph_user_user |>
-  mutate(
-    degree = centrality_degree(), 
-    betweenness = centrality_betweenness(weights = n), 
-    strength = centrality_degree(weights = n)
-  ) |>
-  as_tibble() |> 
-  left_join(oarch_contributors, by = c("user" = "contributor")) |>
-  ggplot(aes(betweenness, contribs)) +
-  geom_point() +
-  scale_x_log10() +
-  scale_y_log10() +
-  geom_smooth(method = "lm")
+#merged
 ```

### 2023-06-14 analysis/99-conclusion.qmd
temporary commit before pull

@@
 While there is an emerging community of practice around open source in archaeology, collaboration remains limited.
-Most work is performed solo and is short-lived.^[**NOTE:** I wonder if we can say something about change over time, for instance, if things are getting better.]
+Most work is performed solo and is short-lived.
+
+**NOTE:** I wonder if we can say something about change over time, for instance, if things are getting better.
 
@@
 This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
+
+future work might include qualitative analysis of issues and comments
+break out and form new kinds of collaborative commitments suited to this kind of work and the media through with it is performed
+relate precarity of central actors with similar phenomena in open source in general
+open source is not a magic bullet, it is people, people who are based in the "real world"

### 2023-06-14 analysis/_01-introduction.qmd

Open archaeology constitutes a series of practices and principles in support greater transparency and access to research processes and outputs. The open science movement operates in various domains of research operations, such as publishing, data-sharing, and education, and continues to expand its reach into the arenas of bibliographic metadata, review processes, and grantsmanship. While these venues for transformation represent relatively novel ways of doing research, academic ventures into open source is more complicated. The open source movement has its own distinct history and developed out of pressures and circumstances beyond those in which the open science movement operates. However, the open source movement directly inspires many aspects of open science, and, crucially, informs how the open science movement has operationalized itself. For instance, the open science movement relies on infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plaintext, diffs, emphasis on formal licencing, the whole hacker aesthetic). But academic open source is distinct from open source that underlies massive projects that literally run the internet (and by extension the global economy), such as the Linux kernel, the Firefox web browser, and openSSL. While they may use similar tools and technical protocols, they are governed by different social and professional pressures. In other words, making your code open source has different meanings, impacts and implications for archaeologists and professional software developers.

This is also complicated by the kind of rhetoric about academic open source that, based on our personal experiences, is discordant with how academics actually _do_ open source. Is open source actually making research processes more transparent and improve research outcomes? Is it actually boosting efficiency by establishing a common store of knowledge and productive code? Is it actually helping to foster new globe-spanning connections and lead to novel research trajectories that would not otherwise come to pass? These outcomes are not a given. They only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results. Beyond simply making their code available on the web, do archaeologists also implement social strategies to advance open source ideals? Does archaeological open source actually help achieve greater transparency, sustainability, and community participation? And if not, what does it actually achieve?

This article presents a survey of archaeological software development with two goals in mind:

(1) we identify *what* kinds of software archaeologists are making;
(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.

Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.

### 2023-06-14 analysis/_03-data_methodology.qmd
Exclude old files from quarto renders
Add headings

@@
+# Methods & Materials <!-- Methods & Materials -->
+<!-- TODO: snappier title? Maybe incoroporating 'open-archaeo'? -->
+
 Our findings are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
@@
 from the API using the R package `gh` [@gh].
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but amongst those that do, GitHub is by far the most popular host.
+This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but amongst those that do, GitHub is by far the most popular host ([@tbl-repo-hosts]).
 
@@
 
-```{r plot-repo-hosts}
-# TODO: turn into table or just text
-# TODO: Indicate the numerical value at the end of each bar in the plot-repo-hosts figure.
-oarch <- mutate(oarch, none = ifelse(is.na(github) &
-                                       is.na(gitlab) &
-                                       is.na(bitbucket) &
-                                       is.na(launchpad) &
-                                       is.na(codeberg),
-                                     "", NA))
-
-oarch %>%
+```{r tbl-repo-hosts}
+#| tbl-cap: Number of open archaeology projects using GitHub and other repository hosts
+oarch |>
+  mutate(none = ifelse(is.na(github) & is.na(gitlab) & is.na(bitbucket) &
+                         is.na(launchpad) & is.na(codeberg), "", NA)) |>
   pivot_longer(c(github, gitlab, bitbucket, launchpad, codeberg, none),
-               names_to = "repo_host", values_to = "repo_url") %>%
-  drop_na(repo_url) %>%
-  group_by(repo_host, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ungroup() %>%
-  mutate(repo_host = fct_reorder(repo_host, total),
-         repo_host = fct_relevel(repo_host, "none")) %>%
-  ggplot(aes(repo_host, n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(title = "Most popular repository hosts",
-       subtitle = "By number of projects, split by primary author",
-       caption = "Data: open-archaeo.info",
-       x = NULL, y = NULL)
-```
-
-```{r table-no-repo}
-# TODO: prettify with gt()
-filter(oarch, none == "") %>%
-  count(category)
+               names_to = "repo_host", values_to = "repo_url") |>
+  drop_na(repo_url) |>
+  count(category, repo_host) |>
+  pivot_wider(names_from = category, values_from = n, values_fill = 0) |>
+  mutate(total = rowSums(across(Guides:`Standalone software`)), 
+         .before = Guides) |>
+  arrange(repo_host == "none", desc(total)) |>
+  mutate(repo_host = recode(repo_host,
+                            github = "GitHub",
+                            codeberg = "Codeberg",
+                            gitlab = "GitLab",
+                            bitbucket = "BitBucket",
+                            launchpad = "Launchpad",
+                            none = "None")) |>
+  gt() |>
+  cols_label(repo_host = "Host", total = "Total") |>
+  tab_spanner("By category", Guides:`Standalone software`)
 ```

### 2023-06-14 analysis/openarchaeo_collaboration.qmd

@@
 ---
-title: "openarchaeoCollaboration"
+title: Open archaeology, open source?
+subtitle: Collaborative practices in an emerging community of archaeological software engineers
 author:
@@
 date: "`r Sys.Date()`"
+abstract: |
+  Surveying the first quarter-century of computer applications in archaeology, Scollar [-@scollar1999] lamented that the field relied almost exclusively on "hand-me-down" tools repurposed from other disciplines.
+25 years later, this is no longer the case: computational archaeologists often find themselves practicing the dual roles of data analyst and research software engineer [@baxter2012], developing and applying new tools that are tailored specifically to archaeological problems and archaeological methods.
+Though this trend can be traced to the very earliest days of the field [@cowgill1967], its most recent manifestation is distinguished by its apparently embrace of practices from free and open source software (FOSS).
+Most prominently, since around 2015, there has been a rapid uptake of workflow tools designed for open source development communities, such as the version control system `git` and associated online source code management platforms (e.g.
+GitHub, GitLab).
+These tools facilitate collaboration among developers and users of FOSS software using patterns that can diverge quite radically from conventional scholarly norms [@tennant2020].
+
+  In this paper, we investigate modes of collaboration in this emerging community of practice using open-archaeo (https://open-archaeo.info/), a curated list of 500+ active open projects in archaeology, augmented where available with data on their activity on GitHub.
+We apply network analysis to map the networks of collaboration that have emerged among archaeologists using GitHub, and conduct a quantitative exploratory analysis to characterize the nature and intensity of these collaborations.
+Our results show an uneven adoption of open source collaborative practices beyond the basic use of `git` as a version control system and GitHub to host source code.
+A majority of repositories have 1–3 contributors, with only a few projects distinguished by an active and diverse developer base.
+More community-oriented features, such as GitHub's 'issues' and 'comments', are used in only a minority of repositories.
+Preliminary findings from ongoing work also shows how collaborative behaviours vary across different subcommunities.
+By analyzing interactions among archaeologists using the GitHub platform, including the ways they contribute to collective code bases, we demonstrate how archaeological software engineering is beginning to foster new kinds of collaborative commitments that blend open source protocols, DIY work ethic, conventional forms of scholarly communication, and cultural practices unique to archaeology.
+format:
+  html:
+    page-layout: article
+    toc: true
+    number-sections: true
+    embed-resources: true
 ---

### 2023-06-14 analysis/_04-what_is_open_archaeology.qmd
Exclude old files from quarto renders
Add headings
Revise fig-cats-plats

@@
-## Breakdown of tags and categories
+# Open archaeology <!-- Results I (description of what's there) -->
+
+Our analysis documented an the widespread presence of open archaeology projects on GitHub and other online platforms.
 
@@
 |
+:Categorisation scheme used by open-archaeo {#tbl-category-defs}
 
@@
 
-## Breakdown of platforms and languages
+## Platforms and languages
 
@@
 
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing "platform" -- for example a package for a programming language like R, or a plugin for an application like QGIS.
+Most projects included in open-archaeo are designed to be used as a package or library atop am existing "platform" -- for example a package for a programming language like R, or a plugin for an application like QGIS ([@fig-cats-plats]).
 The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
 Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
+A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use .
 
-```{r plot-categories}
-oarch <- mutate(oarch,
-                category = recode(category,
-                                  "Specifications, protocols and schemas" = "Specifications etc.",
-                                  .default = category))
-oarch %>%
-  drop_na(category) %>%
-  group_by(category) %>%
-  summarise(n = n()) %>%
+```{r fig-cats-plats}
+#| fig-cap: Number of open archaeology projects by category and platform
+#| fig-subcap: 
+#|   - Categories
+#|   - Platforms
+
+oarch |>
+  drop_na(category) |>
+  group_by(category) |>
+  summarise(n = n()) |>
   ggplot(aes(fct_reorder(category, n), n)) +
   geom_col() +
+  geom_text(aes(label = category), hjust = 0, nudge_y = 1) +
   coord_flip() +
-  labs(
-    title = "Most popular categories",
-    subtitle = "By number of projects",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
+  scale_y_continuous(limits = c(0, 300)) +
+  labs(x = NULL, y = NULL) +
+  theme_oarch() +
+  theme(
+    axis.text.y = element_blank(), 
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_blank()
+  )
+
+oarch |>
+  drop_na(platform) |>
+  count(platform) |>
+  arrange(n) |>
+  ggplot(aes(fct_reorder(platform, n), n)) +
+  geom_col(colour = "#333333") +
+  geom_text(aes(label = platform), hjust = 0, nudge_y = 1) +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 200)) +
+  labs(x = NULL, y = NULL) +
+  theme_oarch() +
+  theme(
+    axis.text.y = element_blank(), 
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_blank()
+  )
 ```
 
-R is overwhelmingly dominant as a platform, closely followed by Python.
+R is overwhelmingly dominant as a platform, closely followed by Python [@fig-cats-plats].
 Plugins for QGIS and mobile apps are also relatively popular.
@@
 ```{r plot-platforms}
-oarch %>%
-  drop_na(platform) %>%
-  group_by(platform, author1_name) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(platform, total), n, group = author1_name)) +
-  geom_col(colour = "#333333") +
-  coord_flip() +
-  labs(
-    title = "Most popular platforms",
-    subtitle = "By number of projects, split by primary author",
-    caption = "Data: open-archaeo.info",
-    x = NULL, y = NULL
-  ) +
-  theme_oarch()
 ```

### 2023-06-21 analysis/_01-introduction.qmd
Some more work on the intro and background

@@
-Open archaeology constitutes a series of practices and principles in support greater transparency and access to research processes and outputs.
-The open science movement operates in various domains of research operations, such as publishing, data-sharing, and education, and continues to expand its reach into the arenas of bibliographic metadata, review processes, and grantsmanship.
-While these venues for transformation represent relatively novel ways of doing research, academic ventures into open source is more complicated.
-The open source movement has its own distinct history and developed out of pressures and circumstances beyond those in which the open science movement operates.
-However, the open source movement directly inspires many aspects of open science, and, crucially, informs how the open science movement has operationalized itself.
-For instance, the open science movement relies on infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plaintext, diffs, emphasis on formal licencing, the whole hacker aesthetic).
-But academic open source is distinct from open source that underlies massive projects that literally run the internet (and by extension the global economy), such as the Linux kernel, the Firefox web browser, and openSSL.
-While they may use similar tools and technical protocols, they are governed by different social and professional pressures.
-In other words, making your code open source has different meanings, impacts and implications for archaeologists and professional software developers.
+# Introduction
 
-This is also complicated by the kind of rhetoric about academic open source that, based on our personal experiences, is discordant with how academics actually _do_ open source.
-Is open source actually making research processes more transparent and improve research outcomes?
+Open archaeology constitutes a series of practices and principles that support greater transparency to research processes and access to data that underlie synthetic outcomes (Beck and Neylon 2012: 479-480).
+As part of the broader open science movement, open archaeology promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, and education, but it continues to expand its reach into the arenas of bibliographic metadata, review processes, and grantsmanship.
+Archaeologists are also actively venturing into open source software development as means of sharing their research processes and creating tools and resources for general use.
+
+However, academic implementation of open source software development models is complicated by the open source movement's distinct history and its nuanced relationship with the open science movement.
+The open source movement developed out of pressures and circumstances that differ from those in which the open science movement operates, and yet the former serves as direct inspiration for the latter.
+For instance, the Open Knowledge Foundation, which publishes a widely accepted defitnion of "open" in the context of scholarly communication, draws directly from the definition of "open source" published by the Open Source Initiative (@openknowledgefoundation2023; @opensourceinitiative2007).
+The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plaintext, diffs, emphasis on formal licencing, the whole hacker aesthetic).
+However, academic work, including the development of academic software, differs signififcantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, the Firefox web browser, and openSSL.
+While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional pressures.
+In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers (Kelty 2008:xx).
+
+The relationship between open science and open source is also complicated by rhetorical claims that, based on our personal experiences, are discordant with how academics actually _do_ open source.
+Does academic open source actually make research processes more transparent and improve research outcomes?
 Is it actually boosting efficiency by establishing a common store of knowledge and productive code?
 Is it actually helping to foster new globe-spanning connections and lead to novel research trajectories that would not otherwise come to pass?
-These outcomes are not a given.
-They only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
-Beyond simply making their code available on the web, do archaeologists also implement social strategies to advance open source ideals?
+Despite hopeful aspirations espoused by open archaeology advocates (Kansa  et al.
+2014; Kintigh et al.
+2015; others), these outcomes are not a given.
+Based on casual observations from our own experiences, we find that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
+Unfortunately, these values are not widespread among archaeologists, even those who are staunch supporters of open science, due to systemic challenges and tensions.
+
+That being said, archaeologists are prolific software developers, as is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources, which records XXX items as of the time of publication.
+But beyond simply making their code available on the web, do archaeologists also implement social strategies to advance open source ideals?
 Does archaeological open source actually help achieve greater transparency, sustainability, and community participation?
@@
 
-Our investigation considers how archaeological software development may be benefiting from or missing out on the affordances that open source development models-- specifically, the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
+Our statistical and network analyses consider how archaeological software development may be benefiting from or missing out on the affordances that open source development models provide, specifically the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
+We found that the vast majority of projects lack any collective effort and die shortly after they are initiated.
+Projects that are maintained by multiple contributors tend to be backed by funded initiatives who employ locally and socially engaged individuals, or are warranted by participation in a scholarly community exhibiting genuine need for a particular kind of resource.
+Moreover, we found that the individuals who play critical roles in supporting the archaeological open source community are precariously employed workers.
+Contrary to popular claims about open source being inherently distributed, resilient, and open-ended, we found that, overall, archaeological open source is actually quite centralized, fragile, and based nearly exclusively on existing professional connections and endeavours.

### 2023-06-21 analysis/_02-background.qmd
Exclude old files from quarto renders
Add headings
Some more work on the intro and background

--- 2023-03-30 analysis/02-background.qmd
@@
-## What is open source?
+# Open science and open source <!-- Background -->
+
+## Open source
 
 Open source is a software development model that prioritizes transparent work processes.
-Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development (Kelty 2008: chapter 3, especially starting at page 99).
+Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development (@kelty2008: chapter 3, especially starting at page 99).
 By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base.
@@
 This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion.
-Personal ego is minimized in favour of co-creating stable and functional outcomes.
-(Coleman 2012)
+Personal ego is minimized in favour of co-creating stable and functional outcomes (@coleman2012).
 
@@
 They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term.
-Open source is more than just putting your code online; to be successful, it requires participation in a social experience.
-(Kelty 2008; O'Neill 2009; Ratto dissertation and derived works)
+Open source is more than just putting your code online; to be successful, it requires participation in a social experience (@kelty2008; Ratto dissertation and derived works).
 
@@
 
-## How does open source relate to open science?
+## Open science
 
 The open science movement comprises a series of practices and principles intended to make research more accessible, transparent, and efficient.
-Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" (https://opendefinition.org/).
+Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" (@openknowledgefoundation2023).
 This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement.
@@
 However, while open science does have the _potential_ to effect radical change, this is not a given.
-The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals.
+The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals (@mirowski2018).
 Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails.
@@
 
-Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain.
+Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain (cf.
+@tennant2020).
 But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce.
@@
 Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out.
-Once a project is completed and credit has been allocated, it is common for scientists to mark their projects as completed and move onto new endeavours.
+Once a project is completed and credit has been allocated, it is common for scientists to mark their projects as completed and move on to new endeavours.
 Open source projects, on the other hand, are motivated by a more practical need for the software to function properly, and contributors may remain actively or sporadically involved to satisfy users' needs.
 
-## Open archaeology at a crossroads
-
-The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings.
-Open science advocates, including archaeologists, have put a lot of effort into developing information infrastructures, or systems designed to facilitate the transmission of data on a global scale.
-And as with open science in general, open archaeology has been wrapped up in discourse on the best tools to use, optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards (cf.
-Kansa and Kansa 2014; Kintigh 2014; Roosevelt 2015).
-This continues to be a prominent part of discourse on the development of open infrastructures (cf.
-Kansa 2022 on unique  identifiers; something about aDNA pipelines; Hacıgüzeller 2021 on CIDOC-CRM (though she is actually very critical); Huvila and his team on formal paradata), but more recent reflections on the use of these infrastructures recognize underlying and systemtic commitments and assumptions that inform how these systems are used and valued by various stakeholders.
-For instance, the pressures and circumstances of being an archaeologist, such as the need to publish stable research outputs, and archaeologists' tendency to move on from projects after they are completed, contribute towards shaping data sharing platforms as delegated archival services, whereupon the obligation to care for data over the long-term is passed on to others.
-In response to criticisms that highlight how these systems add little practical downstream value (Huggett 2018; Sobotkova 2018), open archaeology advocates have begun to prioritize the role that information infrastructures may play in contexts of data reuse, namely by implementing the FAIR principles in order to ensure that data are findable, accessible, interoperable, and reusable (Kansa 2023 in that recent special issue in Advances in Archaeological Practice; one or two additional sources).
-This example is matched by similar efforts to raise the value proposition of using open science tools and systems, such as through an emphasis on enhancing R package development and documentation skills so that code may be run and evaluated in alternative computing environments (cite Marwick and/or Roe), and [another example].
-
-Additionally, archaeologists are becoming more comfortable showcasing and actively discussing undervalued research ouputs such as software and datasets, which are reviewed and published by venues like Peer Community In, the Journal of Archaeological Data, and the Journal of Open Source Software.
-Professional blogs, social media and podcasts are also gaining traction as proper venues for disseminating one's work.
-This is consistent with archaeology's historical willingness to experiment with new methods and media and its casual attitude towards scholarly communication.
-Archaeologists' embrace of methodological plurality and their positions as interdisciplinary scholars scattered across institutional boundaries has prepared them for this kind of pragmatic, improvisatory approach to scholarly communication via the web.
-There is enormous potential for archaeologists to reinvent novel forms of scholarly communication and extend a new set of collaborative commitments that go beyond those typically recognized by scholarly institutions, and which correspond with open science principles.
-But this depends on archaeologists' willingness to continue experimenting and on the community's willingness to buy into novel value regimes.
+The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a genuine desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings.
+However, the predominant concern with implementing best tools to use, adopting optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards (cf.
+@kansa2014; @kintigh2015; @roosevelt2015) distract from fundamental tensions and contradictions regarding the actual value of working in the open.
+For instance, @faniel2013 :299-301, @atici2013 :676-677, @huggett2018, @sobotkova2018, and @opitz2021 demonstrate that to make the reuse of archaeological data feasible and useful in a practical sense, it is necessary to re-introduce social friction that these infrastructures are designed to eliminate.
+In other words, the pressures and circumstances of being an archaeologist and doing archaeological research assert themselves when attempting to make practical use of these infrastructures, and therefore must be accounted for in their design and implementation.
+Here we aim to identify similar sources of dissonance with regards to the promise, potential, and actual implementation of open source software development models among archaeologists.

### 2023-06-27 analysis/_01-introduction.qmd
Refining the intro and background some more.

@@
 
-Open archaeology constitutes a series of practices and principles that support greater transparency to research processes and access to data that underlie synthetic outcomes (Beck and Neylon 2012: 479-480).
-As part of the broader open science movement, open archaeology promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, and education, but it continues to expand its reach into the arenas of bibliographic metadata, review processes, and grantsmanship.
-Archaeologists are also actively venturing into open source software development as means of sharing their research processes and creating tools and resources for general use.
+In their seminal vision for open archaeology, Beck and Neylon (2012: 480-481) identified the movement as comprising a series of principles and practices "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity".
+Open archaeology therefore promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, education, and review processes.
+Archaeologists are also actively engaged in open source software development as means of sharing their research processes and creating tools and resources for general use.
 
-However, academic implementation of open source software development models is complicated by the open source movement's distinct history and its nuanced relationship with the open science movement.
-The open source movement developed out of pressures and circumstances that differ from those in which the open science movement operates, and yet the former serves as direct inspiration for the latter.
-For instance, the Open Knowledge Foundation, which publishes a widely accepted defitnion of "open" in the context of scholarly communication, draws directly from the definition of "open source" published by the Open Source Initiative (@openknowledgefoundation2023; @opensourceinitiative2007).
-The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plaintext, diffs, emphasis on formal licencing, the whole hacker aesthetic).
-However, academic work, including the development of academic software, differs signififcantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, the Firefox web browser, and openSSL.
-While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional pressures.
+However, academic open source has a complicated relationship with open source as practiced by professional software developers, which has its own distinct history and is framed by different objectives, challenges, and value regimes.
+Despite this, the open science movement, within which open archaeology emerged, draws direct inspiration from open source.
+For instance, the Open Knowledge Foundation, which publishes a widely accepted defitnion of "open" in the context of scholarly communication, explicitly refers to the definition of "open source" published by the Open Source Initiative, an authoritative open source advocacy group (@openknowledgefoundation2023; @opensourceinitiative2007).
+The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plaintext, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic).
+However, academic work, including the development of academic software, differs signififcantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL, and the Firefox web browser.
+While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests.
 In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers (Kelty 2008:xx).
@@
 2015; others), these outcomes are not a given.
-Based on casual observations from our own experiences, we find that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
-Unfortunately, these values are not widespread among archaeologists, even those who are staunch supporters of open science, due to systemic challenges and tensions.
+We believe that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
+However, practical circumstances and systemic value regimes that frame what it means to work as an archaeologist presently inhibit the potential for radical transformation, even among open science's most ardent supporters.
 
-That being said, archaeologists are prolific software developers, as is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources, which records XXX items as of the time of publication.
+That being said, archaeologists are prolific software developers.
+This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes XXX records as of the time of publication.
 But beyond simply making their code available on the web, do archaeologists also implement social strategies to advance open source ideals?
@@
 Our statistical and network analyses consider how archaeological software development may be benefiting from or missing out on the affordances that open source development models provide, specifically the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
-We found that the vast majority of projects lack any collective effort and die shortly after they are initiated.
+We found that the vast majority of projects lack any collective effort and activity tends to abruptly end shortly after work is initiated.
 Projects that are maintained by multiple contributors tend to be backed by funded initiatives who employ locally and socially engaged individuals, or are warranted by participation in a scholarly community exhibiting genuine need for a particular kind of resource.

### 2023-06-27 analysis/_02-background.qmd
Refining the intro and background some more.

@@
 By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base.
-Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision.
+Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision (Kelty 2008 on the emacs saga).
 An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.
@@
 Open source has traditionally been referred to as being based on meritocratic principles.
-A good test of whether a contribution should be included in a published software release is whether it is functional.
+A good test of whether a contribution should be included in a published software release is whether it is functional (Kelty 2008: 220).
 Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues.
 This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion.
-Personal ego is minimized in favour of co-creating stable and functional outcomes (@coleman2012).
+Personal ego is minimized in favour of co-creating stable and functional outcomes (@coleman2012; O'Neil 2009: the debian saga).
 
 However, this is not the same as saying that open source is completely anarchic or based on the "wisdom of crowds".
-In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants.
+In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants (O'Neil 2009).
 They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term.
@@
 This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement.
-However, most open science advocates claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.
+However, most open science advocates (including archaeologists, as elicited by Beck and Neylon 2012; Marwick et al.
+2017) claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.
 
@@
 
-Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain (cf.
-@tennant2020).
+Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain (@tennant2020).
 But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce.
@@
 Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out.
-Once a project is completed and credit has been allocated, it is common for scientists to mark their projects as completed and move on to new endeavours.
-Open source projects, on the other hand, are motivated by a more practical need for the software to function properly, and contributors may remain actively or sporadically involved to satisfy users' needs.
+Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours.
+Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software.
 

### 2023-06-27 analysis/_01-introduction.qmd
Properly formatted bibliographic references

@@
 
-In their seminal vision for open archaeology, Beck and Neylon (2012: 480-481) identified the movement as comprising a series of principles and practices "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity".
+In their seminal vision for open archaeology, Beck and Neylon [-@beck2012 480-481] identified the movement as comprising a series of principles and practices "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity".
 Open archaeology therefore promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, education, and review processes.
@@
 Despite this, the open science movement, within which open archaeology emerged, draws direct inspiration from open source.
-For instance, the Open Knowledge Foundation, which publishes a widely accepted defitnion of "open" in the context of scholarly communication, explicitly refers to the definition of "open source" published by the Open Source Initiative, an authoritative open source advocacy group (@openknowledgefoundation2023; @opensourceinitiative2007).
+For instance, the Open Knowledge Foundation [-@openknowledgefoundation2023] publishes a widely accepted defitnion of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group.
 The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plaintext, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic).
@@
 While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests.
-In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers (Kelty 2008:xx).
+In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers [@kelty2008 xx].
 
@@
 Is it actually helping to foster new globe-spanning connections and lead to novel research trajectories that would not otherwise come to pass?
-Despite hopeful aspirations espoused by open archaeology advocates (Kansa  et al.
-2014; Kintigh et al.
-2015; others), these outcomes are not a given.
+Despite hopeful aspirations espoused by open archaeology advocates [see @kansa2014; @kintigh2015], these outcomes are not a given.
 We believe that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.

### 2023-06-27 analysis/_02-background.qmd
Properly formatted bibliographic references

@@
 Open source is a software development model that prioritizes transparent work processes.
-Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development (@kelty2008: chapter 3, especially starting at page 99).
+Initially driven by the idea that computer users should be free to understand and manipulate the software that they install on their computers (e.g., "free software", as initially conceived by the Free Software Foundation), open source has become a means of collaborative software development [@kelty2008: chapter 3, especially starting at page 99].
 By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base.
-Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision (Kelty 2008 on the emacs saga).
+Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision [@kelty2008 on the emacs saga].
 An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.
@@
 Open source has traditionally been referred to as being based on meritocratic principles.
-A good test of whether a contribution should be included in a published software release is whether it is functional (Kelty 2008: 220).
+A good test of whether a contribution should be included in a published software release is whether it is functional [@kelty2008 220].
 Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues.
 This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion.
-Personal ego is minimized in favour of co-creating stable and functional outcomes (@coleman2012; O'Neil 2009: the debian saga).
+Personal ego is minimized in favour of co-creating stable and functional outcomes [@coleman2012; @oneil2009 on the debian saga].
 
 However, this is not the same as saying that open source is completely anarchic or based on the "wisdom of crowds".
-In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants (O'Neil 2009).
+In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants [@oneil2009].
 They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term.
-Open source is more than just putting your code online; to be successful, it requires participation in a social experience (@kelty2008; Ratto dissertation and derived works).
+Open source is more than just putting your code online; to be successful, it requires participation in a social experience [@kelty2008, Ratto dissertation and derived works].
 
@@
 The open science movement comprises a series of practices and principles intended to make research more accessible, transparent, and efficient.
-Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" (@openknowledgefoundation2023).
+Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" [@openknowledgefoundation2023].
 This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement.
-However, most open science advocates (including archaeologists, as elicited by Beck and Neylon 2012; Marwick et al.
-2017) claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.
+However, most open science advocates (including archaeologists, as elicited by @beck2012 and @marwick2017) claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.
 
@@
 However, while open science does have the _potential_ to effect radical change, this is not a given.
-The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals (@mirowski2018).
+The social and institutional contexts in which we do science is firmly embedded within capitalist and neoliberal power structures that reward individualistic competition and do little to actually encourage equitable and accessible research practices, and as such, make it difficult to fully embrace open science ideals [@mirowski2018].
 Moreover, the open science movement, which is dominated by STEM disciplines, prioritizes a grossly simplified and asocial notion of what science is and entails.
@@
 
-Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain (@tennant2020).
+Open science is typically compared with the open source movement in that they both involve a distributed, digitally-mediated and worldwide labour force, who somehow derive rough consensus directed towards assets held in the public domain [@tennant2020].
 But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce.
@@
 The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a genuine desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings.
-However, the predominant concern with implementing best tools to use, adopting optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards (cf.
-@kansa2014; @kintigh2015; @roosevelt2015) distract from fundamental tensions and contradictions regarding the actual value of working in the open.
-For instance, @faniel2013 :299-301, @atici2013 :676-677, @huggett2018, @sobotkova2018, and @opitz2021 demonstrate that to make the reuse of archaeological data feasible and useful in a practical sense, it is necessary to re-introduce social friction that these infrastructures are designed to eliminate.
+However, the predominant concern with implementing best tools to use, adopting optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards [cf.
+@kansa2014; @kintigh2015; @roosevelt2015] distract from fundamental tensions and contradictions regarding the actual value of working in the open.
+For instance, Faniel et al.
+[-@faniel2013 299-301], Atici et al.
+[-@atici2013 676-677], @huggett2018, @sobotkova2018, and @opitz2021 demonstrate that to make the reuse of archaeological data feasible and useful in a practical sense, it is necessary to re-introduce social friction that these infrastructures are designed to eliminate.
 In other words, the pressures and circumstances of being an archaeologist and doing archaeological research assert themselves when attempting to make practical use of these infrastructures, and therefore must be accounted for in their design and implementation.

### 2023-06-27 analysis/_03-data_methodology.qmd
Refining the intro and background some more.
Properly formatted bibliographic references

@@
 
-Our findings are based on an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
+Our study comprises an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
 As of right now it includes 408 items, which are rendered for the web at [open-archaeo.info](open-archaeo.info).
@@
 That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc.
-from the API using the R package `gh` [@gh].
+from the API using the R package `gh` [@csardi2023].
 This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but amongst those that do, GitHub is by far the most popular host ([@tbl-repo-hosts]).

### 2023-06-27 analysis/openarchaeo_collaboration.qmd
Revise fig-cats-plats
Properly formatted bibliographic references

@@
 By analyzing interactions among archaeologists using the GitHub platform, including the ways they contribute to collective code bases, we demonstrate how archaeological software engineering is beginning to foster new kinds of collaborative commitments that blend open source protocols, DIY work ethic, conventional forms of scholarly communication, and cultural practices unique to archaeology.
+bibliography: openarchaeo-collaboration.bib
 format:
@@
 ```{r plot-theme}
-theme_oarch <- function() theme(
-)
+theme_oarch <- function() {
+  theme_minimal() +
+  theme()
+}
 ```

### 2023-09-05 analysis/_99-conclusion.qmd
Exclude old files from quarto renders
Add headings
Added more firm statements of our findings.
Also more explicitly related our work within broader critique of open archaeology.

--- 2023-03-31 analysis/99-conclusion.qmd
@@
-Collaborative open source software development in archaeology has seen a rapid and sustained increase beginning around 2014 or 2015.
-This is marked by a variety of applications and use cases.
-While there is an emerging community of practice around open source in archaeology, collaboration remains limited.
-Most work is performed solo and is short-lived.^[**NOTE:** I wonder if we can say something about change over time, for instance, if things are getting better.]
+# Conclusion
+Archaeology is a "magpie" discipline, using tools and methods that originate from a variety of other disciplines and industries to better understand the past. 
+Moreover, archaeologists are accustomed to working on shoestring budgets, having to adapt their tools to suit specific, non-generalizable situations, and applying a DIY work ethic to create and co-creating with others to get the job done (Batist et al 2021; Morgan and Eve 2012; Caraher 2019).
+It is therefore unsprising that archaeologists are prolific open source software developers.
 
+In this paper we take stock of archaeologists' immense creative capabilities. 
+First, we surveyed what kinds of software archaeologists are making and tracked progression over time. 
+We showed that open source software development in archaeology has seen a rapid and sustained rise beginning around 2014. 
+This is marked by a variety of applications and use cases, including the use of git and GitHub to track and host content other than code. 
+Moreover, archaeologists are very involved in broader scripting ecosystems, as is evident through the predominant creation of R packages and Python libraries designed to process the rich variety of archaeological information.
+At the same time, archaeologists also create standalone software for more intensive tasks that require greater access to system resources or that warrant more complex user interfaces than what R and Python IDEs are capable of providing.
+
+Looking at the distribution of tags, we note that, among items listed as software, repositories tend to be focused on 
+various means of identifying distribution patterns (spatial, temporal, statistical), 
+calibrating data obtained from various instrumental methods (XRF, luminescence dating), 
+supporting specialized finds analysis (zooarchaeology, palaeobotany, archaeogenetics), 
+and supporting the collection and processing of archaeological materials.
+These foci signify gaps in the archaeological toolbox that archaeologists recognized, and have attempted to fill, on their own terms.
+In future studies, it would be interesting to examine how the purposeful expansion of open source tools corresponds with broader methodological trends apparent in publishing patterns or through other expressions of archaeological interest (e.g.
+social media posts, conference sessions).
+
+While there is an emerging community of practice around open source in archaeology, we observed that collaboration remains limited. 
+Most work is performed individually and is short-lived. 
 The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base. 
@@
 Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects. 
-This may relate to our positions as academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
+This may relate to the fact that most developers on this list are academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
+
+However, we seem to have entered a period of critical reflection on the practical ramifications of infrastructures developed to support open science. 
+For instance, Huggett (2022) identified social frictions as a necessary aspect of productive research, which open science infrastructures effectively seek to eliminate. 
+Numerous other studies by Batist (2023), Huvila (2022), Haciguzeller et al.
+(2021), Opitz et al.
+(2021) Sobotkova (2018), Wylie (2017) Atici et al.
+(2013), Faniel et al.
+(2013) (among others) highlighted the need to account for the work involved in preparing and making use of materials shared openly on the web, and which are effectively muted by infrastructures developed in pursuit of open science.
+
+Our network analysis similarly draws attention to the real-world collaborative ties that underpin archaeological open source software development.
+We identified three clusters of close associations between projects:
+the first is centred around a highly specialized supdiscipline exhibiting a very enthusiastic attitude (**ZB NOTE: or maybe even pressure?**) toward working in the open; 
+the second is concerned with developing solutions to a very common and discipline-wide problem, specifically relating to data collection and management in fieldwork; 
+and the third deals with several topic of concern, but represents an emerging community of practice. 
+This third cluster comprises efforts performed by members of a close-knit community, many of whom already collaborate at local research centres or participate in relatively private interest groups. 
+This indicates that open source is firmly embedded within existing social and institutional support and power structures that permeate academic life, both online and offline.

### 2023-09-19 analysis/_03-data_methodology.qmd
Get repo info from GH API
Incl. licenses

@@
   oarch <- mutate(oarch,
-                  gh_langs = map(oarch$gh_repo, gh_lang),
-                  gh_contribs = map(oarch$gh_repo, gh_contrib),
-                  gh_issues = map(oarch$gh_repo, gh_issue),
-                  gh_comments = map(oarch$gh_repo, gh_comment),
-                  gh_commits = map(oarch$gh_repo, gh_commit)
+                  gh_repo_info = map(gh_repo, gh_repo_info, .progress = TRUE),
+                  gh_langs = map(gh_repo, gh_lang, .progress = TRUE),
+                  gh_contribs = map(gh_repo, gh_contrib, .progress = TRUE),
+                  gh_issues = map(gh_repo, gh_issue, .progress = TRUE),
+                  gh_comments = map(gh_repo, gh_comment, .progress = TRUE),
+                  gh_commits = map(gh_repo, gh_commit, .progress = TRUE),
   )
@@
 **NOTE:** Add a breakdown on licensing practices.
+```{r tbl-repo-licenses}
+```
 

### 2023-09-19 R/gh.R

@@ -19,12 +19,50 @@
 #' @name gh_query
 #'
 #' @examples
+#' gh_repo("joeroe/era")
 #' gh_lang("joeroe/era")
 #' gh_contrib("joeroe/era")
 #' gh_issue("joeroe/era")
 #' gh_comment("joeroe/era")
 #' gh_commit("joeroe/era")
 NULL
+
+#' @rdname gh_query
+#' @export
+gh_repo_info <- function(repo) {
+  data <- tibble::tibble(
+    repo = repo,
+    response = list(ghq("GET /repos/{repo}", repo = repo))
+  )
+
+  if (not_all_na(data$response)) {
+    data <- tidyr::hoist(data, response,
+                         url = "html_url",
+                         "name",
+                         "description",
+                         "homepage",
+                         "fork",
+                         "created_at",
+                         "updated_at",
+                         "pushed_at",
+                         "size",
+                         "forks",
+                         stars = "stargazers_count",
+                         watchers = "watchers_count",
+                         "language",
+                         open_issues = "open_issues_count",
+                         "license",
+                         "topics")
+    data <- dplyr::select(data, -response)
+    data <- dplyr::mutate(data,
+                          created_at = lubridate::as_datetime(created_at),
+                          updated_at = lubridate::as_datetime(updated_at),
+                          pushed_at = lubridate::as_datetime(pushed_at),
+    )
+    data
+  }
+  else NA
+}
 
 #' @rdname gh_query
 #' @export

### 2023-09-19 analysis/_03-data_methodology.qmd
Add tbl-repo-licenses

@@
 
-**NOTE:** Add a breakdown on licensing practices.
 ```{r tbl-repo-licenses}
+#| tbl-cap: Summary of licenses used by projects with GitHub repositories.
+oarch |>
+  hoist(gh_repo_info, license = list("license", 1, "name")) |>
+  count(license) |>
+  mutate(license = replace_na(license, "None detected"))
+  arrange(-n) |>
+  gt() |>
+    cols_label(license = "License", n = "N")
 ```

### 2023-09-20 analysis/_03-data_methodology.qmd
Get GH API data in stages
In case you hit the rate limit in the middle.

@@
 if (!file_exists(cached_gh_data)) {
-  oarch <- mutate(oarch,
-                  gh_repo_info = map(gh_repo, gh_repo_info, .progress = TRUE),
-                  gh_langs = map(gh_repo, gh_lang, .progress = TRUE),
-                  gh_contribs = map(gh_repo, gh_contrib, .progress = TRUE),
-                  gh_issues = map(gh_repo, gh_issue, .progress = TRUE),
-                  gh_comments = map(gh_repo, gh_comment, .progress = TRUE),
-                  gh_commits = map(gh_repo, gh_commit, .progress = TRUE),
-  )
+  oarch$gh_repo_info <- map(oarch$gh_repo, gh_repo_info, .progress = TRUE)
+  oarch$gh_langs <- map(oarch$gh_repo, gh_lang, .progress = TRUE)
+  oarch$gh_contribs <- map(oarch$gh_repo, gh_contrib, .progress = TRUE)
+  oarch$gh_issues <- map(oarch$gh_repo, gh_issue, .progress = TRUE)
+  oarch$gh_comments <- map(oarch$gh_repo, gh_comment, .progress = TRUE)
+  oarch$gh_commits <- map(oarch$gh_repo, gh_commit, .progress = TRUE)
   saveRDS(oarch, cached_gh_data)

### 2023-09-20 analysis/_03-data_methodology.qmd
Fix open-archaeo to JOAD release
And rerender manuscript

@@
+---
+bibliography: references.bib
+---
+
 # Methods & Materials <!-- Methods & Materials -->
+
 <!-- TODO: snappier title? Maybe incoroporating 'open-archaeo'? -->
 
-Our study comprises an exploratory analysis of open-archaeo, a list of open source archaeological software and other digital resources.
+Our study comprises an exploratory analysis of open-archaeo [@batist2023], a list of open source archaeological software and other digital resources.
 As of right now it includes 408 items, which are rendered for the web at [open-archaeo.info](open-archaeo.info).
@@
 ```{r data-oarch}
-read_csv("https://raw.githubusercontent.com/zackbatist/open-archaeo/master/open-archaeo.csv") %>%
+read_csv("https://github.com/zackbatist/open-archaeo/raw/2023-08-29/open-archaeo.csv") %>%
   mutate(category = na_if(category, "?")) ->
@@
 For the purpose of this study, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc.
-from the API using the R package `gh` [@csardi2023].
+That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@csardi2023].
 This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but amongst those that do, GitHub is by far the most popular host ([@tbl-repo-hosts]).
@@
   count(license) |>
-  mutate(license = replace_na(license, "None detected"))
+  mutate(license = replace_na(license, "None detected")) |>
   arrange(-n) |>

### 2023-09-20 analysis/_04-open_archaeology.qmd
Rearrange sections a bit

--- 2023-06-14 analysis/_04-what_is_open_archaeology.qmd
@@
-# Open archaeology <!-- Results I (description of what's there) -->
+# Open archaeology <!-- Results I: description of open archaeology projects -->
 
@@
 ```
+
+## History
+
+Archaeological software development activity has increased significantly over the years.
+The following charts portray the cumulative growth of code contributions committed and pushed to GitHub repositories, and the number of GitHub repositories that host archaeological software and resources.
+
+```{r plot-gh-cum}
+oarch %>% 
+  mutate(
+    lumped_category = recode(category,
+                             "Specifications etc." = "Documents",
+                             "Products" = "Documents",
+                             "Guides" = "Documents",
+                             "Lists and datasets" = "Documents",
+                             .default = category),
+    lumped_category = factor(lumped_category, c("Scripts",
+                                                "Documents",
+                                                "Packages and libraries",
+                                                "Standalone software")) ) ->
+  oarch_lumped
+
+p_gh_commits_cum <- oarch_lumped %>%
+  unnest(gh_commits) %>%
+  drop_na(sha) %>%
+  mutate(date = floor_date(datetime, "week")) %>%
+  group_by(lumped_category, date) %>%
+  summarise(n_commits = n(), .groups = "drop_last") %>%
+  mutate(cum_commits = cumsum(n_commits)) %>%
+  ggplot(aes(date, cum_commits,
+             fill = lumped_category,
+             colour = lumped_category,
+             label = lumped_category)) +
+  geom_area() +
+  annotate("segment", 
+           x = ymd("2008-04-01", tz = "UTC"), 
+           xend = ymd("2008-04-01", tz = "UTC"),
+           y = 16000, yend = 4000,
+           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
+  annotate("text", 
+           x = ymd("2008-04-01", tz = "UTC"), y = 19000,
+           label = "GitHub\nlaunched",
+           vjust = 0) +
+  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
+  scale_y_continuous(labels = scales::comma) +
+  scale_colour_bright(guide = guide_none()) +
+  scale_fill_bright() +
+  labs(
+    y = "cumulative commits",
+    x = NULL, fill = NULL
+  ) +
+  theme_oarch() +
+  theme(legend.position = "top")
+
+# TODO: extend to end
+p_gh_repos_cum <- oarch_lumped %>%
+  unnest(gh_commits) %>%
+  drop_na(sha) %>%
+  group_by(repo) %>% 
+  filter(datetime == min(datetime)) %>% # Only first commits
+  mutate(date = floor_date(datetime, "week")) %>%
+  group_by(lumped_category, date) %>%
+  summarise(n_commits = n(), .groups = "drop_last") %>%
+  mutate(cat_total = sum(n_commits)) %>%
+  mutate(cum_commits = cumsum(n_commits)) %>%
+  ggplot(aes(date, cum_commits,
+             fill = lumped_category,
+             colour = lumped_category)) +
+  geom_area() +
+  annotate("segment",
+           x = ymd("2008-04-01", tz = "UTC"),
+           xend = ymd("2008-04-01", tz = "UTC"),
+           y = 40, yend = 10,
+           arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
+  annotate("text",
+           x = ymd("2008-04-01", tz = "UTC"), y = 50,
+           label = "GitHub\nlaunched",
+           vjust = 0) +
+  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
+  scale_y_continuous(labels = scales::comma) +
+  scale_colour_bright(guide = guide_none()) +
+  scale_fill_bright(guide = guide_none()) +
+  labs(
+    y = "cumulative repositories",
+    x = NULL, fill = NULL
+  ) +
+  theme_oarch()
+
+p_gh_commits_cum / p_gh_repos_cum
+```
+
+As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
+But use of git really began to take off around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
+
+Around this time we also see that GitHub starts being used to host documents and scripts.
+This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work in an open and somewhat nerdy way.
+
+---
+
+- Compare changing proportions of each category on a year by year basis
+- Identify temporal trends in the use of licenses

### 2023-09-20 analysis/_07-conclusion.qmd
Rearrange sections a bit

--- 2023-09-05 analysis/_99-conclusion.qmd
@@
 
-In this paper we take stock of archaeologists' immense creative capabilities. 
-First, we surveyed what kinds of software archaeologists are making and tracked progression over time. 
-We showed that open source software development in archaeology has seen a rapid and sustained rise beginning around 2014. 
+In this study we have sought to document an emerging community of open source software engineers in archaeology and to assess the extent to which collaboration within this community resembles conventional 'open source' forms.
+We have operationalised 'open-source collaborative practices' as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
+This excludes a wide range of other potential practices, most obviously the use of other software (though as we demonstrate in section X.X, the git/GitHub combination is overwhelmingly the most popular), but also collaboration through offline or private channels, or forms of collaboration we don't know about.
+We have not directly interviewed those doing the collaborating, though our conclusions do drawn heavily from our experience as members of that community ourselves. 
+Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it (@cite).
+
+These caveats notwithstanding, we have documented that open source software development in archaeology has seen a rapid and sustained rise beginning around 2014. 
 This is marked by a variety of applications and use cases, including the use of git and GitHub to track and host content other than code. 

### 2023-09-20 analysis/openarchaeo_collaboration.qmd
Rearrange sections a bit

@@
 {{< include _03-data_methodology.qmd >}}
-{{< include _04-what_is_open_archaeology.qmd >}}
-{{< include _05-history_of_open_archaeology.qmd >}}
-{{< include _06-project_activity.qmd >}}
-{{< include _07-network.qmd >}}
-{{< include _99-conclusion.qmd >}}
+{{< include _04-open_archaeology.qmd >}}
+{{< include _05-collaboration.qmd >}}
+{{< include _06-network.qmd >}}
+{{< include _07-conclusion.qmd >}}

### 2023-09-20 analysis/_06-network.qmd
Rearrange sections a bit
Add tbl-repo-centrality

--- 2023-06-14 analysis/_07-network.qmd
@@
+# An emerging community of practice? <!-- Results III: network analysis of collaborative communities -->
+
 We constructed and analyzed networks representing users' contributions to GitHub repositories to examine the emergence of collaborative tendencies.
@@
 
-**NOTE:** Create a table containing the nodes with highest betweenness centrality from the repo-repo graph.
-Columns should specify the name of the repository, the number of commits, the category, and tags.
-If possible also include a summary of the pulse as a line chart with x axis as time and y axis as activity.
-This will allow us to identify the characteristics of these repositories that prompt greater integration.
+```{r tbl-repo-centrality}
+#| tbl-cap: The sixteen most central repositories according to the repository–repository network analysis
+oarch_graph_repo_repo |>
+  mutate(centrality = centrality_betweenness()) |>
+  as_tibble() |>
+  arrange(-centrality) |>
+  slice(1:16) |>
+  left_join(oarch, c("repo" = "gh_repo")) |>
+  mutate(n_commits = map_int(gh_commits, nrow)) |>
+  unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
+  select(repo, centrality, category, tags, n_commits) |>
+  gt() |>
+  cols_label(
+    repo = "Repository", 
+    centrality = "Betweenness Centrality",
+    category = "Category",
+    tags = "Tags",
+    n_commits = "Commits"
+  )
+```
 

### 2023-09-20 analysis/_03-data_methodology.qmd
Note TODO
Expand network analysis

@@
 We supplemented this quasi-systematic collection strategy with word-of-mouth contributions made by interested individuals who identified relevant work that we initially overlooked.
+
+<!-- TODO:
+* Introduce git, GitHub, etc. and describe briefly how they work for non-users
+* Define software jargon we use (repo, commit, etc.) – maybe as a glossary table?
+-->
 
@@
 The data and analysis code are available in our research compendium on GitHub, which is also indexed on Zenodo.
+
+```{r data-oarch_contributors}
+oarch |>
+  unnest(gh_contribs) |>
+  drop_na(contributor) |>
+  group_by(contributor) |>
+  mutate(
+    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
+  ) |>
+  group_by(contributor, repo_owner) |>
+  summarise(repos = n(), contribs = sum(contributions)) |>
+  pivot_wider(c(contributor), names_from = repo_owner, 
+              values_from = c(repos, contribs), values_fill = 0) |>
+  mutate(
+    repos = repos_other + repos_own,
+    contribs = contribs_other + contribs_own,
+    .after = contributor
+  ) ->
+  oarch_contributors
+```

### 2023-09-20 analysis/_06-network.qmd
Expand network analysis

--- 2023-06-14 analysis/_07-network.qmd
@@
 
-{r graph-data}
+```{r data-graph}
+# TODO: refactor some of this as functions?
+oarch |>
+  drop_na(gh_repo) |>
+  unite(tags, c(tag1, tag2, tag3, tag4, tag5)) |>
+  select(gh_repo, category, tags, gh_commits, gh_issues, gh_comments) ->
+  oarch_graph
+
+oarch_graph <- bind_rows(
+  oarch_graph |>
+    unnest(gh_commits) |>
+    drop_na(author) |> # Apparently real commits, but we can't connect them
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "commit", n_contrib = n(), .groups = "drop"),
+  oarch_graph |>
+    unnest(gh_comments) |>
+    drop_na(author) |> # Repos with no comments
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "comment", n_contrib = n(), .groups = "drop"),
+  oarch_graph |>
+    unnest(gh_issues) |>
+    drop_na(author) |> # Repos with no issues
+    group_by(repo, author) |>
+    summarise(category = first(category), tags = first(tags),
+              contrib = "issue", n_contrib = n(), .groups = "drop")
+)
+
+# Cleaning
+oarch_graph |>
+  rename(user = author) |>
+  # Remove non-human users
+  filter(!str_detect(user, coll("[bot]"))) |>
+  filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
+                      "gitter-badger", "dependabot-support")) |>
+  # Remove big, non-archaeology projects that flood the network
+  filter(!repo %in% c("Stellarium/stellarium", 
+                     "opengisch/QField",
+                     "lutraconsulting/input")) ->
+  oarch_graph
+```
 
@@
 
-{r plot-graph-repo-user}
+```{r extract-graph-repo-user, eval=FALSE}
+oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
+                                   edges_user_to_repo(oarch_graph))
+```
+
+```{r fig-graph-repo-user, eval=FALSE}
+g_repo_user <- oarch_graph_repo_user |>
+  ggraph(layout = "graphopt") +
+  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
+  geom_node_point(aes(colour = node_type)) +
+  scale_edge_alpha(trans = "log")
+
+#g_repo_user
+girafe(g_repo_user)
+```
 
@@
 
-{r plot-graph-repo-repo}
-{r graph-user-user}
-{r plot-graph-user-user}
-
 ## repo-repo
@@
 
-{r plot-graph-repo-repo-edge-betweenness}
-{r plot-graph-repo-repo-edge-betweenness-dendrogram}
+<!-- TODO: Note that we remove disconnected repos? (and how many?) -->
+
+```{r extract-graph-repo-repo}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
+                                   edges_common_users(oarch_graph),
+                                   directed = FALSE)
+
+oarch_graph_repo_repo |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 1) |>
+  mutate(group = as.character(group_edge_betweenness())) ->
+  oarch_graph_repo_repo
+```
+
+```{r fig-graph-repo-repo, eval=FALSE}
+oarch_graph_repo_repo |>
+  ggraph(layout = "graphopt") +
+  geom_edge_link(aes(alpha = n)) +
+  geom_node_point(aes(colour = category)) +
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common contributors")
+```
+
+```{r fig-graph-repo-repo, eval=FALSE}
+iplot <- oarch_graph_repo_repo |>
+  ggraph(layout = "graphopt") +
+  geom_edge_link(aes(alpha = n)) +
+  #geom_node_point(aes(colour = group)) +
+  #geom_node_text(aes(label = group)) +
+  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
+  scale_edge_alpha(trans = "log") +
+  labs(edge_alpha = "Common contributors")
+
+girafe(ggobj = iplot)
+```
+
+```{r fig-graph-repo-repo-edge-betweenness-dendrogram, eval=FALSE}
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
+                                   edges_common_users(oarch_graph),
+                                   directed = TRUE)
+
+oarch_graph_repo_repo |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 1) |>
+  mutate(group = as.character(group_edge_betweenness())) ->
+  oarch_graph_repo_repo
+
+xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
+xx <- as.dendrogram(xx)
+plot(xx)
+
+ggraph(xx, 'dendrogram', height = height)
+```
 
@@
 
-**NOTE:** Create a table containing the nodes with highest betweenness centrality from the repo-repo graph.
-Columns should specify the name of the repository, the number of commits, the category, and tags.
-If possible also include a summary of the pulse as a line chart with x axis as time and y axis as activity.
-This will allow us to identify the characteristics of these repositories that prompt greater integration.
+```{r tbl-repo-centrality}
+#| tbl-cap: Most central repositories in the repository–repository network
+oarch_graph_repo_repo |>
+  mutate(centrality = centrality_betweenness()) |>
+  as_tibble() |>
+  arrange(-centrality) |>
+  slice(1:10) |>
+  left_join(oarch, c("repo" = "gh_repo")) |>
+  mutate(n_commits = map_int(gh_commits, nrow)) |>
+  unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
+  select(repo, centrality, category, tags, n_commits) |>
+  gt() |>
+  cols_label(
+    repo = "Repository", 
+    centrality = "Betweenness Centrality",
+    category = "Category",
+    tags = "Tags",
+    n_commits = "Commits"
+  )
+```
+
+<!-- TODO: What does tbl-repo-centrality tell us? Can we interrogate it further with a graph? -->
 
@@
 
-**NOTE:** Plot a graph that renders nodes' sizes based on their betweenness centrality.
+<!-- TODO: Note that we remove disconnected users? (and how many?) -->
 
-{r summary-graph-user-user}
+```{r graph-user-user}
+oarch_graph_user_user <- tbl_graph(
+  nodes = nodes_users(oarch_graph), 
+  edges = edges_common_repos(oarch_graph),
+  directed = FALSE
+)
+
+oarch_graph_user_user |>
+  mutate(degree = centrality_degree()) |>
+  filter(degree > 1) |>
+  mutate(
+    centrality = centrality_betweenness(weights = n),
+    group = as.character(group_edge_betweenness())
+  ) ->
+  oarch_graph_user_user
+```
+
+<!--
+ NOTE, ZACK: the graph below is what you requested (nodes' sizes based on their betweenness centrality), but I'm unsure we will ever be able to get these graph plots legible enough to actually include.
+– JR -->
+
+```{r fig-graph-user-user-centrality, eval=TRUE}
+oarch_graph_user_user |> 
+  ggraph(layout = "graphopt") +
+  # geom_edge_link(aes(alpha = n)) +
+  geom_edge_arc(alpha = 0.05) +
+  geom_node_point(aes(size = centrality), 
+                  shape = 21, colour = "#000000", fill = "#ffffff") +
+  # scale_edge_alpha(trans = "log") +
+  # labs(edge_alpha = "Common repositories") +
+  theme_graph()
+```
 
@@
 **NOTE:** Create a table containing the top 10 users based on betweenness centrality values, with columns specifying theur username, real name, job title/affiliation, location, career stage (student, early/mid/late career, tenured/non-tenured, independent, etc).
+
+<!-- JR: I'm not sure about this.
+Either practically (can we reliably determine all this from public info?) or ethically (not necessarily that it is unethical; but we haven't really considered the implications of analysing named individuals in this study) -->
+
+```{r tbl-user-centrality}
+#| tbl-cap: Most central users in the user–user network
+oarch_graph_user_user |>
+  mutate(centrality = centrality_betweenness(weights = n)) |>
+  as_tibble() |> 
+  arrange(-centrality) |>
+  slice(1:10) |>
+  left_join(oarch_contributors, by = c("user" = "contributor")) |>
+  select(user, centrality, repos, contribs_own, contribs_other) |>
+  gt() |>
+  cols_label(
+    user = "User",
+    centrality = "Centrality",
+    contribs_own = "Own repositories",
+    contribs_other = "Others' repositories"
+  ) |>
+  tab_spanner("Contributions", c(contribs_own, contribs_other))
+```
 
@@
 
-**NOTE:** I have no idea what these next two charts depict.
+```{r fig-user-betweenness, eval=FALSE}
+#| fig-cap: Distribution of betweenness centrality values in the user–user network
+oarch_graph_user_user |>
+  mutate(
+    betweenness = centrality_betweenness(weights = n)
+  ) |>
+  data.frame() |>
+  ggplot(aes(x = betweenness)) +
+  geom_density(bw = "nrd") +
+  scale_x_log10(labels = scales::number)
+```
 
-{r plot-user-betweenness-dist}
-{r plot-user-betweenness-activity}
+```{r fig-user-betweenness-activity, eval=FALSE}
+#| fig-cap: Correlation between user's centrality (betweenness) in the network and their total number of contributions
+oarch_graph_user_user |>
+  mutate(
+    degree = centrality_degree(), 
+    betweenness = centrality_betweenness(weights = n), 
+    strength = centrality_degree(weights = n)
+  ) |>
+  as_tibble() |> 
+  left_join(oarch_contributors, by = c("user" = "contributor")) |>
+  ggplot(aes(betweenness, contribs)) +
+  geom_point() +
+  scale_x_log10() +
+  scale_y_log10() +
+  geom_smooth(method = "lm")
+```

### 2023-09-20 analysis/_06-network.qmd
Fix errors and rerender manuscript

@@
     group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "commit", n_contrib = n(), .groups = "drop"),
+    summarise(contrib = "commit", n_contrib = n(), .groups = "drop"),
   oarch_graph |>
@@
     group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "comment", n_contrib = n(), .groups = "drop"),
+    summarise(contrib = "comment", n_contrib = n(), .groups = "drop"),
   oarch_graph |>
@@
     group_by(repo, author) |>
-    summarise(category = first(category), tags = first(tags),
-              contrib = "issue", n_contrib = n(), .groups = "drop")
+    summarise(contrib = "issue", n_contrib = n(), .groups = "drop")
 )
@@
 ```{r extract-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph,category,tags),
+oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
                                    edges_common_users(oarch_graph),
@@
 
-```{r fig-graph-repo-repo, eval=FALSE}
+```{r fig-graph-repo-repo-interactive, eval=FALSE}
 iplot <- oarch_graph_repo_repo |>
@@
   filter(degree > 1) |>
-  mutate(
-    centrality = centrality_betweenness(weights = n),
-    group = as.character(group_edge_betweenness())
-  ) ->
+  mutate(centrality = centrality_betweenness(weights = n)) ->
   oarch_graph_user_user
@@
 
-```{r fig-graph-user-user-centrality, eval=TRUE}
+```{r fig-graph-user-user-centrality, eval=FALSE}
 oarch_graph_user_user |> 

### 2023-09-20 analysis/_01-introduction.qmd
Comment on introduction text

@@
 
-The relationship between open science and open source is also complicated by rhetorical claims that, based on our personal experiences, are discordant with how academics actually _do_ open source.
+The relationship between open science and open source is also complicated by rhetorical claims that have a questionable connection to how academics actually _do_ open source. 
 Does academic open source actually make research processes more transparent and improve research outcomes? 
@@
 Is it actually helping to foster new globe-spanning connections and lead to novel research trajectories that would not otherwise come to pass?
+Basically, is it more than just uploading code and data files to the internet?
+
 Despite hopeful aspirations espoused by open archaeology advocates [see @kansa2014; @kintigh2015], these outcomes are not a given.
-We believe that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
+We believe<!-- Bit weak; can we cite some previous work that indicates this?
+--JR---> that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
 However, practical circumstances and systemic value regimes that frame what it means to work as an archaeologist presently inhibit the potential for radical transformation, even among open science's most ardent supporters.
 
-That being said, archaeologists are prolific software developers.
+<!-- Might be getting a bit ahead of ourselves here.
+--JR -->That being said, archaeologists are prolific software developers.
 This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes XXX records as of the time of publication.

### 2023-09-20 analysis/_04-open_archaeology.qmd
Merge outline into text

@@
-# Open archaeology <!-- Results I: description of open archaeology projects -->
+# Open archaeology
+
+<!-- Results I: description of open archaeology projects
+
+- What kind of projects comprise open archaeology?
+  - Breakdown of tags and categories
+  - Breakdown of platforms and languages
+  - Tags by category
+  - Licenses?
+- How has this changed over time?
+  - Growth of open archaeology by number of commits
+  - Growth of open archaeology by number of repos
+  - Compare changing proportions of each category on a year by year basis
+  - Identify periods associated with different rates of growth, and relate these with different general attitudes in the history of digital archaeology
+-->
 

### 2023-09-20 analysis/_05-collaboration.qmd
Rearrange sections a bit
Merge outline into text

--- 2023-06-14 analysis/_06-project_activity.qmd
@@
+# Collaborative practices
+
+<!-- Results II: description of observed collaborative practices
+
+- How active are projects?
+- How many people are involved?
+- Conversely, how many projects are individual people involved in?
+- What is the level of uptake of GitHub's collaboration features?
+- Has this changed over time?
+
+Figures:
+
+- Charts presenting commit rates and repo longevity
+- Charts presenting the number of contributors and the amount of work that they each contributed
+- Charts presenting the use of issues and comments
+
+-->
+
 But how is this activity distributed across repositories? In this section we assess projects' pulse, or the amount of activity pertaining to a repository.

### 2023-09-20 analysis/_06-network.qmd
Merge outline into text

@@
-# An emerging community of practice? <!-- Results III: network analysis of collaborative communities -->
+# An emerging community of practice?
+
+<!-- Results III: network analysis of collaborative communities
+
+* Are there significant networks? If so, how big and how many?
+  * Only one, it seems!
+
+### [Repo-repo] What forms this network?
+
+* Identify the repos/organisations that are the focus of the most intense collaboration?
+  * Are they thematically or topically linked?
+  * Do they map onto real-world research centres?
+* Is the network a new thing, or just a translation of existing offline networks (e.g. fields, institutions) onto GitHub?
+  * cf. is open source really open?
+
+### [User-user] Who forms this network?
+
+* What proportion of users and repos are connected to this network?
+* What predicts membership in this network?
+* Does centrality in the network correlate with general rate of activity?
+*  Who are the most central nodes in this network? Is there anything they have in common?
+   * High activity?
+   * Certain platforms/categories/languages?
+* Qualitative analysis of top 50? 100? most connected users
+  * Career stage? Precarity? Employer? Location?
+
+-->
 

### 2023-09-20 analysis/_07-conclusion.qmd
Merge outline into text

@@
 # Conclusion
+
+<!--
+
+There is an emerging community of practice around open source in archaeology.
+
+* It is limited and fragmented
+  * A small set of subfields are represented
+  * Probably most people still haven't heard it?
+* Collaboration is limited
+  * Most work remains solo & short-lived
+  * Perhaps because open source norms conflict with academic norms?
+* The members of this community face structural challenges in academia
+-->
+
 Archaeology is a "magpie" discipline, using tools and methods that originate from a variety of other disciplines and industries to better understand the past. 

### 2023-09-20 analysis/_03-data_methodology.qmd
Merge outline into text
Revise first results section
Trying to be a bit more focused on the results rather than questions and definitions.

@@
----
-bibliography: references.bib
----
+# Data and methodology
 
-# Methods & Materials <!-- Methods & Materials -->
+<!--
+* open-archaeo
+* GitHub API
+* Indicate how much is on GitHub or elsewhere
+  -->
 
@@
 While our initial intention was to only list open source software, open-archaeo's scope has expanded to include all software created by and for archaeologists.
-We found that many repositories do not explicitly denote an open source license, or any license at all.
-
-```{r tbl-repo-licenses}
-#| tbl-cap: Summary of licenses used by projects with GitHub repositories.
-oarch |>
-  hoist(gh_repo_info, license = list("license", 1, "name")) |>
-  count(license) |>
-  mutate(license = replace_na(license, "None detected")) |>
-  arrange(-n) |>
-  gt() |>
-    cols_label(license = "License", n = "N")
-```
 

### 2023-09-20 analysis/_04-open_archaeology.qmd
Revise first results section
Trying to be a bit more focused on the results rather than questions and definitions.

--- 2023-09-20 analysis/_04-open_archaeology.qmd
@@
-# Open archaeology <!-- Results I: description of open archaeology projects -->
+---
+bibliography: references.bib
+---
 
-Our analysis documented an the widespread presence of open archaeology projects on GitHub and other online platforms.
+# Open archaeology
 
-We annotated each record with tags that describe aspects of archaeological work that each tool contributes to.
-We assigned tags based on how their maintainers identified their projects.
-The list of tags and their associated definitions is subject to grow or change as archaeologists continue to develop software for additional use cases.
-Right now there are 63 tags.
+```{=html}
+<!-- Results I: description of open archaeology projects
 
-We also categorized records based on how each tool or resource is meant to be accessed or used.
-This gives us a sense of the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-In other words, categories hint at some of the decisions that developers make regarding how they want their tools to be used and accessed.
+- What kind of projects comprise open archaeology?
+  x Breakdown of tags and categories
+  x Breakdown of platforms and languages
+  x Tags by category
+  x Licenses?
+- How has this changed over time?
+  - Growth of open archaeology by number of commits
+  - Growth of open archaeology by number of repos
+  - Compare changing proportions of each category on a year by year basis
+  - Identify periods associated with different rates of growth, and relate these with different general attitudes in the history of digital archaeology
+-->
+```
+As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists, primarily software but also various forms of open document.
+@tbl-categories summarises the categories included.
 
-| Category                              | Kind      | Scope                                                                                                                                                                                                        |
-|:-----------|:-----------|:------------------------------------------------|
-| Packages and libraries                | Software  | Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-|
-| Standalone software                   | Software  | Software that may be operated without needing to first access an underlying platform.
-|
-| Scripts                               | Software  | Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
-|
-| Specifications, protocols and schemas | Documents | A formal data structure or framework intended to be used as a model.
-|
-| Products                              | Documents | Stable outcomes of creative work.
-|
-| Guides                                | Documents | An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.
-|
-| Lists and datasets                    | Documents | A series of consistently organized observations assembled with purpose.
-|
-:Categorisation scheme used by open-archaeo {#tbl-category-defs}
+```{r tbl-categories}
+#| tbl-cap: Categories of open archaeology projects included in open-archaeo
+tribble(
+  ~category,                               ~kind,       ~scope,
+  "Packages and libraries",                "Software", "Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.",
+  "Standalone software",                   "Software", "Software that may be operated without needing to first access an underlying platform.",
+  "Scripts",                               "Software", "Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.",
+  "Specifications, protocols and schemas", "Documents", "A formal data structure or framework intended to be used as a model.",
+  "Products",                              "Documents", "Stable outcomes of creative work.",
+  "Guides",                                "Documents", "An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.",
+  "Lists and datasets",                    "Documents", "A series of consistently organized observations assembled with purpose."
+) |>
+  left_join(oarch, by = "category") |>
+  count(category, kind, scope) |>
+  group_by(kind) |>
+  arrange(-n) |>
+  gt() |>
+  cols_label_with(fn = tools::toTitleCase)
+```
 
-We defined two kinds of categories: software and documents.
-Software comprise collections of instructions that dictate how a computer should operate, and documents are bounded sets of recorded information.
-We define three categories of software and four categories of documents.
-Packages and libraries are sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.
-Platforms are the broader systems upon which packages or libraries are plugged into or are meant to extend upon, such as R, Python, QGIS, and Meshlab, which each have their own protocols to facilitate such integration.
-Standalone software are tools that users may operate without needing to first access an underlying platform.
-The third category of software, scripts, are sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.
+```{r num-p_platform}
+p_platform <- sum(!is.na(oarch$platform)) / nrow(oarch)
+```
 
-The first kind of document, which we refer to as specifications, protocols and schemas, are formal data structures or frameworks intended to be used as models.
-Products are stable outcomes of creative work, loosely defined.
-Guides are educational resources or documented protocols meant to instruct readers how to apply relevant tools or techniques.
-And finally, lists or datasets, including open-archaeo itself, are series of consistently organized observations assembled with purpose.
-
-## Platforms and languages
-
-Our first set of analyses focused on these questions:
-
-* What kinds of software are archaeologists making?
-* What are the general purposes of their software?
-
-Most projects included in open-archaeo are designed to be used as a package or library atop am existing "platform" -- for example a package for a programming language like R, or a plugin for an application like QGIS ([@fig-cats-plats]).
+Most projects (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language or a plugin for an application.
 The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
 Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-A significant amount of projects also comprise of datasets and non-packaged code snippets that have been made available for general use .
+A significant number of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
 
-```{r fig-cats-plats}
-#| fig-cap: Number of open archaeology projects by category and platform
-#| fig-subcap: 
-#|   - Categories
-#|   - Platforms
-
-oarch |>
-  drop_na(category) |>
-  group_by(category) |>
-  summarise(n = n()) |>
-  ggplot(aes(fct_reorder(category, n), n)) +
-  geom_col() +
-  geom_text(aes(label = category), hjust = 0, nudge_y = 1) +
-  coord_flip() +
-  scale_y_continuous(limits = c(0, 300)) +
-  labs(x = NULL, y = NULL) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(), 
-    axis.ticks.y = element_blank(),
-    panel.grid.major.y = element_blank()
-  )
-
+```{r tbl-platforms}
+#| tbl-cap: Platforms and programming languages used by open archaeology projects
 oarch |>
@@
   count(platform) |>
-  arrange(n) |>
-  ggplot(aes(fct_reorder(platform, n), n)) +
-  geom_col(colour = "#333333") +
-  geom_text(aes(label = platform), hjust = 0, nudge_y = 1) +
-  coord_flip() +
-  scale_y_continuous(limits = c(0, 200)) +
-  labs(x = NULL, y = NULL) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_blank(), 
-    axis.ticks.y = element_blank(),
-    panel.grid.major.y = element_blank()
-  )
+  mutate(
+    significant = n > 1,
+    platform = if_else(significant, platform, "Other")
+  ) |>
+  group_by(platform) |>
+  summarise(n = sum(n)) |>
+  arrange(platform == "Other", -n) |>
+  gt() |>
+  cols_label_with(fn = tools::toTitleCase)
 ```
 
-R is overwhelmingly dominant as a platform, closely followed by Python [@fig-cats-plats].
-Plugins for QGIS and mobile apps are also relatively popular.
+```{r num-p_platform_r}
+p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / nrow(oarch)
+```
+
+The statistical programming language R is overwhelmingly the most common platform, representing `r percent(p_platform_r)` of projects in open-archaeo.
+Python, another programming language, is also relatively popular, as are plugins for the open source geographic information system QGIS.
 Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
 AutoCAD, ArcGIS), a number of lesser used programming languages, and a genre consisting of custom forms and spreadsheet templates.
-In the figure below, the darker grey lines are used to break up the bars by primary author, showing that many of the smaller platforms---for example Excel templates---are targeted by only one or two developers.
-The larger platforms tend to be more diverse.
+Many of these are targeted by only one or two developers; the larger platforms tend to be more diverse.
 
-```{r plot-platforms}
-```
+At first glance, the relative popularity of R versus Python is perhaps surprising; Python is regularly ranked as the most popular programming language in the world, with R a distant runner-up.
+However, it accords with the popularity of R as a tool for data analysis in archaeology [@schmidt2020] and other scientific disciplines [@lai2019].
 
+We also annotated each record with 'tags' that describe aspects of archaeological work that each tool contributes to [@fig-tags].
 When we compare categories with tags, we see the general domains that each kind of resource is designed to serve.
@@
 
-```{r plot-tags-cats}
+<!-- TODO: can we say anything more concrete about the tags? -->
+
+```{r fig-tags}
 detail_tags <- c("Instrumental Neutron activation analysis",
@@
 
-## History
+<!-- TODO: expand. -->
+We found that many repositories do not explicitly denote an open source license, or any license at all [@tbl-repo-licenses].
+
+```{r tbl-repo-licenses}
+#| tbl-cap: Summary of licenses used by projects with GitHub repositories.
+oarch |>
+  hoist(gh_repo_info, license = list("license", 1, "name")) |>
+  count(license) |>
+  mutate(license = replace_na(license, "None detected")) |>
+  arrange(-n) |>
+  gt() |>
+    cols_label(license = "License", n = "N")
+```
 
 Archaeological software development activity has increased significantly over the years.
-The following charts portray the cumulative growth of code contributions committed and pushed to GitHub repositories, and the number of GitHub repositories that host archaeological software and resources.
+@fig-github-cumulative shows the cumulative growth of code contributions committed and pushed to GitHub repositories, and the number of GitHub repositories that host archaeological software and resources.
 
-```{r plot-gh-cum}
+```{r fig-github-cumulative}
 oarch %>% 

### 2023-09-20 analysis/openarchaeo_collaboration.qmd
Merge bibliographies
Revise first results section
Trying to be a bit more focused on the results rather than questions and definitions.

@@
 By analyzing interactions among archaeologists using the GitHub platform, including the ways they contribute to collective code bases, we demonstrate how archaeological software engineering is beginning to foster new kinds of collaborative commitments that blend open source protocols, DIY work ethic, conventional forms of scholarly communication, and cultural practices unique to archaeology.
-bibliography: openarchaeo-collaboration.bib
+bibliography: references.bib
 format:
@@
 library("patchwork")
+library("scales")
 library("tidygraph")

### 2023-09-20 analysis/_01-introduction.qmd
Merge outline into text
Modifications and comments.

@@
 # Introduction
+
+<!--
+* There is open archaeology
+* Open archaeology is related to open source
+* Is open archaeology open source, and is open source open?
+  * [Is 'open archaeology' just uploading CSVs?]
+-->
 
@@
 Despite this, the open science movement, within which open archaeology emerged, draws direct inspiration from open source.
-For instance, the Open Knowledge Foundation [-@openknowledgefoundation2023] publishes a widely accepted defitnion of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group.
-The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plaintext, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic).
-However, academic work, including the development of academic software, differs signififcantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL, and the Firefox web browser.
+For instance, the Open Knowledge Foundation [-@openknowledgefoundation2023] publishes a widely accepted definition of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group.
+The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plain text, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic).
+However, academic work, including the development of academic software, differs significantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL, and the Firefox web browser.
 While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests.

### 2023-09-20 analysis/_02-background.qmd
Merge outline into text
Modifications and comments.

@@
-# Open science and open source <!-- Background -->
+# Open science and open source
+
+<!--
+* What is open source?
+* Summary of key publications in open archaeology / DIY archaeology
+* But is any of this real? There seems to be a lack of reflection about what open archaeology is.
+  -->
 
@@
 This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion.
-Personal ego is minimized in favour of co-creating stable and functional outcomes [@coleman2012; @oneil2009 on the debian saga].
+Personal ego is minimized in favour of co-creating stable and functional outcomes [@coleman2012; @oneil2009 on the Debian saga].
 
@@
 
-In other words, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are grossly underappreciated factors that contribute to an open source project's success.
+In other words, as with many so-called "soft skills" that are crucial for academic professional development, additional competencies relating to the maintenance, management, and distribution of software, such as the ability to receive and implement feedback, set and stick with long-term goals, coordinate labour, document work practices, and collaborate with others, are grossly under-appreciated factors that contribute to an open source project's success.
 
@@
 This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement.
-However, most open science advocates (including archaeologists, as elicited by @beck2012 and @marwick2017) claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scienctific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.
+However, most open science advocates (including archaeologists, as elicited by @beck2012 and @marwick2017) claim that they are motivated by a desire to facilitate novel research opportunities, make participation in scientific research more equitable, reclaim science as a public good, and enhance how findings are validated and legitimized.
 
-The idea that scientistis should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production.
+The idea that scientists should generally contribute to a public domain of knowledge without profit motive has led to open science being heralded as revolutionary, community-oriented, and anti-capitalist means of production.
 However, while open science does have the _potential_ to effect radical change, this is not a given.
@@
 Whereas open source emerged from concern of consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible.
-Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (inlcuding: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.
+Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.
 Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience.
@@
 
-The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a genuine desire to facilitate novel research opportunities, to make participation in scienctific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings.
+The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a genuine desire to facilitate novel research opportunities, to make participation in scientific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings.
 However, the predominant concern with implementing best tools to use, adopting optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards [cf.

### 2023-09-20 analysis/_03-data_methodology.qmd
Modifications and comments.

@@
 
-<!-- TODO: snappier title? Maybe incoroporating 'open-archaeo'? -->
+<!-- TODO: snappier title? Maybe incorporating 'open-archaeo'? -->
 
@@
 That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@csardi2023].
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but amongst those that do, GitHub is by far the most popular host ([@tbl-repo-hosts]).
+This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but among those that do, GitHub is by far the most popular host ([@tbl-repo-hosts]).
 

### 2023-09-20 analysis/_04-open_archaeology.qmd
Modifications and comments.

@@
 As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists, primarily software but also various forms of open document.
-@tbl-categories summarises the categories included.
+@tbl-categories summarizes the categories included.
 
@@
 
-<!-- TODO: can we say anything more concrete about the tags? -->
+<!-- TODO: can we say anything more concrete about the tags? 
+ZB response: We can look at the development of tags over time.
+For instance, a chart documenting year-over-year growth of each tag based on date of each project's first commit.
+I imagine this as a stacked bar chart (similar to below) but with segments coded to represent year of first commit.
+This would fit at the end of this section, under fog-github-cumulative.
+We can add charts documenting growth of platforms and licenses too.
+-->
 

### 2023-09-20 analysis/_05-collaboration.qmd
Modifications and comments.

@@
 the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.
-But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterised by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
+But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterized by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
 This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.

### 2023-09-20 analysis/_06-network.qmd
Modifications and comments.

@@
 
-```{r extract-graph-repo-user, eval=FALSE}
+```{r extract-graph-repo-user, eval=TRUE}
 oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
@@
 
-```{r fig-graph-repo-user, eval=FALSE}
+```{r fig-graph-repo-user, eval=TRUE}
 g_repo_user <- oarch_graph_repo_user |>
@@
 We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate nodes, which represent repositories with only a single contributor, we detected 21 clusters.
+Aside from isolate nodes (n = XXX, not visible in our visualizations), which represent repositories with only a single contributor, we detected 21 clusters.
 While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
 
-<!-- TODO: Note that we remove disconnected repos? (and how many?) -->
+<!-- TODO: Note that we remove disconnected repos? (and how many?)
+ZB response: See change in line 95 above -->
 
@@
 
-```{r fig-graph-repo-repo, eval=FALSE}
+```{r fig-graph-repo-repo, eval=TRUE}
+# ZB: I don't think this actually says anything meaningful and could be removed.
 oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
-  geom_edge_link(aes(alpha = n)) +
-  geom_node_point(aes(colour = category)) +
+  ggraph(layout <- create_layout(oarch_graph_repo_repo, layout = 'igraph', algorithm = 'fr')) +
+  geom_edge_link(aes(alpha = 1)) +
+  #geom_node_point(aes(colour = category)) +
   scale_edge_alpha(trans = "log") +
@@
 
-```{r fig-graph-repo-repo-interactive, eval=FALSE}
+```{r fig-graph-repo-repo-interactive, eval=TRUE}
+# Remove the legend.
 iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt") +
+  ggraph(layout = "drl") +
   geom_edge_link(aes(alpha = n)) +
@@
 
-```{r fig-graph-repo-repo-edge-betweenness-dendrogram, eval=FALSE}
+```{r fig-graph-repo-repo-edge-betweenness-dendrogram, eval=TRUE}
 oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
@@
 ```
+<!--
+ZB Note: What we really want out of this is the identification of groups or clusters, as well as the relationships between clusters.
+In theory, the dendrogram provides the best visualization for this information.
+I wonder if we can improve it by colour-coding the algorithmically-identified clusters (which should appear side by side), and making the names of the repositories appear vertically aligned perpendicular to the x axis.
+Additionally, we can colour-code the names of the repositories with their corresponding colours, or indicate that they are part of a cluster using a bracket (like "{" but turned 180 degrees).
+-->
 
-This graph identifies three core clusters characterized by repositories whose contributors commit to projects other than their own.
-The clusters that appear along the peiphery largely correspond with the work of single individuals, and sometimes also their colleagues.
-Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (and perhaps initial) integrations of lone developers into a broader community.
+This graph identifies three core clusters and several peripheral clusters.
+The core clusters are characterized by repositories whose contributors commit to projects other than their own.
+The other peripheral clusters largely correspond with the work of single individuals, and sometimes also their close colleagues.
+Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (or perhaps initial) integration of lone developers into a broader community.
 
@@
 
-<!-- TODO: What does tbl-repo-centrality tell us? Can we interrogate it further with a graph? -->
+<!-- TODO: What does tbl-repo-centrality tell us? Can we interrogate it further with a graph?
+ZB Response: Not much, I think I conceived it more as a means to an end.
+If I recall correctly, my intention was to ascertain whether different kinds (categories, mainly) of repositories are more central.
+One observation is the change in magnitude of betweenness centrality between the two "indexical" entries (open-archaeo and ctv-archaeo) and others in this top 10.
+Perhaps we can detect and visualize additional abrupt "jumps" in magnitude and identify some pattern relating to other variables.
+-->
 
@@
 
-<!--
- NOTE, ZACK: the graph below is what you requested (nodes' sizes based on their betweenness centrality), but I'm unsure we will ever be able to get these graph plots legible enough to actually include.
-– JR -->
+<!-- NOTE, ZACK: the graph below is what you requested (nodes' sizes based on their betweenness centrality), but I'm unsure we will ever be able to get these graph plots legible enough to actually include.
+– JR
+ZB Response: I agree, this visual is useless.
+This, as the other generalized one-mode were meant to simply show that we had indeed split the two-mode into two one-modes.
+They provide little informational value and should be removed.
+-->
 
-```{r fig-graph-user-user-centrality, eval=FALSE}
+```{r fig-graph-user-user-centrality, eval=TRUE}
 oarch_graph_user_user |> 
@@
 
-**NOTE:** Create a table containing the top 10 users based on betweenness centrality values, with columns specifying theur username, real name, job title/affiliation, location, career stage (student, early/mid/late career, tenured/non-tenured, independent, etc).
+**NOTE:** Create a table containing the top 10 users based on betweenness centrality values, with columns specifying their username, real name, job title/affiliation, location, career stage (student, early/mid/late career, tenured/non-tenured, independent, etc).
 
 <!-- JR: I'm not sure about this.
-Either practically (can we reliably determine all this from public info?) or ethically (not necessarily that it is unethical; but we haven't really considered the implications of analysing named individuals in this study) -->
+Either practically (can we reliably determine all this from public info?) or ethically (not necessarily that it is unethical; but we haven't really considered the implications of analysing named individuals in this study)
+ZB Response: As with the table above, this is a means to an end, a way of working out what we can and want to say about this.
+However I think this will require more thought, beyond the scope of what we are currently able to achieve, and may be better suited for another paper (connected with the DIIF grant, perhaps).
+However, see the final note below (line 253) regarding comparison between people who contribute code, things other than code, or both code and non-code, which I think is a viable and valuable addition here.
+-->
 
@@
 ---
-
-```{r fig-user-betweenness, eval=FALSE}
+<!--ZB Note: What is this? -->
+```{r fig-user-betweenness, eval=TRUE}
 #| fig-cap: Distribution of betweenness centrality values in the user–user network
@@
 
-```{r fig-user-betweenness-activity, eval=FALSE}
+```{r fig-user-betweenness-activity, eval=TRUE}
 #| fig-cap: Correlation between user's centrality (betweenness) in the network and their total number of contributions

### 2023-09-20 analysis/_07-conclusion.qmd
Modifications and comments.

@@
 Moreover, archaeologists are accustomed to working on shoestring budgets, having to adapt their tools to suit specific, non-generalizable situations, and applying a DIY work ethic to create and co-creating with others to get the job done (Batist et al 2021; Morgan and Eve 2012; Caraher 2019).
-It is therefore unsprising that archaeologists are prolific open source software developers.
+It is therefore unsurprising that archaeologists are prolific open source software developers.
 
-In this study we have sought to document an emerging community of open source software engineers in archaeology and to assess the extent to which collaboration within this community resembles conventional 'open source' forms.
-We have operationalised 'open-source collaborative practices' as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
+In this study, we documented an emerging community of open source software engineers in archaeology and assessed the extent to which collaboration within this community resembles conventional 'open source' forms.
+We operationalised 'open-source collaborative practices' as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
 This excludes a wide range of other potential practices, most obviously the use of other software (though as we demonstrate in section X.X, the git/GitHub combination is overwhelmingly the most popular), but also collaboration through offline or private channels, or forms of collaboration we don't know about.
-We have not directly interviewed those doing the collaborating, though our conclusions do drawn heavily from our experience as members of that community ourselves. 
-Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it (@cite).
+We have not directly interviewed those doing the collaborating, though our conclusions do draw heavily from our experience as members of that community ourselves. 
+Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it (@ducke2013 10.4324/9781315431932-12; @whallon1972 10.1007/BF02403759).
 

### 2023-09-28 analysis/_06-network.qmd
Updated some explanation, removed irrelevant  content.

@@
 
-We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations.
+We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations (@fig-graph-repo-user).
+We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
 
@@
 
-We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
+## repo-repo
 
-**NOTE:** Combine these figures using patchwork.
-**NOTE:** Integrate {r graph-user-user} and {r plot-graph-user-user} into a single cell, similar to how {r plot-graph-repo-repo} integrates both components.
-
-## repo-repo
-We applied the edge-betweenness community detection method to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate nodes (n = XXX, not visible in our visualizations), which represent repositories with only a single contributor, we detected 21 clusters.
+We applied the edge-betweenness community detection method (Girvan and Newman 2002) to identify clusters of repositories that share common sets of contributors of all kinds.
+Aside from isolate nodes (n = XXX, not visible in our visualizations), which represent repositories with only a single contributor, we detected 21 clusters (@fig-graph-repo-repo-edge-betweenness-dendrogram).
 While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
-
-<!-- TODO: Note that we remove disconnected repos? (and how many?)
-ZB response: See change in line 95 above -->
 
@@
   oarch_graph_repo_repo
-```
-
-```{r fig-graph-repo-repo, eval=TRUE}
-# ZB: I don't think this actually says anything meaningful and could be removed.
-oarch_graph_repo_repo |>
-  ggraph(layout <- create_layout(oarch_graph_repo_repo, layout = 'igraph', algorithm = 'fr')) +
-  geom_edge_link(aes(alpha = 1)) +
-  #geom_node_point(aes(colour = category)) +
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
 ```
@@
 ```{r fig-graph-repo-repo-edge-betweenness-dendrogram, eval=TRUE}
+# ZB Note: What we really want out of this is the identification of groups or clusters, as well as the relationships between clusters.
+In theory, the dendrogram provides the best visualization for this information.
+I wonder if we can improve it by colour-coding the algorithmically-identified clusters (which should appear side by side), and making the names of the repositories appear vertically aligned perpendicular to the x axis.
+Additionally, we can colour-code the names of the repositories with their corresponding colours, or indicate that they are part of a cluster using a bracket (like "{" but turned 180 degrees).
 oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
@@
 ```
-<!--
-ZB Note: What we really want out of this is the identification of groups or clusters, as well as the relationships between clusters.
-In theory, the dendrogram provides the best visualization for this information.
-I wonder if we can improve it by colour-coding the algorithmically-identified clusters (which should appear side by side), and making the names of the repositories appear vertically aligned perpendicular to the x axis.
-Additionally, we can colour-code the names of the repositories with their corresponding colours, or indicate that they are part of a cluster using a bracket (like "{" but turned 180 degrees).
--->
 
@@
 
-**NOTE:** What happens to this interpretation when we differentiate commits from issues and comments?
+<!--  What happens to this interpretation when we differentiate commits from issues and comments? -->
 
@@
 We assembled graphs linking users based on common contributions to the same repositories.
-
-<!-- TODO: Note that we remove disconnected users? (and how many?) -->
+As with the repo-repo network, we excluded isolate nodes (n = xxx), which represent users who only contributed to their own repositories, from the visualization.
 
@@
 
-<!-- NOTE, ZACK: the graph below is what you requested (nodes' sizes based on their betweenness centrality), but I'm unsure we will ever be able to get these graph plots legible enough to actually include.
-– JR
-ZB Response: I agree, this visual is useless.
-This, as the other generalized one-mode were meant to simply show that we had indeed split the two-mode into two one-modes.
-They provide little informational value and should be removed.
--->
-
-```{r fig-graph-user-user-centrality, eval=TRUE}
-oarch_graph_user_user |> 
-  ggraph(layout = "graphopt") +
-  # geom_edge_link(aes(alpha = n)) +
-  geom_edge_arc(alpha = 0.05) +
-  geom_node_point(aes(size = centrality), 
-                  shape = 21, colour = "#000000", fill = "#ffffff") +
-  # scale_edge_alpha(trans = "log") +
-  # labs(edge_alpha = "Common repositories") +
-  theme_graph()
-```
-
-We sought to identify whether certain users, who contribute in certain distinct ways, play different roles in overall network.
-We found that the people with the highest betweenness values are those who primarily produce computational archaeology code as their job.
+We sought to identify whether users who contribute in certain distinct ways play different roles in the overall network.
+We are reluctant to share personal information about specific users without their informed consent, but based on our knowledge of the community we found that the people with the highest betweenness values are those who primarily produce computational archaeology code as their job.
 Moreover, we found that these people tend to be employed under precarious circumstances.
@@
 
-**NOTE:** Create a table containing the top 10 users based on betweenness centrality values, with columns specifying their username, real name, job title/affiliation, location, career stage (student, early/mid/late career, tenured/non-tenured, independent, etc).
-
-<!-- JR: I'm not sure about this.
-Either practically (can we reliably determine all this from public info?) or ethically (not necessarily that it is unethical; but we haven't really considered the implications of analysing named individuals in this study)
-ZB Response: As with the table above, this is a means to an end, a way of working out what we can and want to say about this.
-However I think this will require more thought, beyond the scope of what we are currently able to achieve, and may be better suited for another paper (connected with the DIIF grant, perhaps).
-However, see the final note below (line 253) regarding comparison between people who contribute code, things other than code, or both code and non-code, which I think is a viable and valuable addition here.
--->
-
-```{r tbl-user-centrality}
-#| tbl-cap: Most central users in the user–user network
-oarch_graph_user_user |>
-  mutate(centrality = centrality_betweenness(weights = n)) |>
-  as_tibble() |> 
-  arrange(-centrality) |>
-  slice(1:10) |>
-  left_join(oarch_contributors, by = c("user" = "contributor")) |>
-  select(user, centrality, repos, contribs_own, contribs_other) |>
-  gt() |>
-  cols_label(
-    user = "User",
-    centrality = "Centrality",
-    contribs_own = "Own repositories",
-    contribs_other = "Others' repositories"
-  ) |>
-  tab_spanner("Contributions", c(contribs_own, contribs_other))
-```
+---
 
@@
 
-**NOTE:** Create a table containing the top 10 users based on betweenness centrality values derived from the issues and comments only subnetwork, with columns specifying theur username, real name, job title/affiliation.
+------------------------------------------------------------------------
 
----
 <!--ZB Note: What is this? -->
+
 ```{r fig-user-betweenness, eval=TRUE}

### 2023-09-29 analysis/_04-open_archaeology.qmd
Revised draft of section 04.
Simplified the text, making it less descriptive and more expository. Added captions to figures. Included notes to embed percentages in the text (I could not find the necessary variables).

@@
 ```
-As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists, primarily software but also various forms of open document.
-@tbl-categories summarizes the categories included.
+As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists.
+This primarily constitutes software but also includes various forms of open documents.
+@tbl-categories summarizes the kinds if resources that appear in open-archaeo, and breaks them down into more precise categories.
 
 ```{r tbl-categories}
-#| tbl-cap: Categories of open archaeology projects included in open-archaeo
+#| tbl-cap: Categories of open archaeology projects included in open-archaeo.
+# TODO: Include sum for the "Software" and "Documents" supercategories, and make them appear in bold.
 tribble(
@@
 
-Most projects (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language or a plugin for an application.
+Most resources (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language or a plugin for an application.
 The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
-Others create standalone software that can be run independently of such platforms, for example desktop or web apps.
-A significant number of projects also comprise of datasets and non-packaged code snippets that have been made available for general use.
+Others create standalone software <!-- percent --> that can be run independently of such platforms, for example desktop or web apps.
+A significant number of projects also comprise of datasets <!-- percent --> and non-packaged code snippets <!-- percent --> that have been made available for general use.
 
 ```{r tbl-platforms}
-#| tbl-cap: Platforms and programming languages used by open archaeology projects
+#| tbl-cap: Platforms and programming languages used by open archaeology projects.
 oarch |>
@@
 
-The statistical programming language R is overwhelmingly the most common platform, representing `r percent(p_platform_r)` of projects in open-archaeo.
-Python, another programming language, is also relatively popular, as are plugins for the open source geographic information system QGIS.
+As per @tbl-platforms, The statistical programming language R `r percent(p_platform_r)` is overwhelmingly the most common platform among projects that extend upon existing programming languages and applications.
+This is followed by Python <!-- percent -->, which is another popular scientific scripting language, and the open source geographic information system QGIS <!-- percent -->.
 Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
@@
 We also annotated each record with 'tags' that describe aspects of archaeological work that each tool contributes to [@fig-tags].
+The most common tags unsurprisingly deal with work that naturally benefits from advanced information processing afforded by computers, such as statistical analysis, sample calibration, geographical analysis, data management, and chronological modelling.
+Educational resources and practical guides are also well represented due to the web's usefulness as a medium for sharing and communication.
+
 When we compare categories with tags, we see the general domains that each kind of resource is designed to serve.
@@
 ```{r fig-tags}
+#| fig-cap: Frequency of tags applied to open archaeology projects included in open-archaeo, broken down by category.
 detail_tags <- c("Instrumental Neutron activation analysis",
@@
 ```{r fig-github-cumulative}
+#| fig-cap: Cumulative growth of open archaeological software in terms of number of commits and number of repositories, and broken down by category.
 oarch %>% 
@@
 This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work in an open and somewhat nerdy way.
-
-------------------------------------------------------------------------
-
--   Compare changing proportions of each category on a year by year basis
--   Identify temporal trends in the use of licenses

### 2023-10-11 analysis/_04-open_archaeology.qmd
Merge conflicts

@@
 
-```{=html}
 <!-- Results I: description of open archaeology projects
@@
 ```
+
 As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists.
@@
 
-As per @tbl-platforms, The statistical programming language R `r percent(p_platform_r)` is overwhelmingly the most common platform among projects that extend upon existing programming languages and applications.
-This is followed by Python <!-- percent -->, which is another popular scientific scripting language, and the open source geographic information system QGIS <!-- percent -->.
+`r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used 'platform' by a large margin (@tbl-platforms).
+Python, another programming language, is also relatively popular <!-- TODO: percent -->, as are plugins for the open source geographic information system QGIS <!-- TODO: percent -->.
 Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
@@
 ```{r fig-tags}
-#| fig-cap: Frequency of tags applied to open archaeology projects included in open-archaeo, broken down by category.
-detail_tags <- c("Instrumental Neutron activation analysis",
-                 "Harris Matrix",
-                 "aDNA Simulators",
-                 "Damage profilers",
-                 "Metagenomic classifers",
-                 "aDNA simulators",
-                 "aDNA damage-aware genotypers",
-                 "Low-coverage contamination estimation",
-                 "Low-coverage adapted population-genetics tools",
-                 "aDNA preprocessing",
-                 "aDNA aligner",
-                 "aDNA assembler",
-                 "aDNA toolkit")
-
+#| fig-cap: Tags by category
 oarch %>%
@@
   drop_na(tag) %>%
-  filter(!tag %in% detail_tags) %>%
   mutate(category = recode(category,
-                                  "Specifications etc." = "Documents",
+                                  "Specifications, protocols and schemas" = "Documents",
                                   "Products" = "Documents",
@@
   ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  # facet_wrap(vars(category)) +
   geom_col() +
@@
   labs(
-    title = "Tags by category",
-    caption = "Data: open-archaeo.info",
     x = NULL, y = "Number of projects",

### 2023-10-11 analysis/_04-open_archaeology.qmd
Fix stray chunk boundary

@@
 -->
-```
 

### 2023-10-11 analysis/openarchaeo_collaboration.qmd
Add affiliation URLs
Because why not

@@
     affiliation: Faculty of Information, University of Toronto
+    affiliation-url: https://ischool.utoronto.ca/
     orcid: 0000-0003-0435-508X 
@@
     affiliation: Institute of Archaeological Sciences, University of Bern
+    affiliation-url: https://www.iaw.unibe.ch/
     orcid: 0000-0002-1011-1244

### 2023-10-11 analysis/_05-collaboration.qmd
Add some introductory text to section 5

@@
 
-<!-- Results II: description of observed collaborative practices
+As well as hosting source code, GitHub and other software forges include systems for facilitating collaboration on code and other projects.
+The basic collaborative workflow is inherited from git, which allows multiple users to *commit* (see @tbl-glossary for definitions of this and other git terminology used in this section) code to the repository.
+A user with *commit access* to a repository can change any of its contents at will, so this is usually reserved for the project maintainer and known, trusted collaborators.
+GitHub extends this model with its *pull request* feature, by which any user can offer to contribute code to a repository to which they don't have commit access.
+The maintainer may then choose to *merge* (accept) or *decline* the pull request, facilitating contributions from a wider network of collaborators without the need for permission to be sought in advance.
 
-- How active are projects?
-- How many people are involved?
-- Conversely, how many projects are individual people involved in?
-- What is the level of uptake of GitHub's collaboration features?
-- Has this changed over time?
+GitHub also facilitates collaboration on broader project management tasks, primarily through its *issues* feature.^[Apart from issues, GitHub has a very wide range of project management and social media-like features, including wikis, discussion forums and 'kanban' boards.
+We have not analysed the use of these features here.]
+Unless a repository's maintainer specifically configures it otherwise, any user can create an issue attached to another user's repository, or comment on an existing issue.
+Issues are typically used to log and track bug reports, feature requests, and other comments and suggestions from the project's user base.
+GitHub's pull request feature is also implemented via this system – a pull request is a special type of issue.
 
-Figures:
+Users can also engage with repositories and other users with social media-like features such as *starring* a repository, adding it to a public or private list, or *following* a user.
+These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This timeline was in fact one of the primary ways we compiled open-archaeo.]
+While not constituting a direct form of collaboration, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.
 
-- Charts presenting commit rates and repo longevity
-- Charts presenting the number of contributors and the amount of work that they each contributed
-- Charts presenting the use of issues and comments
+The prototypical collaborative open source project on GitHub has a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through pull requests, and an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
+The data we collected on open-archaeo repositories on GitHub show that uptake of its collaboration features is highly uneven, with very few projects resembling the open source ideal.
+<!-- TODO: note that we don't know if maintainers actually aimed for that model, or leave that for the discussion? -->
 
--->
+* How active are projects
+  * Lifespan and activity rate vary greatly.
+  * Average lifespan, longest lifespan
+  * % of projects with only one commit, under one month, under three months, under a year
+  * NB. when we say lifespan, we mean *development* lifespan
+  * Average commit rate, range of commit rates
 
-But how is this activity distributed across repositories? In this section we assess projects' pulse, or the amount of activity pertaining to a repository.
+* Comparison of activity vs. longevity
+  * Highly active projects (>1 commit per day) tend to be very long-lived or very short-lived (not many in the middle)
+  * Long-lived projects have a range of activity levels, relatively evenly spread
+  * The most "successful" (according to the expectations of open source, i.e. long-lived and active) are also those with large numbers of contributors (up to 150).
+  * The central tendency of the distribution is c. [1000,0.10], i.e. projects that are about three years old and with ~3 commits a month; they are solo projects, or small groups.
+
+* How many people are involved?
+  *  Conversely, how many projects are individual people involved in?
+
+* Overall, what is the level of uptake of GitHub's collaboration features?
+  * Has this changed over time?
 
@@
 
-```{r plot-lifespan}
+```{r fig-lifespan}
+#| fig-cap: Lifespan of repositories based on the date of first and last commit. Points indicate intermediate commits. Only repositories with more than one commit are included.
 oarch %>%
@@
   filter(n > 1) %>%
-  mutate(first_commit = min(datetime),
-         mid_commit = median(datetime),
-         last_commit = max(datetime)) %>%
+  mutate(
+    first_commit = min(datetime, na.rm = TRUE),
+    mid_commit = median(datetime, na.rm = TRUE),
+    last_commit = max(datetime, na.rm = TRUE)
+  ) %>%
   ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
@@
              ymax = last_commit)) +
-  geom_linerange(size = 0.05) +
+  # geom_linerange(size = 0.05) +
   geom_point(size = 0.1) +
@@
   # coord_flip() +
-  labs(
-    title = "Project lifespan",
-    subtitle = glue("Individual commits in GitHub repositories (with more than 1 commit)"),
-    x = NULL, y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
+  labs(x = NULL, y = NULL) +
   theme_oarch() +
@@
 
-```{r lifespan-stats}
+```{r generate-oarch-lifespan}
 oarch %>% 
@@
 
-```{r plot-lifespan-hist}
-# TODO: combine with patchwork
-# TODO: where does oarch_lifespan come from?
+```{r fig-lifespan-hist}
+#| fig-cap: Project lifespan 
 oarch_lifespan %>%
@@
   labs(
-    title = "Project lifespan",
-    subtitle = "Time between first and latest commit, by repository",
-    x = "Days", y = NULL,
-    caption = "Data: open-archaeo.info and GitHub API"
+    x = "Lifespan (days)", y = "Repositories",
   ) +
@@
 
-```{r plot-lifespan-vs-rate}
+```{r fig-lifespan-rate}
+#| fig-cap: Lifespan, commit rate and contributors to repositories. Note that both scales are logarithmic.
+# TODO: add marginal histograms?
 oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n)) +
-  geom_point(shape = 21, fill = NA) +
+  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
+  geom_point(shape = 21, colour = "black", fill = alpha("black", 0.3)) +
   scale_x_log10() +
   scale_y_log10() +
-  labs(
-    title = "Project lifespan vs. average commit rate",
-    x = "Lifespan (days)", y = "Commits per day", size = "Total commits",
-    caption = NULL
-  ) +
-  theme_oarch()
-
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, fill = NA) +
-  scale_x_log10() +
-  scale_y_log10() +
-  labs(
-    title = NULL,
-    x = "Lifespan (days)", y = "Commits per day", size = "Contributors",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
+  labs(x = "Lifespan (days)", y = "Commits per day", size = "Contributors") +
   theme_oarch()
@@
 
-```{r plot-issues-hist}
-# TODO: combine with comments-hist
-oarch %>%
+```{r fig-issues-comments}
+#| fig-cap: Use of issues and comments
+# TODO: include zeroes
+# TODO: why are there no twos?
+fig_issues_hist <- oarch %>%
   unnest(gh_issues) %>%
@@
   scale_x_log10() +
-  labs(
-    title = "Use of issues",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of issues", y = "Repositories"
-  ) +
+  labs(x = "Number of issues", y = "Repositories") +
   theme_oarch()
-```
 
-```{r plot-comments-hist}
-oarch %>%
+fig_comments_hist <- oarch %>%
   unnest(gh_comments) %>%
@@
   scale_x_log10() +
-  labs(
-    title = "Use of comments",
-    caption = "Data: open-archaeo.info and GitHub API",
-    x = "Number of comments", y = "Repositories"
-  ) +
+  labs(x = "Number of comments", y = "Repositories") +
   theme_oarch()
+
+fig_issues_hist + fig_comments_hist
 ```

### 2023-10-11 analysis/_05-collaboration.qmd
Draft results on lifespan/activity

@@
 Users can also engage with repositories and other users with social media-like features such as *starring* a repository, adding it to a public or private list, or *following* a user.
-These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This timeline was in fact one of the primary ways we compiled open-archaeo.]
+These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This feature of GitHub's timeline was one of the primary ways we compiled open-archaeo.]
 While not constituting a direct form of collaboration, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.
@@
 The data we collected on open-archaeo repositories on GitHub show that uptake of its collaboration features is highly uneven, with very few projects resembling the open source ideal.
+The majority of projects are short-lived, with few contributors and a small number of commits.
+Only a fraction of repositories use GitHub's extended collaboration features (pull requests, issues, comments) and those that do use them in a limited way.
 <!-- TODO: note that we don't know if maintainers actually aimed for that model, or leave that for the discussion? -->
-
-* How active are projects
-  * Lifespan and activity rate vary greatly.
-  * Average lifespan, longest lifespan
-  * % of projects with only one commit, under one month, under three months, under a year
-  * NB. when we say lifespan, we mean *development* lifespan
-  * Average commit rate, range of commit rates
-
-* Comparison of activity vs. longevity
-  * Highly active projects (>1 commit per day) tend to be very long-lived or very short-lived (not many in the middle)
-  * Long-lived projects have a range of activity levels, relatively evenly spread
-  * The most "successful" (according to the expectations of open source, i.e. long-lived and active) are also those with large numbers of contributors (up to 150).
-  * The central tendency of the distribution is c. [1000,0.10], i.e. projects that are about three years old and with ~3 commits a month; they are solo projects, or small groups.
-
-* How many people are involved?
-  *  Conversely, how many projects are individual people involved in?
-
-* Overall, what is the level of uptake of GitHub's collaboration features?
-  * Has this changed over time?
-
-In the following figure, each vertical line represents a repository's lifespan, and the points on each line are individual commits.
-The graph is sorted by the date of the most recent commit for each repository, so the ones on the top became inactive earlier and the ones on the bottom represent ongoing work.
+Conversely, a small number of projects (<!-- TODO: do these share any characteristics in terms of platform or subject?
+-->) do appear to fit the aspirational open source model, with a large number of contributors maintaing a long-lived and highly-active repository.
+Archaeologists active on GitHub also make markedly <!-- TODO: verify --> greater use of GitHub's social media-like features (stars, following) than more direct collaborative actions.
 
@@
 #| fig-cap: Lifespan of repositories based on the date of first and last commit. Points indicate intermediate commits. Only repositories with more than one commit are included.
+#TODO: This looks cool, but is it actually more useful than a histogram or two?
 oarch %>%
@@
 
-The span of time between the first and latest commits is visualized more clearly in the following charts.
-Notably, many projects are active for only a short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
-Projects with a short span of activity tend to have very high commit frequency, whereas there is much more variety in long-term projects.
-Among projects with a greater lifespan, those with more total commits have much more frequent activity than those with fewer total commits, which represent the majority of long-lived projects.
-The overall negative trend (seen in the top-right graph) can be explained in terms of natural processes, i.e.
-the only way to fit a high number of commits within the same span as a project with a low number of commits is to commit more frequently.
-But bearing in mind both axes use a logarithmic scale, the upper left portion of the graph does appear to suggest that short-lived projects are characterized by a spree of rapid committing; while longer-lived projects have a more diverse range of workrates.
-This latter pattern in turn appears to be explained by the number of contributors: projects that sustain a high commit rate over a long period of time tend to rely on a large number of contributors, as per the bottom-right graph.
-
 ```{r fig-lifespan-hist}
@@
 ```
+
+The lifespan and activity (commit rate) of repositories in open-archaeo vary greatly (@fig-lifespan).
+Here we refer to the *development* lifespan of a project—measured here as the time between a repository's first and last commit—which is not necessary linked to its use-life.
+The average project lasts <!-- TODO --> days with <!-- TODO --> commits per day.
+<!-- TODO: turn all these percentages into code -->
+Many projects are active for only a short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
+The longest are <!-- TODO -->.
+Commit rates vary between <!-- TODO -->, with peaks at <!-- TODO -->.
 
@@
 
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
-A "contribution", in this case, is a commit, issue, or comment applied by a user to a repository.
+The interaction between project longevity, activity, and number of contributors is multifaceted [@fig-lifespan-rate].
+Highly active projects (one commit per day or more) tend to be either very long-lived or very short-lived; few fall in the centre of the distribution.
+Short-lived projects tend to be characterised by a 'spree' of activity (a high commit rate), while long-lived projects have a broader range of activity profiles.
+The most "successful" projects according to open source norms (i.e. long-lived and active) are with few exceptions those projects with the largest contributor base in our dataset.
+However, the modal project in the centre of the distribution is more modest, lasting around three years, maintained as by an individual or a small group, with around three commits per month.
 
@@
 
-Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
-Of multi-contributor repositories, 92% have a third of contributions made by a single person, 65% have half of all contributions made by a single person, and 36% have three quarters of all contributions made a single person.
+* How many people are involved?
+  *  Conversely, how many projects are individual people involved in?
+
+Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
+A "contribution", in this case, is a commit, issue, or comment applied by a user to a repository.
 
@@
 
-Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-Only 39% of repositories make use of these features.
-Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
+Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
+Of multi-contributor repositories, 92% have a third of contributions made by a single person, 65% have half of all contributions made by a single person, and 36% have three quarters of all contributions made a single person.
 
@@
 ```
+
+Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
+Only 39% of repositories make use of these features.
+Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
+However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
+This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
+
+```{r fig-collaboration-ts}
+# TODO: combined time series of:
+# * average project longevity
+# * commit rate
+# * number of collaborators
+# * use of issues+comments
+```
+
+* Summarise: overall, what is the level of uptake of GitHub's collaboration features?
+  * Has it changed over time? (@fig-collaboration-ts)

### 2023-10-12 analysis/_05-collaboration.qmd
Reorder section 5

@@
 
-GitHub also facilitates collaboration on broader project management tasks, primarily through its *issues* feature.^[Apart from issues, GitHub has a very wide range of project management and social media-like features, including wikis, discussion forums and 'kanban' boards.
-We have not analysed the use of these features here.]
-Unless a repository's maintainer specifically configures it otherwise, any user can create an issue attached to another user's repository, or comment on an existing issue.
-Issues are typically used to log and track bug reports, feature requests, and other comments and suggestions from the project's user base.
-GitHub's pull request feature is also implemented via this system – a pull request is a special type of issue.
-
-Users can also engage with repositories and other users with social media-like features such as *starring* a repository, adding it to a public or private list, or *following* a user.
-These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This feature of GitHub's timeline was one of the primary ways we compiled open-archaeo.]
-While not constituting a direct form of collaboration, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.
-
-The prototypical collaborative open source project on GitHub has a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through pull requests, and an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
-The data we collected on open-archaeo repositories on GitHub show that uptake of its collaboration features is highly uneven, with very few projects resembling the open source ideal.
-The majority of projects are short-lived, with few contributors and a small number of commits.
-Only a fraction of repositories use GitHub's extended collaboration features (pull requests, issues, comments) and those that do use them in a limited way.
-<!-- TODO: note that we don't know if maintainers actually aimed for that model, or leave that for the discussion? -->
-Conversely, a small number of projects (<!-- TODO: do these share any characteristics in terms of platform or subject?
--->) do appear to fit the aspirational open source model, with a large number of contributors maintaing a long-lived and highly-active repository.
-Archaeologists active on GitHub also make markedly <!-- TODO: verify --> greater use of GitHub's social media-like features (stars, following) than more direct collaborative actions.
-
 ```{r fig-lifespan}
@@
 #TODO: This looks cool, but is it actually more useful than a histogram or two?
+#TODO: Remove that one commit in 2001
 oarch %>%
@@
              ymax = last_commit)) +
-  # geom_linerange(size = 0.05) +
+  geom_linerange(linewidth = 0.05) +
   geom_point(size = 0.1) +
-  scale_y_datetime(limits = c(ymd("2005-01-01", tz = "UTC"),
-                              ymd("2021-05-31", tz = "UTC")),
-                   date_breaks = "2 years", date_labels = "%Y") +
+  scale_y_datetime(date_breaks = "2 years", date_labels = "%Y") +
   # coord_flip() +
@@
 
-The lifespan and activity (commit rate) of repositories in open-archaeo vary greatly (@fig-lifespan).
-Here we refer to the *development* lifespan of a project—measured here as the time between a repository's first and last commit—which is not necessary linked to its use-life.
+We measured the lifespan of a repository as the time between the first and latest commit, and its activity as the rate of commits.
+Here therefore refer to the *development* lifespan of a project—measured here as the time between a repository's first and last commit—which is not necessary linked to its use-life.
+The lifespan and activity of repositories in open-archaeo vary greatly (@fig-lifespan).
 The average project lasts <!-- TODO --> days with <!-- TODO --> commits per day.
@@
 
+```{r fig-issues-comments}
+#| fig-cap: Use of issues and comments
+# TODO: include zeroes
+# TODO: why are there no twos?
+fig_issues_hist <- oarch %>%
+  unnest(gh_issues) %>%
+  drop_na(title) %>%
+  group_by(repo) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  ggplot(aes(x = n)) +
+  geom_histogram() +
+  scale_x_log10() +
+  labs(x = "Number of issues", y = "Repositories") +
+  theme_oarch()
+
+fig_comments_hist <- oarch %>%
+  unnest(gh_comments) %>%
+  drop_na(issue_num) %>%
+  group_by(repo) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  ggplot(aes(x = n)) +
+  geom_histogram() +
+  scale_x_log10() +
+  labs(x = "Number of comments", y = "Repositories") +
+  theme_oarch()
+
+fig_issues_hist + fig_comments_hist
+```
+
+Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
+Only 39% of repositories make use of these features.
+Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
+However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
+This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
+
+GitHub also facilitates collaboration on broader project management tasks, primarily through its *issues* feature.^[Apart from issues, GitHub has a very wide range of project management and social media-like features, including wikis, discussion forums and 'kanban' boards.
+We have not analysed the use of these features here.]
+Unless a repository's maintainer specifically configures it otherwise, any user can create an issue attached to another user's repository, or comment on an existing issue.
+Issues are typically used to log and track bug reports, feature requests, and other comments and suggestions from the project's user base.
+GitHub's pull request feature is also implemented via this system – a pull request is a special type of issue.
+
 ```{r plot-gh_contributors-histogram}
@@
 A "contribution", in this case, is a commit, issue, or comment applied by a user to a repository.
-
-**NOTE:** Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic, which are now divided into quarters rather than thirds.
+<!-- TODO: expand -->
 
 ```{r plot-contrib-dist}
+#TODO: Re-calculate these numbers to ensure they are up to date.
+Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic, which are now divided into quarters rather than thirds.
 oarch %>%
@@
 
-```{r fig-issues-comments}
-#| fig-cap: Use of issues and comments
-# TODO: include zeroes
-# TODO: why are there no twos?
-fig_issues_hist <- oarch %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(x = "Number of issues", y = "Repositories") +
-  theme_oarch()
-
-fig_comments_hist <- oarch %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(x = "Number of comments", y = "Repositories") +
-  theme_oarch()
-
-fig_issues_hist + fig_comments_hist
+```{r fig-stars}
+#TODO: stars figure?
 ```
 
-Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-Only 39% of repositories make use of these features.
-Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
+```{r fig-follows}
+#TODO: follows figure?
+```
+
+Another way GitHub users can engage with repositories and other users with social media-like features such as *starring* a repository, adding it to a public or private list, or *following* a user.
+These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This feature of GitHub's timeline was one of the primary ways we compiled open-archaeo.]
+While not constituting a direct form of collaboration, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.
 
@@
   * Has it changed over time? (@fig-collaboration-ts)
+
+The prototypical collaborative open source project on GitHub has a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through pull requests, and an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
+The data we collected on open-archaeo repositories on GitHub show that uptake of its collaboration features is highly uneven, with very few projects resembling the open source ideal.
+The majority of projects are short-lived, with few contributors and a small number of commits.
+Only a fraction of repositories use GitHub's extended collaboration features (pull requests, issues, comments) and those that do use them in a limited way.
+<!-- TODO: note that we don't know if maintainers actually aimed for that model, or leave that for the discussion? -->
+Conversely, a small number of projects (<!-- TODO: do these share any characteristics in terms of platform or subject?
+-->) do appear to fit the aspirational open source model, with a large number of contributors maintaing a long-lived and highly-active repository.
+Archaeologists active on GitHub also make markedly <!-- TODO: verify --> greater use of GitHub's social media-like features (stars, following) than more direct collaborative actions.

### 2023-10-12 analysis/_03-data_methodology.qmd
Add statistics on lifespan and longevity

@@
                                  ))
+
+# TODO: exclude peripheral repos that throw things off
+# * Big boys like QField (see _06-network.qmd)
+# * ArchaeoLines – ends up picking up the whole stellarium repo, which is huge
+# * aDNA??
 ```

### 2023-10-12 analysis/_05-collaboration.qmd
Add statistics on lifespan and longevity

@@
     commit_rate = n / lifespan
-  ) ->
+  ) |>
+  mutate(commit_rate = if_else(is.finite(commit_rate), commit_rate, NA)) ->
   oarch_lifespan
-```
 
-```{r fig-lifespan-hist}
-#| fig-cap: Project lifespan 
-oarch_lifespan %>%
-  ggplot(aes(x = lifespan)) +
-  geom_histogram(binwidth = 100) +
-  scale_x_continuous(breaks = scales::breaks_width(1000)) +
-  labs(
-    x = "Lifespan (days)", y = "Repositories",
-  ) +
-  theme_oarch()
+# Statistics
+n_1commit <- nrow(filter(oarch_lifespan, n == 1))
+
+p_lifespan_30days <- nrow(filter(oarch_lifespan, lifespan < 30)) / nrow(oarch_lifespan)
+p_lifespan_90days <- nrow(filter(oarch_lifespan, lifespan < 90)) / nrow(oarch_lifespan)
+p_lifespan_365days <- nrow(filter(oarch_lifespan, lifespan < 365)) / nrow(oarch_lifespan)
+
+oarch_lifespan |>
+  arrange(-lifespan) |>
+  slice(1:10) |>
+  pull(lifespan) |>
+  range() ->
+  lifespan_top10_range
+
+max_commit_rate <- max(oarch_lifespan$commit_rate, na.rm = TRUE)
+p_commit_rate_1 <- nrow(filter(oarch_lifespan, commit_rate < 1)) / nrow(oarch_lifespan)
 ```
@@
 We measured the lifespan of a repository as the time between the first and latest commit, and its activity as the rate of commits.
-Here therefore refer to the *development* lifespan of a project—measured here as the time between a repository's first and last commit—which is not necessary linked to its use-life.
-The lifespan and activity of repositories in open-archaeo vary greatly (@fig-lifespan).
-The average project lasts <!-- TODO --> days with <!-- TODO --> commits per day.
+Here therefore we refer to the *development* lifespan of a project, which is not necessary related to its use-life.
+By these metrics, the lifespan and activity of repositories in open-archaeo vary greatly (@fig-lifespan).
+The average project lasts `r round(mean(oarch_lifespan$lifespan))` days with `r round(mean(oarch_lifespan$commit_rate, na.rm = TRUE), 2)` commits per day.
 <!-- TODO: turn all these percentages into code -->
-Many projects are active for only a short period of time: about 15% less than a month, 25% less than three months, and 40% less than a year.
-The longest are <!-- TODO -->.
-Commit rates vary between <!-- TODO -->, with peaks at <!-- TODO -->.
+Many projects are active for only a short period of time: about `r percent(p_lifespan_30days)` less than 30 days, `r percent(p_lifespan_90days)` less than 90 days, and `r percent(p_lifespan_365days)` less than a year.
+However, the vast majority (all but `r english(n_1commit)`) do have more than one commit, suggesting that use of GitHub as a pure host for already-finished projects is not common; some degree of iteration, if not collaboration, is almost always present.
+The longest-lived projects have been active for between `english(round(lifespan_top10_range / 365)[1])` and `english(round(lifespan_top10_range / 365)[2])` years.
+The most active projects see up to `round(max_commit_rate)` commits per day, but the majority of repositories (`r percent(p_commit_rate_1)`) receive less one commit per day.
 

### 2023-10-12 analysis/openarchaeo_collaboration.qmd
Add statistics on lifespan and longevity

@@
 
+library("english")
 library("fs")

### 2023-10-12 analysis/_05-collaboration.qmd
Fix inline chunk syntax

@@
 However, the vast majority (all but `r english(n_1commit)`) do have more than one commit, suggesting that use of GitHub as a pure host for already-finished projects is not common; some degree of iteration, if not collaboration, is almost always present.
-The longest-lived projects have been active for between `english(round(lifespan_top10_range / 365)[1])` and `english(round(lifespan_top10_range / 365)[2])` years.
-The most active projects see up to `round(max_commit_rate)` commits per day, but the majority of repositories (`r percent(p_commit_rate_1)`) receive less one commit per day.
+The longest-lived projects have been active for between `r english(round(lifespan_top10_range / 365)[1])` and `r english(round(lifespan_top10_range / 365)[2])` years.
+The most active projects see up to `r round(max_commit_rate)` commits per day, but the majority of repositories (`r percent(p_commit_rate_1)`) receive less one commit per day.
 

### 2023-10-12 analysis/_05-collaboration.qmd
Add stats on issue/comment use

@@
 # Collaborative practices
+
+```{r generate-oarch_github}
+# Subset of dataset with GitHub data
+oarch |> 
+  filter(!is.na(gh_repo_info)) ->
+  oarch_gh
+```
 
@@
 #TODO: Remove that one commit in 2001
-oarch %>%
+oarch_gh %>%
   unnest(gh_commits) %>%
@@
 
-```{r generate-oarch-lifespan}
-oarch %>% 
+```{r generate-oarch_lifespan}
+oarch_gh %>% 
   unnest(gh_commits) %>% 
@@
 # TODO: why are there no twos?
-fig_issues_hist <- oarch %>%
+# TODO: throw stars in here too?
+fig_issues_hist <- oarch_gh %>%
   unnest(gh_issues) %>%
@@
 
-fig_comments_hist <- oarch %>%
+fig_comments_hist <- oarch_gh %>%
   unnest(gh_comments) %>%
@@
 
-fig_issues_hist + fig_comments_hist
+fig_issues_hist / fig_comments_hist
 ```
 
-Perhaps as a result of this, we also find that only a minority of projects make much use of collaborative features of GitHub such as issues and pull requests (left) and comments (right).
-Only 39% of repositories make use of these features.
-Of these repositories with any issues and pull requests, 28% have only 1 and 82% have less than 10.
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity.
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
+```{r generate-issue-comment-stats}
+oarch_gh <- mutate(
+  oarch_gh, 
+  n_issues = map_int(gh_issues, \(x) if(is.data.frame(x)) nrow(x) else 0),
+  n_comments = map_int(gh_comments, \(x) if (is.data.frame(x)) nrow(x) else 0)
+)
+
+p_with_issues <- nrow(filter(oarch_gh, n_issues > 0)) / nrow(oarch_gh)
+p_with_one_issue <- nrow(filter(oarch_gh, n_issues == 1)) / nrow(filter(oarch_gh, n_issues > 0))
+p_with_few_issues <- nrow(filter(oarch_gh, n_issues > 0 & n_issues <= 10)) / nrow(filter(oarch_gh, n_issues > 0))
+```
 
@@
 GitHub's pull request feature is also implemented via this system – a pull request is a special type of issue.
+
+Only `r percent(p_with_issues)` of open-archaeo projects make use of issues/pull requests at all (@fig-issues-comments).
+Of those that do, they are not used very heavily; `r percent(p_with_one_issue)` have only one issue and `r percent(p_with_few_issues)` have ten or less. 
+<!-- TODO: clarify what we mean here? issues tend to have a lot of comments? or active repos have active comment sections? (the latter being a bit of a redundant observation) -->
+However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity. 
+This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
 

### 2023-10-12 analysis/_03-data_methodology.qmd
Add tbl-glossary

@@
 
+```{r tbl-glossary}
+#| tbl-cap: Glossary of git and GitHub terminology
+tribble(
+  ~term, ~definition,
+  "Comment", "Text post attached to an issue, including the first one that describes the issue",
+  "Commit", "Set of changes (addition, alteration, or deletion) to files in a repository that has been recorded by git as one entry in its log",
+  "Commit access", "Ability to make changes to a repository directly, without making a pull request",
+  "Contributor", "User that has made at least one commit to a specified repository",
+  "Follow", "Add activity by another user to a user's timeline",
+  "git", "Open source version control software",
+  "GitHub", "Commercial platform that freely hosts git repositories and provides extended collaboration and social networking features, such as pull requests, issues and stars",
+  "GitLab", "Open source alternative to GitHub",
+  "CodeBerg", "Open source alternative to GitHub",
+  "Issue", "Feature of GitHub that records and tracks a bug report, feature request or other suggestion in a repository",
+  "Maintainer", "Individual that has overall control of a repository, generally assumed to be its primary contributor.
+Repositories can have multiple users with commit access in addition to the maintainer.",
+  "Merge", "Accept a pull request and incorporate its changes into a repository.",
+  "Organization", "Entity representing a group of users, which can also own repositories",
+  "Pull request", "Mechanism by which users that don't have commit access to a repository can contribute to it.
+The repository's maintainer or another user with commit access must decide whether to merge (accept) the changes, or decline them.",
+  "Repository", "Individual project that uses git for version control. Can include a mix of different types of files.",
+  "Star", "GitHub's version of a 'like', applied by users to a repository.",
+  "Timeline", "Chronological feed of GitHub activity from repositories a user has starred and other users they follow.
+Also includes repositories that a user is not following if they are 'trending' or determined relevant by GitHub's algorithm.",
+  "User", "On GitHub, an individual with an account that can own repositories",
+  "Version control", "System for tracking changes (additions, alterations, or deletions) in a set of files, typically but not exclusively computer code"
+) |>
+  gt() |>
+  cols_label_with(fn = tools::toTitleCase) |>
+  fmt_markdown(definition)
+```
+
 <!-- TODO:
 * Introduce git, GitHub, etc. and describe briefly how they work for non-users
-* Define software jargon we use (repo, commit, etc.) – maybe as a glossary table?
 -->

### 2023-10-12 analysis/_05-collaboration.qmd
Add tbl-glossary

@@
 As well as hosting source code, GitHub and other software forges include systems for facilitating collaboration on code and other projects.
-The basic collaborative workflow is inherited from git, which allows multiple users to *commit* (see @tbl-glossary for definitions of this and other git terminology used in this section) code to the repository.
-A user with *commit access* to a repository can change any of its contents at will, so this is usually reserved for the project maintainer and known, trusted collaborators.
-GitHub extends this model with its *pull request* feature, by which any user can offer to contribute code to a repository to which they don't have commit access.
-The maintainer may then choose to *merge* (accept) or *decline* the pull request, facilitating contributions from a wider network of collaborators without the need for permission to be sought in advance.
+The basic collaborative workflow is inherited from git, which allows multiple users to commit (see @tbl-glossary for definitions of this and other git terminology used in this section) code to the repository.
+A user with commit access to a repository can change any of its contents at will, so this is usually reserved for the project maintainer and known, trusted collaborators.
+GitHub extends this model with its pull request feature, by which any user can offer to contribute code to a repository to which they don't have commit access.
+The maintainer may then choose to merge (accept) or decline the pull request, facilitating contributions from a wider network of collaborators without the need for permission to be sought in advance.
 
@@
 
-GitHub also facilitates collaboration on broader project management tasks, primarily through its *issues* feature.^[Apart from issues, GitHub has a very wide range of project management and social media-like features, including wikis, discussion forums and 'kanban' boards.
+GitHub also facilitates collaboration on broader project management tasks, primarily through its issues feature.^[Apart from issues, GitHub has a very wide range of project management and social media-like features, including wikis, discussion forums and 'kanban' boards.
 We have not analysed the use of these features here.]
@@
 
-Another way GitHub users can engage with repositories and other users with social media-like features such as *starring* a repository, adding it to a public or private list, or *following* a user.
+Another way GitHub users can engage with repositories and other users with social media-like features such as starring a repository, adding it to a public or private list, or following a user.
 These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This feature of GitHub's timeline was one of the primary ways we compiled open-archaeo.]

### 2023-10-12 analysis/_03-data_methodology.qmd
Note TODO in methods section
Update tbl-glossary

@@
   ~term, ~definition,
-  "Comment", "Text post attached to an issue, including the first one that describes the issue",
+  "Comment", "On GitHub, Text post attached to an issue, including the first one that describes the issue",
   "Commit", "Set of changes (addition, alteration, or deletion) to files in a repository that has been recorded by git as one entry in its log",
@@
 
+<!-- TODO: add something on EDA and network analysis methodology -->
+
 The data and analysis code are available in our research compendium on GitHub, which is also indexed on Zenodo.

### 2023-10-12 analysis/_03-data_methodology.qmd
Expand GitHub methodology

@@
 Our study comprises an exploratory analysis of open-archaeo [@batist2023], a list of open source archaeological software and other digital resources.
-As of right now it includes 408 items, which are rendered for the web at [open-archaeo.info](open-archaeo.info).
+As of right now it includes <!-- TODO: encode -->408 items, which are rendered for the web at [open-archaeo.info](open-archaeo.info).
 We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, and by tracing connections to other personal, professional, and institutional websites that describe and host additional archaeological software.
@@
   "Follow", "Add activity by another user to a user's timeline",
+  "Forge", "Web-based platform for hosting, distributing and facilitating collaboration on version controlled computer code, e.g. GitHub, GitLab, Codeberg",
   "git", "Open source version control software",
@@
 
-Open-archaeo is a relatively comprehensive list, however it generally overlooks records pertaining to code written before archaeologists started using collaborative software development platforms such as GitHub.
+Open-archaeo is a relatively comprehensive list.
+While our initial intention was to only list open source software, its scope has expanded to include all software created by and for archaeologists.
+Apart from regular updates by its primary maintainer (ZB), it has been expanded by a wider network of contributors and has benefited from the wider range of domain specialisms this has brought.
+However, there are a number of areas that, for practical reasons, are not well-covered.
+<!-- TODO: next three sentences are a bit repeatative, can we more concisely state what OA doesn't cover well? -->
+Code written before archaeologists started using collaborative software development platforms such as GitHub is generally missing.
 It also lacks references to code that is not available on the web.
@@
 
-For the purpose of this study, we augmented the core open-archaeo dataset with data obtained from GitHub's public API.
-That is, for each project in open-archaeo with a GitHub repository, we added information on individual contributors, commits, issues, etc. from the API using the R package `gh` [@csardi2023].
-This does reduce our sample size slightly; about an eighth of projects don't use version control, at least not publicly, but among those that do, GitHub is by far the most popular host ([@tbl-repo-hosts]).
-
 ```{r data-github}
@@
 
-```{r tbl-repo-hosts}
-#| tbl-cap: Number of open archaeology projects using GitHub and other repository hosts
-oarch |>
+```{r data-oarch_forges}
+oarch_forges <- oarch |>
   mutate(none = ifelse(is.na(github) & is.na(gitlab) & is.na(bitbucket) &
@@
   pivot_wider(names_from = category, values_from = n, values_fill = 0) |>
-  mutate(total = rowSums(across(Guides:`Standalone software`)), 
-         .before = Guides) |>
+  mutate(
+    total = rowSums(across(Guides:`Standalone software`)), 
+    p_total = total / sum(total),
+    .before = Guides
+  ) |>
   arrange(repo_host == "none", desc(total)) |>
@@
                             launchpad = "Launchpad",
-                            none = "None")) |>
-  gt() |>
+                            none = "None"))
+
+p_no_vc <- filter(oarch_forges, repo_host == "None")[["p_total"]]
+p_no_github <- sum(filter(oarch_forges, repo_host != "None" & repo_host != "GitHub")$p_total)
+```
+
+```{r tbl-forges}
+#| tbl-cap: Software forges used by open-archaeo projects
+gt(oarch_forges) |>
   cols_label(repo_host = "Host", total = "Total") |>
+  cols_merge_n_pct(total, p_total) |>
+  fmt_percent(p_total) |>
   tab_spanner("By category", Guides:`Standalone software`)
@@
 
-While our initial intention was to only list open source software, open-archaeo's scope has expanded to include all software created by and for archaeologists.
+We augmented the open-archaeo's list of with projects with information on its activity on GitHub, where applicable.
+This included metadata on the GitHub repository and the programming languages it uses, as well as detailed data on each repository's individual contributors, commits, issues and comments (see @tbl-glossary for a glossary of git/GitHub terminology used in this paper).
+This data was collected via GitHub's public API using the R package `gh` [@csardi2023].
+We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
+This means that projects that don't use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data included in open-archaeo itself.
 
@@
 The data and analysis code are available in our research compendium on GitHub, which is also indexed on Zenodo.
-
-```{r data-oarch_contributors}
-oarch |>
-  unnest(gh_contribs) |>
-  drop_na(contributor) |>
-  group_by(contributor) |>
-  mutate(
-    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
-  ) |>
-  group_by(contributor, repo_owner) |>
-  summarise(repos = n(), contribs = sum(contributions)) |>
-  pivot_wider(c(contributor), names_from = repo_owner, 
-              values_from = c(repos, contribs), values_fill = 0) |>
-  mutate(
-    repos = repos_other + repos_own,
-    contribs = contribs_other + contribs_own,
-    .after = contributor
-  ) ->
-  oarch_contributors
-```

### 2023-10-12 analysis/_05-collaboration.qmd
Expand GitHub methodology

@@
 
-```{r generate-oarch_github}
+```{r data-oarch_github}
 # Subset of dataset with GitHub data
@@
 
-```{r generate-oarch_lifespan}
+```{r data-oarch_lifespan}
 oarch_gh %>% 
@@
 
-```{r generate-issue-comment-stats}
+```{r data-issue-comment-stats}
 oarch_gh <- mutate(

### 2023-10-12 analysis/_06-network.qmd
Expand GitHub methodology

@@
 -->
+
+```{r data-oarch_contributors}
+oarch |>
+  unnest(gh_contribs) |>
+  drop_na(contributor) |>
+  group_by(contributor) |>
+  mutate(
+    repo_owner = if_else(str_extract(repo, "[^/]+") == contributor, "own", "other")
+  ) |>
+  group_by(contributor, repo_owner) |>
+  summarise(repos = n(), contribs = sum(contributions)) |>
+  pivot_wider(c(contributor), names_from = repo_owner, 
+              values_from = c(repos, contribs), values_fill = 0) |>
+  mutate(
+    repos = repos_other + repos_own,
+    contribs = contribs_other + contribs_own,
+    .after = contributor
+  ) ->
+  oarch_contributors
+```
 

### 2023-10-12 analysis/_01-introduction.qmd
Mostly writing uodates.
Updated/corrected bibliographic references. Updated certain parts of the text for clarity. Some other scattered notes and superficial changes.

@@
 Despite this, the open science movement, within which open archaeology emerged, draws direct inspiration from open source.
-For instance, the Open Knowledge Foundation [-@openknowledgefoundation2023] publishes a widely accepted definition of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group.
+For instance, the Open Knowledge Foundation [-@openknowledgefoundation2015] publishes a widely accepted definition of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group.
 The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plain text, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic).
@@
 While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests.
-In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers [@kelty2008 xx].
+In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers [@ratto2007; @kelty2008 chapter 9].
 
@@
 --JR -->That being said, archaeologists are prolific software developers.
-This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes XXX records as of the time of publication.
+This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes <!-- TODO: encode -->XXX records, as of the time of this article's publication.
 But beyond simply making their code available on the web, do archaeologists also implement social strategies to advance open source ideals?
@@
 
-(1) we identify *what* kinds of software archaeologists are making;
-(2) and evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
+(1) we identify *what* kinds of software archaeologists are making; and
+(2) we evaluate *how* archaeologists create these tools, with particular emphasis on practices of collaboration.
 

### 2023-10-12 analysis/_02-background.qmd
Mostly writing uodates.
Updated/corrected bibliographic references. Updated certain parts of the text for clarity. Some other scattered notes and superficial changes.

@@
 By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base.
-Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision [@kelty2008 on the emacs saga].
+Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision [see @kelty2008 on the emacs saga for a quintessential example].
 An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.
 
-Open source has traditionally been referred to as being based on meritocratic principles.
+Open source has traditionally been referred to as being based on meritocratic principles [@raymond1999 39].
 A good test of whether a contribution should be included in a published software release is whether it is functional [@kelty2008 220].
-Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues.
+Moreover, with more eyes looking over a code base, it is easier to identify flaws with a contribution and flag potential bugs or security issues [@raymond1999 27-30].
 This is all done in the spirit of producing functional code, and in ideal circumstances faulty contributions will be corrected before inclusion.
-Personal ego is minimized in favour of co-creating stable and functional outcomes [@coleman2012; @oneil2009 on the Debian saga].
+Personal ego is minimized in favour of co-creating stable and functional outcomes [@raymond1999 39-41].
 
@@
 They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term.
-Open source is more than just putting your code online; to be successful, it requires participation in a social experience [@kelty2008, Ratto dissertation and derived works].
+Open source is more than just putting your code online; to be successful, it requires participation in a social experience [@ratto2003; @kelty2008].
 
@@
 The open science movement comprises a series of practices and principles intended to make research more accessible, transparent, and efficient.
-Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" [@openknowledgefoundation2023].
+Although the concept of "open" is somewhat nebulous, in terms of its abstract definition and with regards to what real-world applications count as being open, one commonly-cited definition describes content that "can be freely used, modified, and shared by anyone for any purpose" [@openknowledgefoundation2015].
 This definition does not state what open is for, how to be open, or any sort of social or discursive framing behind the open movement.
@@
 [-@faniel2013 299-301], Atici et al.
-[-@atici2013 676-677], @huggett2018, @sobotkova2018, and @opitz2021 demonstrate that to make the reuse of archaeological data feasible and useful in a practical sense, it is necessary to re-introduce social friction that these infrastructures are designed to eliminate.
+[-@atici2013 676-677], Huggett [-@huggett2018; -@huggett2022], @sobotkova2018, @opitz2021, @haciguzeller2021, and @batist2023a demonstrate that to make the reuse of archaeological data feasible and useful in a practical sense, it is necessary to re-introduce social friction that these infrastructures are designed to eliminate.
 In other words, the pressures and circumstances of being an archaeologist and doing archaeological research assert themselves when attempting to make practical use of these infrastructures, and therefore must be accounted for in their design and implementation.

### 2023-10-12 analysis/_03-data_methodology.qmd
Mostly writing uodates.
Updated/corrected bibliographic references. Updated certain parts of the text for clarity. Some other scattered notes and superficial changes.

@@
 Apart from regular updates by its primary maintainer (ZB), it has been expanded by a wider network of contributors and has benefited from the wider range of domain specialisms this has brought.
-However, there are a number of areas that, for practical reasons, are not well-covered.
-<!-- TODO: next three sentences are a bit repeatative, can we more concisely state what OA doesn't cover well? -->
-Code written before archaeologists started using collaborative software development platforms such as GitHub is generally missing.
-It also lacks references to code that is not available on the web.
-Open-archaeo is also limited by the experiences of its primary maintainers, and we welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to contribute.
+However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
+The dataset is also limited by the experiences of its primary maintainers.
+We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
 
@@
 
-We augmented the open-archaeo's list of with projects with information on its activity on GitHub, where applicable.
+Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API using the R package `gh` [@csardi2023].
 This included metadata on the GitHub repository and the programming languages it uses, as well as detailed data on each repository's individual contributors, commits, issues and comments (see @tbl-glossary for a glossary of git/GitHub terminology used in this paper).
-This data was collected via GitHub's public API using the R package `gh` [@csardi2023].
 We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
-This means that projects that don't use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data included in open-archaeo itself.
+This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data included in open-archaeo itself.
 
@@
 
-The data and analysis code are available in our research compendium on GitHub, which is also indexed on Zenodo.
+The data and analysis code are available in our research compendium on GitHub, which is also indexed on Zenodo. <!-- TODO: add hyperlinks to these resources -->

### 2023-10-12 analysis/_04-open_archaeology.qmd
Mostly writing uodates.
Updated/corrected bibliographic references. Updated certain parts of the text for clarity. Some other scattered notes and superficial changes.

@@
 
-`r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used 'platform' by a large margin (@tbl-platforms).
+`r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used platform by a large margin (@tbl-platforms).
 Python, another programming language, is also relatively popular <!-- TODO: percent -->, as are plugins for the open source geographic information system QGIS <!-- TODO: percent -->.
@@
 
-We also annotated each record with 'tags' that describe aspects of archaeological work that each tool contributes to [@fig-tags].
+We also annotated each record with "tags" that describe aspects of archaeological work that each tool contributes to [@fig-tags].
 The most common tags unsurprisingly deal with work that naturally benefits from advanced information processing afforded by computers, such as statistical analysis, sample calibration, geographical analysis, data management, and chronological modelling.

### 2023-10-12 analysis/_05-collaboration.qmd
Mostly writing uodates.
Updated/corrected bibliographic references. Updated certain parts of the text for clarity. Some other scattered notes and superficial changes.

@@
 ```{r fig-issues-comments}
-#| fig-cap: Use of issues and comments
+#| fig-cap: Use of issues and comments.
 # TODO: include zeroes
@@
 ```{r plot-gh_contributors-histogram}
+#| fig-cap: Number of contributors.
 oarch %>%
@@
 
-* How many people are involved?
-  *  Conversely, how many projects are individual people involved in?
-
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\].
+Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\] (@plot-gh_contributors-histogram).
 A "contribution", in this case, is a commit, issue, or comment applied by a user to a repository.
@@
 ```{r plot-contrib-dist}
+#| fig-cap: Distribution of contributions in multi-contributor projects.
 #TODO: Re-calculate these numbers to ensure they are up to date.
@@
 
-Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly.
+Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly (@plot-contrib-dist).
 Of multi-contributor repositories, 92% have a third of contributions made by a single person, 65% have half of all contributions made by a single person, and 36% have three quarters of all contributions made a single person.
@@
 Conversely, a small number of projects (<!-- TODO: do these share any characteristics in terms of platform or subject?
--->) do appear to fit the aspirational open source model, with a large number of contributors maintaing a long-lived and highly-active repository.
+-->) do appear to fit the aspirational open source model, with a large number of contributors maintaining a long-lived and highly-active repository.
 Archaeologists active on GitHub also make markedly <!-- TODO: verify --> greater use of GitHub's social media-like features (stars, following) than more direct collaborative actions.

### 2023-10-12 analysis/openarchaeo_collaboration.qmd
Mostly writing uodates.
Updated/corrected bibliographic references. Updated certain parts of the text for clarity. Some other scattered notes and superficial changes.

@@
 
-  In this paper, we investigate modes of collaboration in this emerging community of practice using open-archaeo (https://open-archaeo.info/), a curated list of 500+ active open projects in archaeology, augmented where available with data on their activity on GitHub.
+  In this paper, we investigate modes of collaboration in this emerging community of practice using open-archaeo [open-archaeo.info](open-archaeo.info), a curated list of 500+ active open projects in archaeology, augmented where available with data on their activity on GitHub.
 We apply network analysis to map the networks of collaboration that have emerged among archaeologists using GitHub, and conduct a quantitative exploratory analysis to characterize the nature and intensity of these collaborations.

### 2023-10-12 analysis/_03-data_methodology.qmd
Added description of git and github
Also tried to integrate this with our data collection methods. Feel free to make changes / rip into this, since it's just an initial draft.

@@
 As of right now it includes <!-- TODO: encode -->408 items, which are rendered for the web at [open-archaeo.info](open-archaeo.info).
-We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, and by tracing connections to other personal, professional, and institutional websites that describe and host additional archaeological software.
-This entailed manually crawling through users' profiles, particularly those who identify as archaeologists or who have contributed to tools that pertain to archaeological work.
-We supplemented this quasi-systematic collection strategy with word-of-mouth contributions made by interested individuals who identified relevant work that we initially overlooked.
+
+We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, which are designed to facilitate open source software development by hosting public git repositories.
+Git is a protocol designed to facilitate open and distributed contributions to a common code base. <!-- mention other version control systems like SVN and mercurial maybe? -->
+It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
+Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a branch in which they apply their modifications, and then request that their branch be merged into the central code base.
+After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent branches intact.
+
+Git is also designed to facilitate code review.
+All modifications are tracked as "diffs", which highlight additions or deletions to the code base, including changes within individual files.
+Typically, a contributor will group a series of changes into a more comprehensive "commit" based on a specific task or part of a workflow.
+Commits are always accompanied by a message, in which the contributor (ideally) describes the reason and context for the changes included in the commit.
+Moreover, git assigns each commit a unique identifier and identifies the contributor by name and email address to ensure some degree of public accountability.
+
+In addition to serving as hosting providers for public git repositories, platforms like as GitHub, GitLab and Codeberg support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
+They also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, identify certain repositories as "favourites" and maintain a public-facing profile that includes personally-identifying information (e.g.
+profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
+Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
+However, these additional features are meant to complement and enhance the experience of contributing to open source projects, but are not actually part of the git protocol.
+See @tbl-glossary for a glossary of terminology used to describe many of these concepts and features, and which we use throughout this paper.
+
+We rely heavily on the social networking features provided by GitHub, GitLab and Codeberg to maintain open-archaeo.
+More specifically, we update open-archaeo by manually crawling through archaeologists' profiles on these platforms, as well as on other personal, professional, and institutional websites that describe and host additional archaeological software.
+We supplement this quasi-systematic collection strategy with word-of-mouth contributions made by interested individuals who identified relevant work that we initially overlooked.
+
+Open-archaeo is a relatively comprehensive list.
+While our initial intention was to only list open source software, its scope has expanded to include all software created by and for archaeologists.
+Apart from regular updates by its primary maintainer (ZB), it has been expanded by a wider network of contributors and has benefited from the wider range of domain specialisms this has brought.
+However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
+The dataset is also limited by the experiences of its primary maintainers.
+We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
 
@@
 ```
-
-<!-- TODO:
-* Introduce git, GitHub, etc. and describe briefly how they work for non-users
--->
-
-Open-archaeo is a relatively comprehensive list.
-While our initial intention was to only list open source software, its scope has expanded to include all software created by and for archaeologists.
-Apart from regular updates by its primary maintainer (ZB), it has been expanded by a wider network of contributors and has benefited from the wider range of domain specialisms this has brought.
-However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
-The dataset is also limited by the experiences of its primary maintainers.
-We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
 
@@
 Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API using the R package `gh` [@csardi2023].
-This included metadata on the GitHub repository and the programming languages it uses, as well as detailed data on each repository's individual contributors, commits, issues and comments (see @tbl-glossary for a glossary of git/GitHub terminology used in this paper).
+This included metadata on the GitHub repository and the programming languages it uses, as well as detailed data on each repository's individual contributors, commits, issues and comments.
 We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).

### 2023-10-13 analysis/_03-data_methodology.qmd
Fill in blanks in section 3

@@
 
-<!--
-* open-archaeo
-* GitHub API
-* Indicate how much is on GitHub or elsewhere
-  -->
+```{r data-oarch}
+read_csv("https://github.com/zackbatist/open-archaeo/raw/2023-08-29/open-archaeo.csv") |>
+  mutate(category = na_if(category, "?")) ->
+  oarch
 
-<!-- TODO: snappier title? Maybe incorporating 'open-archaeo'? -->
+# Extract clean repos from GitHub URLs
+oarch |>
+  mutate(
+    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
+  ) ->
+  oarch
 
-Our study comprises an exploratory analysis of open-archaeo [@batist2023], a list of open source archaeological software and other digital resources.
-As of right now it includes <!-- TODO: encode -->408 items, which are rendered for the web at [open-archaeo.info](open-archaeo.info).
+# Amalgamate categories into fewer rough categories
+# TODO: redo
+oarch <- mutate(oarch, rough_category = recode(category,
+                                 "R" = "Packages (R, Python, etc.)",
+                                 "Web apps" = "Standalone apps",
+                                 "Datasets" = "Analyses & datasets",
+                                 "Tutorials" = "Lists & tutorials",
+                                 "Python" = "Packages (R, Python, etc.)",
+                                 "Desktop apps" = "Standalone apps",
+                                 "QGIS" = "Plugins",
+                                 "Lists" = "Lists & tutorials",
+                                 "Mobile apps" = "Standalone apps",
+                                 "Analyses" = "Analyses & datasets",
+                                 "Spreadsheets" = "Forms & spreadsheets",
+                                 "Command line tools" = "Standalone apps",
+                                 "Plugins" = "Plugins",
+                                 "ArcGis" = "Plugins",
+                                 "Filemaker" = "Plugins",
+                                 "ODK" = "Forms & spreadsheets",
+                                 "Lisp" = "Packages (R, Python, etc.)",
+                                 .default = "Other",
+                                 .missing = "Other"
+                                 ))
 
-We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, which are designed to facilitate open source software development by hosting public git repositories.
-Git is a protocol designed to facilitate open and distributed contributions to a common code base. <!-- mention other version control systems like SVN and mercurial maybe? -->
-It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
-Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a branch in which they apply their modifications, and then request that their branch be merged into the central code base.
-After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent branches intact.
+# TODO: exclude peripheral repos that throw things off
+# * Big boys like QField (see _06-network.qmd)
+# * ArchaeoLines – ends up picking up the whole stellarium repo, which is huge
+# * aDNA??
+```
 
-Git is also designed to facilitate code review.
-All modifications are tracked as "diffs", which highlight additions or deletions to the code base, including changes within individual files.
-Typically, a contributor will group a series of changes into a more comprehensive "commit" based on a specific task or part of a workflow.
-Commits are always accompanied by a message, in which the contributor (ideally) describes the reason and context for the changes included in the commit.
-Moreover, git assigns each commit a unique identifier and identifies the contributor by name and email address to ensure some degree of public accountability.
-
-In addition to serving as hosting providers for public git repositories, platforms like as GitHub, GitLab and Codeberg support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
-They also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, identify certain repositories as "favourites" and maintain a public-facing profile that includes personally-identifying information (e.g.
-profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
-Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
-However, these additional features are meant to complement and enhance the experience of contributing to open source projects, but are not actually part of the git protocol.
-See @tbl-glossary for a glossary of terminology used to describe many of these concepts and features, and which we use throughout this paper.
-
-We rely heavily on the social networking features provided by GitHub, GitLab and Codeberg to maintain open-archaeo.
-More specifically, we update open-archaeo by manually crawling through archaeologists' profiles on these platforms, as well as on other personal, professional, and institutional websites that describe and host additional archaeological software.
-We supplement this quasi-systematic collection strategy with word-of-mouth contributions made by interested individuals who identified relevant work that we initially overlooked.
-
-Open-archaeo is a relatively comprehensive list.
-While our initial intention was to only list open source software, its scope has expanded to include all software created by and for archaeologists.
-Apart from regular updates by its primary maintainer (ZB), it has been expanded by a wider network of contributors and has benefited from the wider range of domain specialisms this has brought.
-However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
-The dataset is also limited by the experiences of its primary maintainers.
-We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
+We present an exploratory analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources.
 
@@
 
-```{r data-oarch}
-read_csv("https://github.com/zackbatist/open-archaeo/raw/2023-08-29/open-archaeo.csv") %>%
-  mutate(category = na_if(category, "?")) ->
-  oarch
+We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, which are designed to facilitate open source software development by hosting public git repositories.
+Git is a protocol designed to facilitate open and distributed contributions to a common code base. <!-- mention other version control systems like SVN and mercurial maybe? -->
+It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
+Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a branch in which they apply their modifications, and then request that their branch be merged into the central code base.
+After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent branches intact.
 
-# Extract clean repos from GitHub URLs
-oarch %>%
-  mutate(
-    gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
+Git is also designed to facilitate code review.
+All modifications are tracked as "diffs", which highlight additions or deletions to the code base, including changes within individual files.
+Typically, a contributor will group a series of changes into a more comprehensive "commit" based on a specific task or part of a workflow.
+Commits are always accompanied by a message, in which the contributor (ideally) describes the reason and context for the changes included in the commit.
+Moreover, git assigns each commit a unique identifier and identifies the contributor by name and email address to ensure some degree of public accountability.
 
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
+In addition to serving as hosting providers for public git repositories, platforms like as GitHub, GitLab and Codeberg support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
+They also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, identify certain repositories as "favourites" and maintain a public-facing profile that includes personally-identifying information (e.g.
+profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
+Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
+However, these additional features are meant to complement and enhance the experience of contributing to open source projects, but are not actually part of the git protocol.
+See @tbl-glossary for a glossary of terminology used to describe many of these concepts and features, and which we use throughout this paper.
 
-# TODO: exclude peripheral repos that throw things off
-# * Big boys like QField (see _06-network.qmd)
-# * ArchaeoLines – ends up picking up the whole stellarium repo, which is huge
-# * aDNA??
-```
+We rely heavily on the social networking features provided by GitHub, GitLab and Codeberg to maintain open-archaeo.
+More specifically, we update open-archaeo by manually crawling through archaeologists' profiles on these platforms, as well as on other personal, professional, and institutional websites that describe and host additional archaeological software.
+We supplement this quasi-systematic collection strategy with word-of-mouth contributions made by interested individuals who identified relevant work that we initially overlooked.
+
+Open-archaeo is a relatively comprehensive list.
+While our initial intention was to only list open source software, its scope has expanded to include all software created by and for archaeologists.
+Apart from regular updates by its primary maintainer (ZB), it has been expanded by a wider network of contributors and has benefited from the wider range of domain specialisms this has brought.
+However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
+The dataset is also limited by the experiences of its primary maintainers.
+We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
 
@@
 
-The data and analysis code are available in our research compendium on GitHub, which is also indexed on Zenodo. <!-- TODO: add hyperlinks to these resources -->
+The quantitative analyses and figures presented here were generated with `r version$version.string` (@Manual).
+The full data and code is available in the compendium that accompanies this paper (@Compendium).

### 2023-10-13 analysis/_03-data_methodology.qmd
Revise section 5
Getting close!

@@
   "Forge", "Web-based platform for hosting, distributing and facilitating collaboration on version controlled computer code, e.g. GitHub, GitLab, Codeberg",
+  "Fork", "Copy of a repository owned by another user; forking is a prerequisite to making a pull request",
   "git", "Open source version control software",

### 2023-10-13 analysis/_05-collaboration.qmd
Revise section 5
Getting close!

@@
 A user with commit access to a repository can change any of its contents at will, so this is usually reserved for the project maintainer and known, trusted collaborators.
-GitHub extends this model with its pull request feature, by which any user can offer to contribute code to a repository to which they don't have commit access.
-The maintainer may then choose to merge (accept) or decline the pull request, facilitating contributions from a wider network of collaborators without the need for permission to be sought in advance.
+GitHub extends this model with its pull request feature, by which any user can fork a repository to which they don't have commit access, make changes, then offer to contribute those changes back to the original repository.
+The maintainer can choose to merge (accept) or decline the pull request, facilitating contributions from a wider network of collaborators without the need for permission to be sought in advance.
 
 ```{r fig-lifespan}
-#| fig-cap: Lifespan of repositories based on the date of first and last commit. Points indicate intermediate commits. Only repositories with more than one commit are included.
-#TODO: This looks cool, but is it actually more useful than a histogram or two?
-#TODO: Remove that one commit in 2001
-oarch_gh %>%
-  unnest(gh_commits) %>%
-  drop_na(datetime) %>%
-  group_by(item_name) %>%
-  add_tally() %>%
-  filter(n > 1) %>%
+#| fig-cap: Lifespan of repositories. Each point indicates a commit; excludes repositories with only one commit.
+oarch_gh |>
+  unnest(gh_commits) |>
+  drop_na(datetime) |>
+  filter(datetime > ymd("2001-01-02")) |> # Apparent error
+  group_by(item_name) |>
+  add_tally() |>
+  filter(n > 1) |>
   mutate(
@@
     last_commit = max(datetime, na.rm = TRUE)
-  ) %>%
+  ) |>
   ggplot(aes(x = fct_reorder(item_name, last_commit, .desc = FALSE),
@@
   scale_y_datetime(date_breaks = "2 years", date_labels = "%Y") +
-  # coord_flip() +
   labs(x = NULL, y = NULL) +
@@
 The average project lasts `r round(mean(oarch_lifespan$lifespan))` days with `r round(mean(oarch_lifespan$commit_rate, na.rm = TRUE), 2)` commits per day.
-<!-- TODO: turn all these percentages into code -->
 Many projects are active for only a short period of time: about `r percent(p_lifespan_30days)` less than 30 days, `r percent(p_lifespan_90days)` less than 90 days, and `r percent(p_lifespan_365days)` less than a year.
@@
 
-```{r fig-issues-comments}
-#| fig-cap: Use of issues and comments.
-# TODO: include zeroes
-# TODO: why are there no twos?
-# TODO: throw stars in here too?
-fig_issues_hist <- oarch_gh %>%
-  unnest(gh_issues) %>%
-  drop_na(title) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(x = "Number of issues", y = "Repositories") +
-  theme_oarch()
+```{r data-collab-features}
+oarch_gh |> 
+  hoist(gh_repo_info, "stars", "forks", .remove = FALSE) |>
+  mutate(
+    issues = map_int(gh_issues, \(x) if(is.data.frame(x)) nrow(x) else 0),
+    comments = map_int(gh_comments, \(x) if (is.data.frame(x)) nrow(x) else 0)
+  ) ->
+  oarch_gh
 
-fig_comments_hist <- oarch_gh %>%
-  unnest(gh_comments) %>%
-  drop_na(issue_num) %>%
-  group_by(repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(x = "Number of comments", y = "Repositories") +
-  theme_oarch()
+p_with_issues <- n_filter(oarch_gh, issues > 0) / nrow(oarch_gh)
+p_with_one_issue <- n_filter(oarch_gh, issues == 1) / n_filter(oarch_gh, issues > 0)
+p_with_few_issues <- n_filter(oarch_gh, issues > 0 & issues <= 10) / n_filter(oarch_gh, issues > 0)
 
-fig_issues_hist / fig_comments_hist
+p_with_forks <- n_filter(oarch_gh, forks > 0) / nrow(oarch_gh)
+p_with_stars <- n_filter(oarch_gh, stars > 0) / nrow(oarch_gh)
+
+oarch_gh |> 
+  select(-description, -comments) |> 
+  unnest(gh_issues) |>
+  drop_na(comments) |>
+  count(comments > 0) |>
+  mutate(p = n / sum(n)) |>
+  filter(`comments > 0`) |>
+  pull(p) ->
+  p_issues_with_comments 
 ```
 
-```{r data-issue-comment-stats}
-oarch_gh <- mutate(
-  oarch_gh, 
-  n_issues = map_int(gh_issues, \(x) if(is.data.frame(x)) nrow(x) else 0),
-  n_comments = map_int(gh_comments, \(x) if (is.data.frame(x)) nrow(x) else 0)
-)
-
-p_with_issues <- nrow(filter(oarch_gh, n_issues > 0)) / nrow(oarch_gh)
-p_with_one_issue <- nrow(filter(oarch_gh, n_issues == 1)) / nrow(filter(oarch_gh, n_issues > 0))
-p_with_few_issues <- nrow(filter(oarch_gh, n_issues > 0 & n_issues <= 10)) / nrow(filter(oarch_gh, n_issues > 0))
+```{r fig-collab-features}
+#| fig-cap: Use of GitHub collaboration features (forks, issues, comments, stars) across repositories
+oarch_gh |>
+  pivot_longer(c(forks, issues, comments, stars), 
+               names_to = "feature", values_to = "n") |>
+  mutate(
+    feature = factor(feature, c("stars", "comments", "issues", "forks")),
+    feature = recode(feature, issues = "Issues and pull requests"),
+    feature = fct_relabel(feature, str_to_sentence)
+  ) |>
+  ggplot(aes(n, feature)) +
+  geom_boxplot() +
+  scale_x_log10() +
+  labs(x = NULL, y = NULL) +
+  theme_oarch()
 ```
@@
 GitHub's pull request feature is also implemented via this system – a pull request is a special type of issue.
+According to the data we collected from the GitHub API, these features are not widely used by open-archaeo projects (@fig-collab-features).
+Only `r percent(p_with_forks)` of repositories have been forked at least once and only `r percent(p_with_issues)` of repositories make use of issues/pull requests.
+Those repositories that do use issues do not not use them very extensively; `r percent(p_with_one_issue)` have only one issue and `r percent(p_with_few_issues)` have ten or less.
 
-Only `r percent(p_with_issues)` of open-archaeo projects make use of issues/pull requests at all (@fig-issues-comments).
-Of those that do, they are not used very heavily; `r percent(p_with_one_issue)` have only one issue and `r percent(p_with_few_issues)` have ten or less. 
-<!-- TODO: clarify what we mean here? issues tend to have a lot of comments? or active repos have active comment sections? (the latter being a bit of a redundant observation) -->
-However, of all repositories that have any comments posted in reply to issues and pull requests, there is actually a significant degree of activity. 
-This seems to suggest that people are willing to respond to issues raised by others, but may not necessarily want to take initiative, for a variety of reasons that may be more effectively investigated through qualitative analysis of the conversations in the issue threads.
+Another way GitHub users can engage with repositories and other users is with social media-like features such as starring a repository, commenting on an existing issue, or following a user.
+These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This feature of GitHub's timeline was one of the primary ways we compiled open-archaeo.]
+While not as a direct a contribution as pull requests or issues, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.
+These features are more used more widely than forks, issues and pull requests (@fig-collab-features): `r percent(p_with_stars)` of repositories have at least one star and, in those repositories that use issues, `r percent(p_issues_with_comments)` of them has received at least one additional comment.
 
-```{r plot-gh_contributors-histogram}
-#| fig-cap: Number of contributors.
-oarch %>%
-  drop_na(gh_repo) %>%
-  unnest(gh_contribs) %>%
-  group_by(gh_repo) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  ggplot(aes(x = n)) +
-  geom_histogram() +
-  scale_x_log10() +
-  labs(
-    title = "Repository contributors",
-    x = "Number of contributors", y = "Repositories",
-    caption = "Data: open-archaeo.info and GitHub API"
-  ) +
-  theme_oarch()
-```
-
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\] (@plot-gh_contributors-histogram).
-A "contribution", in this case, is a commit, issue, or comment applied by a user to a repository.
-<!-- TODO: expand -->
-
-```{r plot-contrib-dist}
-#| fig-cap: Distribution of contributions in multi-contributor projects.
-#TODO: Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic, which are now divided into quarters rather than thirds.
-oarch %>%
-  unnest(gh_contribs) %>%
-  drop_na(contributor) %>%
-  group_by(item_name) %>%
-  mutate(n_contributors = n()) %>%
-  filter(n_contributors > 1) %>%
+```{r fig-contributors}
+#| fig-cap: Distribution of contributions in multi-contributor projects
+oarch |>
+  unnest(gh_contribs) |>
+  drop_na(contributor) |>
+  group_by(item_name) |>
+  mutate(n_contributors = n()) |>
+  filter(n_contributors > 1) |>
   mutate(big_boy = ifelse(contributions == max(contributions),
-                          "Leading contributor", "Other contributors")) %>% 
-  group_by(item_name, big_boy) %>%
-  summarise(contributions = sum(contributions)) %>%
+                          "Leading contributor", "Other contributors")) |> 
+  group_by(item_name, big_boy) |>
+  summarise(contributions = sum(contributions)) |>
   mutate(
@@
     p_big_boy = first(p_contributions)
-  ) %>%
-  ungroup() %>%
-  mutate(item_name = fct_reorder(item_name, p_big_boy)) %>% 
+  ) |>
+  ungroup() |>
+  mutate(item_name = fct_reorder(item_name, p_big_boy, .desc = TRUE)) |> 
   ggplot(aes(x = item_name, y = contributions,
-             fill = fct_relevel(big_boy, c("Other contributors",
-                                           "Leading contributor")))) +
-  geom_col(position = "fill", colour = "white", size = 0.1) +
+             fill = ordered(big_boy, c("Other contributors", "Leading contributor")))) +
+  geom_col(position = "fill", linewidth = 0, width = 1) +
   geom_hline(yintercept = c(0.25, 0.5, 0.75), alpha = 0.5) +
   scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast(guide = guide_legend(reverse = TRUE)) +
-  labs(title = "Repository contributors by proportion of contributions",
-       x = NULL, y = "Contributions",
-       fill = NULL,
-       caption = "Data: open-archaeo.info and GitHub API") +
-  coord_flip() +
+  scale_fill_highcontrast() +
+  labs(x = "Individual repositories", y = "Contributions", fill = NULL) +
   theme_oarch() +
   theme(
-    axis.text.y = element_blank(),
-    axis.ticks.y = element_blank()
+    axis.text.x = element_blank(),
+    axis.ticks.x = element_blank(),
   )
@@
 
+<!-- TODO: introduce notion of contributors -->
+<!--
+#TODO: Re-calculate these numbers to ensure they are up to date.
+Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic, which are now divided into quarters rather than thirds.
+-->
+Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\] (@plot-gh_contributors-histogram).
+A "contribution", in this case, is a commit, issue, or comment applied by a user to a repository.
 Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly (@plot-contrib-dist).
 Of multi-contributor repositories, 92% have a third of contributions made by a single person, 65% have half of all contributions made by a single person, and 36% have three quarters of all contributions made a single person.
+<!-- TODO: expand description of contributors -->
 
-```{r fig-stars}
-#TODO: stars figure?
-```
-
-```{r fig-follows}
-#TODO: follows figure?
-```
-
-Another way GitHub users can engage with repositories and other users with social media-like features such as starring a repository, adding it to a public or private list, or following a user.
-These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This feature of GitHub's timeline was one of the primary ways we compiled open-archaeo.]
-While not constituting a direct form of collaboration, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.
-
-```{r fig-collaboration-ts}
-# TODO: combined time series of:
-# * average project longevity
-# * commit rate
-# * number of collaborators
-# * use of issues+comments
-```
-
-* Summarise: overall, what is the level of uptake of GitHub's collaboration features?
-  * Has it changed over time? (@fig-collaboration-ts)
-
-The prototypical collaborative open source project on GitHub has a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through pull requests, and an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
-The data we collected on open-archaeo repositories on GitHub show that uptake of its collaboration features is highly uneven, with very few projects resembling the open source ideal.
+The prototypical collaborative open source project on GitHub has a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through forks and pull requests, plus an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
+<!-- TODO: note that we don't know if maintainers actually aimed for that model, or leave that for the discussion? -->
+Only a small number of open-archaeo projects (<!-- TODO: do these share any characteristics in terms of platform or subject? -->) approach this model.
 The majority of projects are short-lived, with few contributors and a small number of commits.
-Only a fraction of repositories use GitHub's extended collaboration features (pull requests, issues, comments) and those that do use them in a limited way.
-<!-- TODO: note that we don't know if maintainers actually aimed for that model, or leave that for the discussion? -->
-Conversely, a small number of projects (<!-- TODO: do these share any characteristics in terms of platform or subject?
--->) do appear to fit the aspirational open source model, with a large number of contributors maintaining a long-lived and highly-active repository.
-Archaeologists active on GitHub also make markedly <!-- TODO: verify --> greater use of GitHub's social media-like features (stars, following) than more direct collaborative actions.
+Use of GitHub's collaboration features is also generally low (@fig-collab-features), although the data also shows a divergence between the uptake of features that facilitate direct code contributions (forks, issues, pull requests), which have markedly zero-skewed distributions, versus more indirect, social media-like features (comments, stars), which are moderately well-used.
+We hypothesise that this shows a preference for passive/reactive rather than active/proactive engagement with others work – a point we will return to in the conclusion.

### 2023-10-13 R/utils.R

@@ -1,0 +1,5 @@
+#' @export
+#' @noRd
+n_filter <- function(.data, ...) {
+  nrow(dplyr::filter(.data, ...))
+}

### 2023-10-13 analysis/_05-collaboration.qmd
Update results on distribution of contributions

@@
 
-```{r fig-contributors}
-#| fig-cap: Distribution of contributions in multi-contributor projects
-oarch |>
+```{r data-contributions}
+oarch_contributions <- oarch_gh |>
+  unnest(gh_contribs) |>
+  drop_na(contributor) |>
+  group_by(item_name) |>
+  summarise(n_contributors = n())
+
+p_one_contrib <- n_filter(oarch_contributions, n_contributors == 1) / nrow(oarch_contributions)
+
+oarch_contribs_by_repo <- oarch_gh |>
   unnest(gh_contribs) |>
@@
   ) |>
-  ungroup() |>
+  ungroup()
+
+p_big_boy_50 <- n_filter(oarch_contribs_by_repo, p_big_boy > 0.5) / nrow(oarch_contribs_by_repo)
+p_big_boy_75 <- n_filter(oarch_contribs_by_repo, p_big_boy > 0.75) / nrow(oarch_contribs_by_repo)
+```
+
+```{r fig-contributions}
+#| fig-cap: Distribution of contributions in multi-contributor projects
+oarch_contribs_by_repo |>
   mutate(item_name = fct_reorder(item_name, p_big_boy, .desc = TRUE)) |> 
@@
 
-<!-- TODO: introduce notion of contributors -->
-<!--
-#TODO: Re-calculate these numbers to ensure they are up to date.
-Translate the first point to be about how many repos have a quarter of contributions by a single person, so that it corresponds with the vertical lines on the graphic, which are now divided into quarters rather than thirds.
--->
-Perhaps unfortunately then, we also find that nearly 60% of projects have only a single contributor, and projects with more than a dozen are very rare indeed \[\<3%\] (@plot-gh_contributors-histogram).
-A "contribution", in this case, is a commit, issue, or comment applied by a user to a repository.
-Furthermore, when we break down the contributions made by all contributors in multi-contributor projects, we also see that work tends to be distributed very unevenly (@plot-contrib-dist).
-Of multi-contributor repositories, 92% have a third of contributions made by a single person, 65% have half of all contributions made by a single person, and 36% have three quarters of all contributions made a single person.
-<!-- TODO: expand description of contributors -->
+Perhaps unsurprisingly, given the low uptake of GitHub's collaborative features, `r percent(p_one_contrib)` of open-archaeo repositories only contain commits from a single user.
+Even in the minority of projects that have more than one contributor, work (as measured by number of recorded commits) is distributed highly unevenly (@fig-contributions).
+The lead maintainer almost always does the lion share of the work: they are responsible for more than half of commits in `r percent(p_big_boy_50)` of projects and more than three-quarters in `r percent(p_big_boy_50)`.
+<!-- TODO: what do we make of this observation? -->
 

### 2023-10-13 analysis/_03-data_methodology.qmd
Updated the descriptions and interpretations of findings.
Also some minor updates to branch/fork terminology.

@@
   ~term, ~definition,
-  "Comment", "On GitHub, Text post attached to an issue, including the first one that describes the issue",
+  "Comment", "On GitHub, text post attached to an issue, including the first one that describes the issue",
   "Commit", "Set of changes (addition, alteration, or deletion) to files in a repository that has been recorded by git as one entry in its log",
@@
 It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
-Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a branch in which they apply their modifications, and then request that their branch be merged into the central code base.
-After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent branches intact.
+Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a fork in which they apply their modifications, and then request that their fork be merged into the central code base.
+After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent forks intact.
 
@@
 
-In addition to serving as hosting providers for public git repositories, platforms like as GitHub, GitLab and Codeberg support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
-They also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, identify certain repositories as "favourites" and maintain a public-facing profile that includes personally-identifying information (e.g.
+In addition to serving as hosting providers for public git repositories, platforms like GitHub, GitLab and Codeberg support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
+They also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as favourites, and maintain a public-facing profile that includes personally-identifying information (e.g.
 profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.

### 2023-10-13 analysis/_04-open_archaeology.qmd
Updated the descriptions and interpretations of findings.
Also some minor updates to branch/fork terminology.

@@
 
-<!-- TODO: expand. -->
-We found that many repositories do not explicitly denote an open source license, or any license at all [@tbl-repo-licenses].
+To enact their mandate of ensuring that anyone can access and modify software and other creative works, the open source and open science movements encourage developers and scientists to adopt open licenses.
+Licenses are legally-bindings statements that stipulate how a creative work can be accessed and used.
+Proprietary licenses usually require explicit permission to be granted so that the work can be accessed or modified, usually in exchange for financial compensation.
+Open licenses, on the other hand, are more permissive, and allow anyone to use creative works without such harsh restrictions.
+While it is certainly possible to write your own license, it is very common to simply use one of several standardized open licenses (see [choosealicense.com](https://choosealicense.com for a more comprehensive overview of various options)).
+Some licenses, like GNU, MIT and Apache, are explicitly suited for distributing software, and specify certain use cases that are afforded by digital media.
+Other licenses, like the Creative Commons variants, are more suited to other kinds of creative works like books, articles, movies, music, photographs, and websites.
+The Creative Commons licenses also include clauses that cater to academic or creative sensibilities, such as requirements to attribute credit to the original authors, to restrict commercial use, and to propagate similar restrictions in derivative works.
+
+We found that the projects that host their code on GitHub use a wide variety of standard licenses.
+However, nearly half of all repositories <!-- TODO: encode the percentage --> do not explicitly denote an open source license, or any license at all [@tbl-repo-licenses].
+Maintainers may have assumed that making the code available, without explicitly stating permitting use, is enough to allow unrestricted access to the repository's contents.
+This is evocative of a similar ignorance among academics regarding the unsustainable use of GitHub as a long-term code and data hosting platform (@milliken2021; @escamilla2022; @escamilla2023).
+<!-- TODO: Expand with additional info after seeing what is in the "other" category -->
 

### 2023-10-13 analysis/_05-collaboration.qmd
Updated the descriptions and interpretations of findings.
Also some minor updates to branch/fork terminology.

@@
 As well as hosting source code, GitHub and other software forges include systems for facilitating collaboration on code and other projects.
-The basic collaborative workflow is inherited from git, which allows multiple users to commit (see @tbl-glossary for definitions of this and other git terminology used in this section) code to the repository.
+The basic collaborative workflow is inherited from git, which allows multiple users to commit code to the repository (see @tbl-glossary for definitions of this and other git terminology used in this section).
 A user with commit access to a repository can change any of its contents at will, so this is usually reserved for the project maintainer and known, trusted collaborators.
@@
 While not as a direct a contribution as pull requests or issues, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.
-These features are more used more widely than forks, issues and pull requests (@fig-collab-features): `r percent(p_with_stars)` of repositories have at least one star and, in those repositories that use issues, `r percent(p_issues_with_comments)` of them has received at least one additional comment.
+These features are used more widely than forks, issues and pull requests (@fig-collab-features): `r percent(p_with_stars)` of repositories have at least one star and, in those repositories that use issues, `r percent(p_issues_with_comments)` of them has received at least one additional comment.
 
@@
 The lead maintainer almost always does the lion share of the work: they are responsible for more than half of commits in `r percent(p_big_boy_50)` of projects and more than three-quarters in `r percent(p_big_boy_50)`.
-<!-- TODO: what do we make of this observation? -->
+This may be attributed to the steep learning curve commonly attributed to working with git.
+While git can be a great way to track changes and manage distributed contributions to a common code base, it can also be unwieldy in situations when multiple users (especially those with less experience using git for collaborative purposes) are expected to contribute within short spans of time.
+This meshes with our prior observations that projects which tend to exhibiting higher commit rates have less contributors.
+Additionally, our analyses neglects to account for contributions that are not tracked via git or GitHub.
+Those who do not code may provide creative guidance or feedback during in-person meetings, via email, or using alternative online messaging or social media platforms.
+A more focused qualitative assessment of these non-coding and supportive work practices would shed more light on the totality of effort that goes into producing and maintaining open source projects.
 
-The prototypical collaborative open source project on GitHub has a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through forks and pull requests, plus an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
-<!-- TODO: note that we don't know if maintainers actually aimed for that model, or leave that for the discussion? -->
-Only a small number of open-archaeo projects (<!-- TODO: do these share any characteristics in terms of platform or subject? -->) approach this model.
-The majority of projects are short-lived, with few contributors and a small number of commits.
+<!-- TODO: Abrupt transition, need to revise -->
+The prototypical and idyllic collaborative open source project ingrained into the common consciousness hypothetically comprises a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through forks and pull requests, plus an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
+It is unclear whether archaeological software developers actually aim to operate following this model, or whether it is even suitable for supporting what open science aims to achieve.
+However it is clear that only a small number of open-archaeo projects (<!-- TODO: do these share any characteristics in terms of platform or subject? -->) operate according to this model.
+The majority of projects are in fact short-lived, with few contributors and a small number of commits.
 Use of GitHub's collaboration features is also generally low (@fig-collab-features), although the data also shows a divergence between the uptake of features that facilitate direct code contributions (forks, issues, pull requests), which have markedly zero-skewed distributions, versus more indirect, social media-like features (comments, stars), which are moderately well-used.

### 2023-10-16 analysis/_04-open_archaeology.qmd
Adding some more info on tags and categories.

@@
 As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists.
-This primarily constitutes software but also includes various forms of open documents.
-@tbl-categories summarizes the kinds if resources that appear in open-archaeo, and breaks them down into more precise categories.
+It includes both software and documents, but not research compendiums.^[see https://github.com/benmarwick/ctv-archaeology#publications-that-include-r-code for a similar list of archaeology publications that include R code]
+We annotated each record with categories and tags that describe what aspect of archaeological research each tool or resource was meant to address.
+We categorized all records based on how each tool or resource is meant to be accessed or used, and assigned tags based on how developers identified their projects' purpose and scope.
+See [batist2023] for a more comprehensive overview of the tags and categories applied to open-archaeo.
 
@@
 
+In our breakdown of open-archaeo by category (see @tbl-categories), we demonstrate the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
 Most resources (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language or a plugin for an application.
@@
 
-We also annotated each record with "tags" that describe aspects of archaeological work that each tool contributes to [@fig-tags].
+Our analysis of tags highlights aspects of archaeological work that software developers are inclined to contribute to [@fig-tags].
 The most common tags unsurprisingly deal with work that naturally benefits from advanced information processing afforded by computers, such as statistical analysis, sample calibration, geographical analysis, data management, and chronological modelling.

### 2023-10-16 analysis/_03-data_methodology.qmd
Added overview of our methods, and their relevance.
Also updated the reference list.

@@
 We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, which are designed to facilitate open source software development by hosting public git repositories.
-Git is a protocol designed to facilitate open and distributed contributions to a common code base. <!-- mention other version control systems like SVN and mercurial maybe? -->
+Git is a protocol designed to facilitate open and distributed contributions to a common code base. <!-- mention other version control systems like SVN and mercurial, maybe? -->
 It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
@@
 
-<!-- TODO: add something on EDA and network analysis methodology -->
+The open-archaeo dataset and the supplemental data from the GitHub API provide us with a rich resource to explore the makeup of the archaeological software development scene.
+We apply exploratory data analysis methods to break down and visualize the data, so that we may ascertain overall trends [@tukey1977].
+In Section 4, <!-- encode section, see https://quarto.org/docs/authoring/cross-references.html#sections --> our focus is on examining the general state of open source archaeological software and resource development.
+In Section 5, <!-- encode section --> we refine our analysis to examine development processes, with specific focus on collaborative experiences.
+Finally, in Section 6, <!-- encode section --> we apply network analysis methods to investigate the formation of broader collaborative communities.
+Our analyses combine to support our objectives of understanding what kinds of software and resources archaeologists are making, and how they create these tools in response to specific needs and use-cases, as afforded by offline social and professional connections, and within the context of an emerging community of practice.
 

### 2023-10-17 analysis/_03-data_methodology.qmd
Filter non-archaeology items

@@
 
+```{r data-oarch-filter}
+oarch_filter <- read_csv(here("analysis/data/raw_data/open-archaeo_filter.csv"))
+n_items_removed <- nrow(filter(oarch_filter, !item_name %in% oarch$item_name))
+oarch |>
+  filter(!item_name %in% oarch_filter$item_name) ->
+  oarch
+
+# Exclude from GitHub API data collection because does not have own repo
+oarch$gh_repo[oarch$item_name == "ArchaeoLines"] <- NA
+```
+
 Open-archaeo is a relatively comprehensive list.
@@
 We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
+
+We excluded `r n_items_removed` records from open-archaeo from this analysis because they did not fit our "made by archaeologists, for archaeologists" criteria.
 

### 2023-10-17 analysis/_06-network.qmd
Filter non-archaeology items

@@
   filter(!user %in% c("ghost", "codecov-io", "coveralls", "codecov-commenter",
-                      "gitter-badger", "dependabot-support")) |>
-  # Remove big, non-archaeology projects that flood the network
-  filter(!repo %in% c("Stellarium/stellarium", 
-                     "opengisch/QField",
-                     "lutraconsulting/input")) ->
+                      "gitter-badger", "dependabot-support")) ->
   oarch_graph

### 2023-10-17 analysis/_04-open_archaeology.qmd
Update license data
Adding Zack's manual coding of repos GitHub couldn't automatically recognise; and simplifying the table by abbreviating license names and merging different versions of the same license

@@
 
-We found that the projects that host their code on GitHub use a wide variety of standard licenses.
-However, nearly half of all repositories <!-- TODO: encode the percentage --> do not explicitly denote an open source license, or any license at all [@tbl-repo-licenses].
-Maintainers may have assumed that making the code available, without explicitly stating permitting use, is enough to allow unrestricted access to the repository's contents.
-This is evocative of a similar ignorance among academics regarding the unsustainable use of GitHub as a long-term code and data hosting platform (@milliken2021; @escamilla2022; @escamilla2023).
-<!-- TODO: Expand with additional info after seeing what is in the "other" category -->
+```{r data-licenses}
+oarch_licenses <- read_csv(here("analysis/data/raw_data/open-archaeo_licenses.csv"))
+oarch_licenses |>
+  select(gh_repo = url, other_license = license) |>
+  right_join(oarch, by = "gh_repo") |>
+  hoist(gh_repo_info, license = list("license", 1, "spdx_id")) |>
+  mutate(license = if_else(license == "NOASSERTION", other_license, license)) |>
+  select(-other_license) ->
+  oarch
+```
 
 ```{r tbl-repo-licenses}
-#| tbl-cap: Summary of licenses used by projects with GitHub repositories.
+#| tbl-cap: Summary of licenses used by projects with GitHub repositories
+# TODO: group small categories into "other"
 oarch |>
-  hoist(gh_repo_info, license = list("license", 1, "name")) |>
+  mutate(license = str_remove(license, "-?\\d\\.\\d$")) |>
   count(license) |>
@@
 ```
+
+We found that the projects that host their code on GitHub use a wide variety of standard licenses.
+However, nearly half of all repositories <!-- TODO: encode the percentage --> do not explicitly denote an open source license, or any license at all [@tbl-repo-licenses].
+Maintainers may have assumed that making the code available, without explicitly stating permitting use, is enough to allow unrestricted access to the repository's contents.
+This is evocative of a similar ignorance among academics regarding the unsustainable use of GitHub as a long-term code and data hosting platform (@milliken2021; @escamilla2022; @escamilla2023).
+<!-- TODO: Expand with additional info after seeing what is in the "other" category -->
 

### 2023-10-17 analysis/_04-open_archaeology.qmd
Expand a bit on licenses
Written in haste...

@@
   oarch
+
+p_license_gpl <- n_filter(oarch, str_detect(license, "GPL")) / n_filter(oarch, !is.na(license))
+p_license_mit <- n_filter(oarch, str_detect(license, "MIT")) / n_filter(oarch, !is.na(license))
+p_license_cc <- n_filter(oarch, str_detect(license, "CC")) / n_filter(oarch, !is.na(license))
 ```
@@
 
-We found that the projects that host their code on GitHub use a wide variety of standard licenses.
-However, nearly half of all repositories <!-- TODO: encode the percentage --> do not explicitly denote an open source license, or any license at all [@tbl-repo-licenses].
-Maintainers may have assumed that making the code available, without explicitly stating permitting use, is enough to allow unrestricted access to the repository's contents.
-This is evocative of a similar ignorance among academics regarding the unsustainable use of GitHub as a long-term code and data hosting platform (@milliken2021; @escamilla2022; @escamilla2023).
-<!-- TODO: Expand with additional info after seeing what is in the "other" category -->
+Roughly half of open-archaeo repositories are accompanied by an explicit license [@tbl-repo-licenses].
+Two common free software licenses account for the majority of these: the GNU General Public License (GPL, `r percent(p_license_gpl)`) and the MIT License (`r percent(p_license_mit)`).
+These differ primarily in the restricts they place on reuse: the MIT License aims to be maximally permissive, while the GPL is a 'copyleft' license that specifies that all derivative works must be distributed under similar terms (in other words, it prohibits the use of FOSS software within non-FOSS software, [@cite?]).
+Archaeologists' preference for the more restrictive of these two licenses is the reverse of the general trend seen in open source projects on GitHub [@Balter2015].
+Creative Commons licenses are a distant third place (`r percent(p_license_cc)` of repositories), in contrast to their widespread use for other forms of scholarly output [@cite?].
+Many repositories do not specify a license; their maintainers may have assumed that making the code available, without explicitly stating permitting use, is enough to allow unrestricted access to the repository's contents.
+This is evocative of a similar ignorance among academics regarding the unsustainable use of GitHub as a long-term code and data hosting platform [@milliken2021; @escamilla2022; @escamilla2023].
+<!-- TODO: Seems like a pair of slightly uncharitable assumptions? -JR -->
+<!-- TODO: (ZB) Expand with additional info after seeing what is in the "other" category -->
 

### 2023-10-17 analysis/_03-data_methodology.qmd
Minor edits to the text.
Focused on the text concerning licensing. Fixed an innacurate encoding in Section 5. Updated the bib file. Removed irrelevant or completed TODOs and comments. Other very minor spelling and grammar tweaks.

@@
 
-We present an exploratory analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources.
+We present an exploratory analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` <!-- TODO: make this consistent with what appears in Section 4, based on the filter --> pieces of open source archaeological software and other digital resources.
 
@@
 We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, which are designed to facilitate open source software development by hosting public git repositories.
-Git is a protocol designed to facilitate open and distributed contributions to a common code base. <!-- mention other version control systems like SVN and mercurial, maybe? -->
+Git is a protocol designed to facilitate open and distributed contributions to a common code base.
 It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
@@
 
-Git is also designed to facilitate code review.
+Git is also designed to facilitate code review and version control.
 All modifications are tracked as "diffs", which highlight additions or deletions to the code base, including changes within individual files.

### 2023-10-17 analysis/_04-open_archaeology.qmd
Minor edits to the text.
Focused on the text concerning licensing. Fixed an innacurate encoding in Section 5. Updated the bib file. Removed irrelevant or completed TODOs and comments. Other very minor spelling and grammar tweaks.

@@
 As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists.
-It includes both software and documents, but not research compendiums.^[see https://github.com/benmarwick/ctv-archaeology#publications-that-include-r-code for a similar list of archaeology publications that include R code]
+It includes both software and documents, but not research compendiums.^[see [https://github.com/benmarwick/ctv-archaeology#publications-that-include-r-code](https://github.com/benmarwick/ctv-archaeology#publications-that-include-r-code) for a similar list of archaeology publications that include R code]
 We annotated each record with categories and tags that describe what aspect of archaeological research each tool or resource was meant to address.
 We categorized all records based on how each tool or resource is meant to be accessed or used, and assigned tags based on how developers identified their projects' purpose and scope.
-See [batist2023] for a more comprehensive overview of the tags and categories applied to open-archaeo.
+See [@batist2023] for a more comprehensive overview of the tags and categories applied to open-archaeo.
 
@@
 To enact their mandate of ensuring that anyone can access and modify software and other creative works, the open source and open science movements encourage developers and scientists to adopt open licenses.
-Licenses are legally-bindings statements that stipulate how a creative work can be accessed and used.
+Licenses are legally-binding statements that stipulate how a creative work can be accessed and used.
 Proprietary licenses usually require explicit permission to be granted so that the work can be accessed or modified, usually in exchange for financial compensation.
@@
 Two common free software licenses account for the majority of these: the GNU General Public License (GPL, `r percent(p_license_gpl)`) and the MIT License (`r percent(p_license_mit)`).
-These differ primarily in the restricts they place on reuse: the MIT License aims to be maximally permissive, while the GPL is a 'copyleft' license that specifies that all derivative works must be distributed under similar terms (in other words, it prohibits the use of FOSS software within non-FOSS software, [@cite?]).
-Archaeologists' preference for the more restrictive of these two licenses is the reverse of the general trend seen in open source projects on GitHub [@Balter2015].
-Creative Commons licenses are a distant third place (`r percent(p_license_cc)` of repositories), in contrast to their widespread use for other forms of scholarly output [@cite?].
-Many repositories do not specify a license; their maintainers may have assumed that making the code available, without explicitly stating permitting use, is enough to allow unrestricted access to the repository's contents.
-This is evocative of a similar ignorance among academics regarding the unsustainable use of GitHub as a long-term code and data hosting platform [@milliken2021; @escamilla2022; @escamilla2023].
-<!-- TODO: Seems like a pair of slightly uncharitable assumptions? -JR -->
-<!-- TODO: (ZB) Expand with additional info after seeing what is in the "other" category -->
+These differ primarily in the restrictions they place on reuse: the MIT License aims to be maximally permissive, while the GPL is a 'copyleft' license that specifies that all derivative works must be distributed under similar terms (in other words, it prohibits the use of FOSS software within non-FOSS software, [@dusollier2007]).
+Archaeologists' preference for the more restrictive of these two licenses is the reverse of the general trend seen in open source projects on GitHub [@balter2015].
+Creative Commons licenses are a distant third place (`r percent(p_license_cc)` of repositories), in contrast to their widespread use for other forms of scholarly output [@kim2007].
+Many repositories do not specify a license; given a documented misconception among academics that GitHub can serve as a sustainable and long-term code and data hosting platform [@milliken2021; @escamilla2022; @escamilla2023], it is possible that many maintainers whose work is included in open-archaeo similarly assumed that making their work available, without explicitly stating permissible use, is enough to allow unrestricted access to the repository's contents.
+However, we can not verify this potential explanation given the methods we currently employ, and more discursive qualitative research is needed to explore the rationales behind such decisions.
 

### 2023-10-17 analysis/_05-collaboration.qmd
Minor edits to the text.
Focused on the text concerning licensing. Fixed an innacurate encoding in Section 5. Updated the bib file. Removed irrelevant or completed TODOs and comments. Other very minor spelling and grammar tweaks.

@@
 Even in the minority of projects that have more than one contributor, work (as measured by number of recorded commits) is distributed highly unevenly (@fig-contributions).
-The lead maintainer almost always does the lion share of the work: they are responsible for more than half of commits in `r percent(p_big_boy_50)` of projects and more than three-quarters in `r percent(p_big_boy_50)`.
+The lead maintainer almost always does the lion share of the work: they are responsible for more than half of commits in `r percent(p_big_boy_50)` of projects and more than three-quarters in `r percent(p_big_boy_75)`.
 This may be attributed to the steep learning curve commonly attributed to working with git.
@@
 
-<!-- TODO: Abrupt transition, need to revise -->
 The prototypical and idyllic collaborative open source project ingrained into the common consciousness hypothetically comprises a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through forks and pull requests, plus an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
 It is unclear whether archaeological software developers actually aim to operate following this model, or whether it is even suitable for supporting what open science aims to achieve.
-However it is clear that only a small number of open-archaeo projects (<!-- TODO: do these share any characteristics in terms of platform or subject? -->) operate according to this model.
+However it is clear that only a small number of open-archaeo projects operate according to this model.
 The majority of projects are in fact short-lived, with few contributors and a small number of commits.

### 2024-01-10 analysis/_07-conclusion.qmd
Some minor updates to the conclusion

--- 2023-09-20 analysis/_07-conclusion.qmd
@@
 
-Archaeology is a "magpie" discipline, using tools and methods that originate from a variety of other disciplines and industries to better understand the past. 
-Moreover, archaeologists are accustomed to working on shoestring budgets, having to adapt their tools to suit specific, non-generalizable situations, and applying a DIY work ethic to create and co-creating with others to get the job done (Batist et al 2021; Morgan and Eve 2012; Caraher 2019).
-It is therefore unsprising that archaeologists are prolific open source software developers.
+Archaeology is a "magpie" discipline, in that it assembles tools and methods that originate from a variety of other disciplines and industries to better understand the past [@wylie1989; @nilssonstutz2018].
+Moreover, archaeologists are accustomed to working on shoestring budgets, having to adapt their tools to suit specific, non-generalizable situations, and applying a DIY work ethic to create and co-create with others [@morgan2012;  @caraher2019; @batist2021].
+The freedom to reuse, modify and extend code for a variety of purposes is a boon for the development of improvized tech stacks and workflows called up in contexts of archaeological fieldwork and analysis, and archaeologists share with open source software developers a pragmatic mindset with regards to planning collective work and evaluating outcomes [@scollar1999].
+It is therefore unsurprising that archaeologists are prolific users and developers of open source software.
 
-In this study we have sought to document an emerging community of open source software engineers in archaeology and to assess the extent to which collaboration within this community resembles conventional 'open source' forms.
-We have operationalised 'open-source collaborative practices' as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
-This excludes a wide range of other potential practices, most obviously the use of other software (though as we demonstrate in section X.X, the git/GitHub combination is overwhelmingly the most popular), but also collaboration through offline or private channels, or forms of collaboration we don't know about.
-We have not directly interviewed those doing the collaborating, though our conclusions do drawn heavily from our experience as members of that community ourselves. 
-Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it (@cite).
+Our goal in this study was to investigate the under-explored research practices involved in research software engineering. 
+We sought to identify not only _what_ kinds of software archaeologists are making, but _how_ archaeologists create these tools as part of a broader community of practice. 
+Our emphasis on the collaborative experiences involved in open source software development emerged from our experience maintaining open-archaeo, through which we observed that making one's code openly available on the web does not necessarily garner the benefits often touted by open science advocates, namely that source code can be audited, forked, and appropriated for alternative use cases, which are effectively social and collaborative experiences.
+
+To investigate these concerns, we operationalised open-source collaborative experiences as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
+This excludes a wide range of other potential practices, most obviously the use of other means of version-control and code-hosting (though as we demonstrate in <!-- encode the section --> Section 3, the git/GitHub combination is overwhelmingly the most popular).
+It also excludes collaboration through offline or private channels, or forms of collaboration we don't know about. 
+We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
+Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it [@ducke2013 ; @whallon1972].
 
@@
 Our analysis also shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host. 
-Generally speaking, we believe that this is because people do not want to step on other people's shoes by raising issues or intruding on other people's projects. 
+Generally speaking, we believe that this is because people do not want to step on other people's toes by raising issues or intruding on other people's projects. 
 This may relate to the fact that most developers on this list are academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
 
-However, we seem to have entered a period of critical reflection on the practical ramifications of infrastructures developed to support open science. 
-For instance, Huggett (2022) identified social frictions as a necessary aspect of productive research, which open science infrastructures effectively seek to eliminate. 
-Numerous other studies by Batist (2023), Huvila (2022), Haciguzeller et al.
-(2021), Opitz et al.
-(2021) Sobotkova (2018), Wylie (2017) Atici et al.
-(2013), Faniel et al.
-(2013) (among others) highlighted the need to account for the work involved in preparing and making use of materials shared openly on the web, and which are effectively muted by infrastructures developed in pursuit of open science.
-
+<!-- This entire paragraph is subject to change based on revised network analysis -->
 Our network analysis similarly draws attention to the real-world collaborative ties that underpin archaeological open source software development.
@@
 This indicates that open source is firmly embedded within existing social and institutional support and power structures that permeate academic life, both online and offline.
+
+There is thus little evidence to support the notion that archaeologists benefit from the positive outcomes that are commonly argued to be the natural results of open source development models -- namely, greater degrees of transparency, extensibility and participatory action.
+While opening the source code may facilitate these positive outcomes as necessary preconditional factors, we argue that this only amounts to establishing the _potential_ for people to put these values into practice.
+Moreover, we argue that the objectives and circumstances that frame archaeological practice significantly influence how far archaeologists (and academics in general) are willing to push for these values, and limit the ability for archaeologists to do open source in ways that resemble more mainstream open source projects.
+For instance, successful open source projects like the Linux kernel or the Firefox web browser are driven by collective and popular interest in ensuring that code remains functional, and the code base is therefore constantly in flux and bears an accumulating list of contributing members; this significantly differs from the organizational principles that govern archaeological work, namely that a project's leaders runs the whole show from the top down and operationalize all other actors as instruments of their central directives.
+Moreover, archaeological projects ultimately seek to produce stable textual outcomes bearing clear delineation of authorship and that require no upkeep whatsoever. 
+Sustaining an open source project is simply not compatible with the factors that currently drive the momentum behind archaeological work.
+
+As such, we advocate for more focused attention on specific disciplinary norms and institutional support structures that inform how knowledge is created and validated, and how varied contributions to the scholarly enterprise are mediated, credited, and valued.
+These factors, which are often ignored in open science discourse, vary from discipline to discipline and are rarely accounted for in the infrastructures themselves and in the policies that dictate or guide their implementation.
+However these factors should be considered as equal, if not more important, elements contributing to the success of open science, relative to the technical apparatus.

### 2024-01-24 analysis/_03-data_methodology.qmd
Replace gt with tinytable
Avoiding the awkward dependency on V8

@@
 ```{r data-oarch}
+# TODO: update and replace with local file before submission
 read_csv("https://github.com/zackbatist/open-archaeo/raw/2023-08-29/open-archaeo.csv") |>
@@
 ) |>
-  gt() |>
-  cols_label_with(fn = tools::toTitleCase) |>
-  fmt_markdown(definition)
+  rename_with(tools::toTitleCase) |>
+  tt() |>
+  format_tt(2, markdown = TRUE)
 ```
@@
 #| tbl-cap: Software forges used by open-archaeo projects
-gt(oarch_forges) |>
-  cols_label(repo_host = "Host", total = "Total") |>
-  cols_merge_n_pct(total, p_total) |>
-  fmt_percent(p_total) |>
-  tab_spanner("By category", Guides:`Standalone software`)
+oarch_forges |>
+  mutate(
+    total = glue("{total} ({format(p_total * 100, trim = TRUE, digits = 1)}%)")
+  ) |>
+  select(-p_total) |>
+  rename(Host = repo_host, Total = total) |>
+  tt()
 ```

### 2024-01-24 analysis/_04-open_archaeology.qmd
Replace gt with tinytable
Avoiding the awkward dependency on V8

@@
   arrange(-n) |>
-  gt() |>
-  cols_label_with(fn = tools::toTitleCase)
+  rename_with(tools::toTitleCase) |>
+  tt()
 ```
@@
   arrange(platform == "Other", -n) |>
-  gt() |>
-  cols_label_with(fn = tools::toTitleCase)
+  rename_with(tools::toTitleCase) |>
+  tt()
 ```
@@
   arrange(-n) |>
-  gt() |>
-    cols_label(license = "License", n = "N")
+  rename_with(tools::toTitleCase) |>
+  tt()
 ```

### 2024-01-24 analysis/_06-network.qmd
Replace gt with tinytable
Avoiding the awkward dependency on V8

@@
   select(repo, centrality, category, tags, n_commits) |>
-  gt() |>
-  cols_label(
-    repo = "Repository", 
-    centrality = "Betweenness Centrality",
-    category = "Category",
-    tags = "Tags",
-    n_commits = "Commits"
-  )
+  rename(repository = repo, commits = n_commits) |>
+  rename_with(tools::toTitleCase) |>
+  tt()
 ```

### 2024-01-24 analysis/openarchaeo_collaboration.qmd
Replace gt with tinytable
Avoiding the awkward dependency on V8

@@
 library("ggraph")
-library("gt")
 library("here")
@@
 library("tidyverse")
+library("tinytable")
 ```

### 2024-01-27 analysis/_06-network.qmd
Expand network analysis

@@
 # An emerging community of practice?
-
-<!-- Results III: network analysis of collaborative communities
-
-* Are there significant networks? If so, how big and how many?
-  * Only one, it seems!
-
-### [Repo-repo] What forms this network?
-
-* Identify the repos/organisations that are the focus of the most intense collaboration?
-  * Are they thematically or topically linked?
-  * Do they map onto real-world research centres?
-* Is the network a new thing, or just a translation of existing offline networks (e.g. fields, institutions) onto GitHub?
-  * cf. is open source really open?
-
-### [User-user] Who forms this network?
-
-* What proportion of users and repos are connected to this network?
-* What predicts membership in this network?
-* Does centrality in the network correlate with general rate of activity?
-*  Who are the most central nodes in this network? Is there anything they have in common?
-   * High activity?
-   * Certain platforms/categories/languages?
-* Qualitative analysis of top 50? 100? most connected users
-  * Career stage? Precarity? Employer? Location?
-
--->
 
@@
   group_by(contributor, repo_owner) |>
-  summarise(repos = n(), contribs = sum(contributions)) |>
-  pivot_wider(c(contributor), names_from = repo_owner, 
+  summarise(repos = n(), contribs = sum(contributions), .groups = "drop_last") |>
+  pivot_wider(id_cols = c(contributor), names_from = repo_owner, 
               values_from = c(repos, contribs), values_fill = 0) |>
@@
 
-We constructed and analyzed networks representing users' contributions to GitHub repositories to examine the emergence of collaborative tendencies.
-We sought to identify clusters of repositories that share common sets of contributors, and the different kinds of contributions that draw the network together.
+By contributing to shared repositories—whether with code (commits), issues, or comments—archaeologists using GitHub form a collaborative network which we can map using data from the GitHub API.
+Here we consider two facets of this network: repositories connected by common contributors (the repository–repository graph), and users connected by contributions to common repositories (the user–user graph).
+In both cases, number of contributions constitutes a natural measure of the strength or weight of the connection, which can be further broken down by type of contribution (commit, issue/pull request, or comment).
 
 ```{r data-graph}
-# TODO: refactor some of this as functions?
 oarch |>
@@
   oarch_graph
+
+# Repo-user
+oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
+                                   edges_user_to_repo(oarch_graph))
+
+# Repo-repo
+tbl_graph(nodes_repos(oarch_graph),
+          edges_common_users(oarch_graph),
+          directed = FALSE) |>
+  mutate(
+    degree = centrality_degree(),
+    betweenness = centrality_betweenness(weights = n),
+    subgraph = as_factor(group_components()),
+    group = as_factor(group_edge_betweenness())
+  ) ->
+  oarch_graph_repo_repo
+
+# User-user
+tbl_graph(nodes = nodes_users(oarch_graph), 
+          edges = edges_common_repos(oarch_graph),
+          directed = FALSE) |>
+  mutate(
+    degree = centrality_degree(),
+    betweenness = centrality_betweenness(weights = n),
+    subgraph = as_factor(group_components())
+  ) ->
+  oarch_graph_user_user
+
+# Network stats
+graph_n_repos_total <- length(oarch_graph_repo_repo)
+graph_n_users_total <- length(oarch_graph_user_user)
+
+subgraphs_n_repos <- oarch_graph_repo_repo |>
+  as_tibble() |>
+  count(subgraph)
+
+subgraphs_n_users <- oarch_graph_user_user |>
+  as_tibble() |>
+  count(subgraph)
+
+clusters_n_repos <- oarch_graph_repo_repo |>
+  as_tibble() |>
+  count(group)
+
+# Filter out isolates
+oarch_graph_repo_repo <- filter(oarch_graph_repo_repo, degree > 1)
+oarch_graph_user_user <- filter(oarch_graph_user_user, degree > 1)
+
+# Network stats (ignoring isolates)
+graph_n_repos <- length(oarch_graph_repo_repo)
+graph_n_users <- length(oarch_graph_user_user)
 ```
 
-We started by constructing a graph connecting users to the repositories that they contributed to, accounting for commits, issues, and comments as distinct kinds of relations (@fig-graph-repo-user).
-We then extracted two one-mode networks from this, one connecting repositories by common users, and the other connecting users by common repositories.
+```{r fig-graph-repo-user, eval=FALSE}
+#| fig-cap: Interactive repository–user graph.
+# TODO: is there a khroma scale that can accomodate this many groups?
+g_repo_repo <- oarch_graph_repo_user |>
+  ggraph(layout = "graphopt", mass = 100) +
+  geom_edge_link(aes(group = contrib, alpha = n_contrib)) +
+  geom_point_interactive(aes(x, y, colour = node_type, tooltip = node)) + 
+  scale_edge_alpha(trans = "log", guide = guide_none()) +
+  scale_colour_discrete(guide = guide_none()) +
+  labs(edge_alpha = "Common contributors") +
+  theme(panel.background = element_rect(fill = "white"))
 
-```{r extract-graph-repo-user, eval=TRUE}
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
+girafe(ggobj = g_repo_repo)
 ```
 
-```{r fig-graph-repo-user, eval=TRUE}
-g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt") +
-  geom_edge_fan(aes(colour = contrib, alpha = n_contrib)) +
-  geom_node_point(aes(colour = node_type)) +
-  scale_edge_alpha(trans = "log")
+```{r fig-graph-repo-repo}
+#| fig-cap: Interactive repository–repository graph.
+Repositories are connected by contributions by common users, with darker edges indicating more common contributors.
+Node colour indicates cluster membership according to .
+# TODO: is there a khroma scale that can accomodate this many groups?
+g_repo_repo <- oarch_graph_repo_repo |>
+  ggraph(layout = "graphopt", mass = 100) +
+  geom_edge_link(aes(alpha = n)) +
+  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
+  scale_edge_alpha(trans = "log", guide = guide_none()) +
+  scale_colour_discrete(guide = guide_none()) +
+  labs(edge_alpha = "Common contributors") +
+  theme(panel.background = element_rect(fill = "white"))
 
-#g_repo_user
-girafe(g_repo_user)
+girafe(ggobj = g_repo_repo)
 ```
 
-## repo-repo
+```{r fig-graph-user-user, eval=FALSE}
+#| fig-cap: Interactive user–user graph.
+Users are connected by contributions by common repositories, with darker edges indicating more common repositories.
+Node colour indicates cluster membership according to the edge–betweenness method [@Girvan Newman2002].
+# TODO: is there a khroma scale that can accomodate this many groups?
+g_repo_repo <- oarch_graph_user_user |>
+  filter(degree > 170) |>
+  ggraph(layout = "graphopt", mass = 100) +
+  geom_edge_link(aes(alpha = n)) +
+  geom_point_interactive(aes(x, y, tooltip = user)) + 
+  scale_edge_alpha(trans = "log", guide = guide_none()) +
+  scale_colour_discrete(guide = guide_none()) +
+  labs(edge_alpha = "Common contributors") +
+  theme(panel.background = element_rect(fill = "white"))
 
-We applied the edge-betweenness community detection method (Girvan and Newman 2002) to identify clusters of repositories that share common sets of contributors of all kinds.
-Aside from isolate nodes (n = XXX, not visible in our visualizations), which represent repositories with only a single contributor, we detected 21 clusters (@fig-graph-repo-repo-edge-betweenness-dendrogram).
-While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
-
-```{r extract-graph-repo-repo}
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = FALSE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
+girafe(ggobj = g_repo_repo)
 ```
 
-```{r fig-graph-repo-repo-interactive, eval=TRUE}
-# Remove the legend.
-iplot <- oarch_graph_repo_repo |>
-  ggraph(layout = "drl") +
-  geom_edge_link(aes(alpha = n)) +
-  #geom_node_point(aes(colour = group)) +
-  #geom_node_text(aes(label = group)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log") +
-  labs(edge_alpha = "Common contributors")
+<!-- Is there a network? How big is it? -->
+Our data shows that there is a significant network of archaeologists collaborating on GitHub.
+`r percent(graph_n_repos / graph_n_repos_total)` of repositories and `r percent(graph_n_users / graph_n_users_total)` of users in our dataset are connected to at least one other repository/user.
+Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-repo).
 
-girafe(ggobj = iplot)
-```
-
-```{r fig-graph-repo-repo-edge-betweenness-dendrogram, eval=TRUE}
+```{r fig-graph-repo-repo-edge-betweenness-dendrogram}
 # ZB Note: What we really want out of this is the identification of groups or clusters, as well as the relationships between clusters.
@@
 Additionally, we can colour-code the names of the repositories with their corresponding colours, or indicate that they are part of a cluster using a bracket (like "{" but turned 180 degrees).
-oarch_graph_repo_repo <- tbl_graph(nodes_repos(oarch_graph),
-                                   edges_common_users(oarch_graph),
-                                   directed = TRUE)
-
-oarch_graph_repo_repo |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(group = as.character(group_edge_betweenness())) ->
-  oarch_graph_repo_repo
-
+# @ZB: do we need both this and the graph viz. with clusters, though?
 xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
@@
 
-ggraph(xx, 'dendrogram', height = height)
+# TODO: not working
+#ggraph(xx, 'dendrogram', height = height)
 ```
 
-This graph identifies three core clusters and several peripheral clusters.
+<!-- Clustering within the network -->
+To explore the topography of this network, we applied the edge–betweenness community detection method [@GirvanNewman2002] to delimit `r nrow(filter(clusters_n_repos, n > 1))` distinct clusters in the repository–repository network (@fig-graph-repo-repo-edge-betweenness-dendrogram).
+While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
 The core clusters are characterized by repositories whose contributors commit to projects other than their own. 
@@
 
-```{r tbl-repo-centrality}
-#| tbl-cap: Most central repositories in the repository–repository network
+<!-- @ZB: Can you check that this is still valid, after we removed a lot of the genetics entries? -->
+The three core clusters, have distinct characters.
+One has a focus on archaeogenetics, which consists of a very well established collaborative network and general reliance on data modelling and processing tools. 
+A second cluster is mostly centred on fieldwork-oriented data collection tools, and particularly the emergence of well-funded and well-supported dominant platforms that attract more attention than other independent projects scattered across the network.
+The third and most significant cluster includes a smorgasbord of projects whose contributors share varied interests. 
+The emphasis in this latter cluster is on the formation of a central software development community rather than on any specific topic of work. 
+Many of the projects represented in this third cluster emerge from underlying professional partnerships, namely research labs (e.g. ISAA-Kiel) and special interest groups (SSLA).
+
+```{r tbl-repo-betweenness}
+#| tbl-cap: Most central repositories (weighted node betweenness) in the repository–repository network
 oarch_graph_repo_repo |>
-  mutate(centrality = centrality_betweenness()) |>
+  mutate(betweenness = centrality_betweenness()) |>
   as_tibble() |>
-  arrange(-centrality) |>
+  arrange(-betweenness) |>
   slice(1:10) |>
@@
   unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
-  select(repo, centrality, category, tags, n_commits) |>
+  select(repo, betweenness, category, tags, n_commits) |>
   rename(repository = repo, commits = n_commits) |>
@@
 
-<!-- TODO: What does tbl-repo-centrality tell us? Can we interrogate it further with a graph?
-ZB Response: Not much, I think I conceived it more as a means to an end.
-If I recall correctly, my intention was to ascertain whether different kinds (categories, mainly) of repositories are more central.
-One observation is the change in magnitude of betweenness centrality between the two "indexical" entries (open-archaeo and ctv-archaeo) and others in this top 10.
-Perhaps we can detect and visualize additional abrupt "jumps" in magnitude and identify some pattern relating to other variables.
--->
+```{r fig-repo-betweenness-age-length}
+#| fig-cap: Log-log correlation between repository centrality (weighted node betweenness), age (left) and length (right)
+left <- oarch_graph_repo_repo |>
+  as_tibble() |> 
+  left_join(oarch, by = c("repo" = "gh_repo")) |>
+  mutate(
+    age = map_int(gh_commits, \(x) min(today() - date(x$datetime)))
+  ) |>
+  ggplot(aes(age, betweenness)) +
+  geom_point() +
+  scale_x_log10() +
+  scale_y_log10() +
+  geom_smooth(method = "glm") +
+  labs(x = "Repository age (days)", y = "Centrality (betweenness)") +
+  theme_oarch()
 
-The three core clusters have their own distinct character.
-One has a focus on archaeogenetics, which consists of a very well established collaborative network and general reliance on data modelling and processing tools.
-A second cluster is mostly centred on fieldwork-oriented data collection tools, and particularly the emergence of well-funded and well-supported dominant platforms that attract more attention than other independent projects scattered across the network.
-The third and most significant cluster includes a schmorgasborg of projects whose contributors share varied interests.
-The emphasis in this latter cluster is on the formation of a central software development community rather than on any specific topic of work.
-Many of the projects represented in this third cluster emerge from underlying professional partnerships, namely research labs (e.g.
-ISAA-Kiel) and special interest groups (SSLA).
+right <- oarch_graph_repo_repo |>
+  as_tibble() |> 
+  left_join(oarch, by = c("repo" = "gh_repo")) |>
+  mutate(
+    length = map_int(gh_commits, nrow)
+  ) |>
+  ggplot(aes(length, betweenness)) +
+  geom_point() +
+  scale_x_log10() +
+  scale_y_log10() +
+  geom_smooth(method = "glm") +
+  labs(x = "Number of commits", y = NULL) +
+  theme_oarch()
 
-<!--  What happens to this interpretation when we differentiate commits from issues and comments? -->
-
-## user-user
-
-We assembled graphs linking users based on common contributions to the same repositories.
-As with the repo-repo network, we excluded isolate nodes (n = xxx), which represent users who only contributed to their own repositories, from the visualization.
-
-```{r graph-user-user}
-oarch_graph_user_user <- tbl_graph(
-  nodes = nodes_users(oarch_graph), 
-  edges = edges_common_repos(oarch_graph),
-  directed = FALSE
-)
-
-oarch_graph_user_user |>
-  mutate(degree = centrality_degree()) |>
-  filter(degree > 1) |>
-  mutate(centrality = centrality_betweenness(weights = n)) ->
-  oarch_graph_user_user
+left + right
 ```
 
-We sought to identify whether users who contribute in certain distinct ways play different roles in the overall network.
-We are reluctant to share personal information about specific users without their informed consent, but based on our knowledge of the community we found that the people with the highest betweenness values are those who primarily produce computational archaeology code as their job.
-Moreover, we found that these people tend to be employed under precarious circumstances.
-Although precarious employment is part of our sad reality, in the context of developing and maintaining open source software, this presents a serious source of risk.
-The people who occupy central positions in these networks are crucial community members that make the network whole, and if they are either unable to continue on in their contributions or decide to leave archaeology entirely, then the overall network would fragment.
+<!-- Predictors of repo centrality -->
+The repositories most central to the network as a whole [@tbl-repo-betweenness] include several lists and directories, including open-archaeo itself.
+Four of the top ten most central repositories (ISAAKiel/mortAAR, ISAAKiel/shapAAR, ISAAKiel/quantAAR, ropensci/c14bazAAR) are associated with the ISAA-Kiel group, emphasising the importance of existing 'real world' collaborations in structuring the network.
+Three (ropensci/c14bazAAR, ahb108/rcarbon, Chronomodel/chronomodel) are R packages relating to radiocarbon data and modelling, probably reflecting the fact that this has long been a prominent focus of statistical software development in archaeology.
+Repository centrality is predicted by the total number of commits it has received but, somewhat surprisingly, it is younger repositories rather than older ones that tend to be more central.
+Tentatively, we interpret this as an indication that the network has become more connected over time, but we leave a fuller analysis of temporal trends in collaborative activity to future work.
 
----
+```{r fig-repo-betweenness-category-platform}
+#| fig-cap: Mean repository centrality (weighted node betweenness) by open-archaeo category (top) and platform (bottom)
+top <- oarch_graph_repo_repo |>
+  as_tibble() |> 
+  left_join(oarch, by = c("repo" = "gh_repo")) |>
+  group_by(category) |>
+  summarise(betweenness = mean(betweenness)) |>
+  mutate(category = fct_reorder(category, betweenness)) |>
+  ggplot(aes(category, betweenness)) +
+  geom_col() +
+  scale_y_log10() +
+  coord_flip() +
+  labs(y = NULL, x = NULL) +
+  theme_oarch()
 
+bottom <- oarch_graph_repo_repo |>
+  as_tibble() |> 
+  left_join(oarch, by = c("repo" = "gh_repo")) |>
+  drop_na(platform) |>
+  group_by(platform) |>
+  summarise(betweenness = mean(betweenness)) |>
+  arrange(-betweenness) |>
+  slice(1:7) |>
+  mutate(platform = fct_reorder(platform, betweenness)) |>
+  ggplot(aes(platform, betweenness)) +
+  geom_col() +
+  scale_y_log10() +
+  coord_flip() +
+  labs(y = "Centrality (weighted betweenness)", x = NULL) +
+  theme_oarch()
+
+top / bottom
+```
+
+The highest mean centrality is seen in repositories that contain lists/datasets, standalone software, or packages/libraries, and in repositories based on Python, R or QGIS [@fig-repo-betweenness-category-platform].
+Interestingly, these trends depart from the observed popularity of different categories/platform in the open-archaeo dataset as a whole (see section XXX<!-- TODO: How do you do section references in Quarto?
+-->): standalone software is more central than packages/libraries, even though there are more of the latter by a significant margin.
+Similarly, Python packages tend to be more central than R packages, despite being the minority language.
+<!-- @ZB: I think this is interesting, but I couldn't really think of a general observation to make here. Maybe you can? -->
+
+```{r fig-user-betweenness-contribs, eval=TRUE}
+#| fig-cap: Log-log correlation between user centrality (weighted node betweenness) in the user-user graph and total number of contributions
+oarch_graph_user_user |>
+  as_tibble() |> 
+  left_join(oarch_contributors, by = c("user" = "contributor")) |>
+  ggplot(aes(betweenness, contribs)) +
+  geom_point() +
+  scale_x_log10() +
+  scale_y_log10() +
+  geom_smooth(method = "glm") +
+  labs(x = "Centrality (betweenness)", y = "Total contributions") +
+  theme_oarch()
+```
+
+<!-- Predictors of user centrality -->
+Centrality to the user–user graph is weakly predicted by a user's overall rate of activity, as measured by their total number of contributions [@fig-user-betweenness-contribs]. 
+<!-- @ZB: Have softened the following, not because I think it's wrong, just because I'm a bit uncomfortable making strong claims without a reproducible analysis to back it up. 
+Also N.B.: I don't think they write code for a living. I don't know of anybody that rights code for a living in archaeology. I think it's the opposite, actually. -->
+We did not collect demographic data on users that appear in our dataset, but based on our own knowledge of the community we can observe that those highly central to the network tend to be employed in (junior) academic positions, or in a few cases in cultural heritage authorities, rather than specifically as research software engineers.
+Such positions tend to not to actively reward or encourage software development, at least not on a par with more traditional academic outputs, and are increasingly precarious.
+This obviously poses a serious risk to the sustainability and growth of open source software in archaeology: if the people who occupy central positions in the network cease to be active, then it is likely that the overall network would fragment.
+Assessing and mitigating this risk should be a high priority for future research in this area.
+
+<!-- TODO: this should be backed up with a table or figure. The analysis seems to have got lost somewhere. Or just cut? -->
 We also applied the same betweenness centrality algorithm on a subnetwork whose links are based only on issues and comments, and not code contributions.
@@
 Although it is out of this study's scope a qualitative analysis of issues and comments may yield more insight on the kinds of contributions that each of these participants make.
-
-------------------------------------------------------------------------
-
-<!--ZB Note: What is this? -->
-
-```{r fig-user-betweenness, eval=TRUE}
-#| fig-cap: Distribution of betweenness centrality values in the user–user network
-oarch_graph_user_user |>
-  mutate(
-    betweenness = centrality_betweenness(weights = n)
-  ) |>
-  data.frame() |>
-  ggplot(aes(x = betweenness)) +
-  geom_density(bw = "nrd") +
-  scale_x_log10(labels = scales::number)
-```
-
-```{r fig-user-betweenness-activity, eval=TRUE}
-#| fig-cap: Correlation between user's centrality (betweenness) in the network and their total number of contributions
-oarch_graph_user_user |>
-  mutate(
-    degree = centrality_degree(), 
-    betweenness = centrality_betweenness(weights = n), 
-    strength = centrality_degree(weights = n)
-  ) |>
-  as_tibble() |> 
-  left_join(oarch_contributors, by = c("user" = "contributor")) |>
-  ggplot(aes(betweenness, contribs)) +
-  geom_point() +
-  scale_x_log10() +
-  scale_y_log10() +
-  geom_smooth(method = "lm")
-```

### 2024-01-27 analysis/_04-open_archaeology.qmd
Resolve TODOs in section 4

@@
 p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / nrow(oarch)
+p_platform_python <- sum(oarch$platform == "Python", na.rm = TRUE) / nrow(oarch)
+p_platform_qgis <- sum(oarch$platform == "QGIS", na.rm = TRUE) / nrow(oarch)
 ```
@@
 `r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used platform by a large margin (@tbl-platforms).
-Python, another programming language, is also relatively popular <!-- TODO: percent -->, as are plugins for the open source geographic information system QGIS <!-- TODO: percent -->.
+Python, another programming language, is also relatively popular (`r percent(p_platform_python)`), as are plugins for the open source geographic information system QGIS (`r percent(p_platform_qgis)`).
 Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
@@
 These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.
-
-<!-- TODO: can we say anything more concrete about the tags? 
-ZB response: We can look at the development of tags over time.
-For instance, a chart documenting year-over-year growth of each tag based on date of each project's first commit.
-I imagine this as a stacked bar chart (similar to below) but with segments coded to represent year of first commit.
-This would fit at the end of this section, under fog-github-cumulative.
-We can add charts documenting growth of platforms and licenses too.
--->
 

### 2024-01-30 analysis/_03-data_methodology.qmd
Mostly minor but necessary fixes to code and text
Provided a little explanation for the manual licenses file, included a missing value to the licenses file, corrected some citations, added section slugs and encoded links to sections, removed various completed or unnecessary todos and notes, removed dendrogram, verified and updated characterization of clusters, fixed references to graph objects

@@
-# Data and methodology
+# Data and methodology {#sec-data-methodology}
 
@@
 
-We present an exploratory analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` <!-- TODO: make this consistent with what appears in Section 4, based on the filter --> pieces of open source archaeological software and other digital resources.
+We present an exploratory analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources.
 
@@
 We apply exploratory data analysis methods to break down and visualize the data, so that we may ascertain overall trends [@tukey1977].
-In Section 4, <!-- encode section, see https://quarto.org/docs/authoring/cross-references.html#sections --> our focus is on examining the general state of open source archaeological software and resource development.
-In Section 5, <!-- encode section --> we refine our analysis to examine development processes, with specific focus on collaborative experiences.
-Finally, in Section 6, <!-- encode section --> we apply network analysis methods to investigate the formation of broader collaborative communities.
+In @sec-open-archaeology, our focus is on examining the general state of open source archaeological software and resource development.
+In @sec-collaboration, we refine our analysis to examine development processes, with specific focus on collaborative experiences.
+Finally, in @sec-network, we apply network analysis methods to investigate the formation of broader collaborative communities.
 Our analyses combine to support our objectives of understanding what kinds of software and resources archaeologists are making, and how they create these tools in response to specific needs and use-cases, as afforded by offline social and professional connections, and within the context of an emerging community of practice.
 
-The quantitative analyses and figures presented here were generated with `r version$version.string` (@Manual).
-The full data and code is available in the compendium that accompanies this paper (@Compendium).
+The quantitative analyses and figures presented here were generated with `r version$version.string` [@rcoreteam2023].
+The full data and code is available in the compendium that accompanies this paper [@roe2023] <!-- reference should be updated to reflect the actual version number and date -->.

### 2024-01-30 analysis/_04-open_archaeology.qmd
Mostly minor but necessary fixes to code and text
Provided a little explanation for the manual licenses file, included a missing value to the licenses file, corrected some citations, added section slugs and encoded links to sections, removed various completed or unnecessary todos and notes, removed dendrogram, verified and updated characterization of clusters, fixed references to graph objects

@@
 
-# Open archaeology
+# Open archaeology {#sec-open-archaeology}
 
@@
 #| tbl-cap: Categories of open archaeology projects included in open-archaeo.
-# TODO: Include sum for the "Software" and "Documents" supercategories, and make them appear in bold.
 tribble(
@@
 ```{r data-licenses}
+# open-archaeo_licenses.csv contains a list of licenses that had to be manually verified.
+This is necessary due to the fact that GitHub and CRAN have different conventions for naming the files containing the license.
 oarch_licenses <- read_csv(here("analysis/data/raw_data/open-archaeo_licenses.csv"))
@@
 #| tbl-cap: Summary of licenses used by projects with GitHub repositories
-# TODO: group small categories into "other"
 oarch |>

### 2024-01-30 analysis/_06-network.qmd
Mostly minor but necessary fixes to code and text
Provided a little explanation for the manual licenses file, included a missing value to the licenses file, corrected some citations, added section slugs and encoded links to sections, removed various completed or unnecessary todos and notes, removed dendrogram, verified and updated characterization of clusters, fixed references to graph objects

@@
-# An emerging community of practice?
+# An emerging community of practice? {#sec-network}
 
@@
 
-```{r fig-graph-repo-user, eval=FALSE}
+```{r fig-graph-repo-user}
 #| fig-cap: Interactive repository–user graph.
 # TODO: is there a khroma scale that can accomodate this many groups?
-g_repo_repo <- oarch_graph_repo_user |>
+g_repo_user <- oarch_graph_repo_user |>
   ggraph(layout = "graphopt", mass = 100) +
@@
 
-girafe(ggobj = g_repo_repo)
+girafe(ggobj = g_repo_user)
 ```
@@
 Repositories are connected by contributions by common users, with darker edges indicating more common contributors.
-Node colour indicates cluster membership according to .
+Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002].
+Excludes isolate nodes.
 # TODO: is there a khroma scale that can accomodate this many groups?
@@
 
-```{r fig-graph-user-user, eval=FALSE}
+```{r fig-graph-user-user}
 #| fig-cap: Interactive user–user graph.
 Users are connected by contributions by common repositories, with darker edges indicating more common repositories.
-Node colour indicates cluster membership according to the edge–betweenness method [@Girvan Newman2002].
+Node colour indicates cluster membership according to the edge–betweenness method [girvan2002].
+Excludes isolate nodes.
 # TODO: is there a khroma scale that can accomodate this many groups?
-g_repo_repo <- oarch_graph_user_user |>
+g_user_user <- oarch_graph_user_user |>
   filter(degree > 170) |>
@@
   scale_colour_discrete(guide = guide_none()) +
-  labs(edge_alpha = "Common contributors") +
+  labs(edge_alpha = "Common repositories") +
   theme(panel.background = element_rect(fill = "white"))
 
-girafe(ggobj = g_repo_repo)
+girafe(ggobj = g_user_user)
 ```
@@
 Our data shows that there is a significant network of archaeologists collaborating on GitHub.
-`r percent(graph_n_repos / graph_n_repos_total)` of repositories and `r percent(graph_n_users / graph_n_users_total)` of users in our dataset are connected to at least one other repository/user.
+`r percent(graph_n_repos / graph_n_repos_total)` of repositories and `r percent(graph_n_users / graph_n_users_total)` of users in our dataset are connected to at least one other repository or user.
 Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-repo).
 
-```{r fig-graph-repo-repo-edge-betweenness-dendrogram}
-# ZB Note: What we really want out of this is the identification of groups or clusters, as well as the relationships between clusters.
-In theory, the dendrogram provides the best visualization for this information.
-I wonder if we can improve it by colour-coding the algorithmically-identified clusters (which should appear side by side), and making the names of the repositories appear vertically aligned perpendicular to the x axis.
-Additionally, we can colour-code the names of the repositories with their corresponding colours, or indicate that they are part of a cluster using a bracket (like "{" but turned 180 degrees).
-# @ZB: do we need both this and the graph viz. with clusters, though?
-xx <- igraph::edge.betweenness.community(oarch_graph_repo_repo)
-xx <- as.dendrogram(xx)
-plot(xx)
-
-# TODO: not working
-#ggraph(xx, 'dendrogram', height = height)
-```
-
 <!-- Clustering within the network -->
-To explore the topography of this network, we applied the edge–betweenness community detection method [@GirvanNewman2002] to delimit `r nrow(filter(clusters_n_repos, n > 1))` distinct clusters in the repository–repository network (@fig-graph-repo-repo-edge-betweenness-dendrogram).
-While many of these clusters are interconnected, some discrete components containing between 2-7 repositories appear as distinct from a primary core.
-The core clusters are characterized by repositories whose contributors commit to projects other than their own. 
-The other peripheral clusters largely correspond with the work of single individuals, and sometimes also their close colleagues.
+To explore the topography of this network, we applied the edge–betweenness community detection method [@girvan2002] to delimit `r nrow(filter(clusters_n_repos, n > 1))` distinct clusters in the repository–repository network.
+While many of these clusters are interconnected, some discrete components containing between 2-20 repositories appear as distinct from a primary core.
+The core cluster is characterized by repositories whose contributors commit to projects other than their own, and it includes a smorgasbord of projects whose contributors share varied interests.
+This indicates the formation of a central software development community rather than a focus on any specific topic of work.
+The other peripheral clusters situated around the core largely correspond with the work of single individuals, and sometimes also their close colleagues.
+Many of the projects represented in the core emerge from underlying professional partnerships, namely research labs (e.g. ISAA-Kiel) and special interest groups (SSLA).
 Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (or perhaps initial) integration of lone developers into a broader community.
-
-<!-- @ZB: Can you check that this is still valid, after we removed a lot of the genetics entries? -->
-The three core clusters, have distinct characters.
-One has a focus on archaeogenetics, which consists of a very well established collaborative network and general reliance on data modelling and processing tools. 
-A second cluster is mostly centred on fieldwork-oriented data collection tools, and particularly the emergence of well-funded and well-supported dominant platforms that attract more attention than other independent projects scattered across the network.
-The third and most significant cluster includes a smorgasbord of projects whose contributors share varied interests. 
-The emphasis in this latter cluster is on the formation of a central software development community rather than on any specific topic of work. 
-Many of the projects represented in this third cluster emerge from underlying professional partnerships, namely research labs (e.g. ISAA-Kiel) and special interest groups (SSLA).
 
@@
 The highest mean centrality is seen in repositories that contain lists/datasets, standalone software, or packages/libraries, and in repositories based on Python, R or QGIS [@fig-repo-betweenness-category-platform].
-Interestingly, these trends depart from the observed popularity of different categories/platform in the open-archaeo dataset as a whole (see section XXX<!-- TODO: How do you do section references in Quarto?
--->): standalone software is more central than packages/libraries, even though there are more of the latter by a significant margin.
+Interestingly, these trends depart from the observed popularity of different categories/platform in the open-archaeo dataset as a whole (see @sec-open-archaeology): standalone software is more central than packages/libraries, even though there are more of the latter by a significant margin.
 Similarly, Python packages tend to be more central than R packages, despite being the minority language.
 <!-- @ZB: I think this is interesting, but I couldn't really think of a general observation to make here. Maybe you can? -->
+<!-- Possibly that R packages are easier to spin up solo, whereas standalone tends to have more financial/institutional backing, and thus greater potential to support teams of developers -->
 
@@
 Centrality to the user–user graph is weakly predicted by a user's overall rate of activity, as measured by their total number of contributions [@fig-user-betweenness-contribs]. 
-<!-- @ZB: Have softened the following, not because I think it's wrong, just because I'm a bit uncomfortable making strong claims without a reproducible analysis to back it up. 
-Also N.B.: I don't think they write code for a living. I don't know of anybody that rights code for a living in archaeology. I think it's the opposite, actually. -->
 We did not collect demographic data on users that appear in our dataset, but based on our own knowledge of the community we can observe that those highly central to the network tend to be employed in (junior) academic positions, or in a few cases in cultural heritage authorities, rather than specifically as research software engineers.

### 2024-01-31 analysis/_06-network.qmd
Add a couple of references

@@
 We did not collect demographic data on users that appear in our dataset, but based on our own knowledge of the community we can observe that those highly central to the network tend to be employed in (junior) academic positions, or in a few cases in cultural heritage authorities, rather than specifically as research software engineers.
-Such positions tend to not to actively reward or encourage software development, at least not on a par with more traditional academic outputs, and are increasingly precarious.
+Such positions tend to not to actively reward or encourage software development, at least not on a par with more traditional academic outputs [@baxter2012], and are increasingly precarious [@Cornelius-BellBell2021].
 This obviously poses a serious risk to the sustainability and growth of open source software in archaeology: if the people who occupy central positions in the network cease to be active, then it is likely that the overall network would fragment.

### 2024-01-31 analysis/_06-network.qmd
Export graphs as gexf

@@
 graph_n_users <- length(oarch_graph_user_user)
+
+# Export as gexf
+write.gexf(igraph.to.gexf(oarch_graph_repo_user), 
+           output = here("analysis/data/derived_data/oarch_repo_user.gexf"))
+write.gexf(igraph.to.gexf(oarch_graph_repo_repo), 
+           output = here("analysis/data/derived_data/oarch_repo_repo.gexf"))
+write.gexf(igraph.to.gexf(oarch_graph_user_user), 
+           output = here("analysis/data/derived_data/oarch_user_user.gexf"))
 ```

### 2024-01-31 analysis/openarchaeo_collaboration.qmd
Export graphs as gexf

--- 2023-10-12 analysis/openarchaeo_collaboration.qmd
@@
 library("patchwork")
+library("rgexf")
 library("scales")

### 2024-02-01 analysis/_06-network.qmd
Worked on the networks section.
Added more tangible insights to the text, experimented with new layouts (repo-user looks good now), refactored the way the repo-user graph is generated.

@@
 # Repo-user
-oarch_graph_repo_user <- tbl_graph(nodes_repos_users(oarch_graph),
-                                   edges_user_to_repo(oarch_graph))
+tbl_graph(nodes_repos_users(oarch_graph),
+          edges_user_to_repo(oarch_graph),
+          directed = FALSE) |>
+          mutate(
+            degree = centrality_degree(),
+            betweenness = centrality_betweenness(),
+            subgraph = as_factor(group_components()),
+            group = as_factor(group_edge_betweenness())
+            ) ->
+            oarch_graph_repo_user
 
@@
     subgraph = as_factor(group_components())
+    # group = as_factor(group_edge_betweenness()) # <- very slow!
   ) ->
@@
 
-```{r fig-graph-repo-user}
+```{r fig-graph-repo-user, eval=FALSE}
 #| fig-cap: Interactive repository–user graph.
@@
 g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "graphopt", mass = 100) +
+  ggraph(layout = "fr", repulserad = 10, area = 10000, niter = 5000, grid = "nogrid") +
   geom_edge_link(aes(group = contrib, alpha = n_contrib)) +
-  geom_point_interactive(aes(x, y, colour = node_type, tooltip = node)) + 
+  geom_point_interactive(aes(x, y, colour = group, shape=node_type, tooltip = node)) + 
   scale_edge_alpha(trans = "log", guide = guide_none()) +
@@
 
-```{r fig-graph-repo-repo}
+```{r fig-graph-repo-repo, eval=FALSE}
 #| fig-cap: Interactive repository–repository graph.
@@
 
-```{r fig-graph-user-user}
+```{r fig-graph-user-user, eval=FALSE}
 #| fig-cap: Interactive user–user graph.
@@
   filter(degree > 170) |>
-  ggraph(layout = "graphopt", mass = 100) +
+  ggraph(layout = "fr", repulserad = 10, area = 100000, niter = 10000000, grid = "nogrid") +
   geom_edge_link(aes(alpha = n)) +
@@
 `r percent(graph_n_repos / graph_n_repos_total)` of repositories and `r percent(graph_n_users / graph_n_users_total)` of users in our dataset are connected to at least one other repository or user.
-Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-repo).
+Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-repo and @fig-graph-user-user).
 
@@
 The core cluster is characterized by repositories whose contributors commit to projects other than their own, and it includes a smorgasbord of projects whose contributors share varied interests.
-This indicates the formation of a central software development community rather than a focus on any specific topic of work.
-The other peripheral clusters situated around the core largely correspond with the work of single individuals, and sometimes also their close colleagues.
-Many of the projects represented in the core emerge from underlying professional partnerships, namely research labs (e.g. ISAA-Kiel) and special interest groups (SSLA).
-Peripheral clusters that are connected to core clusters by only a few relationships represent the sole (or perhaps initial) integration of lone developers into a broader community.
+
+Likewise, we applied the same edge-betweenness community detection method to identify clusters in the user-user network, which reveal collaborative groupings.
+We again see a complementary primary core connected to several more peripheral clusters, which are internally-cohesive and exhibit few connections with other peripheral clusters.
+The central core bridges all the peripheral clusters.
+The central core is not uniform, and comprises several relatively discrete sub-groups representing collaborative sub-communities.
+While these sub-groups are internally cohesive, they exhibit enough connections to other members of the central core so as to not be considered as separate or peripheral clusters.
+
+In both the repository-repository and user-user networks, the peripheral clusters correspond with either the connections surrounding specific projects or the series of repositories created by single individuals and sometimes also their close colleagues.
+On the other hand, the central cores exhibit greater internal variety that may correspond with social connections and the formation of a complex software development community.
+This is evident through the fact that many of the connections represented in the cores emerge from underlying professional partnerships, namely research labs (e.g.
+ISAA-Kiel) and special interest groups (CAA-SSLA).
+Peripheral clusters that are connected to the central core by only a few relationships represent the sole (or perhaps initial) integration of lone developers into a broader community.
 
@@
 Interestingly, these trends depart from the observed popularity of different categories/platform in the open-archaeo dataset as a whole (see @sec-open-archaeology): standalone software is more central than packages/libraries, even though there are more of the latter by a significant margin.
-Similarly, Python packages tend to be more central than R packages, despite being the minority language.
+This may be due to the fact that many packages are developed to support specific practices or use-cases (often inspired by personal need), or are designed to run relatively stable statistical functions that need not change over time.
+These are therefore relatively stable and require little additional input after release.
+On the other hand, as discussed in @sec-open-archaeology, standalone software tend to integrate multiple system components and may evolve over time to add new features or support new workflows.
+Moreover, standalone software are generally rooted in longer-term and community-held objectives, and their development may therefore be backed by insitutions with funding and resources to support developers.
+
+Despite being the minority language, Python packages tend to be more central than R packages, on average.
+This might relate to the different ways in which R and Python developers call dependencies ???
 <!-- @ZB: I think this is interesting, but I couldn't really think of a general observation to make here. Maybe you can? -->
-<!-- Possibly that R packages are easier to spin up solo, whereas standalone tends to have more financial/institutional backing, and thus greater potential to support teams of developers -->
+<!-- I wonder what the median is? Also: cltk is a python script, and is the massive pink blob in the repo-repo and user-user networks, so that might skew things slightly -->
 

### 2024-02-02 R/forceatlas2.R
Add ForceAtlas2 layout and licenses

@@ -1,0 +1,181 @@
+#' ForceAtlas2 layout
+#'
+#' Copied from <https://github.com/analyxcompany/ForceAtlas2> (licensed GPL>=2)
+#'
+#' @param graph An igraph network or a data frame of three columns: source, target, and weights.
+#' @param directed Logical. TRUE if the network is directed. Ignored if graph is an igraph object.
+#' @param iterations Number of iterations to be performed.
+#' @param linlog Logical. If TRUE the algorithm uses logarithmic attraction force `F <- log (1+F)`
+#' @param pos A data frame or matrix (NumberOfNodes x dimension) of the initial locations of points. If NULL the initial positions are random.
+#' @param nohubs Logical. If TRUE nodes with high indegree have more central position than nodes with outdegree (for directed graphs).
+#' @param k Is the repel constant: the greater the constant k the stronger the repulsion force between points.
+#' @param gravity Gravity constant: indicates how strongly the nodes should be attracted to the center of gravity.
+#' @param ks Speed constant: the greater the value of `ks` the more movement the nodes make under the acting forces.
+#' @param ksmax Limits the speed from above.
+#' @param delta Modify attraction force, the weights are raised to the power of `delta`
+#' @param center Center of gravity.
+#' @param tolerance Tolerance to swinging constant.
+#' @param dim Dimension of the positions.
+#' @param plotstep is the frequency of plotting intermediate iterations.
+#' @param plotlabels Logical. If TRUE the labels should be included in the intermediate interations plot
+#' @return When graph is an igraph object, the function returns a matrix with the positions. If graph is a data frame, this is a data frame with the names of the nodes and the positions.
+#' @examples
+#' library(igraph)
+#' g <- graph.ring(50)
+#' layout <- layout.forceatlas2(g, iterations=3000, plotstep=100)
+#' plot(g, layout=layout)
+#'
+#' @source
+#' Jacomy M, Venturini T, Heymann S, Bastian M (2014) ForceAtlas2, a Continuous Graph Layout Algorithm for Handy Network Visualization Designed for the Gephi Software. PLoS ONE 9(6): e98679. <https://doi.org/10.1371/journal.pone.0098679>
+#'
[…]

### 2024-02-02 analysis/_03-data_methodology.qmd
Revise figures and tables
Removing superfluous code, consistent style, consistent captions and labels, etc.

@@
   oarch
-
-# Amalgamate categories into fewer rough categories
-# TODO: redo
-oarch <- mutate(oarch, rough_category = recode(category,
-                                 "R" = "Packages (R, Python, etc.)",
-                                 "Web apps" = "Standalone apps",
-                                 "Datasets" = "Analyses & datasets",
-                                 "Tutorials" = "Lists & tutorials",
-                                 "Python" = "Packages (R, Python, etc.)",
-                                 "Desktop apps" = "Standalone apps",
-                                 "QGIS" = "Plugins",
-                                 "Lists" = "Lists & tutorials",
-                                 "Mobile apps" = "Standalone apps",
-                                 "Analyses" = "Analyses & datasets",
-                                 "Spreadsheets" = "Forms & spreadsheets",
-                                 "Command line tools" = "Standalone apps",
-                                 "Plugins" = "Plugins",
-                                 "ArcGis" = "Plugins",
-                                 "Filemaker" = "Plugins",
-                                 "ODK" = "Forms & spreadsheets",
-                                 "Lisp" = "Packages (R, Python, etc.)",
-                                 .default = "Other",
-                                 .missing = "Other"
-                                 ))
 
@@
   tt() |>
-  format_tt(2, markdown = TRUE)
+  format_tt(2, markdown = TRUE) |>
+  style_tt(j = 1, italic = TRUE)
 ```
@@
   drop_na(repo_url) |>
-  count(category, repo_host) |>
-  pivot_wider(names_from = category, values_from = n, values_fill = 0) |>
-  mutate(
-    total = rowSums(across(Guides:`Standalone software`)), 
-    p_total = total / sum(total),
-    .before = Guides
-  ) |>
-  arrange(repo_host == "none", desc(total)) |>
+  count(repo_host) |>
+  mutate(p = n / sum(n)) |>
+  arrange(repo_host == "none", desc(n)) |>
   mutate(repo_host = recode(repo_host,
@@
 
-p_no_vc <- filter(oarch_forges, repo_host == "None")[["p_total"]]
-p_no_github <- sum(filter(oarch_forges, repo_host != "None" & repo_host != "GitHub")$p_total)
+p_no_vc <- filter(oarch_forges, repo_host == "None")$p
+p_no_github <- sum(filter(oarch_forges, repo_host != "None" & repo_host != "GitHub")$p)
 ```
@@
 ```{r tbl-forges}
-#| tbl-cap: Software forges used by open-archaeo projects
+#| tbl-cap: Software forges used by open archaeology projects
 oarch_forges |>
-  mutate(
-    total = glue("{total} ({format(p_total * 100, trim = TRUE, digits = 1)}%)")
-  ) |>
-  select(-p_total) |>
-  rename(Host = repo_host, Total = total) |>
+  mutate(p = paste0(format(p * 100, trim = TRUE, digits = 1), "%")) |>
+  rename(Host = repo_host, `%` = p) |>
   tt()

### 2024-02-02 analysis/_04-open_archaeology.qmd
Revise figures and tables
Removing superfluous code, consistent style, consistent captions and labels, etc.

@@
 ```{r tbl-categories}
-#| tbl-cap: Categories of open archaeology projects included in open-archaeo.
+#| tbl-cap: Categories of open archaeology projects
 tribble(
@@
   left_join(oarch, by = "category") |>
-  count(category, kind, scope) |>
-  group_by(kind) |>
-  arrange(-n) |>
+  count(kind, category, scope) |>
+  mutate(
+    p = n / sum(n),
+    p = paste0(format(p * 100, trim = TRUE, digits = 1), "%")
+  ) |>
+  arrange(desc(kind), -n) |>
+  select(-kind) |>
   rename_with(tools::toTitleCase) |>
-  tt()
+  rename(`%` = p) |>
+  tt() |>
+  group_tt(list("Software" = 1, "Documents" = 4))
+```
+
+```{r data-merge-categories}
+# Merge smaller categories for subsequent analyses
+oarch <- mutate(
+  oarch,
+  category = recode(
+    category,
+    "Specifications, protocols and schemas" = "Other documents",
+    "Products" = "Other documents",
+    "Guides" = "Other documents",
+    .default = category
+  ),
+  category = factor(
+    category, 
+    c("Packages and libraries", 
+      "Standalone software",
+      "Scripts",
+      "Lists and datasets",
+      "Other documents")
+  )
+)
 ```
@@
 ```{r tbl-platforms}
-#| tbl-cap: Platforms and programming languages used by open archaeology projects.
+#| tbl-cap: Platforms and programming languages used by open archaeology projects
 oarch |>
@@
   summarise(n = sum(n)) |>
+  mutate(
+    p = n / sum(n),
+    p = paste0(format(p * 100, trim = TRUE, digits = 1), "%")
+  ) |>
   arrange(platform == "Other", -n) |>
@@
 
-Our analysis of tags highlights aspects of archaeological work that software developers are inclined to contribute to [@fig-tags].
+Our analysis of tags highlights aspects of archaeological work that software developers are inclined to contribute to [@tbl-tags].
 The most common tags unsurprisingly deal with work that naturally benefits from advanced information processing afforded by computers, such as statistical analysis, sample calibration, geographical analysis, data management, and chronological modelling.
@@
 
-```{r fig-tags}
-#| fig-cap: Tags by category
-oarch %>%
+```{r tbl-tags}
+#| tbl-cap: Themes of open archaeology projects
+oarch |>
   pivot_longer(c(tag1, tag2, tag3, tag4, tag5), values_to = "tag",
-               values_drop_na = TRUE) %>%
-  drop_na(tag) %>%
-  mutate(category = recode(category,
-                                  "Specifications, protocols and schemas" = "Documents",
-                                  "Products" = "Documents",
-                                  "Guides" = "Documents",
-                                  "Lists and datasets" = "Documents",
-                                  .default = category)) %>%
-  group_by(tag, category) %>%
-  summarise(n = n(), .groups = "drop_last") %>%
-  mutate(total = sum(n)) %>%
-  ggplot(aes(fct_reorder(tag, total), n, fill = category)) +
-  geom_col() +
-  scale_fill_bright() +
-  coord_flip() +
-  labs(
-    x = NULL, y = "Number of projects",
-    fill = NULL
-  ) +
-  theme_oarch() +
-  theme(
-    axis.text.y = element_text(size = 7)
-  )
+               values_drop_na = TRUE) |>
+  drop_na(tag) |>
+  group_by(tag, category) |>
+  summarise(n = n(), .groups = "drop_last") |>
+  mutate(total = sum(n)) |>
+  pivot_wider(names_from = category, values_from = n) |>
+  mutate(across(everything(), \(x) replace_na(x, 0))) |>
+  arrange(-total) |>
+  rename(theme = tag) |>
+  rename_with(tools::toTitleCase) |>
+  tt()
 ```
@@
 ```{r data-licenses}
-# open-archaeo_licenses.csv contains a list of licenses that had to be manually verified.
-This is necessary due to the fact that GitHub and CRAN have different conventions for naming the files containing the license.
+# open-archaeo_licenses.csv contains a list of licenses that had to be manually 
+# verified. This is necessary due to the fact that GitHub and CRAN have 
+# different conventions for naming the files containing the license.
 oarch_licenses <- read_csv(here("analysis/data/raw_data/open-archaeo_licenses.csv"))
@@
 ```{r tbl-repo-licenses}
-#| tbl-cap: Summary of licenses used by projects with GitHub repositories
+#| tbl-cap: Licenses used by open archaeology projects on GitHub
 oarch |>
@@
   count(license) |>
-  mutate(license = replace_na(license, "None detected")) |>
+  mutate(
+    license = replace_na(license, "None detected"),
+    p = n / sum(n),
+    p = paste0(format(p * 100, trim = TRUE, digits = 1), "%")
+  ) |>
   arrange(-n) |>
   rename_with(tools::toTitleCase) |>
+  rename(`%` = p) |>
   tt()
@@
 ```{r fig-github-cumulative}
-#| fig-cap: Cumulative growth of open archaeological software in terms of number of commits and number of repositories, and broken down by category.
-oarch %>% 
-  mutate(
-    lumped_category = recode(category,
-                             "Specifications etc." = "Documents",
-                             "Products" = "Documents",
-                             "Guides" = "Documents",
-                             "Lists and datasets" = "Documents",
-                             .default = category),
-    lumped_category = factor(lumped_category, c("Scripts",
-                                                "Documents",
-                                                "Packages and libraries",
-                                                "Standalone software")) ) ->
-  oarch_lumped
-
-p_gh_commits_cum <- oarch_lumped %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  mutate(date = floor_date(datetime, "week")) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
+#| fig-cap: Growth of open archaeology projects on GitHub
+p_gh_commits_cum <- oarch |>
+  unnest(gh_commits) |>
+  drop_na(sha) |>
+  mutate(date = floor_date(datetime, "week")) |>
+  group_by(category, date) |>
+  summarise(n_commits = n(), .groups = "drop_last") |>
+  mutate(cum_commits = cumsum(n_commits)) |>
   ggplot(aes(date, cum_commits,
-             fill = lumped_category,
-             colour = lumped_category,
-             label = lumped_category)) +
+             fill = fct_rev(category),
+             colour = fct_rev(category))) +
   geom_area() +
@@
            vjust = 0) +
-  scale_x_datetime(date_breaks = "2 years", date_labels = "%Y") +
+  scale_x_datetime(limits = c(ymd("2006-12-31", tz = "UTC"), now("UTC")),
+                   date_breaks = "2 years", 
+                   date_labels = "%Y",
+                   expand = c(0,0)) +
   scale_y_continuous(labels = scales::comma) +
@@
   labs(
-    y = "cumulative commits",
+    y = "Cumulative commits",
     x = NULL, fill = NULL
@@
 
-# TODO: extend to end
-p_gh_repos_cum <- oarch_lumped %>%
-  unnest(gh_commits) %>%
-  drop_na(sha) %>%
-  group_by(repo) %>% 
-  filter(datetime == min(datetime)) %>% # Only first commits
-  mutate(date = floor_date(datetime, "week")) %>%
-  group_by(lumped_category, date) %>%
-  summarise(n_commits = n(), .groups = "drop_last") %>%
-  mutate(cat_total = sum(n_commits)) %>%
-  mutate(cum_commits = cumsum(n_commits)) %>%
+p_gh_repos_cum <- oarch |>
+  unnest(gh_commits) |>
+  drop_na(sha) |>
+  group_by(repo) |> 
+  filter(datetime == min(datetime)) |> # Only first commits
+  mutate(date = floor_date(datetime, "week")) |>
+  group_by(category, date) |>
+  summarise(n_commits = n(), .groups = "drop_last") |>
+  mutate(cat_total = sum(n_commits)) |>
+  mutate(cum_commits = cumsum(n_commits)) |>
+  select(category, date, cum_commits) |>
+  group_by(category) |>
+  nest() |>
+  mutate(data = map(data, \(x) { # Extend the range of each category to today
+    bind_rows(x, list(date = today(), cum_commits = max(x$cum_commits)))
+  })) |>
+  unnest(c(data)) |>
   ggplot(aes(date, cum_commits,
-             fill = lumped_category,
-             colour = lumped_category)) +
+             fill = fct_rev(category),
+             colour = fct_rev(category))) +
   geom_area() +
@@
   labs(
-    y = "cumulative repositories",
+    y = "Cumulative repositories",
     x = NULL, fill = NULL

### 2024-02-02 analysis/_05-collaboration.qmd
Mostly minor but necessary fixes to code and text
Provided a little explanation for the manual licenses file, included a missing value to the licenses file, corrected some citations, added section slugs and encoded links to sections, removed various completed or unnecessary todos and notes, removed dendrogram, verified and updated characterization of clusters, fixed references to graph objects
Revise figures and tables
Removing superfluous code, consistent style, consistent captions and labels, etc.

@@
-# Collaborative practices
+# Collaborative practices {#sec-collaboration}
 
@@
 ```{r fig-lifespan}
-#| fig-cap: Lifespan of repositories. Each point indicates a commit; excludes repositories with only one commit.
+#| fig-cap: Lifespan of open archaeology repositories. Each point indicates a commit; excludes repositories with only one commit.
 oarch_gh |>
@@
 ```{r fig-lifespan-rate}
-#| fig-cap: Lifespan, commit rate and contributors to repositories. Note that both scales are logarithmic.
-# TODO: add marginal histograms?
+#| fig-cap: Lifespan and commit rate of open archaeology repositories
 oarch_lifespan %>%
-  ggplot(aes(x = lifespan, y = commit_rate, size = n_contribs)) +
-  geom_point(shape = 21, colour = "black", fill = alpha("black", 0.3)) +
+  ggplot(aes(x = lifespan, y = commit_rate, fill = category, 0.3, size = n_contribs)) +
+  geom_point(shape = 21, colour = "black", alpha = 0.7) +
   scale_x_log10() +
   scale_y_log10() +
-  labs(x = "Lifespan (days)", y = "Commits per day", size = "Contributors") +
+  scale_fill_bright() +
+  scale_size(trans = "log10") +
+  labs(x = "Lifespan (days)", y = "Commits per day", 
+       fill = "Category", size = "Contributors") +
   theme_oarch()
@@
 ```{r fig-collab-features}
-#| fig-cap: Use of GitHub collaboration features (forks, issues, comments, stars) across repositories
+#| fig-cap: Use of GitHub collaboration features in open archaeology repositories
 oarch_gh |>
@@
 ```{r fig-contributions}
-#| fig-cap: Distribution of contributions in multi-contributor projects
+#| fig-cap: Distribution of contributions in multi-contributor open archaeology repositories
 oarch_contribs_by_repo |>
@@
   scale_y_continuous(labels = scales::percent) +
-  scale_fill_highcontrast() +
+  scale_fill_bright() +
   labs(x = "Individual repositories", y = "Contributions", fill = NULL) +

### 2024-02-02 analysis/_06-network.qmd
Revise figures and tables
Removing superfluous code, consistent style, consistent captions and labels, etc.

@@
 
-```{r fig-graph-repo-user, eval=FALSE}
-#| fig-cap: Interactive repository–user graph.
+```{r fig-graph-repo-user}
+#| fig-cap: Graph of open archaeology repositories and users connected by contributions.
+Darker edges indicate a great number of contributions.
+Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002].
+Excludes isolate nodes.
 # TODO: is there a khroma scale that can accomodate this many groups?
 g_repo_user <- oarch_graph_repo_user |>
-  ggraph(layout = "fr", repulserad = 10, area = 10000, niter = 5000, grid = "nogrid") +
+  ggraph(layout = "fr", niter = 5000, grid = "nogrid") +
   geom_edge_link(aes(group = contrib, alpha = n_contrib)) +
@@
   scale_colour_discrete(guide = guide_none()) +
-  labs(edge_alpha = "Common contributors") +
-  theme(panel.background = element_rect(fill = "white"))
+  labs(shape = NULL) +
+  theme(panel.background = element_rect(fill = "white"), legend.position = "bottom")
 
@@
 
-```{r fig-graph-repo-repo, eval=FALSE}
-#| fig-cap: Interactive repository–repository graph.
-Repositories are connected by contributions by common users, with darker edges indicating more common contributors.
+```{r fig-graph-repo-repo}
+#| fig-cap: Graph of open archaeology repositories connected by common contributors.
+Darker edges indicate a great number of common contributors.
 Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002].
@@
   scale_colour_discrete(guide = guide_none()) +
-  labs(edge_alpha = "Common contributors") +
   theme(panel.background = element_rect(fill = "white"))
@@
 ```{r fig-graph-user-user, eval=FALSE}
-#| fig-cap: Interactive user–user graph.
-Users are connected by contributions by common repositories, with darker edges indicating more common repositories.
-Node colour indicates cluster membership according to the edge–betweenness method [girvan2002].
+#| fig-cap: Graph of open archaeology users connected by contributions to common repositories.
+Darker edges indicate a great number of common repositories.
+Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002].
 Excludes isolate nodes.
@@
   filter(degree > 170) |>
-  ggraph(layout = "fr", repulserad = 10, area = 100000, niter = 10000000, grid = "nogrid") +
+  ggraph(layout = "fr", niter = 10000000, grid = "nogrid") +
   geom_edge_link(aes(alpha = n)) +
@@
   scale_colour_discrete(guide = guide_none()) +
-  labs(edge_alpha = "Common repositories") +
   theme(panel.background = element_rect(fill = "white"))
@@
 `r percent(graph_n_repos / graph_n_repos_total)` of repositories and `r percent(graph_n_users / graph_n_users_total)` of users in our dataset are connected to at least one other repository or user.
-Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-repo and @fig-graph-user-user).
+Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-user and @fig-graph-repo-repo).
 
@@
 ```{r tbl-repo-betweenness}
-#| tbl-cap: Most central repositories (weighted node betweenness) in the repository–repository network
+#| tbl-cap: Repositories ranked by centrality to the repository–repository network. Centrality is measured by node betweenness weighted by number of contributions.
 oarch_graph_repo_repo |>
-  mutate(betweenness = centrality_betweenness()) |>
   as_tibble() |>
@@
   unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
-  select(repo, betweenness, category, tags, n_commits) |>
+  transmute(rank = row_number(), repo, category, tags, n_commits) |>
   rename(repository = repo, commits = n_commits) |>
@@
 ```{r fig-repo-betweenness-age-length}
-#| fig-cap: Log-log correlation between repository centrality (weighted node betweenness), age (left) and length (right)
+#| fig-cap: Repository centrality by age (left) and length (right). Centrality is measured by node betweenness weighted by number of contributions.
 left <- oarch_graph_repo_repo |>
@@
   geom_smooth(method = "glm") +
-  labs(x = "Repository age (days)", y = "Centrality (betweenness)") +
+  labs(x = "Repository age (days)", y = "Centrality") +
   theme_oarch()
@@
 ```{r fig-repo-betweenness-category-platform}
-#| fig-cap: Mean repository centrality (weighted node betweenness) by open-archaeo category (top) and platform (bottom)
+#| fig-cap: Mean repository centrality by category (top) and platform (bottom). Centrality is measured by node betweenness weighted by number of contributions.
 top <- oarch_graph_repo_repo |>
@@
   geom_col() +
-  scale_y_log10() +
+  scale_y_log10(labels = NULL) +
   coord_flip() +
-  labs(y = "Centrality (weighted betweenness)", x = NULL) +
+  labs(y = "Centrality →", x = NULL) +
   theme_oarch()
@@
 
-```{r fig-user-betweenness-contribs, eval=TRUE}
-#| fig-cap: Log-log correlation between user centrality (weighted node betweenness) in the user-user graph and total number of contributions
+```{r fig-user-betweenness-contribs}
+#| fig-cap: User centrality by total number of contributions. Centrality is measured by node betweenness weighted by number of contributions.
 oarch_graph_user_user |>
@@
   left_join(oarch_contributors, by = c("user" = "contributor")) |>
-  ggplot(aes(betweenness, contribs)) +
+  ggplot(aes(contribs, betweenness)) +
   geom_point() +
@@
   geom_smooth(method = "glm") +
-  labs(x = "Centrality (betweenness)", y = "Total contributions") +
+  labs(x = "Total contributions", y = "Centrality") +
   theme_oarch()

### 2024-02-02 analysis/openarchaeo_collaboration.qmd
Revise figures and tables
Removing superfluous code, consistent style, consistent captions and labels, etc.

--- 2023-10-12 analysis/openarchaeo_collaboration.qmd
@@
 library("fs")
-library("geomtextpath")
+library("ggdist")
 library("ggiraph")
 library("ggraph")
-library("gt")
 library("here")

### 2024-02-02 analysis/_05-collaboration.qmd
Clarify fig-collab-features caption

@@
 ```{r fig-collab-features}
-#| fig-cap: Use of GitHub collaboration features in open archaeology repositories
+#| fig-cap: Box plot showing use of GitHub collaboration features in open archaeology repositories
 oarch_gh |>

### 2024-02-02 analysis/_04-open_archaeology.qmd
Encode percentages in section 4

@@
 
-```{r data-merge-categories}
+```{r data-categories-platforms}
 # Merge smaller categories for subsequent analyses
@@
 )
-```
 
-```{r num-p_platform}
-p_platform <- sum(!is.na(oarch$platform)) / nrow(oarch)
+n_platform <- sum(!is.na(oarch$platform))
+p_platform <- n_platform / nrow(oarch)
+
+n_standalone <- nrow(filter(oarch, category == "Standalone software"))
+p_standalone <- n_standalone / nrow(oarch)
+
+n_datasets <- nrow(filter(oarch, tag1 == "Datasets" | tag2 == "Datasets" | 
+                            tag3 == "Datasets" | tag4 == "Datasets" | 
+                            tag5 == "Datasets"))
+p_datasets <- n_datasets / nrow(oarch)
+
+n_scripts <- nrow(filter(oarch, category == "Scripts"))
+p_scripts <- n_scripts / nrow(oarch)
 ```
@@
 The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
-Others create standalone software <!-- percent --> that can be run independently of such platforms, for example desktop or web apps.
-A significant number of projects also comprise of datasets <!-- percent --> and non-packaged code snippets <!-- percent --> that have been made available for general use.
+Others create standalone software (`r percent(p_standalone)`) that can be run independently of such platforms, for example desktop or web apps.
+A significant number of projects also comprise of datasets (`r percent(p_datasets)`) and non-packaged code snippets (`r percent(p_scripts)`) that have been made available for general use.
 

### 2024-02-02 analysis/_06-network.qmd
Only show largest clusters in graph viz

@@
             subgraph = as_factor(group_components()),
-            group = as_factor(group_edge_betweenness())
+            cluster = as_factor(group_edge_betweenness())
             ) ->
@@
     subgraph = as_factor(group_components()),
-    group = as_factor(group_edge_betweenness())
+    cluster = as_factor(group_edge_betweenness())
   ) ->
@@
     subgraph = as_factor(group_components())
-    # group = as_factor(group_edge_betweenness()) # <- very slow!
+    # cluster = as_factor(group_edge_betweenness()) # <- very slow!
   ) ->
@@
   as_tibble() |>
-  count(group)
+  count(cluster)
 
@@
 Darker edges indicate a great number of contributions.
-Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002].
+Node colour indicates membership of the largest clusters according to the edge–betweenness method [@girvan2002].
 Excludes isolate nodes.
-# TODO: is there a khroma scale that can accomodate this many groups?
 g_repo_user <- oarch_graph_repo_user |>
+  group_by(cluster) |>
+  mutate(cluster_n = n()) |>
+  ungroup() |>
+  mutate(
+    cluster_rank = dense_rank(desc(cluster_n)),
+    cluster = if_else(cluster_rank < 7, cluster, NA)
+  ) |>
   ggraph(layout = "fr", niter = 5000, grid = "nogrid") +
   geom_edge_link(aes(group = contrib, alpha = n_contrib)) +
-  geom_point_interactive(aes(x, y, colour = group, shape=node_type, tooltip = node)) + 
+  geom_point_interactive(aes(x, y, colour = cluster, shape=node_type, tooltip = node)) + 
   scale_edge_alpha(trans = "log", guide = guide_none()) +
@@
 Darker edges indicate a great number of common contributors.
-Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002].
+Node colour indicates membership of the largest clusters according to the edge–betweenness method [@girvan2002].
 Excludes isolate nodes.
-# TODO: is there a khroma scale that can accomodate this many groups?
 g_repo_repo <- oarch_graph_repo_repo |>
-  ggraph(layout = "graphopt", mass = 100) +
+  group_by(cluster) |>
+  mutate(cluster_n = n()) |>
+  ungroup() |>
+  mutate(
+    cluster_rank = dense_rank(desc(cluster_n)),
+    cluster = if_else(cluster_rank < 7, cluster, NA)
+  ) |>
+  ggraph(layout = "fr", niter = 5000, grid = "nogrid") +
   geom_edge_link(aes(alpha = n)) +
-  geom_point_interactive(aes(x, y, colour = group, tooltip = repo)) + 
+  geom_point_interactive(aes(x, y, colour = cluster, tooltip = repo)) + 
   scale_edge_alpha(trans = "log", guide = guide_none()) +
@@
 Darker edges indicate a great number of common repositories.
-Node colour indicates cluster membership according to the edge–betweenness method [@girvan2002].
+Node colour indicates membership of the largest clusters according to the edge–betweenness method [@girvan2002].
 Excludes isolate nodes.
-# TODO: is there a khroma scale that can accomodate this many groups?
 g_user_user <- oarch_graph_user_user |>

### 2024-02-02 analysis/_03-data_methodology.qmd
Remove obsolete TODO

@@
   oarch
-
-# TODO: exclude peripheral repos that throw things off
-# * Big boys like QField (see _06-network.qmd)
-# * ArchaeoLines – ends up picking up the whole stellarium repo, which is huge
-# * aDNA??
 ```

### 2024-02-02 analysis/_06-network.qmd
Remove cltk and re-enable user-user graph

@@
     betweenness = centrality_betweenness(weights = n),
-    subgraph = as_factor(group_components())
-    # cluster = as_factor(group_edge_betweenness()) # <- very slow!
+    subgraph = as_factor(group_components()),
+    cluster = as_factor(group_edge_betweenness())
   ) ->
@@
     cluster_rank = dense_rank(desc(cluster_n)),
-    cluster = if_else(cluster_rank < 7, cluster, NA)
+    cluster = if_else(cluster_rank <= 7, cluster, NA)
   ) |>
@@
 
-```{r fig-graph-user-user, eval=FALSE}
+```{r fig-graph-user-user}
 #| fig-cap: Graph of open archaeology users connected by contributions to common repositories.
@@
 g_user_user <- oarch_graph_user_user |>
-  filter(degree > 170) |>
-  ggraph(layout = "fr", niter = 10000000, grid = "nogrid") +
+  group_by(cluster) |>
+  mutate(cluster_n = n()) |>
+  ungroup() |>
+  mutate(
+    cluster_rank = dense_rank(desc(cluster_n)),
+    cluster = if_else(cluster_rank <= 7, cluster, NA)
+  ) |>
+  ggraph(layout = "fr", niter = 5000, grid = "nogrid") +
   geom_edge_link(aes(alpha = n)) +
-  geom_point_interactive(aes(x, y, tooltip = user)) + 
+  geom_point_interactive(aes(x, y, colour = cluster, tooltip = user)) + 
   scale_edge_alpha(trans = "log", guide = guide_none()) +

### 2024-02-02 analysis/openarchaeo_collaboration.qmd
Remove unused package

--- 2024-01-24 analysis/openarchaeo_collaboration.qmd
@@
 library("fs")
-library("geomtextpath")
 library("ggiraph")

### 2024-02-02 R/forceatlas2.R
Update dependencies and fix R CMD check notes

@@ -1,181 +1,0 @@
-#' ForceAtlas2 layout
-#'
-#' Copied from <https://github.com/analyxcompany/ForceAtlas2> (licensed GPL>=2)
-#'
-#' @param graph An igraph network or a data frame of three columns: source, target, and weights.
-#' @param directed Logical. TRUE if the network is directed. Ignored if graph is an igraph object.
-#' @param iterations Number of iterations to be performed.
-#' @param linlog Logical. If TRUE the algorithm uses logarithmic attraction force `F <- log (1+F)`
-#' @param pos A data frame or matrix (NumberOfNodes x dimension) of the initial locations of points. If NULL the initial positions are random.
-#' @param nohubs Logical. If TRUE nodes with high indegree have more central position than nodes with outdegree (for directed graphs).
-#' @param k Is the repel constant: the greater the constant k the stronger the repulsion force between points.
-#' @param gravity Gravity constant: indicates how strongly the nodes should be attracted to the center of gravity.
-#' @param ks Speed constant: the greater the value of `ks` the more movement the nodes make under the acting forces.
-#' @param ksmax Limits the speed from above.
-#' @param delta Modify attraction force, the weights are raised to the power of `delta`
-#' @param center Center of gravity.
-#' @param tolerance Tolerance to swinging constant.
-#' @param dim Dimension of the positions.
-#' @param plotstep is the frequency of plotting intermediate iterations.
-#' @param plotlabels Logical. If TRUE the labels should be included in the intermediate interations plot
-#' @return When graph is an igraph object, the function returns a matrix with the positions. If graph is a data frame, this is a data frame with the names of the nodes and the positions.
-#' @examples
-#' library(igraph)
-#' g <- graph.ring(50)
-#' layout <- layout.forceatlas2(g, iterations=3000, plotstep=100)
-#' plot(g, layout=layout)
-#'
-#' @source
-#' Jacomy M, Venturini T, Heymann S, Bastian M (2014) ForceAtlas2, a Continuous Graph Layout Algorithm for Handy Network Visualization Designed for the Gephi Software. PLoS ONE 9(6): e98679. <https://doi.org/10.1371/journal.pone.0098679>
-#'
[…]

### 2024-02-02 R/gh.R

@@ -19,7 +19,7 @@
 #' @name gh_query
 #'
 #' @examples
-#' gh_repo("joeroe/era")
+#' gh_repo_info("joeroe/era")
 #' gh_lang("joeroe/era")
 #' gh_contrib("joeroe/era")
 #' gh_issue("joeroe/era")
@@ -201,7 +201,7 @@ ghq <- function(endpoint, ...) {
   if (any(is.na(list(...)))) NA
   else {
     res <- tryCatch(
-      gh(endpoint, ..., .limit = Inf),
+      gh::gh(endpoint, ..., .limit = Inf),
       http_error_404 = gh_warn_404
     )
     if (length(res) > 0) res
@@ -209,6 +209,8 @@ ghq <- function(endpoint, ...) {
   }
 }
 
+#' keyword internal
+#' @noRd
 gh_warn_404 <- function(c) {
   w <- stringr::str_extract(as.character(c), "URL not found:.*$")
   rlang::warn(paste0("GitHub API ", w), class = "http_error_404")

### 2024-02-02 R/graph.R

@@ -1,61 +1,73 @@
+#' Functions for extracting a graph from GitHub API data
+#'
+#' @param oarch Data frame containing data from the GitHub API
+#' @param ... Passed to [dplyr::select()]
+#'
+#' @rdname gh_graph
 #' @export
-nodes_repos_users <- function(oarch_graph) {
-  oarch_graph |>
-    pivot_longer(c(repo, user), names_to = "node_type", values_to = "node") |>
-    select(node, node_type) |>
-    distinct()
+nodes_repos_users <- function(oarch) {
+  oarch |>
+    tidyr::pivot_longer(c(repo, user), names_to = "node_type", values_to = "node") |>
+    dplyr::select(node, node_type) |>
+    dplyr::distinct()
 }
 
+#' @rdname gh_graph
 #' @export
-nodes_repos <- function(oarch_graph, ...) {
-  oarch_graph |>
-    select(repo, ...) |>
-    distinct()
+nodes_repos <- function(oarch, ...) {
+  oarch |>
+    dplyr::select(repo, ...) |>
+    dplyr::distinct()
 }
 
+#' @rdname gh_graph
 #' @export
-nodes_users <- function(oarch_graph, ...) {
-  oarch_graph |>
-    select(user, ...) |>
-    distinct()
+nodes_users <- function(oarch, ...) {
+  oarch |>
+    dplyr::select(user, ...) |>
+    dplyr::distinct()
 }
 
+#' @rdname gh_graph
 #' @export
-edges_user_to_repo <- function(oarch_graph) {
-  oarch_graph |>
-    rename(to = repo, from = user)
+edges_user_to_repo <- function(oarch) {
+  oarch |>
+    dplyr::rename(to = repo, from = user)
 }
 
+#' @rdname gh_graph
 #' @export
-edges_repo_to_user <- function(oarch_graph) {
-  oarch_graph |>
-    rename(to = user, from = repo)
+edges_repo_to_user <- function(oarch) {
+  oarch |>
+    dplyr::rename(to = user, from = repo)
 }
 
+#' @rdname gh_graph
 #' @export
-edges_common_users <- function(oarch_graph) {
-  oarch_graph |>
-    left_join(oarch_graph, by = "user") |>
-    select(repo.x, user, repo.y) |>
-    filter(repo.x != repo.y) |>
-    mutate(repo.xy = map2_chr(repo.x, repo.y,
[…]

### 2024-02-02 R/utils.R

@@ -1,5 +1,9 @@
+#' Count number of rows in a filtered data frame
+#'
+#' @param .data Data frame
+#' @param ... Filters passed to [dplyr::filter()]
+#'
 #' @export
-#' @noRd
 n_filter <- function(.data, ...) {
   nrow(dplyr::filter(.data, ...))
 }

### 2024-02-02 analysis/_03-data_methodology.qmd
Add links to tables

@@
 # TODO: update and replace with local file before submission
-read_csv("https://github.com/zackbatist/open-archaeo/raw/2023-08-29/open-archaeo.csv") |>
+read_csv("https://github.com/zackbatist/open-archaeo/raw/master/open-archaeo.csv") |>
   mutate(category = na_if(category, "?")) ->
@@
 oarch_forges |>
-  mutate(p = paste0(format(p * 100, trim = TRUE, digits = 1), "%")) |>
+  mutate(
+    repo_host = recode(
+      repo_host,
+      "GitHub" = "[GitHub](https://github.com/)",
+      "Codeberg" = "[Codeberg](https://codeberg.org/)",
+      "GitLab" = "[GitLab](https://about.gitlab.com/)",
+      "BitBucket" = "[Bitbucket](https://bitbucket.org/product/)",
+      "Launchpad" = "[Launchpad](https://launchpad.net/)"
+    ),
+    p = paste0(format(p * 100, trim = TRUE, digits = 1), "%")
+  ) |>
   rename(Host = repo_host, `%` = p) |>
-  tt()
+  tt() |>
+  format_tt(1, markdown = TRUE)
 ```

### 2024-02-02 analysis/_04-open_archaeology.qmd
Add links to tables

@@
 tribble(
-  ~category,                               ~kind,       ~scope,
-  "Packages and libraries",                "Software", "Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.",
-  "Standalone software",                   "Software", "Software that may be operated without needing to first access an underlying platform.",
-  "Scripts",                               "Software", "Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.",
-  "Specifications, protocols and schemas", "Documents", "A formal data structure or framework intended to be used as a model.",
-  "Products",                              "Documents", "Stable outcomes of creative work.",
-  "Guides",                                "Documents", "An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.",
-  "Lists and datasets",                    "Documents", "A series of consistently organized observations assembled with purpose."
+  ~category, ~url,                               ~kind,       ~scope,
+  "Packages and libraries", "https://open-archaeo.info/categories/packages-and-libraries/",                "Software", "Sets of functions assembled with clear purpose, and made accessible using standards established by an underlying platform.",
+  "Standalone software", "https://open-archaeo.info/categories/standalone-software/",                   "Software", "Software that may be operated without needing to first access an underlying platform.",
+  "Scripts", "https://open-archaeo.info/categories/scripts/",                               "Software", "Sets of pragmatically assembled mutable functions, often lacking complete documentation or adherence to protocols that would otherwise facilitate secondary use outside their original contexts of creation.",
+  "Specifications, protocols and schemas", "https://open-archaeo.info/categories/specifications-protocols-and-schemas/", "Documents", "A formal data structure or framework intended to be used as a model.",
+  "Products", "https://open-archaeo.info/categories/products/",                              "Documents", "Stable outcomes of creative work.",
+  "Guides", "https://open-archaeo.info/categories/guides/",                                "Documents", "An educational resource or documented protocol meant to instruct readers how to apply relevant tools or techniques.",
+  "Lists and datasets", "https://open-archaeo.info/categories/lists-and-datasets/",                    "Documents", "A series of consistently organized observations assembled with purpose."
 ) |>
   left_join(oarch, by = "category") |>
-  count(kind, category, scope) |>
+  count(kind, category, url, scope) |>
   mutate(
+    category = glue("[{category}]({url})"),
     p = n / sum(n),
@@
   tt() |>
-  group_tt(list("Software" = 1, "Documents" = 4))
+  group_tt(list("Software" = 1, "Documents" = 4)) |>
+  format_tt(1, markdown = TRUE)
 ```
@@
   pivot_wider(names_from = category, values_from = n) |>
-  mutate(across(everything(), \(x) replace_na(x, 0))) |>
+  mutate(
+    tag = glue("[{tag}](https://open-archaeo.info/{slugify(tag)})"),
+    across(everything(), \(x) replace_na(x, 0))
+  ) |>
   arrange(-total) |>
@@
 #| tbl-cap: Licenses used by open archaeology projects on GitHub
+# TODO: Add links to licenses?
 oarch |>

### 2024-02-02 analysis/_05-collaboration.qmd
Add links to tables

@@
 #| fig-cap: Lifespan of open archaeology repositories. Each point indicates a commit; excludes repositories with only one commit.
-oarch_gh |>
+fig_lifespan <- oarch_gh |>
   unnest(gh_commits) |>
@@
   )
+
+girafe(fig_lifespan)
 ```

### 2024-02-02 R/utils.R

@@ -7,3 +7,15 @@
 n_filter <- function(.data, ...) {
   nrow(dplyr::filter(.data, ...))
 }
+
+#' Mimic hugo's slugification algorithm
+#'
+#' @param x Character vector to slugify
+#'
+#' @export
+slugify <- function(x) {
+  x |>
+    tolower() |>
+    stringr::str_remove_all(",") |>
+    stringr::str_replace_all("\\W", "-")
+}

### 2024-02-02 analysis/_03-data_methodology.qmd
Update open-archaeo and GitHub data

@@
 ```{r data-oarch}
-# TODO: update and replace with local file before submission
-read_csv("https://github.com/zackbatist/open-archaeo/raw/master/open-archaeo.csv") |>
-  mutate(category = na_if(category, "?")) ->
-  oarch
+# open-archaeo version 2024-02-02
+# https://doi.org/10.5281/zenodo.10611198
+oarch <- read_csv(here("analysis/data/raw_data/open-archaeo_2024-02-02.csv"))
 
 # Extract clean repos from GitHub URLs
-oarch |>
-  mutate(
+oarch <- mutate(
+  oarch,
   gh_repo = str_match(oarch$github, "^https://github.com/([\\w-\\.]+/[\\w-\\.]+)/?.*?$")[,2]
-  ) ->
-  oarch
+)
 ```
@@
 
-# Exclude from GitHub API data collection because does not have own repo
+# Exclude from GitHub API data collection because they part of a larger
+# repository (or forked from one)
 oarch$gh_repo[oarch$item_name == "ArchaeoLines"] <- NA
+oarch$gh_repo[oarch$gh_repo == "carpentries-incubator/R-archaeology-lesson"] <- NA
 ```
@@
 oarch <- readRDS(cached_gh_data)
+
+n_repos <- n_filter(oarch, !is.na(gh_repo))
+n_commits <- oarch |> drop_na(gh_repo) |> unnest(gh_commits) |> nrow()
+n_issues <- oarch |> drop_na(gh_repo) |> unnest(gh_issues) |> nrow()
+n_comments <- oarch |> drop_na(gh_repo) |> unnest(gh_comments) |> nrow()
+n_contributors <- oarch |> drop_na(gh_repo) |> unnest(gh_contribs) |> 
+  distinct(contributor) |> nrow()
 ```
@@
 
-Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API using the R package `gh` [@csardi2023].
-This included metadata on the GitHub repository and the programming languages it uses, as well as detailed data on each repository's individual contributors, commits, issues and comments.
+Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API (application programming interface).
 We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
 This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data included in open-archaeo itself.
+In total, we analysed data on `r n_repos` repositories, comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
 
-The open-archaeo dataset and the supplemental data from the GitHub API provide us with a rich resource to explore the makeup of the archaeological software development scene.
-We apply exploratory data analysis methods to break down and visualize the data, so that we may ascertain overall trends [@tukey1977].
+Taken together the open-archaeo directory and the supplemental data from the GitHub API provide a rich resource to explore the nature of collaborative software engineering in archaeology.
+Here we employ exploratory data analysis [@tukey1977] to identify and characterise overall trends visible in this rich dataset.
 In @sec-open-archaeology, our focus is on examining the general state of open source archaeological software and resource development.
@@
 The quantitative analyses and figures presented here were generated with `r version$version.string` [@rcoreteam2023].
-The full data and code is available in the compendium that accompanies this paper [@roe2023] <!-- reference should be updated to reflect the actual version number and date -->.
+The full data and code is available in the compendium that accompanies this paper [@roe2023] <!-- TODO: reference should be updated to reflect the actual version number and date -->.

### 2024-02-02 analysis/_04-open_archaeology.qmd
Update open-archaeo and GitHub data

@@
   arrange(desc(kind), -n) |>
-  select(-kind) |>
+  select(category, scope, n, p) |>
   rename_with(tools::toTitleCase) |>
@@
   mutate(
-    tag = glue("[{tag}](https://open-archaeo.info/{slugify(tag)})"),
+    tag = glue("[{tag}](https://open-archaeo.info/tags/{slugify(tag)})"),
     across(everything(), \(x) replace_na(x, 0))
@@
   rename_with(tools::toTitleCase) |>
-  tt()
+  tt() |>
+  format_tt(1, markdown = TRUE)
 ```
@@
 #| fig-cap: Growth of open archaeology projects on GitHub
+dt_last_commit <- oarch |> unnest(gh_commits) |> pull(datetime) |> max(na.rm = TRUE)
+
 p_gh_commits_cum <- oarch |>
@@
   mutate(cum_commits = cumsum(n_commits)) |>
+  group_by(category) |>
+  nest() |>
+  mutate(data = map(data, \(x) { # Extend the range of each category to end
+    bind_rows(x, list(date = dt_last_commit, cum_commits = max(x$cum_commits)))
+  })) |>
+  unnest(c(data)) |>
   ggplot(aes(date, cum_commits,
@@
            vjust = 0) +
-  scale_x_datetime(limits = c(ymd("2006-12-31", tz = "UTC"), now("UTC")),
+  scale_x_datetime(limits = c(ymd("2006-12-31", tz = "UTC"), dt_last_commit),
                    date_breaks = "2 years", 
@@
   group_by(category, date) |>
-  summarise(n_commits = n(), .groups = "drop_last") |>
-  mutate(cat_total = sum(n_commits)) |>
-  mutate(cum_commits = cumsum(n_commits)) |>
-  select(category, date, cum_commits) |>
+  summarise(n_repos = n(), .groups = "drop_last") |>
+  mutate(cat_total = sum(n_repos)) |>
+  mutate(cum_repos = cumsum(n_repos)) |>
+  select(category, date, cum_repos) |>
   group_by(category) |>
   nest() |>
-  mutate(data = map(data, \(x) { # Extend the range of each category to today
-    bind_rows(x, list(date = today(), cum_commits = max(x$cum_commits)))
+  mutate(data = map(data, \(x) { # Extend the range of each category to end
+    bind_rows(x, list(date = dt_last_commit, cum_repos = max(x$cum_repos)))
   })) |>
   unnest(c(data)) |>
-  ggplot(aes(date, cum_commits,
+  ggplot(aes(date, cum_repos,
              fill = fct_rev(category),
@@
 But use of git really began to take off around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
+<!-- TODO: Update and note plateau -->
 

### 2024-02-02 analysis/_06-network.qmd
Update open-archaeo and GitHub data

@@
     subgraph = as_factor(group_components()),
-    cluster = as_factor(group_edge_betweenness())
+    cluster = as_factor(group_edge_betweenness()) # <- slow!
   ) ->
@@
   geom_edge_link(aes(group = contrib, alpha = n_contrib)) +
-  geom_point_interactive(aes(x, y, colour = cluster, shape=node_type, tooltip = node)) + 
+  geom_point_interactive(aes(x, y, colour = cluster, shape = node_type, 
+                             tooltip = node)) + 
   scale_edge_alpha(trans = "log", guide = guide_none()) +
@@
   labs(shape = NULL) +
-  theme(panel.background = element_rect(fill = "white"), legend.position = "bottom")
+  theme(panel.background = element_rect(fill = "white"), 
+        legend.position = "bottom")
 
@@
 <!-- Predictors of repo centrality -->
+<!-- TODO: Needs an update. These repos are no longer in the top 10, and it's probably best not to be so specific given how volatile that is. -->
 The repositories most central to the network as a whole [@tbl-repo-betweenness] include several lists and directories, including open-archaeo itself.

### 2024-02-02 analysis/openarchaeo_collaboration.qmd
Update open-archaeo and GitHub data

@@
 library("ggraph")
+library("glue")
 library("here")

### 2024-02-02 analysis/_01-introduction.qmd
Mostly minor but necessary fixes to code and text
Provided a little explanation for the manual licenses file, included a missing value to the licenses file, corrected some citations, added section slugs and encoded links to sections, removed various completed or unnecessary todos and notes, removed dendrogram, verified and updated characterization of clusters, fixed references to graph objects
Updates to the text and general touch ups.
Moved stuff on git and github to a new background subsection, moved data caveats from conclusion to methods, updated some interpretations in section 6, update i

@@
-# Introduction
+# Introduction {#sec-introduction}
 
@@
 --JR -->That being said, archaeologists are prolific software developers.
-This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes <!-- TODO: encode -->XXX records, as of the time of this article's publication.
+This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes 501 records, as of the time of this article's publication.
 But beyond simply making their code available on the web, do archaeologists also implement social strategies to advance open source ideals?

### 2024-02-02 analysis/_02-background.qmd
Mostly minor but necessary fixes to code and text
Provided a little explanation for the manual licenses file, included a missing value to the licenses file, corrected some citations, added section slugs and encoded links to sections, removed various completed or unnecessary todos and notes, removed dendrogram, verified and updated characterization of clusters, fixed references to graph objects
Updates to the text and general touch ups.
Moved stuff on git and github to a new background subsection, moved data caveats from conclusion to methods, updated some interpretations in section 6, update i

@@
-# Open science and open source
+# Open science and open source {#sec-background}
 
@@
 In other words, the pressures and circumstances of being an archaeologist and doing archaeological research assert themselves when attempting to make practical use of these infrastructures, and therefore must be accounted for in their design and implementation.
-Here we aim to identify similar sources of dissonance with regards to the promise, potential, and actual implementation of open source software development models among archaeologists.
+In this paper we aim to identify similar sources of dissonance with regards to the promise, potential, and actual implementation of open source software development models among archaeologists.
+
+## Open Source Infrastructure
+
+Open source is an inherently internet-based development model and is supported by technical infrastructures which facilitate global distribution of labour and code.
+Here we provide a brief overview of key technologies that archaeologists have come to rely on as they develop open source software.
+
+Chief among these is git, a protocol designed to facilitate open and distributed contributions to a common code base.
+It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
+Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a fork in which they apply their modifications, and then request that their fork be merged into the central code base.
+After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent forks intact.
+
+Git is also designed to facilitate code review and version control.
+All modifications are tracked as "diffs", which highlight additions or deletions to the code base, including changes within individual files.
+Typically, a contributor will group a series of changes into a more comprehensive "commit" based on a specific task or part of a workflow.
+Commits are always accompanied by a message, in which the contributor (ideally) describes the reason and context for the changes included in the commit.
+Moreover, git assigns each commit a unique identifier and identifies the contributor by name and email address to ensure some degree of public accountability.
+
+Forges, or collaborative software development platforms like GitHub, GitLab and Codeberg are designed to facilitate open source software development by hosting public git repositories.
+However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
+These platforms also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as favourites, and maintain a public-facing profile that includes personally-identifying information (e.g.
+profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
+Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
+While these additional features are meant to complement and enhance the experience of contributing to open source projects, they are not actually part of the git protocol.
+See @tbl-glossary for a glossary of terminology used to describe many of these concepts and features, and which we use throughout this paper.
+
+```{r tbl-glossary}
+#| tbl-cap: Glossary of git and GitHub terminology
+tribble(
+  ~term, ~definition,
+  "Comment", "On GitHub, text post attached to an issue, including the first one that describes the issue",
+  "Commit", "Set of changes (addition, alteration, or deletion) to files in a repository that has been recorded by git as one entry in its log",
+  "Commit access", "Ability to make changes to a repository directly, without making a pull request",
+  "Contributor", "User that has made at least one commit to a specified repository",
+  "Follow", "Add activity by another user to a user's timeline",
+  "Forge", "Web-based platform for hosting, distributing and facilitating collaboration on version controlled computer code, e.g. GitHub, GitLab, Codeberg",
+  "Fork", "Copy of a repository owned by another user; forking is a prerequisite to making a pull request",
+  "git", "Open source version control software",
+  "GitHub", "Commercial platform that freely hosts git repositories and provides extended collaboration and social networking features, such as pull requests, issues and stars",
+  "GitLab", "Open source alternative to GitHub",
+  "CodeBerg", "Open source alternative to GitHub",
+  "Issue", "Feature of GitHub that records and tracks a bug report, feature request or other suggestion in a repository",
+  "Maintainer", "Individual that has overall control of a repository, generally assumed to be its primary contributor.
+Repositories can have multiple users with commit access in addition to the maintainer.",
+  "Merge", "Accept a pull request and incorporate its changes into a repository.",
+  "Organization", "Entity representing a group of users, which can also own repositories",
+  "Pull request", "Mechanism by which users that don't have commit access to a repository can contribute to it.
+The repository's maintainer or another user with commit access must decide whether to merge (accept) the changes, or decline them.",
+  "Repository", "Individual project that uses git for version control. Can include a mix of different types of files.",
+  "Star", "GitHub's version of a 'like', applied by users to a repository.",
+  "Timeline", "Chronological feed of GitHub activity from repositories a user has starred and other users they follow.
+Also includes repositories that a user is not following if they are 'trending' or determined relevant by GitHub's algorithm.",
+  "User", "On GitHub, an individual with an account that can own repositories",
+  "Version control", "System for tracking changes (additions, alterations, or deletions) in a set of files, typically but not exclusively computer code"
+) |>
+  rename_with(tools::toTitleCase) |>
+  tt() |>
+  format_tt(2, markdown = TRUE) |>
+  style_tt(j = 1, italic = TRUE)
+```

### 2024-02-02 analysis/_03-data_methodology.qmd
Updates to the text and general touch ups.
Moved stuff on git and github to a new background subsection, moved data caveats from conclusion to methods, updated some interpretations in section 6, update interpretation based on cumulative growth figure.

@@
 
-```{r tbl-glossary}
-#| tbl-cap: Glossary of git and GitHub terminology
-tribble(
-  ~term, ~definition,
-  "Comment", "On GitHub, text post attached to an issue, including the first one that describes the issue",
-  "Commit", "Set of changes (addition, alteration, or deletion) to files in a repository that has been recorded by git as one entry in its log",
-  "Commit access", "Ability to make changes to a repository directly, without making a pull request",
-  "Contributor", "User that has made at least one commit to a specified repository",
-  "Follow", "Add activity by another user to a user's timeline",
-  "Forge", "Web-based platform for hosting, distributing and facilitating collaboration on version controlled computer code, e.g. GitHub, GitLab, Codeberg",
-  "Fork", "Copy of a repository owned by another user; forking is a prerequisite to making a pull request",
-  "git", "Open source version control software",
-  "GitHub", "Commercial platform that freely hosts git repositories and provides extended collaboration and social networking features, such as pull requests, issues and stars",
-  "GitLab", "Open source alternative to GitHub",
-  "CodeBerg", "Open source alternative to GitHub",
-  "Issue", "Feature of GitHub that records and tracks a bug report, feature request or other suggestion in a repository",
-  "Maintainer", "Individual that has overall control of a repository, generally assumed to be its primary contributor.
-Repositories can have multiple users with commit access in addition to the maintainer.",
-  "Merge", "Accept a pull request and incorporate its changes into a repository.",
-  "Organization", "Entity representing a group of users, which can also own repositories",
-  "Pull request", "Mechanism by which users that don't have commit access to a repository can contribute to it.
-The repository's maintainer or another user with commit access must decide whether to merge (accept) the changes, or decline them.",
-  "Repository", "Individual project that uses git for version control. Can include a mix of different types of files.",
-  "Star", "GitHub's version of a 'like', applied by users to a repository.",
-  "Timeline", "Chronological feed of GitHub activity from repositories a user has starred and other users they follow.
-Also includes repositories that a user is not following if they are 'trending' or determined relevant by GitHub's algorithm.",
-  "User", "On GitHub, an individual with an account that can own repositories",
-  "Version control", "System for tracking changes (additions, alterations, or deletions) in a set of files, typically but not exclusively computer code"
-) |>
-  rename_with(tools::toTitleCase) |>
-  tt() |>
-  format_tt(2, markdown = TRUE) |>
-  style_tt(j = 1, italic = TRUE)
-```
-
-We compiled the dataset by browsing collaborative software development platforms such as GitHub, GitLab and Codeberg, which are designed to facilitate open source software development by hosting public git repositories.
-Git is a protocol designed to facilitate open and distributed contributions to a common code base.
-It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
-Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a fork in which they apply their modifications, and then request that their fork be merged into the central code base.
-After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent forks intact.
-
-Git is also designed to facilitate code review and version control.
-All modifications are tracked as "diffs", which highlight additions or deletions to the code base, including changes within individual files.
-Typically, a contributor will group a series of changes into a more comprehensive "commit" based on a specific task or part of a workflow.
-Commits are always accompanied by a message, in which the contributor (ideally) describes the reason and context for the changes included in the commit.
-Moreover, git assigns each commit a unique identifier and identifies the contributor by name and email address to ensure some degree of public accountability.
-
-In addition to serving as hosting providers for public git repositories, platforms like GitHub, GitLab and Codeberg support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
-They also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as favourites, and maintain a public-facing profile that includes personally-identifying information (e.g.
-profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
-Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
-However, these additional features are meant to complement and enhance the experience of contributing to open source projects, but are not actually part of the git protocol.
-See @tbl-glossary for a glossary of terminology used to describe many of these concepts and features, and which we use throughout this paper.
-
-We rely heavily on the social networking features provided by GitHub, GitLab and Codeberg to maintain open-archaeo.
+We compiled the dataset by browsing collaborative software development platforms, relying heavily on their social networking features.
 More specifically, we update open-archaeo by manually crawling through archaeologists' profiles on these platforms, as well as on other personal, professional, and institutional websites that describe and host additional archaeological software.
@@
 We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
-
-We excluded `r n_items_removed` records from open-archaeo from this analysis because they did not fit our "made by archaeologists, for archaeologists" criteria.
 
@@
 Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API (application programming interface).
+In total, we analysed data on `r n_repos` repositories, comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
+
 We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
 This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data included in open-archaeo itself.
-In total, we analysed data on `r n_repos` repositories, comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
+We also exclude collaboration through offline or private channels, or forms of collaboration we do not know about. 
+We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
+Our earliest data is from 2005 and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it [@ducke2013 ; @whallon1972].
 
-Taken together the open-archaeo directory and the supplemental data from the GitHub API provide a rich resource to explore the nature of collaborative software engineering in archaeology.
+These caveats notwithstanding, the open-archaeo directory and the supplemental data from the GitHub API provide a rich resource to explore the nature of collaborative software engineering in archaeology.
 Here we employ exploratory data analysis [@tukey1977] to identify and characterise overall trends visible in this rich dataset.

### 2024-02-02 analysis/_04-open_archaeology.qmd
Updates to the text and general touch ups.
Moved stuff on git and github to a new background subsection, moved data caveats from conclusion to methods, updated some interpretations in section 6, update interpretation based on cumulative growth figure.

@@
 
-As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
-But use of git really began to take off around 2014--2015, when we see an uptick in the rate of growth that has continued ever since.
-<!-- TODO: Update and note plateau -->
-
+As we can see <!-- this is no longer evident in the figure -->, archaeologists have been using git from even before GitHub was launched in 2008.
+But use of git really began to take off around 2014--2015, when we see an uptick in the rate of growth.
 Around this time we also see that GitHub starts being used to host documents and scripts.
 This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work in an open and somewhat nerdy way.
+
+However, since around 2022, the number of new GitHub repositories has significantly plateaud, while the cumulative number of commits has continued to rise.
+This may relate to a general emphasis on maintaining existing code and working on established projects, rather than spinning up new ones.

### 2024-02-02 analysis/_06-network.qmd
Updates to the text and general touch ups.
Moved stuff on git and github to a new background subsection, moved data caveats from conclusion to methods, updated some interpretations in section 6, update interpretation based on cumulative growth figure.

@@
 Node colour indicates membership of the largest clusters according to the edge–betweenness method [@girvan2002].
-Excludes isolate nodes.
+Excludes repositories with only a single user.
 g_repo_repo <- oarch_graph_repo_repo |>
@@
 Node colour indicates membership of the largest clusters according to the edge–betweenness method [@girvan2002].
-Excludes isolate nodes.
+Excludes users that have only contributed to a single repository.
 g_user_user <- oarch_graph_user_user |>
@@
 The central core bridges all the peripheral clusters.
-The central core is not uniform, and comprises several relatively discrete sub-groups representing collaborative sub-communities.
-While these sub-groups are internally cohesive, they exhibit enough connections to other members of the central core so as to not be considered as separate or peripheral clusters.
+The central core is not uniform, and comprises several relatively discrete sub-clusters representing collaborative sub-communities.
+While these sub-clusters are internally cohesive, they exhibit enough connections to other members of the central core so as to not be considered as separate or peripheral clusters.
 
@@
 Despite being the minority language, Python packages tend to be more central than R packages, on average.
-This might relate to the different ways in which R and Python developers call dependencies ???
-<!-- @ZB: I think this is interesting, but I couldn't really think of a general observation to make here. Maybe you can? -->
-<!-- I wonder what the median is? Also: cltk is a python script, and is the massive pink blob in the repo-repo and user-user networks, so that might skew things slightly -->
+Python projects in archaeology tend to be large, highly collaborative, and either infrastructural or of interest to other fields, so they have disproportionately high centrality.
+Our findings support the notion that there are significant development patterns when working with these two languages, and further analysis to qualify this observation is warranted.
 
@@
 We did not collect demographic data on users that appear in our dataset, but based on our own knowledge of the community we can observe that those highly central to the network tend to be employed in (junior) academic positions, or in a few cases in cultural heritage authorities, rather than specifically as research software engineers.
-Such positions tend to not to actively reward or encourage software development, at least not on a par with more traditional academic outputs [@baxter2012], and are increasingly precarious [@Cornelius-BellBell2021].
+Such positions tend to not to actively reward or encourage software development, at least not on a par with more traditional academic outputs [@baxter2012], and are increasingly precarious [@cornelius-bell2021].
 This obviously poses a serious risk to the sustainability and growth of open source software in archaeology: if the people who occupy central positions in the network cease to be active, then it is likely that the overall network would fragment.
 Assessing and mitigating this risk should be a high priority for future research in this area.
-
-<!-- TODO: this should be backed up with a table or figure. The analysis seems to have got lost somewhere. Or just cut? -->
-We also applied the same betweenness centrality algorithm on a subnetwork whose links are based only on issues and comments, and not code contributions.
-In this subnetwork, people who commit less code have higher betweenness scores.
-However, many of the people with high betweenness from the graph representing all contributions also appear here.
-These people who appear in both of these lists have a tendency to contribute as both committers and as commenters.
-This list also includes a series of contributors who never or rarely commit code.
-Although it is out of this study's scope a qualitative analysis of issues and comments may yield more insight on the kinds of contributions that each of these participants make.

### 2024-02-02 analysis/_07-conclusion.qmd
Updates to the text and general touch ups.
Moved stuff on git and github to a new background subsection, moved data caveats from conclusion to methods, updated some interpretations in section 6, update interpretation based on cumulative growth figure.

@@
 
-Archaeology is a "magpie" discipline, in that it assembles tools and methods that originate from a variety of other disciplines and industries to better understand the past [@wylie1989; @nilssonstutz2018].
-Moreover, archaeologists are accustomed to working on shoestring budgets, having to adapt their tools to suit specific, non-generalizable situations, and applying a DIY work ethic to create and co-create with others [@morgan2012;  @caraher2019; @batist2021].
-The freedom to reuse, modify and extend code for a variety of purposes is a boon for the development of improvized tech stacks and workflows called up in contexts of archaeological fieldwork and analysis, and archaeologists share with open source software developers a pragmatic mindset with regards to planning collective work and evaluating outcomes [@scollar1999].
-It is therefore unsurprising that archaeologists are prolific users and developers of open source software.
-
 Our goal in this study was to investigate the under-explored research practices involved in research software engineering. 
@@
 To investigate these concerns, we operationalised open-source collaborative experiences as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
-This excludes a wide range of other potential practices, most obviously the use of other means of version-control and code-hosting (though as we demonstrate in <!-- encode the section --> Section 3, the git/GitHub combination is overwhelmingly the most popular).
-It also excludes collaboration through offline or private channels, or forms of collaboration we don't know about. 
-We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
-Our earliest data is from XXX and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it [@ducke2013 ; @whallon1972].
-
-These caveats notwithstanding, we have documented that open source software development in archaeology has seen a rapid and sustained rise beginning around 2014. 
+With this data, we documented that open source software development in archaeology has seen a rapid and sustained rise beginning around 2014. 
 This is marked by a variety of applications and use cases, including the use of git and GitHub to track and host content other than code. 
@@
 
-<!-- This entire paragraph is subject to change based on revised network analysis -->
 Our network analysis similarly draws attention to the real-world collaborative ties that underpin archaeological open source software development.
-We identified three clusters of close associations between projects:
-the first is centred around a highly specialized supdiscipline exhibiting a very enthusiastic attitude (**ZB NOTE: or maybe even pressure?**) toward working in the open; 
-the second is concerned with developing solutions to a very common and discipline-wide problem, specifically relating to data collection and management in fieldwork; 
-and the third deals with several topic of concern, but represents an emerging community of practice. 
-This third cluster comprises efforts performed by members of a close-knit community, many of whom already collaborate at local research centres or participate in relatively private interest groups. 
-This indicates that open source is firmly embedded within existing social and institutional support and power structures that permeate academic life, both online and offline.
+We identify a core cluster representing a series of collaboratibe ties among members of a archaeological software developer community of practice.
+This core exhibits complexity that correspond with social patterns.
+Indeed, we demonstrate that real-world social connections and institutional support structures are strong predictors of centrality, with sub-clusters appearing within the central core mapping onto established partnerships.
+This indicates that archaeological open source is firmly embedded within existing power structures that permeate academic life, both online and offline.
 
@@
 Moreover, we argue that the objectives and circumstances that frame archaeological practice significantly influence how far archaeologists (and academics in general) are willing to push for these values, and limit the ability for archaeologists to do open source in ways that resemble more mainstream open source projects.
-For instance, successful open source projects like the Linux kernel or the Firefox web browser are driven by collective and popular interest in ensuring that code remains functional, and the code base is therefore constantly in flux and bears an accumulating list of contributing members; this significantly differs from the organizational principles that govern archaeological work, namely that a project's leaders runs the whole show from the top down and operationalize all other actors as instruments of their central directives.
+For instance, successful open source projects like the Linux kernel or the Firefox web browser are driven by collective and popular interest in ensuring that code remains functional, and the code base is therefore constantly in flux and bears an accumulating list of contributing members.
+This significantly differs from the organizational principles that govern archaeological work, namely that a project's leaders runs the whole show from the top down and operationalize all other actors as instruments of their central directives.
 Moreover, archaeological projects ultimately seek to produce stable textual outcomes bearing clear delineation of authorship and that require no upkeep whatsoever. 

### 2024-02-03 analysis/_01-introduction.qmd
General updates to the text, and also replaced the main data file.
Text modifications throughout. Replaced the data file according to the new release (https://github.com/zackbatist/open-archaeo/releases/tag/2024-02-03). Revised the abstract. Verified and corrected the bibliographic references.

@@
 The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plain text, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic).
-However, academic work, including the development of academic software, differs significantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL, and the Firefox web browser.
+However, academic work, including the development of academic software, differs significantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL and the Firefox web browser.
 While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests.
@@
 Despite hopeful aspirations espoused by open archaeology advocates [see @kansa2014; @kintigh2015], these outcomes are not a given.
-We believe<!-- Bit weak; can we cite some previous work that indicates this?
---JR---> that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
+As per @nguyen2022, @pownall2023 and @leonelli2023 we believe that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
 However, practical circumstances and systemic value regimes that frame what it means to work as an archaeologist presently inhibit the potential for radical transformation, even among open science's most ardent supporters.
 
-<!-- Might be getting a bit ahead of ourselves here.
---JR -->That being said, archaeologists are prolific software developers.
-This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes 501 records, as of the time of this article's publication.
+That being said, archaeologists are prolific software developers.
+This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes 495 records, as of the time of this article's publication.
 But beyond simply making their code available on the web, do archaeologists also implement social strategies to advance open source ideals?

### 2024-02-03 analysis/_03-data_methodology.qmd
General updates to the text, and also replaced the main data file.
Text modifications throughout. Replaced the data file according to the new release (https://github.com/zackbatist/open-archaeo/releases/tag/2024-02-03). Revised the abstract. Verified and corrected the bibliographic references.

@@
 ```{r data-oarch}
-# open-archaeo version 2024-02-02
-# https://doi.org/10.5281/zenodo.10611198
-oarch <- read_csv(here("analysis/data/raw_data/open-archaeo_2024-02-02.csv"))
+# open-archaeo version 2024-02-03
+# https://doi.org/10.5281/zenodo.10613694
+oarch <- read_csv(here("analysis/data/raw_data/open-archaeo_2024-02-03.csv"))
 

### 2024-02-03 analysis/_04-open_archaeology.qmd
General updates to the text, and also replaced the main data file.
Text modifications throughout. Replaced the data file according to the new release (https://github.com/zackbatist/open-archaeo/releases/tag/2024-02-03). Revised the abstract. Verified and corrected the bibliographic references.

@@
 
-Our analysis of tags highlights aspects of archaeological work that software developers are inclined to contribute to [@tbl-tags].
+Our analysis of tags highlights aspects of archaeological work that software developers are inclined to contribute to (@tbl-tags).
 The most common tags unsurprisingly deal with work that naturally benefits from advanced information processing afforded by computers, such as statistical analysis, sample calibration, geographical analysis, data management, and chronological modelling.
@@
 
-Roughly half of open-archaeo repositories are accompanied by an explicit license [@tbl-repo-licenses].
+Roughly half of open-archaeo repositories are accompanied by an explicit license (@tbl-repo-licenses).
 Two common free software licenses account for the majority of these: the GNU General Public License (GPL, `r percent(p_license_gpl)`) and the MIT License (`r percent(p_license_mit)`).
-These differ primarily in the restrictions they place on reuse: the MIT License aims to be maximally permissive, while the GPL is a 'copyleft' license that specifies that all derivative works must be distributed under similar terms (in other words, it prohibits the use of FOSS software within non-FOSS software, [@dusollier2007]).
+These differ primarily in the restrictions they place on reuse: the MIT License aims to be maximally permissive, while the GPL is a 'copyleft' license that specifies that all derivative works must be distributed under similar terms (in other words, it prohibits the use of open source software within non-open software, [@dusollier2007]).
 Archaeologists' preference for the more restrictive of these two licenses is the reverse of the general trend seen in open source projects on GitHub [@balter2015].

### 2024-02-03 analysis/_06-network.qmd
General updates to the text, and also replaced the main data file.
Text modifications throughout. Replaced the data file according to the new release (https://github.com/zackbatist/open-archaeo/releases/tag/2024-02-03). Revised the abstract. Verified and corrected the bibliographic references.

@@
 <!-- Predictors of repo centrality -->
-<!-- TODO: Needs an update. These repos are no longer in the top 10, and it's probably best not to be so specific given how volatile that is. -->
-The repositories most central to the network as a whole [@tbl-repo-betweenness] include several lists and directories, including open-archaeo itself.
-Four of the top ten most central repositories (ISAAKiel/mortAAR, ISAAKiel/shapAAR, ISAAKiel/quantAAR, ropensci/c14bazAAR) are associated with the ISAA-Kiel group, emphasising the importance of existing 'real world' collaborations in structuring the network.
-Three (ropensci/c14bazAAR, ahb108/rcarbon, Chronomodel/chronomodel) are R packages relating to radiocarbon data and modelling, probably reflecting the fact that this has long been a prominent focus of statistical software development in archaeology.
-Repository centrality is predicted by the total number of commits it has received but, somewhat surprisingly, it is younger repositories rather than older ones that tend to be more central.
+The repositories most central to the network as a whole (@tbl-repo-betweenness) include three lists and directories, including open-archaeo itself.
+Three relate to making large data repositories accessible for analysis, and one is a very well-supported field recording application.
+Community input is therefore centred on infrastructural projects, including those which index and publicize available tools and resources.
+Moreover, three relate to radiocarbon data modelling and two relate to palaeoenvironment reconstruction, which reflects the fact that these have long been prominent foci of statistical software development in archaeology.
+
+Repository centrality is predicted by the total number of commits it has received but, somewhat surprisingly, younger repositories rather than older ones tend to be more central (@fig-repo-betweenness-age-length).
 Tentatively, we interpret this as an indication that the network has become more connected over time, but we leave a fuller analysis of temporal trends in collaborative activity to future work.
@@
 
-The highest mean centrality is seen in repositories that contain lists/datasets, standalone software, or packages/libraries, and in repositories based on Python, R or QGIS [@fig-repo-betweenness-category-platform].
-Interestingly, these trends depart from the observed popularity of different categories/platform in the open-archaeo dataset as a whole (see @sec-open-archaeology): standalone software is more central than packages/libraries, even though there are more of the latter by a significant margin.
+When comparing across categories and platforms, the highest mean centrality is seen in repositories that contain lists and datasets, standalone software, or packages and libraries, and in repositories based on Python, R or QGIS (@fig-repo-betweenness-category-platform).
+Interestingly, these trends depart from the observed popularity of different categories and platforms in the open-archaeo dataset as a whole (see @sec-open-archaeology): standalone software is more central than packages/libraries, even though there are more of the latter by a significant margin.
 This may be due to the fact that many packages are developed to support specific practices or use-cases (often inspired by personal need), or are designed to run relatively stable statistical functions that need not change over time.
@@
 <!-- Predictors of user centrality -->
-Centrality to the user–user graph is weakly predicted by a user's overall rate of activity, as measured by their total number of contributions [@fig-user-betweenness-contribs]. 
+Centrality to the user–user graph is weakly predicted by a user's overall rate of activity, as measured by their total number of contributions (@fig-user-betweenness-contribs). 
 We did not collect demographic data on users that appear in our dataset, but based on our own knowledge of the community we can observe that those highly central to the network tend to be employed in (junior) academic positions, or in a few cases in cultural heritage authorities, rather than specifically as research software engineers.

### 2024-02-03 analysis/_07-conclusion.qmd
General updates to the text, and also replaced the main data file.
Text modifications throughout. Replaced the data file according to the new release (https://github.com/zackbatist/open-archaeo/releases/tag/2024-02-03). Revised the abstract. Verified and corrected the bibliographic references.

@@
 Our network analysis similarly draws attention to the real-world collaborative ties that underpin archaeological open source software development.
-We identify a core cluster representing a series of collaboratibe ties among members of a archaeological software developer community of practice.
-This core exhibits complexity that correspond with social patterns.
-Indeed, we demonstrate that real-world social connections and institutional support structures are strong predictors of centrality, with sub-clusters appearing within the central core mapping onto established partnerships.
+We identify a core cluster representing a series of collaborative ties among members of an archaeological software engineering community of practice.
+This core exhibits complexity that corresponds with social patterns, such as the presence of various sub-clusters representing interconnected interest or affinity groups.
+Indeed, we demonstrate that real-world social connections and institutional support structures are strong predictors of centrality, since these sub-clusters are representative of established professional partnerships.
 This indicates that archaeological open source is firmly embedded within existing power structures that permeate academic life, both online and offline.
@@
 Moreover, we argue that the objectives and circumstances that frame archaeological practice significantly influence how far archaeologists (and academics in general) are willing to push for these values, and limit the ability for archaeologists to do open source in ways that resemble more mainstream open source projects.
-For instance, successful open source projects like the Linux kernel or the Firefox web browser are driven by collective and popular interest in ensuring that code remains functional, and the code base is therefore constantly in flux and bears an accumulating list of contributing members.
+For instance, successful open source projects like the Linux kernel, openSSL the Firefox web browser are driven by collective and popular interest in ensuring that code remains functional, and the code base is therefore constantly in flux and bears an accumulating list of contributing members.
 This significantly differs from the organizational principles that govern archaeological work, namely that a project's leaders runs the whole show from the top down and operationalize all other actors as instruments of their central directives.

### 2024-02-03 analysis/openarchaeo_collaboration.qmd
General updates to the text, and also replaced the main data file.
Text modifications throughout. Replaced the data file according to the new release (https://github.com/zackbatist/open-archaeo/releases/tag/2024-02-03). Revised the abstract. Verified and corrected the bibliographic references.

@@
 author:
-  - name: Zack Batist
+  - name: Zachary Batist
     url: https://zackbatist.info
-    affiliation: Faculty of Information, University of Toronto
-    affiliation-url: https://ischool.utoronto.ca/
+    affiliation: Independent Scholar
     orcid: 0000-0003-0435-508X 
@@
   Surveying the first quarter-century of computer applications in archaeology, Scollar [-@scollar1999] lamented that the field relied almost exclusively on "hand-me-down" tools repurposed from other disciplines.
-25 years later, this is no longer the case: computational archaeologists often find themselves practicing the dual roles of data analyst and research software engineer [@baxter2012], developing and applying new tools that are tailored specifically to archaeological problems and archaeological methods.
-Though this trend can be traced to the very earliest days of the field [@cowgill1967], its most recent manifestation is distinguished by its apparently embrace of practices from free and open source software (FOSS).
-Most prominently, since around 2015, there has been a rapid uptake of workflow tools designed for open source development communities, such as the version control system `git` and associated online source code management platforms (e.g.
+25 years later, this is no longer the case: computational archaeologists often find themselves practicing the dual roles of data analyst and research software engineer [@baxter2012; @schmidt2020], developing and applying new tools that are tailored specifically to archaeological problems and archaeological methods.
+Though this trend can be traced to the very earliest days of the field [@cowgill1967], its most recent manifestation is distinguished by its apparent embrace of practices from free and open source software .
+Most prominently, since around 2015, there has been a rapid uptake of workflow tools designed for open source development communities, such as the version control system git and associated online source code management platforms (e.g.
 GitHub, GitLab).
-These tools facilitate collaboration among developers and users of FOSS software using patterns that can diverge quite radically from conventional scholarly norms [@tennant2020].
+These tools facilitate collaboration among developers and users of open source software using patterns that can diverge quite radically from conventional scholarly norms [@tennant2020].
 
-  In this paper, we investigate modes of collaboration in this emerging community of practice using open-archaeo [open-archaeo.info](open-archaeo.info), a curated list of 500+ active open projects in archaeology, augmented where available with data on their activity on GitHub.
+  In this paper, we investigate modes of collaboration in this emerging community of practice using open-archaeo [open-archaeo.info](open-archaeo.info), a curated list of archaeological software, augmented where available with data on their activity on GitHub.
   We apply network analysis to map the networks of collaboration that have emerged among archaeologists using GitHub, and conduct a quantitative exploratory analysis to characterize the nature and intensity of these collaborations.
-Our results show an uneven adoption of open source collaborative practices beyond the basic use of `git` as a version control system and GitHub to host source code.
-A majority of repositories have 1–3 contributors, with only a few projects distinguished by an active and diverse developer base.
-More community-oriented features, such as GitHub's 'issues' and 'comments', are used in only a minority of repositories.
-Preliminary findings from ongoing work also shows how collaborative behaviours vary across different subcommunities.
-By analyzing interactions among archaeologists using the GitHub platform, including the ways they contribute to collective code bases, we demonstrate how archaeological software engineering is beginning to foster new kinds of collaborative commitments that blend open source protocols, DIY work ethic, conventional forms of scholarly communication, and cultural practices unique to archaeology.
+  Our results show uneven adoption of open source collaborative practices beyond the basic use of git as a version control system and GitHub to host source code.
+  For instance, a majority of repositories have 1–3 contributors, with only a few projects distinguished by an active and diverse developer base.
+  Furthermore, community-oriented features, such as GitHub's 'issues' and 'comments', are used in only a minority of repositories.
+  Our network analysis similarly demonstrates that archaeological open source is firmly embedded within existing power structures that permeate academic life, and that apparent connections are representative of established professional partnerships.
+  As such, we find little evidence to support the notion that archaeologists benefit from the potential positive outcomes that are commonly argued to be the natural results of open source development models – namely, greater degrees of transparency, extensibility and participatory action.
 bibliography: references.bib

### 2024-02-04 analysis/_03-data_methodology.qmd
Revised with 2024-02-04 version.
Updated the section on average centrality by platform, and another very minor reference correction.

@@
 # https://doi.org/10.5281/zenodo.10613694
-oarch <- read_csv(here("analysis/data/raw_data/open-archaeo_2024-02-03.csv"))
+oarch <- read_csv(here("analysis/data/raw_data/open-archaeo_2024-02-04.csv"))
 

### 2024-02-04 analysis/_06-network.qmd
Revised with 2024-02-04 version.
Updated the section on average centrality by platform, and another very minor reference correction.

@@
 
-Despite being the minority language, Python packages tend to be more central than R packages, on average.
-Python projects in archaeology tend to be large, highly collaborative, and either infrastructural or of interest to other fields, so they have disproportionately high centrality.
-Our findings support the notion that there are significant development patterns when working with these two languages, and further analysis to qualify this observation is warranted.
+Despite being a minority language, Blender packages are more central than all other package platforms on average, but this is a statistical anomaly caused by uneven sampling.
+R is naturally the platform with the next highest average centrality since it serves as a _lingua franca_ that draws developers from across the discipline.
+Many of the QGIS plugins add various specialized features to the extensible GIS platform, and are therefore developed by interdisciplinary teams, which explains its high rank.
+Python projects, which tend to be infrastructural or are of interest to members of other fields, are also highly ranked in terms of average centrality.
+It should be noted that while Blender and QGIS plugins are written using the Python language, our intent while categorizing platforms was to get a sense of the developer ecosystems in which archaeological software engineers participate, rather than to simply gauge the popularity of different languages [@batist2023 2].
+Our findings support the notion that there are significant development patterns when working across different languages and platforms, and further analysis to qualify these observation is warranted.
 

### 2024-02-06 analysis/_01-introduction.qmd
Minor edits to readability and text flow

@@
 # Introduction {#sec-introduction}
-
-<!--
-* There is open archaeology
-* Open archaeology is related to open source
-* Is open archaeology open source, and is open source open?
-  * [Is 'open archaeology' just uploading CSVs?]
--->
 
@@
 
-However, academic open source has a complicated relationship with open source as practiced by professional software developers, which has its own distinct history and is framed by different objectives, challenges, and value regimes.
-Despite this, the open science movement, within which open archaeology emerged, draws direct inspiration from open source.
-For instance, the Open Knowledge Foundation [-@openknowledgefoundation2015] publishes a widely accepted definition of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group.
-The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plain text, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic).
-However, academic work, including the development of academic software, differs significantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL and the Firefox web browser.
-While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests.
-In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers [@ratto2007; @kelty2008 chapter 9].
-
-The relationship between open science and open source is also complicated by rhetorical claims that have a questionable connection to how academics actually _do_ open source. 
+The relationship between open science and open source is complicated by rhetorical claims that have a questionable connection to how academics actually _do_ open source. 
 Does academic open source actually make research processes more transparent and improve research outcomes? 
@@
 Is it actually helping to foster new globe-spanning connections and lead to novel research trajectories that would not otherwise come to pass?
-Basically, is it more than just uploading code and data files to the internet?
+Basically, is there more to it than just uploading code and data files to the internet?
 
@@
 
-That being said, archaeologists are prolific software developers.
-This is made evident by open-archaeo, a comprehensive list of open source archaeological software and resources that indexes 495 records, as of the time of this article's publication.
+There is no question that archaeologists are prolific software developers [@batist2023].
 But beyond simply making their code available on the web, do archaeologists also implement social strategies to advance open source ideals?
@@
 
-Our statistical and network analyses consider how archaeological software development may be benefiting from or missing out on the affordances that open source development models provide, specifically the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
-We found that the vast majority of projects lack any collective effort and activity tends to abruptly end shortly after work is initiated.
-Projects that are maintained by multiple contributors tend to be backed by funded initiatives who employ locally and socially engaged individuals, or are warranted by participation in a scholarly community exhibiting genuine need for a particular kind of resource.
-Moreover, we found that the individuals who play critical roles in supporting the archaeological open source community are precariously employed workers.
-Contrary to popular claims about open source being inherently distributed, resilient, and open-ended, we found that, overall, archaeological open source is actually quite centralized, fragile, and based nearly exclusively on existing professional connections and endeavours.
+We use quantitative analysis to consider how archaeological software development may be benefiting from, or missing out on, the affordances that open source development models provide, specifically the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.

### 2024-02-06 analysis/_02-background.qmd
Minor edits to readability and text flow

@@
 
-<!--
-* What is open source?
-* Summary of key publications in open archaeology / DIY archaeology
-* But is any of this real? There seems to be a lack of reflection about what open archaeology is.
-  -->
+Academic open source has a complicated relationship with open source as practiced by professional software developers, which has its own distinct history and is framed by different objectives, challenges, and value regimes.
+Despite this, the open science movement, within which open archaeology emerged, draws direct inspiration from open source.
+For instance, the Open Knowledge Foundation [-@openknowledgefoundation2015] publishes a widely accepted definition of "open" in the context of scholarly communication that explicitly refers to the definition of "open source" published by the Open Source Initiative [-@opensourceinitiative2007], an authoritative open source advocacy group.
+The open science movement further mimics open source by operationalizing scholarly communication through technical infrastructures and protocols that closely resemble systems and processes designed to develop open source software (e.g., the use of plain text, line-resolution version control, emphasis on formal licencing, the general hacker aesthetic).
+However, academic work, including the development of academic software, differs significantly from the work involved in massive open source projects that literally run the internet, such as the Linux kernel, openSSL and the Firefox web browser.
+While they may use similar tools and technical protocols to manage coding operations, the open science and open source movements are governed by different social and professional warrants and interests.
+In other words, publishing code openly on the web has different meanings, impacts and implications for archaeologists and professional software developers [@ratto2007; @kelty2008 chapter 9].
 
@@
 Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.
-Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience.
+<!-- TODO: Some strong claims here.
+Can we add citations?
+-->Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience.
 Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of tenure and promotion.
@@
 
-## Open Source Infrastructure
+## Git and GitHub
 
@@
 Here we provide a brief overview of key technologies that archaeologists have come to rely on as they develop open source software.
-
-Chief among these is git, a protocol designed to facilitate open and distributed contributions to a common code base.
-It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
-Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a fork in which they apply their modifications, and then request that their fork be merged into the central code base.
-After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent forks intact.
-
-Git is also designed to facilitate code review and version control.
-All modifications are tracked as "diffs", which highlight additions or deletions to the code base, including changes within individual files.
-Typically, a contributor will group a series of changes into a more comprehensive "commit" based on a specific task or part of a workflow.
-Commits are always accompanied by a message, in which the contributor (ideally) describes the reason and context for the changes included in the commit.
-Moreover, git assigns each commit a unique identifier and identifies the contributor by name and email address to ensure some degree of public accountability.
-
-Forges, or collaborative software development platforms like GitHub, GitLab and Codeberg are designed to facilitate open source software development by hosting public git repositories.
-However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
-These platforms also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as favourites, and maintain a public-facing profile that includes personally-identifying information (e.g.
-profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
-Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
-While these additional features are meant to complement and enhance the experience of contributing to open source projects, they are not actually part of the git protocol.
-See @tbl-glossary for a glossary of terminology used to describe many of these concepts and features, and which we use throughout this paper.
+See @tbl-glossary a glossary of the git-, GitHub- and software engineering-related terminology which we use here and throughout this paper.
 
@@
 ```
+
+Chief among these is git, a protocol designed to facilitate open and distributed contributions to a common code base.
+It operates by providing mechanisms for synchronizing communal, web-based public repositories with local iterations stored on contributors' private workstations.
+Contributors who volunteer or are assigned to develop, inspect, or revise a specific aspect of a code base download a copy of the public repository into their own work environment, create a fork in which they apply their modifications, and then request that their fork be merged into the central code base.
+After a public repository's maintainers decide to merge the proposed changes into the communal code base, other developers may use git to download these changes while keeping their own independent forks intact.
+
+Git is also designed to facilitate code review and version control.
+All modifications are tracked as "diffs", which highlight additions or deletions to the code base, including changes within individual files.
+Typically, a contributor will group a series of changes into a more comprehensive "commit" based on a specific task or part of a workflow.
+Commits are always accompanied by a message, in which the contributor (ideally) describes the reason and context for the changes included in the commit.
+Moreover, git assigns each commit a unique identifier and identifies the contributor by name and email address to ensure some degree of public accountability.
+
+Software forges—collaborative web platforms like GitHub, GitLab and Codeberg—are designed to facilitate open source software development by hosting public git repositories.
+However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
+These platforms also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as a combine bookmarking and 'like' feature, and maintain a public-facing profile that includes personally-identifying information (e.g.
+profile picture, username, real name, employer or affiliation), references to all public activity on the platform, and links to the user's other social media profiles.
+Code-sharing platforms thus serve as comprehensive developer portfolios and community networking resources.
+While these additional features are meant to complement and enhance the experience of contributing to open source projects, they are not actually part of the git protocol.

### 2024-02-06 analysis/_03-data_methodology.qmd
Minor edits to readability and text flow

@@
 ```{r data-oarch}
-# open-archaeo version 2024-02-03
-# https://doi.org/10.5281/zenodo.10613694
-oarch <- read_csv(here("analysis/data/raw_data/open-archaeo_2024-02-04.csv"))
+# open-archaeo version 2024-02-06
+# https://doi.org/10.5281/zenodo.10625236
+oarch <- read_csv(here("analysis/data/raw_data/open-archaeo_2024-02-06.csv"))
 

### 2024-02-06 analysis/_04-open_archaeology.qmd
Revised with 2024-02-04 version.
Updated the section on average centrality by platform, and another very minor reference correction.
Minor edits to readability and text flow

@@
 
-<!-- Results I: description of open archaeology projects
-
-- What kind of projects comprise open archaeology?
-  x Breakdown of tags and categories
-  x Breakdown of platforms and languages
-  x Tags by category
-  x Licenses?
-- How has this changed over time?
-  - Growth of open archaeology by number of commits
-  - Growth of open archaeology by number of repos
-  - Compare changing proportions of each category on a year by year basis
-  - Identify periods associated with different rates of growth, and relate these with different general attitudes in the history of digital archaeology
--->
-
 As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists.
@@
 We annotated each record with categories and tags that describe what aspect of archaeological research each tool or resource was meant to address.
-We categorized all records based on how each tool or resource is meant to be accessed or used, and assigned tags based on how developers identified their projects' purpose and scope.
-See [@batist2023] for a more comprehensive overview of the tags and categories applied to open-archaeo.
+We also categorized all records based on how each tool or resource is meant to be accessed or used, and assigned tags based on how developers identified their projects' purpose and scope.
+See @batist2023 for a more comprehensive overview of the tags and categories applied to open-archaeo.
 
@@
 Most resources (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language or a plugin for an application.
-The designers of this code are basically creating additional functions within the base platform that are useful for archaeological purposes.
+Essentially such projects create additional functions within the base platform that are useful for archaeological purposes.
 Others create standalone software (`r percent(p_standalone)`) that can be run independently of such platforms, for example desktop or web apps.
@@
 
-Our analysis of tags highlights aspects of archaeological work that software developers are inclined to contribute to (@tbl-tags).
-The most common tags unsurprisingly deal with work that naturally benefits from advanced information processing afforded by computers, such as statistical analysis, sample calibration, geographical analysis, data management, and chronological modelling.
+Our analysis of thematic tags highlights aspects of archaeological work that software developers are inclined to contribute to (@tbl-tags).
+The most common themes are work that naturally benefits from advanced information processing afforded by computers, such as statistical analysis, sample calibration, geographical analysis, data management, and chronological modelling.
 Educational resources and practical guides are also well represented due to the web's usefulness as a medium for sharing and communication.
 
-When we compare categories with tags, we see the general domains that each kind of resource is designed to serve.
+When we compare categories with thematic tags, we see the general domains that each kind of resource is designed to serve.
 We see that packages are fairly common across the board.
@@
 Two common free software licenses account for the majority of these: the GNU General Public License (GPL, `r percent(p_license_gpl)`) and the MIT License (`r percent(p_license_mit)`).
-These differ primarily in the restrictions they place on reuse: the MIT License aims to be maximally permissive, while the GPL is a 'copyleft' license that specifies that all derivative works must be distributed under similar terms (in other words, it prohibits the use of open source software within non-open software, [@dusollier2007]).
-Archaeologists' preference for the more restrictive of these two licenses is the reverse of the general trend seen in open source projects on GitHub [@balter2015].
+These differ primarily in the restrictions they place on reuse: the MIT License aims to be maximally permissive, while the GPL is a 'copyleft' license that specifies that all derivative works must be distributed under similar terms (in other words, it prohibits the use of open source software within non-open software, [@dusollier2007].
+Interestingly, archaeologists' preference for the more restrictive of these two licenses is the reverse of the general trend seen in open source projects on GitHub [@balter2015].
 Creative Commons licenses are a distant third place (`r percent(p_license_cc)` of repositories), in contrast to their widespread use for other forms of scholarly output [@kim2007].

### 2024-02-06 analysis/_05-collaboration.qmd
General updates to the text, and also replaced the main data file.
Text modifications throughout. Replaced the data file according to the new release (https://github.com/zackbatist/open-archaeo/releases/tag/2024-02-03). Revised the abstract. Verified and corrected the bibliographic references.
Minor edits to readability and text flow

--- 2024-02-02 analysis/_05-collaboration.qmd
@@
 
-The interaction between project longevity, activity, and number of contributors is multifaceted [@fig-lifespan-rate].
+The interaction between project longevity, activity, and number of contributors is multifaceted (@fig-lifespan-rate).
 Highly active projects (one commit per day or more) tend to be either very long-lived or very short-lived; few fall in the centre of the distribution.
@@
 
-The prototypical and idyllic collaborative open source project ingrained into the common consciousness hypothetically comprises a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through forks and pull requests, plus an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
+The prototypical open source project comprises a core group of developers (often a single maintainer) that regularly commit new code, a wider network of collaborators that contribute through forks and pull requests, plus an active user base that create and comment on issues, who have indicated their support for the project by starring its repository.
 It is unclear whether archaeological software developers actually aim to operate following this model, or whether it is even suitable for supporting what open science aims to achieve.

### 2024-02-06 analysis/_06-network.qmd
Minor edits to readability and text flow

@@
 <!-- Clustering within the network -->
-To explore the topography of this network, we applied the edge–betweenness community detection method [@girvan2002] to delimit `r nrow(filter(clusters_n_repos, n > 1))` distinct clusters in the repository–repository network.
+We delimited `r nrow(filter(clusters_n_repos, n > 1))` distinct clusters that outline the topography of the repository–repository network.
 While many of these clusters are interconnected, some discrete components containing between 2-20 repositories appear as distinct from a primary core.
@@
 
-Likewise, we applied the same edge-betweenness community detection method to identify clusters in the user-user network, which reveal collaborative groupings.
+Clustering also reveals distinct collaborative networks within the user-user graph.
 We again see a complementary primary core connected to several more peripheral clusters, which are internally-cohesive and exhibit few connections with other peripheral clusters.
@@
 On the other hand, the central cores exhibit greater internal variety that may correspond with social connections and the formation of a complex software development community.
-This is evident through the fact that many of the connections represented in the cores emerge from underlying professional partnerships, namely research labs (e.g.
-ISAA-Kiel) and special interest groups (CAA-SSLA).
+This is evident through the fact that many of the connections represented in the cores emerge from more conventional professional networks, e.g.
+[ISAAKiel](https://isaakiel.github.io/), a working group centered around the University of Kiel), or [CAA-SSLA](https://sslarch.github.io/), a special interest group of the international scholarly society 'Computational and Quantitative Applications in Archaeology' (CAA) focused on scientific programming.
 Peripheral clusters that are connected to the central core by only a few relationships represent the sole (or perhaps initial) integration of lone developers into a broader community.
@@
 Python projects, which tend to be infrastructural or are of interest to members of other fields, are also highly ranked in terms of average centrality.
-It should be noted that while Blender and QGIS plugins are written using the Python language, our intent while categorizing platforms was to get a sense of the developer ecosystems in which archaeological software engineers participate, rather than to simply gauge the popularity of different languages [@batist2023 2].
+^[While Blender and QGIS plugins are written using the Python language, our intent while categorizing platforms was to get a sense of the developer ecosystems in which archaeological software engineers participate, rather than to simply gauge the popularity of different languages [@batist2023 2].]
 Our findings support the notion that there are significant development patterns when working across different languages and platforms, and further analysis to qualify these observation is warranted.

### 2024-02-06 analysis/_07-conclusion.qmd
Minor edits to readability and text flow

@@
 
-Looking at the distribution of tags, we note that, among items listed as software, repositories tend to be focused on 
+Turning to the thematic distribution of open archaeology projects, we note that, among items listed as software, repositories tend to be focused on 
 various means of identifying distribution patterns (spatial, temporal, statistical), 
@@
 
+Overall, we found that the vast majority of projects lack any collective effort and activity tends to abruptly end shortly after work is initiated.
+Projects that are maintained by multiple contributors tend to be backed by funded initiatives who employ locally and socially engaged individuals, or are warranted by participation in a scholarly community exhibiting genuine need for a particular kind of resource.
+Moreover, we found that the individuals who play critical roles in supporting the archaeological open source community are precariously employed workers.
+Contrary to popular claims about open source being inherently distributed, resilient, and open-ended, we found that, overall, archaeological open source is actually quite centralized, fragile, and based nearly exclusively on existing professional connections and endeavours.
+
 There is thus little evidence to support the notion that archaeologists benefit from the positive outcomes that are commonly argued to be the natural results of open source development models -- namely, greater degrees of transparency, extensibility and participatory action.

### 2024-02-06 analysis/_03-data_methodology.qmd
Update to open-archaeo 2024-02-6

@@
 We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
-Our earliest data is from 2005 and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it [@ducke2013 ; @whallon1972].
+Our earliest data is from 2005 and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it [@ducke2013; @whallon1972].
 

### 2024-02-06 analysis/_04-open_archaeology.qmd
Adjust range of fig-github-cumulative

@@
            vjust = 0) +
-  scale_x_datetime(limits = c(ymd("2006-12-31", tz = "UTC"), dt_last_commit),
+  scale_x_datetime(limirs = c(ymd("2004-12-31", tz = "UTC"), dt_last_commit),
                    date_breaks = "2 years", 
@@
            vjust = 0) +
-  scale_x_datetime(limits = c(ymd("2006-12-31", tz = "UTC"), dt_last_commit),
+  scale_x_datetime(limits = c(ymd("2004-12-31", tz = "UTC"), dt_last_commit),
                    date_breaks = "2 years", 
@@
 
-As we can see <!-- this is no longer evident in the figure -->, archaeologists have been using git from even before GitHub was launched in 2008.
+As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
 But use of git really began to take off around 2014--2015, when we see an uptick in the rate of growth.

### 2024-02-06 analysis/_03-data_methodology.qmd
Remove filter
None of the items it filters are in open-archaeo any more

@@
 ```{r data-oarch-filter}
-oarch_filter <- read_csv(here("analysis/data/raw_data/open-archaeo_filter.csv"))
-n_items_removed <- nrow(filter(oarch_filter, !item_name %in% oarch$item_name))
-oarch |>
-  filter(!item_name %in% oarch_filter$item_name) ->
-  oarch
-
 # Exclude from GitHub API data collection because they part of a larger

### 2024-02-06 analysis/_02-background.qmd
Add references

@@
 Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.
-<!-- TODO: Some strong claims here.
-Can we add citations?
--->Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs, and rarely anticipate use of their work beyond the limited scope of their primary targeted audience.
-Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of tenure and promotion.
+Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs.
+Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of resume-building, tenure and promotion [@mirowski2018; @dorta-gonzalez2021].
 Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out.
-Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours.
-Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software.
+Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours [@kelty2008: 271-275].
+Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software [@kelty2008: 278-281; @coleman2012: 116-122; @hippel2003].
 

### 2024-02-06 analysis/_04-open_archaeology.qmd
Reknit

@@
            vjust = 0) +
-  scale_x_datetime(limirs = c(ymd("2004-12-31", tz = "UTC"), dt_last_commit),
+  scale_x_datetime(limits = c(ymd("2004-12-31", tz = "UTC"), dt_last_commit),
                    date_breaks = "2 years", 

### 2024-02-07 analysis/_06-network.qmd
Revise graph aesthetics

@@
   geom_edge_link(aes(group = contrib, alpha = n_contrib)) +
-  geom_point_interactive(aes(x, y, colour = cluster, shape = node_type, 
-                             tooltip = node)) + 
-  scale_edge_alpha(trans = "log", guide = guide_none()) +
+  geom_point_interactive(
+    aes(x, y, colour = cluster, shape = node_type, tooltip = node),
+    alpha = 0.8
+  ) + 
+  scale_edge_alpha(trans = "log", guide = guide_none(), range = c(0.05, 0.25)) +
   scale_colour_discrete(guide = guide_none()) +
@@
 Node colour indicates membership of the largest clusters according to the edge–betweenness method [@girvan2002].
-Excludes repositories with only a single user.
+Excludes isolate nodes.
 g_repo_repo <- oarch_graph_repo_repo |>
@@
   geom_edge_link(aes(alpha = n)) +
-  geom_point_interactive(aes(x, y, colour = cluster, tooltip = repo)) + 
-  scale_edge_alpha(trans = "log", guide = guide_none()) +
+  geom_point_interactive(
+    aes(x, y, colour = cluster, tooltip = repo),
+    alpha = 0.8
+  ) + 
+  scale_edge_alpha(guide = guide_none(), range = c(0.05, 0.25)) +
   scale_colour_discrete(guide = guide_none()) +
@@
 Node colour indicates membership of the largest clusters according to the edge–betweenness method [@girvan2002].
-Excludes users that have only contributed to a single repository.
+Excludes isolate nodes.
 g_user_user <- oarch_graph_user_user |>
@@
   geom_edge_link(aes(alpha = n)) +
-  geom_point_interactive(aes(x, y, colour = cluster, tooltip = user)) + 
-  scale_edge_alpha(trans = "log", guide = guide_none()) +
+  geom_point_interactive(
+    aes(x, y, colour = cluster, tooltip = user),
+    alpha = 0.8
+  ) + 
+  scale_edge_alpha(guide = guide_none(), range = c(0.05, 0.25)) +
   scale_colour_discrete(guide = guide_none()) +
@@
 
-<!-- Is there a network? How big is it? -->
 Our data shows that there is a significant network of archaeologists collaborating on GitHub.
@@
 
-<!-- Clustering within the network -->
 We delimited `r nrow(filter(clusters_n_repos, n > 1))` distinct clusters that outline the topography of the repository–repository network.
@@
   scale_x_log10() +
-  scale_y_log10() +
+  scale_y_log10(labels = NULL) +
   geom_smooth(method = "glm") +
-  labs(x = "Repository age (days)", y = "Centrality") +
+  labs(x = "Days", y = "Log centrality →") +
   theme_oarch()
@@
   scale_x_log10() +
-  scale_y_log10() +
+  scale_y_log10(labels = NULL) +
   geom_smooth(method = "glm") +
-  labs(x = "Number of commits", y = NULL) +
+  labs(x = "Commits", y = NULL) +
   theme_oarch()
@@
 
-<!-- Predictors of repo centrality -->
 The repositories most central to the network as a whole (@tbl-repo-betweenness) include three lists and directories, including open-archaeo itself.
@@
   geom_col() +
-  scale_y_log10() +
+  scale_y_log10(labels = NULL) +
   coord_flip() +
@@
   coord_flip() +
-  labs(y = "Centrality →", x = NULL) +
+  labs(y = "Log centrality →", x = NULL) +
   theme_oarch()
@@
 
-<!-- Predictors of user centrality -->
 Centrality to the user–user graph is weakly predicted by a user's overall rate of activity, as measured by their total number of contributions (@fig-user-betweenness-contribs). 

### 2024-02-07 analysis/_03-data_methodology.qmd
Fix knit problems

@@
 
-```{r data-oarch_forges}
+```{r data-oarch-forges}
 oarch_forges <- oarch |>

### 2024-02-07 analysis/_04-open_archaeology.qmd
Fix knit problems

@@
 
-```{r num-p_platform_r}
+```{r data-platforms}
 p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / nrow(oarch)

### 2024-02-07 analysis/_05-collaboration.qmd
Fix knit problems

@@
 
-```{r data-oarch_github}
+```{r data-oarch-github}
 # Subset of dataset with GitHub data
@@
 
-```{r data-oarch_lifespan}
+```{r data-oarch-lifespan}
 oarch_gh %>% 

### 2024-02-07 analysis/_06-network.qmd
Fix knit problems

@@
 
-```{r data-oarch_contributors}
+```{r data-oarch-contributors}
 oarch |>
@@
 
-girafe(ggobj = g_repo_user)
+if(knitr::is_html_output()) girafe(ggobj = g_repo_user) else g_repo_user
 ```
@@
 
-girafe(ggobj = g_repo_repo)
+if(knitr::is_html_output()) girafe(ggobj = g_repo_repo) else g_repo_user
 ```
@@
 
-girafe(ggobj = g_user_user)
+if(knitr::is_html_output()) girafe(ggobj = g_user_user) else g_user_user
 ```

### 2024-02-07 analysis/openarchaeo_collaboration.qmd
Cache chunk output
Fix knit problems

@@
     embed-resources: true
+  pdf:
+    documentclass: article
+    pagesize: a4
+    toc: true
+    number-sections: true
+    extra_dependencies: [ tabularray ]
+    includes:
+      in_header: fix-tables.tex # workaround for https://tex.stackexchange.com/questions/611786/misplaced-noalign-because-input-before-booktabs-rule
+execute:
+  cache: true
+  echo: false
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = FALSE,
-	message = FALSE,
-	warning = FALSE
-)
+```{r setup}
+options("tinytable_tabularray_placement" = "H")
+library("openarchaeoCollaboration") # this package, install with devtools::build()
 ```
 
-```{r dependencies, include=FALSE}
-library("openarchaeoCollaboration") # this package, install with devtools::build()
-
+```{r dependencies}
 library("english")

### 2024-02-07 analysis/_02-background.qmd
update references in background

@@
 Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.
-Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs.
+Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs [@kling2003].
 Additionally, open science is bounded by the professional contexts in which science operates, and as such, produces outputs that can be easily credited to specific sets of individuals for reasons of resume-building, tenure and promotion [@mirowski2018; @dorta-gonzalez2021].
-Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out.
-Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours [@kelty2008: 271-275].
+Open science projects whose contributions are supported by research funding also face sustainability concerns, as participants lose motivation to contribute once funding runs out [@carver2022; @adema2021].
+Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours [@kelty2008: 271-275; @howison2013].
 Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software [@kelty2008: 278-281; @coleman2012: 116-122; @hippel2003].

### 2024-02-07 analysis/openarchaeo_collaboration.qmd
update references in background

@@
     embed-resources: true
-  pdf:
-    documentclass: article
-    pagesize: a4
-    toc: true
-    number-sections: true
-    extra_dependencies: [ tabularray ]
-    includes:
-      in_header: fix-tables.tex # workaround for https://tex.stackexchange.com/questions/611786/misplaced-noalign-because-input-before-booktabs-rule
 execute:
-  cache: true
   echo: false

### 2024-02-07 analysis/_03-data_methodology.qmd
Revise conclusions text

@@
 
-We present an exploratory analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources.
+We present an exploratory quantitative analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources maintained primarily by one of us (ZB) since 2018.
 
@@
 However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
-The dataset is also limited by the experiences of its primary maintainers.
-We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
+The dataset is also limited by the experiences of its primary maintainers.^[We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
+Instructions for contributing to open-archaeo can be found at <https://github.com/zackbatist/open-archaeo>.]
 
@@
 Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API (application programming interface).
-In total, we analysed data on `r n_repos` repositories, comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
+Our analysis incorporates data on `r n_repos` repositories, comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
 
 We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
-This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data included in open-archaeo itself.
-We also exclude collaboration through offline or private channels, or forms of collaboration we do not know about. 
+This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data compiled in open-archaeo itself.
+We also cannot include collaboration through offline or private channels, or forms of collaboration we do not know about. 
 We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
@@
 These caveats notwithstanding, the open-archaeo directory and the supplemental data from the GitHub API provide a rich resource to explore the nature of collaborative software engineering in archaeology.
-Here we employ exploratory data analysis [@tukey1977] to identify and characterise overall trends visible in this rich dataset.
+Here we employ exploratory data analysis [*sensu* @tukey1977] to identify and describe overall patterns visible in this rich dataset.
 In @sec-open-archaeology, our focus is on examining the general state of open source archaeological software and resource development.

### 2024-02-07 analysis/_06-network.qmd
Revise conclusions text

@@
 The central core bridges all the peripheral clusters.
-The central core is not uniform, and comprises several relatively discrete sub-clusters representing collaborative sub-communities.
-While these sub-clusters are internally cohesive, they exhibit enough connections to other members of the central core so as to not be considered as separate or peripheral clusters.
+The central core is not uniform, and comprises several relatively discrete clusters representing collaborative sub-communities.
+While these clusters are internally cohesive, they exhibit enough connections to other members of the central core so as to not be considered as separate or peripheral clusters.
 

### 2024-02-07 analysis/_07-conclusion.qmd
Revise conclusions text

--- 2024-02-03 analysis/_07-conclusion.qmd
@@
 
-<!--
-
-There is an emerging community of practice around open source in archaeology.
-
-* It is limited and fragmented
-  * A small set of subfields are represented
-  * Probably most people still haven't heard it?
-* Collaboration is limited
-  * Most work remains solo & short-lived
-  * Perhaps because open source norms conflict with academic norms?
-* The members of this community face structural challenges in academia
--->
-
-Our goal in this study was to investigate the under-explored research practices involved in research software engineering. 
+Our goal in this study was to investigate the under-explored research practices involved in research software engineering in archaeology.
 We sought to identify not only _what_ kinds of software archaeologists are making, but _how_ archaeologists create these tools as part of a broader community of practice. 
@@
 To investigate these concerns, we operationalised open-source collaborative experiences as the use of certain features of git and GitHub visible to us in data from the GitHub API. 
-With this data, we documented that open source software development in archaeology has seen a rapid and sustained rise beginning around 2014. 
+With this data, we documented that open source software development in archaeology has seen a rapid and sustained rise beginning around 2014 (@fig-github-cumulative). 
 This is marked by a variety of applications and use cases, including the use of git and GitHub to track and host content other than code. 
@@
 At the same time, archaeologists also create standalone software for more intensive tasks that require greater access to system resources or that warrant more complex user interfaces than what R and Python IDEs are capable of providing.
+These tools tends to be focused on various means of identifying distribution patterns (spatial, temporal, statistical), calibrating data obtained from various instrumental methods (XRF, luminescence dating), supporting specialized finds analysis (zooarchaeology, palaeobotany, archaeogenetics), and supporting the collection and processing of archaeological materials.
+These foci signify gaps in the archaeological toolbox that archaeologists recognized, and have attempted to fill, on their own terms.
 
-Looking at the distribution of tags, we note that, among items listed as software, repositories tend to be focused on 
-various means of identifying distribution patterns (spatial, temporal, statistical), 
-calibrating data obtained from various instrumental methods (XRF, luminescence dating), 
-supporting specialized finds analysis (zooarchaeology, palaeobotany, archaeogenetics), 
-and supporting the collection and processing of archaeological materials.
-These foci signify gaps in the archaeological toolbox that archaeologists recognized, and have attempted to fill, on their own terms.
-In future studies, it would be interesting to examine how the purposeful expansion of open source tools corresponds with broader methodological trends apparent in publishing patterns or through other expressions of archaeological interest (e.g.
-social media posts, conference sessions).
+There is an emerging community of practice around open source research software in archaeology. 
+All but a handful of the GitHub repositories we analysed have more than commit, showing that archaeologists use it for ongoing work rather than merely to upload finished products.
+They relatively frequently make use of the 'star' and 'comment' features to engage with others' repositories (@fig-collab-features) and, via these and other shared contributions, we can trace a collaborative network that includes the majority of archaeologists active on GitHub (see @sec-network).
 
-While there is an emerging community of practice around open source in archaeology, we observed that collaboration remains limited. 
-Most work is performed individually and is short-lived. 
+On the other hand, we found that the forms and intensity of collaboration remains limited.
+Most work is performed individually (@fig-contributions) and is short-lived (@fig-lifespan; @fig-lifespan-rate). 
 The vast majority of repositories have 1--3 contributors, with only a few distinguished by an active and diverse developer base. 
 Our analysis also shows an uneven use of git and GitHub's extended features, beyond their basic usage as a version control system and repository host. 
-Generally speaking, we believe that this is because people do not want to step on other people's toes by raising issues or intruding on other people's projects. 
+While GitHub's more passive collaborative features (stars, comments) are commonly used, those that involve direct engagement with repository content (issues, forks, pull requests) are not (@fig-collab-features);
+perhaps because people do not want to 'step on toes' or be seen to be intruding on others' projects. 
 This may relate to the fact that most developers on this list are academics who hold different values relative to the designers of open source development environments, regarding how collaboration should occur, for example, when dealing with how projects and ideas are 'owned' by individuals or communities, and how work should be iteratively improved upon.
 
-Our network analysis similarly draws attention to the real-world collaborative ties that underpin archaeological open source software development.
+Our network analysis (@sec-network) similarly draws attention to the real-world collaborative ties that underpin archaeological open source software development.
 We identify a core cluster representing a series of collaborative ties among members of an archaeological software engineering community of practice.
-This core exhibits complexity that corresponds with social patterns, such as the presence of various sub-clusters representing interconnected interest or affinity groups.
-Indeed, we demonstrate that real-world social connections and institutional support structures are strong predictors of centrality, since these sub-clusters are representative of established professional partnerships.
-This indicates that archaeological open source is firmly embedded within existing power structures that permeate academic life, both online and offline.
+This core exhibits complexity that corresponds with social patterns, such as the presence of various clusters representing interconnected interest or affinity groups.
+Indeed, we have found that 'real-world' social connections and institutional support structures are strong predictors of centrality, since these clusters are representative of established professional partnerships.
+This suggests that archaeological open source is firmly embedded within existing power structures that permeate academic life, both online and offline.
+Similarly, we found that the individuals who play critical roles in supporting the archaeological open source community are precariously employed workers. 
+Far from open source being inherently distributed, resilient, and open-ended, this indicates that research software engineering is actually quite centralized, fragile, and based heavily on existing professional connections and endeavours.
 
-There is thus little evidence to support the notion that archaeologists benefit from the positive outcomes that are commonly argued to be the natural results of open source development models -- namely, greater degrees of transparency, extensibility and participatory action.
+These findings call into question the notion that archaeologists benefit from the positive outcomes that are commonly argued to be the natural results of open source development models -- namely, greater degrees of extensibility and participatory action.
 While opening the source code may facilitate these positive outcomes as necessary preconditional factors, we argue that this only amounts to establishing the _potential_ for people to put these values into practice.
 Moreover, we argue that the objectives and circumstances that frame archaeological practice significantly influence how far archaeologists (and academics in general) are willing to push for these values, and limit the ability for archaeologists to do open source in ways that resemble more mainstream open source projects.
-For instance, successful open source projects like the Linux kernel, openSSL the Firefox web browser are driven by collective and popular interest in ensuring that code remains functional, and the code base is therefore constantly in flux and bears an accumulating list of contributing members.
-This significantly differs from the organizational principles that govern archaeological work, namely that a project's leaders runs the whole show from the top down and operationalize all other actors as instruments of their central directives.
+For instance, successful open source projects like the Linux kernel, openSSL, or the Firefox web browser are driven by collective and popular interest in ensuring that code remains functional, and the code base is therefore constantly in flux and bears an accumulating list of contributing members.
+This differs from the organizational principles that govern much archaeological work, namely where a director or directors (of a field project, research group, etc.) sets the goals and orientation of the group and commissions and manages other actors accordingly.
 Moreover, archaeological projects ultimately seek to produce stable textual outcomes bearing clear delineation of authorship and that require no upkeep whatsoever. 

### 2024-02-07 analysis/openarchaeo_collaboration.qmd
Revise conclusions text

@@
   Surveying the first quarter-century of computer applications in archaeology, Scollar [-@scollar1999] lamented that the field relied almost exclusively on "hand-me-down" tools repurposed from other disciplines.
-25 years later, this is no longer the case: computational archaeologists often find themselves practicing the dual roles of data analyst and research software engineer [@baxter2012; @schmidt2020], developing and applying new tools that are tailored specifically to archaeological problems and archaeological methods.
-Though this trend can be traced to the very earliest days of the field [@cowgill1967], its most recent manifestation is distinguished by its apparent embrace of practices from free and open source software .
+Twenty five years later, this is no longer the case: computational archaeologists often find themselves practicing the dual roles of data analyst and research software engineer [@baxter2012; @schmidt2020], developing and applying new tools that are tailored specifically to archaeological problems and archaeological methods.
+Though this trend can be traced to the very earliest days of the field [@cowgill1967], its most recent manifestation is distinguished by its apparent embrace of practices from free and open source software.
 Most prominently, since around 2015, there has been a rapid uptake of workflow tools designed for open source development communities, such as the version control system git and associated online source code management platforms (e.g.
@@
 
-  In this paper, we investigate modes of collaboration in this emerging community of practice using open-archaeo [open-archaeo.info](open-archaeo.info), a curated list of archaeological software, augmented where available with data on their activity on GitHub.
-  We apply network analysis to map the networks of collaboration that have emerged among archaeologists using GitHub, and conduct a quantitative exploratory analysis to characterize the nature and intensity of these collaborations.
-  Our results show uneven adoption of open source collaborative practices beyond the basic use of git as a version control system and GitHub to host source code.
-  For instance, a majority of repositories have 1–3 contributors, with only a few projects distinguished by an active and diverse developer base.
-  Furthermore, community-oriented features, such as GitHub's 'issues' and 'comments', are used in only a minority of repositories.
-  Our network analysis similarly demonstrates that archaeological open source is firmly embedded within existing power structures that permeate academic life, and that apparent connections are representative of established professional partnerships.
-  As such, we find little evidence to support the notion that archaeologists benefit from the potential positive outcomes that are commonly argued to be the natural results of open source development models – namely, greater degrees of transparency, extensibility and participatory action.
+  In this paper, we investigate modes of collaboration in this emerging community of practice using 'open-archaeo', a curated list of archaeological software, and data on the activity of associated GitHub repositories and users.
+  We conduct an exploratory quantitative analysis to characterize the nature and intensity of these collaborations and map the collaborative networks that emerge from them.
+  We document uneven adoption of open source collaborative practices beyond the basic use of git as a version control system and GitHub to host source code.
+  Most projects do make use of collaborative features and, through shared contributions, we can can trace a collaborative network that includes the majority of archaeologists active on GitHub.
+  However, a majority of repositories have 1–3 contributors, with only a few projects distinguished by an active and diverse developer base.
+  Direct collaboration on code or other repository content—as opposed to the more passive, social media-style interaction that GitHub supports—remains very limited.
+  In other words, there is little evidence that archaeologists' adoption of open source *tools* (git and GitHub) has been accompanied by the decentralized, participatory forms of collaboration that characterise other open source communities.
+  On the contrary, our results indicate that research software engineering in archaeology remains largely embedded in conventional professional norms and organizational structures of academia.
 bibliography: references.bib
@@
   echo: false
+  message: false
+  warning: false
 ---

### 2024-03-14 analysis/_01-introduction.qmd
Started implementing changes posed by the peer-reviewer.

@@
 
-In their seminal vision for open archaeology, Beck and Neylon [-@beck2012 480-481] identified the movement as comprising a series of principles and practices "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity".
-Open archaeology therefore promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, education, and review processes.
-Archaeologists are also actively engaged in open source software development as means of sharing their research processes and creating tools and resources for general use.
+In a 2012 special issue of World Archaeology, Beck and Neylon [-@beck2012 480-481] idenfified open archaeology as comprising a series of principles and practices "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity".
+They situate open archaeology as a disciplinary branch of the open science movement, which promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, education, and review processes.
+@ducke2012 similarly argued that archaeologists should more actively engage in open source software development, which involves making source code available for anyone to freely access, modify or reuse, and which complements a common vision of open science as a communal, community-driven, self-correcting and wholly-transparent mode of production.
 
-The relationship between open science and open source is complicated by rhetorical claims that have a questionable connection to how academics actually _do_ open source. 
+However, the relationship between open science and open source is complicated by rhetorical claims that have a questionable connection to how academics actually _do_ open source. 
 Does academic open source actually make research processes more transparent and improve research outcomes? 
@@
 
-Despite hopeful aspirations espoused by open archaeology advocates [see @kansa2014; @kintigh2015], these outcomes are not a given.
+Despite these hopeful aspirations espoused by open archaeology advocates [see @kansa2014; @kintigh2015], these beneficial outcomes are not a given.
 As per @nguyen2022, @pownall2023 and @leonelli2023 we believe that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
@@
 
-This article presents a survey of archaeological software development with two goals in mind:
+This article presents a survey of open source archaeological software development with two goals in mind:
 

### 2024-03-14 analysis/_02-background.qmd
Started implementing changes posed by the peer-reviewer.

@@
 By putting one's code on the web without restriction on how it may be used or manipulated, this encourages creativity to flourish as people contribute to help improve the code base.
-Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision [see @kelty2008 on the emacs saga for a quintessential example].
+Software thus emerges from the coordinated labour of worldwide volunteers, who shape the product according to the collective vision.
 An open code base may also be used to support alternative projects whose missions diverge from the original plan, and an entire project may be "forked", or taken in a new direction if contributors are dissatisfied with how core developers run things.
@@
 However, this is not the same as saying that open source is completely anarchic or based on the "wisdom of crowds".
-In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms or social mediation to help delegate and vet contributions made by distributed participants [@oneil2009].
+In fact, successful open source projects incorporate complex organizational structures, governance strategies, and forms of social mediation to help delegate and vet contributions made by distributed participants [@oneil2009].
 They rely on, rather than eschew, institutional support structures, in order to motivate work, keep volunteer maintainers involved, and generally ensure that the project can be sustained over the long-term.
@@
 Namely, it considers science as the accumulation and assembly of a species-level understanding of the world, which is not held by any one individual but is stored in seemingly value-neutral and disembodied media, facts and observations.
-This is manifested by digital telecommunications systems that host files, document processes, facilitate co-working opportunities, and perform automated processes.
+This is manifested by information and communication technologies that host files, document processes, facilitate co-working opportunities, and perform automated processes (e.g.
+digital repositories, reproducible notebooks, automated evaluation of research findings, systems for ensuring stable references to consistent identifiers, etc).
 However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.
@@
 But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce.
-Whereas open source emerged from concern of consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible.
+Whereas open source emerged from concerns over consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible.
 Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.
@@
 Once a project is completed, papers have been published, and credit has been allocated, it is common for scientists to mark their projects as finished and move on to new endeavours [@kelty2008: 271-275; @howison2013].
+Scientists may archive their work in an institutional repository, at which point the work enters a stasis state that invites reference to the completed work, but which precludes any potential for the work to be updated, modified, or directly built upon.
 Open source projects, on the other hand, are motivated by a more practical need for the software to function properly in perpetuity, and contributors may remain actively or sporadically involved to satisfy users' needs, or to direct users to derivative and functional forks of abandoned software [@kelty2008: 278-281; @coleman2012: 116-122; @hippel2003].
@@
 Software forges—collaborative web platforms like GitHub, GitLab and Codeberg—are designed to facilitate open source software development by hosting public git repositories.
-However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.
+However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.^[Much of this relies on proprietary software owned and managed by commercial organisations, and there has been some controversy about the take-over of open source platforms by for-profit entities (see XXXX)].
 These platforms also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as a combine bookmarking and 'like' feature, and maintain a public-facing profile that includes personally-identifying information (e.g.

### 2024-03-14 analysis/_03-data_methodology.qmd
Updating reference to the compendium
Started implementing changes posed by the peer-reviewer.

@@
 
-We present an exploratory quantitative analysis of open-archaeo [[open-archaeo.info](open-archaeo.info), @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources maintained primarily by one of us (ZB) since 2018.
+We present an exploratory quantitative analysis of open-archaeo [open-archaeo.info, @batist2023], a directory of `r nrow(oarch)` pieces of open source archaeological software and other digital resources maintained primarily by one of us (ZB) since 2018.
 
@@
 However, open-archaeo generally lacks software written before archaeologists started using collaborative software development platforms such as GitHub, and software that is not shared on the web at all.
+It also includes numerous non-software resources, as well as software developed and distributed without the use of software forges.
 The dataset is also limited by the experiences of its primary maintainers.^[We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
@@
 Where applicable, we obtained more detailed information about each repository's contents and contribution histories from the GitHub API (application programming interface).
-Our analysis incorporates data on `r n_repos` repositories, comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
+Our analysis incorporates data on `r n_repos` repositories^[We excluded three GitHub repositories from the analysis for technical reasons.
+For example, https://github.com/carpentries-incubator/R-archaeology-lesson is a repository within the scope of open-archaeo, but it was forked from a pre-existing repository that is not (https://github.com/datacarpentry/R-ecology-lesson) and thus includes in its commit history irrelevant data from the parent repository.], comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
 
 We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
-This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or host it elsewhere (`r percent(p_no_github)` of the total), are excluded from these parts of the analysis, though we were still able to perform an analysis of their contents and authorship from the data compiled in open-archaeo itself.
-We also cannot include collaboration through offline or private channels, or forms of collaboration we do not know about. 
+This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or which develop and host code on platforms other than GitHub (`r percent(p_no_github)` of the total), were excluded from parts of the analysis that examine *how* archaelogists develop software.
+However, we were still able to draw from all records to ascertain the general composition of open-archaeo, and by extension, to address *what* kinds of software and resources archaeologists make.
+
+That being said, we cannot account for practices that occur through offline or private channels, or forms of collaboration we do not know about. 
 We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
@@
 The quantitative analyses and figures presented here were generated with `r version$version.string` [@rcoreteam2023].
-The full data and code is available in the compendium that accompanies this paper [@roe2023] <!-- TODO: reference should be updated to reflect the actual version number and date -->.
+The full data and code is available in the compendium that accompanies this paper [@roe2024].

### 2024-03-14 analysis/_04-open_archaeology.qmd
Started implementing changes posed by the peer-reviewer.

@@
 `r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used platform by a large margin (@tbl-platforms).
-Python, another programming language, is also relatively popular (`r percent(p_platform_python)`), as are plugins for the open source geographic information system QGIS (`r percent(p_platform_qgis)`).
+Python, another programming language, is also relatively popular (`r percent(p_platform_python)`), as are plugins for the open source geographic information system QGIS (`r percent(p_platform_qgis)`).^[While Blender and QGIS plugins are written using the Python language, our intent while categorizing platforms was to get a sense of the developer ecosystems in which archaeological software engineers participate, rather than to simply gauge the popularity of different languages [@batist2023 2].]
 Beyond that, there is a rather fragmented landscape of plugins for other desktop software (e.g.
@@
 Tags that are notable for having a higher proportion of standalone software include archaeogenetics, data management, 3D modelling, photogrammetry, drivers and IO, and simulations or agent based modelling.
-These tools may require greater access to system resources, or may require more complex user interfaces than are more complex than what R or Python IDEs (integrated development environments) tend to provide.
+These tools may require greater access to system resources, or may require more complex user interfaces than what R or Python IDEs (integrated development environments) tend to provide.
 
 ```{r tbl-tags}
-#| tbl-cap: Themes of open archaeology projects
+#| tbl-cap: Themes of open archaeology projects. See https://github.com/zackbatist/open-archaeo/blob/master/tags.md for a more comprehensive description of each tag's scope.
 oarch |>
@@
 
-However, since around 2022, the number of new GitHub repositories has significantly plateaud, while the cumulative number of commits has continued to rise.
+However, since around 2022, the number of new GitHub repositories has significantly plateaued, while the cumulative number of commits has continued to rise.
 This may relate to a general emphasis on maintaining existing code and working on established projects, rather than spinning up new ones.

### 2024-03-14 analysis/_05-collaboration.qmd
Started implementing changes posed by the peer-reviewer.

@@
 While git can be a great way to track changes and manage distributed contributions to a common code base, it can also be unwieldy in situations when multiple users (especially those with less experience using git for collaborative purposes) are expected to contribute within short spans of time.
-This meshes with our prior observations that projects which tend to exhibiting higher commit rates have less contributors.
-Additionally, our analyses neglects to account for contributions that are not tracked via git or GitHub.
+This meshes with our prior observations that projects exhibiting higher commit rates have less contributors.
+Additionally, our analyses neglect to account for contributions that are not tracked via git or GitHub.
 Those who do not code may provide creative guidance or feedback during in-person meetings, via email, or using alternative online messaging or social media platforms.

### 2024-03-14 analysis/_06-network.qmd
Started implementing changes posed by the peer-reviewer.

@@
 Python projects, which tend to be infrastructural or are of interest to members of other fields, are also highly ranked in terms of average centrality.
-^[While Blender and QGIS plugins are written using the Python language, our intent while categorizing platforms was to get a sense of the developer ecosystems in which archaeological software engineers participate, rather than to simply gauge the popularity of different languages [@batist2023 2].]
 Our findings support the notion that there are significant development patterns when working across different languages and platforms, and further analysis to qualify these observation is warranted.

### 2024-03-18 analysis/_05-collaboration.qmd
Some further revisions to the text of section 5.

@@
 Use of GitHub's collaboration features is also generally low (@fig-collab-features), although the data also shows a divergence between the uptake of features that facilitate direct code contributions (forks, issues, pull requests), which have markedly zero-skewed distributions, versus more indirect, social media-like features (comments, stars), which are moderately well-used.
-We hypothesise that this shows a preference for passive/reactive rather than active/proactive engagement with others work – a point we will return to in the conclusion.
+
+We hypothesise that these findings show a preference for passive/reactive rather than active/proactive engagement with others' work.
+This attitude is not conducive to achieving the desired outcomes of opening source code, namely enabling greater engagement, reuse and critique.
+The technical apparatus provides the _potential_ for archaeologists to perform these actions, but these are inherently social activities, and as such, are circumscribed by social norms, including expectations and taboos surrounding the permeability of a grant-funded project's boundaries, or whether or not it is proper to actively engage or interfere with the work directed under another the aegis of another project, during or after the project's completion.
+Our findings therefore suggest that there is no strong will to radically re-arrange the collaborative commitments that govern research practice.
+If we want to make open source effective in relation to the aforementioned goals, we also need to foster a culture that supports active, pragmatic and humble critique, and which instills a de-territorialized attitude concerning what it means to contribute to collective knowledge.
+This means fighting against the pathological power-relations that scaffold all aspects of academic life, and not fooling ourselves into believing that technical solutions (i.e.
+using git) will on their own resolve wicked social problems.

### 2024-03-18 analysis/_06-network.qmd
Additional revisions to section 6.

@@
 This is evident through the fact that many of the connections represented in the cores emerge from more conventional professional networks, e.g.
-[ISAAKiel](https://isaakiel.github.io/), a working group centered around the University of Kiel), or [CAA-SSLA](https://sslarch.github.io/), a special interest group of the international scholarly society 'Computational and Quantitative Applications in Archaeology' (CAA) focused on scientific programming.
+the [ISAAKiel](https://isaakiel.github.io/) group based at the University of Kiel or [CAA-SSLA](https://sslarch.github.io/), a special interest group of the international scholarly society 'Computational and Quantitative Applications in Archaeology' (CAA) focused on scientific programming, and which clusters around users 'nevrome' and 'martinHinz', who are core members of both organizations.
 Peripheral clusters that are connected to the central core by only a few relationships represent the sole (or perhaps initial) integration of lone developers into a broader community.

### 2024-03-21 analysis/_01-introduction.qmd
Work more references into introduction

--- 2024-02-06 analysis/_01-introduction.qmd
@@
 
-In their seminal vision for open archaeology, Beck and Neylon [-@beck2012 480-481] identified the movement as comprising a series of principles and practices "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity".
-Open archaeology therefore promotes more thoughtful scholarly communication practices, most notably in operations relating to publishing, data-sharing, education, and review processes.
-Archaeologists are also actively engaged in open source software development as means of sharing their research processes and creating tools and resources for general use.
+A 2012 special issue of *World Archaeology* marked the coming of age of 'open archaeology',
+a new, digitally-oriented archaeology "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity" [@beck2012 480-481],
+situated within the wider the open science movement [@lake2012 and @beck2012].
+In the same issue, @ducke2012 specifically called attention to software, calling on archaeologists to more actively engage in open source practices – making software with source code available for anyone to freely access, modify or reuse.
+Open source, though originating outside academy, emerged as an important component and complement of open science, sharing its vision for a communal, self-correcting and transparent mode of knowledge production.
+Where before archaeologists had relied almost exclusively on "hand-me-down" software re-purposed from other fields [@scollar1999], open archaeology envisaged a community of archaeological research software engineers oriented around open source development practices and tools.
 
-The relationship between open science and open source is complicated by rhetorical claims that have a questionable connection to how academics actually _do_ open source. 
+More than decade on, we are in a position to ask whether this hopeful early rhetoric of open archaeology has been borne out in actually-existing open source research software engineering in the field.
 Does academic open source actually make research processes more transparent and improve research outcomes? 
@@
 Is it actually helping to foster new globe-spanning connections and lead to novel research trajectories that would not otherwise come to pass?
-Basically, is there more to it than just uploading code and data files to the internet?
+In other words, is there more to 'open archaeology' than just uploading text files to the internet?
 
-Despite hopeful aspirations espoused by open archaeology advocates [see @kansa2014; @kintigh2015], these outcomes are not a given.
-As per @nguyen2022, @pownall2023 and @leonelli2023 we believe that these outcomes only actually arise in contexts where participants adhere to and are motivated by warrants, professional norms, and governance strategies that encourage these results.
-However, practical circumstances and systemic value regimes that frame what it means to work as an archaeologist presently inhibit the potential for radical transformation, even among open science's most ardent supporters.
+The aspirations of open archaeology's early proponents [e.g.
+@kansa2014; @kintigh2015] were tempered by notes of caution that "the reward structures in academic and professional archaeology do little to incentivise participation in Open Archaeology" [@lake2012, echoing @beck2012, @kansa2012, @huggett2012 and @limp2011].
+Following recent work by @nguyen2022, @pownall2023 and @leonelli2023, we believe that the outcomes listed above indeed only arise in contexts where there are organization structures, governance strategies, and professional norms that encourage them.
+Thus practical circumstances and systemic value regimes that frame what it means to work as an archaeologist presently inhibit the potential for radical transformation, even among open science's most ardent supporters.
 

### 2024-03-21 analysis/_06-network.qmd
Put names to usernames

--- 2024-03-14 analysis/_06-network.qmd
@@
 This is evident through the fact that many of the connections represented in the cores emerge from more conventional professional networks, e.g.
-[ISAAKiel](https://isaakiel.github.io/), a working group centered around the University of Kiel), or [CAA-SSLA](https://sslarch.github.io/), a special interest group of the international scholarly society 'Computational and Quantitative Applications in Archaeology' (CAA) focused on scientific programming.
+the [ISAAKiel](https://isaakiel.github.io/) group based at the University of Kiel or [CAA-SSLA](https://sslarch.github.io/), a special interest group of the international scholarly society 'Computational and Quantitative Applications in Archaeology' (CAA) focused on scientific programming, and which clusters around users 'nevrome' (Clemens Schmid) and 'martinHinz' (Martin Hinz), who are core members of both organizations.
 Peripheral clusters that are connected to the central core by only a few relationships represent the sole (or perhaps initial) integration of lone developers into a broader community.

### 2024-03-21 analysis/_03-data_methodology.qmd
Add appendices

@@
   saveRDS(oarch, cached_gh_data)
+  # TODO: save as CSVs too
 }

### 2024-03-21 analysis/_99-appendices.qmd
Add appendices

# Data availability {.appendix}

The results presented here are based on the directory of open archaeology projects maintained at <https://open-archaeo.info> [@batist2023].
The specific version we used is available from Zenodo at <https://doi.org/10.5281/zenodo.10625236> as well as in the compendium associated with this paper [@roe2024, https://zenodo.org/doi/10.5281/zenodo.8393043], in CSV format under `analysis/raw_data`.

Further data on activity in the git repositories associated with these projects was obtained via the GitHub API.
This data is available from Zenodo in the compendium associated with this paper [@roe2024, https://zenodo.org/doi/10.5281/zenodo.8393043], in RData and CSV format under `analysis/derived_data`.

# Acknowledgements {.appendix}

The impetus for this work was a session at the Computer Applications and Quantitative Methods in Archaeology (CAA) virtual annual meeting in Cyprus in June 2021, 'Tools for the Revolution: developing packages for scientific programming in archaeology'.
We are grateful to the conference organizers as well as to the (co-)convenors of the session, Martin Hinz and Clemens Schmid, for facilitating it.
Two anonymous reviewers offered feedback on our initial conference abstract.
The session was organized by the CAA special interest group on 'scientific scripting languages in archaeology' (CAA-SSLA, https://sslarch.github.io/), and we must thank its members—especially Sophie Schmidt, Clemens Schmid, Martin Hinz and Ben Marwick—for the encouragement in developing the work further.

Our original manuscript benefited greatly from the insights of the anonymous reviewer for *Internet Archaeology*.
We are also grateful to *IA* editor Judith Winters for managing the process and facilitating the interactive figures.

JR was supported by the Swiss National Science Foundation (SNSF Project #198152) for the duration of the production of this work.

<!-- TODO: Zack: anything to add? -->

### 2024-03-21 analysis/openarchaeo_collaboration.qmd
Add appendices

@@
 {{< include _07-conclusion.qmd >}}
+{{< include _99-appendices.qmd >}}

### 2024-03-21 analysis/_04-open_archaeology.qmd
Replace link to GitHub with JOAD citation
I think that's the more permanent reference?
Fix citation
This is intended to refer to Kansa's chapter in 'Critical Archaeology in the Digital Age: Proceedings of the 12th IEMA Visiting Scholar’s Conference'

@@
 ```{r tbl-tags}
-#| tbl-cap: Themes of open archaeology projects. See https://github.com/zackbatist/open-archaeo/blob/master/tags.md for a more comprehensive description of each tag's scope.
+#| tbl-cap: Themes of open archaeology projects. See Batist and Roe [-@batist2023 table 1] for a description of each tag's scope.
 oarch |>
@@
 
-As we can see, archaeologists have been using git from even before GitHub was launched in 2008.
-But use of git really began to take off around 2014--2015, when we see an uptick in the rate of growth.
-Around this time we also see that GitHub starts being used to host documents and scripts.
-This may represent a recognition of GitHub's ability to track things other than code, and a willingness to experiment with version control systems as a medium for disseminating work in an open and somewhat nerdy way.
+Archaeologists have been using git from the early 2000s, before GitHub was launched,^[Commits dated to before GitHub's launch in 2008 come from repositories that were created earlier and later imported to GitHub's platform.
+We have excluded commits that have obviously erroneous dates, e.g.
+`2001-01-01` in projects started in the mid–late 2010s.].
+but it saw a marked increase in popularity c. 2014--2015.
+From that point and until recently there was an exponential uptake of GitHub by archaeologists,
+but while we were preparing this paper (c. 2022) the first signs of a diminishing growth have appeared,
+with the number of cumulative commits continuing to rise but the number of repositories hitting a sharp plateau.
+This could be explained by market exhaustion, a shift in emphasis to maintaining existing code and working on established projects, and/or growing doubts about the appropriateness and sustainability of GitHub following its acquisition by Microsoft [@kansa2022b].^[Concerns we very much share.]
 
-However, since around 2022, the number of new GitHub repositories has significantly plateaued, while the cumulative number of commits has continued to rise.
-This may relate to a general emphasis on maintaining existing code and working on established projects, rather than spinning up new ones.
+GitHub's entry into the digital archaeology mainstream in 2014--2015 also marks the point at which we see it being used for things other than packaged source code (e.g. documents and scripts).
+This suggests that the 'early adopters' of GitHub, pre-2014, were more directly embedded in existing (open source) software development communities,
+while those that came later also saw version control systems as a potential medium for dissemination and archiving.
+It may also reflect a general move towards git- and GitHub-based workflows by archaeologists attracted to open, participatory, and/or generally 'nerdy' academic practices.

### 2024-03-21 analysis/_99-appendices.qmd
Shorten acknowledgements and fix links

@@
 The results presented here are based on the directory of open archaeology projects maintained at <https://open-archaeo.info> [@batist2023].
-The specific version we used is available from Zenodo at <https://doi.org/10.5281/zenodo.10625236> as well as in the compendium associated with this paper [@roe2024, https://zenodo.org/doi/10.5281/zenodo.8393043], in CSV format under `analysis/raw_data`.
+The specific version we used is available from Zenodo at <https://doi.org/10.5281/zenodo.10625236> as well as in the compendium associated with this paper [@roe2024, https://doi.org/10.5281/zenodo.8393043], in CSV format under `analysis/raw_data`.
 
 Further data on activity in the git repositories associated with these projects was obtained via the GitHub API.
-This data is available from Zenodo in the compendium associated with this paper [@roe2024, https://zenodo.org/doi/10.5281/zenodo.8393043], in RData and CSV format under `analysis/derived_data`.
+This data is available from Zenodo in the compendium associated with this paper [@roe2024, <https://doi.org/10.5281/zenodo.8393043>], in RData and CSV format under `analysis/derived_data`.
 
@@
 
-The impetus for this work was a session at the Computer Applications and Quantitative Methods in Archaeology (CAA) virtual annual meeting in Cyprus in June 2021, 'Tools for the Revolution: developing packages for scientific programming in archaeology'.
-We are grateful to the conference organizers as well as to the (co-)convenors of the session, Martin Hinz and Clemens Schmid, for facilitating it.
-Two anonymous reviewers offered feedback on our initial conference abstract.
-The session was organized by the CAA special interest group on 'scientific scripting languages in archaeology' (CAA-SSLA, https://sslarch.github.io/), and we must thank its members—especially Sophie Schmidt, Clemens Schmid, Martin Hinz and Ben Marwick—for the encouragement in developing the work further.
+The impetus for this work was a session at the Computer Applications and Quantitative Methods in Archaeology (CAA) virtual annual meeting in Cyprus in June 2021.
+Our thanks to Martin Hinz, Clemens Schmid, Sophie Schmidt and other members of the 'scientific scripting languages in archaeology (CAA-SSLA, <https://sslarch.github.io/>) for organizing the session and encouraging us to develop the work further.
 

### 2024-03-29 analysis/_02-background.qmd
More updates in response to reviewer comments.

@@
 Whereas open source emerged from concerns over consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible.
-Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.
+Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.^[Though, participatory or community-based research does exist and is an exception to this generalization (see @morgan2012).
+It requires active effort to do well and should not be considered a passive by-product of posting one's research on the web.]
 Moreover, whereas open source projects often attract participants with varied stakes in the software and use cases in mind, open science projects are typically bounded by small communities of specialists with very particular needs [@kling2003].
@@
 Software forges—collaborative web platforms like GitHub, GitLab and Codeberg—are designed to facilitate open source software development by hosting public git repositories.
-However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.^[Much of this relies on proprietary software owned and managed by commercial organisations, and there has been some controversy about the take-over of open source platforms by for-profit entities (see XXXX)].
+However, they also support common software developer and project management practices, such as issue and bug tracking, code-commenting, task management, identity and permissions management, web publishing, vulnerability detection, creation and maintenance of metadata, and financial sponsorship.^[Much of this relies on proprietary software owned and managed by commercial organisations, and there has been some controversy surrounding the take-over of open source platforms by for-profit entities (see @saunders2022 and @brembs2023)].
 These platforms also implement standard social media functions, like the ability to follow projects and individual users to receive updates on their activities, "star" certain repositories as a combine bookmarking and 'like' feature, and maintain a public-facing profile that includes personally-identifying information (e.g.

### 2024-03-29 analysis/_03-data_methodology.qmd
More updates in response to reviewer comments.

@@
 We opted to only collect repository data from GitHub because it is the most popular forge platform used by open-archaeo projects (@tbl-forges).
-This means that projects that do not use version control (`r percent(p_no_vc)` of the total), or which develop and host code on platforms other than GitHub (`r percent(p_no_github)` of the total), were excluded from parts of the analysis that examine *how* archaelogists develop software.
+This means that we excluded projects that do not use version control (`r percent(p_no_vc)` of the total) or which develop and host code on platforms other than GitHub (`r percent(p_no_github)` of the total) from the parts of the analysis that examine *how* archaelogists develop software.
 However, we were still able to draw from all records to ascertain the general composition of open-archaeo, and by extension, to address *what* kinds of software and resources archaeologists make.

### 2024-03-29 analysis/_05-collaboration.qmd
Comment on new conclusion in section 5
More updates in response to reviewer comments.

--- 2024-03-14 analysis/_05-collaboration.qmd
@@
 Use of GitHub's collaboration features is also generally low (@fig-collab-features), although the data also shows a divergence between the uptake of features that facilitate direct code contributions (forks, issues, pull requests), which have markedly zero-skewed distributions, versus more indirect, social media-like features (comments, stars), which are moderately well-used.
-We hypothesise that this shows a preference for passive/reactive rather than active/proactive engagement with others work – a point we will return to in the conclusion.
+
+These findings show a preference for passive/reactive rather than active/proactive engagement with others' work, which is not conducive to achieving the desired outcomes of opening source code, namely enabling greater engagement, reuse and critique.
+While limited time or technical capability may be contributing factors (which should be targeted in more focused investigations), we believe that social norms, such as expectations and taboos surrounding the permeability of a grant-funded project's boundaries, or whether it is proper to actively engage or interfere with the work directed under another the aegis of another project, play a very significant role.
+Our network analysis of collaborative ties, presented in the following section, corroborates this claim.

### 2024-03-29 analysis/_07-conclusion.qmd
More updates in response to reviewer comments.

@@
 As such, we advocate for more focused attention on specific disciplinary norms and institutional support structures that inform how knowledge is created and validated, and how varied contributions to the scholarly enterprise are mediated, credited, and valued.
-These factors, which are often ignored in open science discourse, vary from discipline to discipline and are rarely accounted for in the infrastructures themselves and in the policies that dictate or guide their implementation.
-However these factors should be considered as equal, if not more important, elements contributing to the success of open science, relative to the technical apparatus.
+In other words, if we want to make open source effective in relation to the aforementioned goals of encouraging greater inclusivity, transparency, and productivity, we also need to foster a culture that supports active, pragmatic and humble critique, and which instills a de-territorialized attitude concerning what it means to contribute to collective knowledge [@morgan2015; @ducke2015].
+This means fighting against the pathological power-relations that scaffold all aspects of academic life, and not fooling ourselves into believing that technical solutions (i.e.
+using git) will, on their own, resolve the wicked social problems that lie at the heart of scientific research practice.

### 2024-04-09 analysis/_02-background.qmd
Additional changes based on reviewer comments.

@@
 digital repositories, reproducible notebooks, automated evaluation of research findings, systems for ensuring stable references to consistent identifiers, etc).
-However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science.
+However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science [@leonelli2023: 23-24].
 

### 2024-04-09 analysis/_07-conclusion.qmd
Additional changes based on reviewer comments.

@@
 While opening the source code may facilitate these positive outcomes as necessary preconditional factors, we argue that this only amounts to establishing the _potential_ for people to put these values into practice.
-Moreover, we argue that the objectives and circumstances that frame archaeological practice significantly influence how far archaeologists (and academics in general) are willing to push for these values, and limit the ability for archaeologists to do open source in ways that resemble more mainstream open source projects.
+We argue that the objectives and circumstances that frame archaeological practice significantly influence how far archaeologists (and academics in general) are willing to push for these values, and limit the ability for archaeologists to do open source in ways that resemble more mainstream open source projects.
 For instance, successful open source projects like the Linux kernel, openSSL, or the Firefox web browser are driven by collective and popular interest in ensuring that code remains functional, and the code base is therefore constantly in flux and bears an accumulating list of contributing members.

### 2024-04-15 analysis/_02-background.qmd
Add archaeological examples

@@
 While these additional features are meant to complement and enhance the experience of contributing to open source projects, they are not actually part of the git protocol.
+
+As a concrete example of how git and GitHub is used by archaeologists, we can take <https://open-archaeo.info> itself.
+The website at that address was at the time of writing generated from a *git* source repository containing the data on the individual entries, a set of HTML templates, and some R scripts that translate between them.
+A copy of this repository can be downloaded and worked on locally by anyone, who will have access not just to the current state of the source code but its full history through the git *version control* software.
+The authoritative version is hosted on *GitHub* at <https://github.com/zackbatist/open-archaeo>.
+The GitHub URL indicates the primary *maintainer* of the repository, 'zackbatist' – that is, one of us (Zack) – who created and has ultimate control over it.
+However, various other *contributors* (such as the other of us, Joe) have added entries, corrected entries, or added functionality to the website using the *fork* and *pull request* features of GitHub.
+The basic unit of this type of contribution is the *commit*, which is a discrete set of changes (e.g. adding an entry) associated with one contributor at a single point of time.
+Others have contributed by raising *issues* describing problems with or suggestions for the project, leaving *comments* on these issues, or more loosely by using GitHub's social media features (*stars* and *following*).
+The full history of all these types of contributions to open-archaeo can be accessed through GitHub – or, as we use here, its programmatic API.
+This basic workflow is the same whether the project in question is a document like open-archaeo, or a piece of software, or a research paper; though actual patterns of collaboration vary markedly, as we will see.

### 2024-04-15 analysis/_04-open_archaeology.qmd
Add archaeological examples

@@
 In our breakdown of open-archaeo by category (see @tbl-categories), we demonstrate the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-Most resources (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language or a plugin for an application.
+Most resources (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language (e.g.
+radiocarbon calibration is implemented in R in the package 'rcarbon' or Python in the package 'iosacal') or a plugin for an application (e.g.
+'ArchaeoAstroInsight' adds tools for measuring astronomical alignments to QGIS).
 Essentially such projects create additional functions within the base platform that are useful for archaeological purposes.
@@
 with the number of cumulative commits continuing to rise but the number of repositories hitting a sharp plateau.
-This could be explained by market exhaustion, a shift in emphasis to maintaining existing code and working on established projects, and/or growing doubts about the appropriateness and sustainability of GitHub following its acquisition by Microsoft [@kansa2022b].^[Concerns we very much share.]
+This could be explained by market exhaustion, a shift in emphasis to maintaining existing code and working on established projects, and/or growing doubts about the appropriateness and sustainability of GitHub following its acquisition by Microsoft [@kansa2022].^[Concerns we very much share.]
 

### 2024-04-15 analysis/_04-open_archaeology.qmd
Correct description of fig. 1
The pre-2008ies have disappeared at some point.

@@
 
-Archaeologists have been using git from the early 2000s, before GitHub was launched,^[Commits dated to before GitHub's launch in 2008 come from repositories that were created earlier and later imported to GitHub's platform.
-We have excluded commits that have obviously erroneous dates, e.g.
+Archaeologists have been using git from at least the late 2000s, shortly after GitHub was launched,^[We excluded commits that have obviously erroneous dates, e.g.
 `2001-01-01` in projects started in the mid–late 2010s.].

### 2024-04-15 analysis/_06-network.qmd
Add zoom to graphs
Add links to Table 7

@@
 
+```{r setup-girafe}
+set_girafe_defaults(
+  opts_zoom = opts_zoom(min = 1, max = 9) 
+)
+```
+
 ```{r fig-graph-repo-user}
@@
   unite(tags, tag1:tag5, sep = "; ", na.rm = TRUE) |>
+  mutate(repo = glue("[{repo}](https://github.com/{repo})")) |>
   transmute(rank = row_number(), repo, category, tags, n_commits) |>
@@
   rename_with(tools::toTitleCase) |>
-  tt()
+  tt() |>
+  format_tt(2, markdown = TRUE)
 ```

### 2024-04-15 analysis/_99-appendices.qmd
Minor fix to appendix.

@@
 The impetus for this work was a session at the Computer Applications and Quantitative Methods in Archaeology (CAA) virtual annual meeting in Cyprus in June 2021.
-Our thanks to Martin Hinz, Clemens Schmid, Sophie Schmidt and other members of the 'scientific scripting languages in archaeology (CAA-SSLA, <https://sslarch.github.io/>) for organizing the session and encouraging us to develop the work further.
+Our thanks to Martin Hinz, Clemens Schmid, Sophie Schmidt and other members of the Scientific Scripting Languages in Archaeology Special Interest Group (CAA-SSLA, <https://sslarch.github.io/>) for organizing the session and encouraging us to develop the work further.
 
@@
 JR was supported by the Swiss National Science Foundation (SNSF Project #198152) for the duration of the production of this work.
-
-<!-- TODO: Zack: anything to add? -->

### 2024-04-15 analysis/_99-appendices.qmd
Fix link

@@
 The results presented here are based on the directory of open archaeology projects maintained at <https://open-archaeo.info> [@batist2023].
-The specific version we used is available from Zenodo at <https://doi.org/10.5281/zenodo.10625236> as well as in the compendium associated with this paper [@roe2024, https://doi.org/10.5281/zenodo.8393043], in CSV format under `analysis/raw_data`.
+The specific version we used is available from Zenodo at <https://doi.org/10.5281/zenodo.10625236> as well as in the compendium associated with this paper [@roe2024, <https://doi.org/10.5281/zenodo.8393043>], in CSV format under `analysis/raw_data`.
 

### 2024-05-23 analysis/_01-introduction.qmd
Revisions based on editorial comments.

@@
 A 2012 special issue of *World Archaeology* marked the coming of age of 'open archaeology',
-a new, digitally-oriented archaeology "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity" [@beck2012 480-481],
+a new, digitally orientated archaeology "predicated on promoting open redistribution and access to the data, processes and syntheses generated within the archaeological domain" with the aim of "maximizing transparency, reuse and engagement while maintaining professional probity" [@beck2012 480-481],
 situated within the wider the open science movement [@lake2012 and @beck2012].
-In the same issue, @ducke2012 specifically called attention to software, calling on archaeologists to more actively engage in open source practices – making software with source code available for anyone to freely access, modify or reuse.
+In the same issue, @ducke2012 specifically drew attention to software -- the programs and other operating information used by a computer to analyse archaeological information -- calling on archaeologists to more actively engage in open-source practices – making software with source code available for anyone to freely access, modify or reuse.
 Open source, though originating outside academy, emerged as an important component and complement of open science, sharing its vision for a communal, self-correcting and transparent mode of knowledge production.
-Where before archaeologists had relied almost exclusively on "hand-me-down" software re-purposed from other fields [@scollar1999], open archaeology envisaged a community of archaeological research software engineers oriented around open source development practices and tools.
+Where before archaeologists had relied almost exclusively on "hand-me-down" software re-purposed from other fields [@scollar1999], open archaeology envisaged a community of archaeological research software engineers oriented around open-source development practices and tools.
 
-More than decade on, we are in a position to ask whether this hopeful early rhetoric of open archaeology has been borne out in actually-existing open source research software engineering in the field.
+More than decade on, we are in a position to ask whether this hopeful early rhetoric of open archaeology has been borne out in actually existing open-source research software engineering in the field.
 Does academic open source actually make research processes more transparent and improve research outcomes? 
@@
 
-We use quantitative analysis to consider how archaeological software development may be benefiting from, or missing out on, the affordances that open source development models provide, specifically the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
+We use quantitative analysis to consider how archaeological software development may be benefiting from, or missing out on, the affordances that open-source development models provide, specifically the value added through working as part of a broader community of invested stakeholders, processes of iterative improvement, and increased code transparency.
+As such, our work examines whether archaeologists are harnessing the collaborative potential that the open science movement ascribes to the use of open source software and resources.

### 2024-05-23 analysis/_02-background.qmd
Revisions based on editorial comments.

@@
 digital repositories, reproducible notebooks, automated evaluation of research findings, systems for ensuring stable references to consistent identifiers, etc).
-However, these systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science [@leonelli2023: 23-24].
+
+This culminates in an obsessive concern with digital workflows pertaining to legal and logistical issues; for many proponents of open science, publishing is largely considered the business of typesetting and copyright law, which could be rendered moot by using automated publishing workflows and by encouraging use of open licensing agreements (cf.
+Foster and Deardorff 2017; Harnad 1998).
+Viewed as merely technical systems, these could be resolved through technical means.
+However academic publishing and ownership involve social arrangements that serve to stabilise knowledge, grant authority to validated claims, and enable science to move forward (Kelty et al.
+2008: 274-275).
+In other words, technocentric visions of publication workflows tend to ignore the fact that publication is a cultural phenomenon, whereby projects are made complete and knowledge claims are articulated, credited, and rendered accountable to the people who proposed them.
+However, technological systems have become so emblematic of open science that the use of these tools and resources designed to _support_ open science is often mistaken for _actually doing_ open science [@leonelli2023: 23-24].
 
@@
 But they differ in terms of the contexts in which they operate, the stakeholders involved, and the kinds of outcomes they produce.
-Whereas open source emerged from concerns over consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science comes out of a desire to make research practices more transparent and accessible.
+Whereas open source emerged from concerns over consumer rights and then developed as a means of maintaining resilient and collectively motivated projects, open science is driven by a warrant to make research practices more transparent and accessible.
 Open source is performed by professional and hobbyist software developers alike, and participants contribute in a wide variety of ways (including: programming, writing documentation, translating software and documentation, bug reporting, and financial support), but, in open science, scientists are usually the only participants actively involved in creating and maintaining contributions.^[Though, participatory or community-based research does exist and is an exception to this generalization (see @morgan2012).
@@
 
-The adoption of open source development models among archaeologists is generally informed by the broader open science movement, which is motivated by a genuine desire to facilitate novel research opportunities, to make participation in scientific research more equitable, to reclaim science as a public good, and to enhance the means of validating findings.
+The adoption of open source development models among archaeologists is generally informed by the broader open science movement.
 However, the predominant concern with implementing best tools to use, adopting optimal data processing pipelines, and tying into global, web-based infrastructures, protocols and standards [cf.
@@
 [-@atici2013 676-677], Huggett [-@huggett2018; -@huggett2022], @sobotkova2018, @opitz2021, @haciguzeller2021, and @batist2023a demonstrate that to make the reuse of archaeological data feasible and useful in a practical sense, it is necessary to re-introduce social friction that these infrastructures are designed to eliminate.
-In other words, the pressures and circumstances of being an archaeologist and doing archaeological research assert themselves when attempting to make practical use of these infrastructures, and therefore must be accounted for in their design and implementation.
-In this paper we aim to identify similar sources of dissonance with regards to the promise, potential, and actual implementation of open source software development models among archaeologists.
+In other words, the pressures and circumstances of being an archaeologist and doing archaeological research -- such as the inherent subjectivities involved in characterising finds, or the reliance on analysts’ reputation and academic pedigree to establish trust in the data they produce -- assert themselves when attempting to make practical use of these infrastructures, and therefore must be accounted for in their design and implementation.
+In this articlle we draw attention to similar sources of dissonance with regards to the promise, potential, and actual implementation of open source software development models among archaeologists by analysing the collaborative milieu in which archaeological software are actually being built and relating our findings to broader observations about how archaeologists tend to collaborate in practice.
 

### 2024-05-23 analysis/_03-data_methodology.qmd
Revisions based on editorial comments.

@@
 More specifically, we update open-archaeo by manually crawling through archaeologists' profiles on these platforms, as well as on other personal, professional, and institutional websites that describe and host additional archaeological software.
-We supplement this quasi-systematic collection strategy with word-of-mouth contributions made by interested individuals who identified relevant work that we initially overlooked.
+We supplement this quasi-systematic collection strategy with word-of-mouth contributions made by interested individuals, who reached out via email, social media or at conferences to identify relevant work that we initially overlooked, including work that they created themselves.
 
@@
 It also includes numerous non-software resources, as well as software developed and distributed without the use of software forges.
-The dataset is also limited by the experiences of its primary maintainers.^[We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
+Open-archaeo is also limited by the experiences of its primary maintainers, which affects the dataset's overall scope and how comprehensively it covers various domains of archaeological research.^[We welcome anyone, especially domain specialists who are familiar with the kinds of tools commonly used in their specific fields, to help fill in these gaps.
 Instructions for contributing to open-archaeo can be found at <https://github.com/zackbatist/open-archaeo>.]
@@
 ```{r tbl-forges}
-#| tbl-cap: Software forges used by open archaeology projects
+#| tbl-cap: Software forges used by open archaeology projects, as per open-archaeo
 oarch_forges |>
@@
 Our analysis incorporates data on `r n_repos` repositories^[We excluded three GitHub repositories from the analysis for technical reasons.
-For example, https://github.com/carpentries-incubator/R-archaeology-lesson is a repository within the scope of open-archaeo, but it was forked from a pre-existing repository that is not (https://github.com/datacarpentry/R-ecology-lesson) and thus includes in its commit history irrelevant data from the parent repository.], comprising `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
+For example, https://github.com/carpentries-incubator/R-archaeology-lesson is a repository within the scope of open-archaeo, but it was forked from a pre-existing repository that is not (https://github.com/datacarpentry/R-ecology-lesson) and thus includes in its commit history irrelevant data from the parent repository.], comprising `r n_commits` commits and `r n_issues` issues/pull requests from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
 
@@
 That being said, we cannot account for practices that occur through offline or private channels, or forms of collaboration we do not know about. 
-We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves. 
+We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves.^[JR is an extremely prolific archaeological open source software developer, who actively builds code, teaches programming methods, and conducts code reviews.
+Through his maintenance of open-archaeo over the past 5+ years, ZB has honed an extensive understanding of the people who create archaeological software and the institutional ecology that supports their work.]
 Our earliest data is from 2005 and our study can say little about collaborative software development in archaeology before this point, though we know there was a significant amount of it [@ducke2013; @whallon1972].
@@
 Finally, in @sec-network, we apply network analysis methods to investigate the formation of broader collaborative communities.
-Our analyses combine to support our objectives of understanding what kinds of software and resources archaeologists are making, and how they create these tools in response to specific needs and use-cases, as afforded by offline social and professional connections, and within the context of an emerging community of practice.
+Our analyses combine to support our objectives of understanding what kinds of software and resources archaeologists are making, how they create these tools in response to specific needs and use-cases, and how this work is situated within the context of an emerging community of practice.
 

### 2024-05-23 analysis/_04-open_archaeology.qmd
Revisions based on editorial comments.

--- 2024-04-15 analysis/_04-open_archaeology.qmd
@@
 As of writing, open-archaeo catalogues `r nrow(oarch)` resources created by and for archaeologists.
-It includes both software and documents, but not research compendiums.^[see [https://github.com/benmarwick/ctv-archaeology#publications-that-include-r-code](https://github.com/benmarwick/ctv-archaeology#publications-that-include-r-code) for a similar list of archaeology publications that include R code]
-We annotated each record with categories and tags that describe what aspect of archaeological research each tool or resource was meant to address.
-We also categorized all records based on how each tool or resource is meant to be accessed or used, and assigned tags based on how developers identified their projects' purpose and scope.
+It includes both software and documents, but not research compendiums (collections of digital resources, including data, code, and documentation, which accompany or enhance a scientific publication; see [https://research-compendium.science](https://research-compendium.science).).^[see [https://github.com/benmarwick/ctv-archaeology#publications-that-include-r-code](https://github.com/benmarwick/ctv-archaeology#publications-that-include-r-code) for a similar list of archaeology publications that include R code]
+Each record in open-archaeo is is assigned to a category based on how the tool or resource is meant to be accessed or used, and is annotated with tags that describe what aspect of archaeological research each itemtool or resource was meant to address.
+Tags are ascribed based on how developers identified their projects’ purpose and scope, and each record can have multiple tags.
 See @batist2023 for a more comprehensive overview of the tags and categories applied to open-archaeo.
@@
 
-n_platform <- sum(!is.na(oarch$platform))
-p_platform <- n_platform / nrow(oarch)
+n_packages <- nrow(filter(oarch, category == "Packages and libraries"))
+p_packages <- n_packages / nrow(oarch)
 
@@
 
-n_datasets <- nrow(filter(oarch, tag1 == "Datasets" | tag2 == "Datasets" | 
-                            tag3 == "Datasets" | tag4 == "Datasets" | 
-                            tag5 == "Datasets"))
+n_scripts <- nrow(filter(oarch, category == "Scripts"))
+p_scripts <- n_scripts / nrow(oarch)
+
+n_datasets <- nrow(filter(oarch, category == "Lists and datasets"))
 p_datasets <- n_datasets / nrow(oarch)
 
-n_scripts <- nrow(filter(oarch, category == "Scripts"))
-p_scripts <- n_scripts / nrow(oarch)
+# n_platform <- sum(!is.na(oarch$platform))
+# p_platform <- n_platform / nrow(oarch)
+
+# n_datasets <- nrow(filter(oarch, tag1 == "Datasets" | tag2 == "Datasets" | 
+#                             tag3 == "Datasets" | tag4 == "Datasets" | 
+#                             tag5 == "Datasets"))
+# p_datasets <- n_datasets / nrow(oarch)
 ```
 
-In our breakdown of open-archaeo by category (see @tbl-categories), we demonstrate the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-Most resources (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language (e.g.
+From our breakdown of open-archaeo by category (see @tbl-categories), we can infer the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
+Most resources (`r percent(p_packages)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language (e.g.
 radiocarbon calibration is implemented in R in the package 'rcarbon' or Python in the package 'iosacal') or a plugin for an application (e.g.
@@
 ```
-
 ```{r data-platforms}
-p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / nrow(oarch)
-p_platform_python <- sum(oarch$platform == "Python", na.rm = TRUE) / nrow(oarch)
-p_platform_qgis <- sum(oarch$platform == "QGIS", na.rm = TRUE) / nrow(oarch)
+  p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / sum(!is.na(oarch$platform))
+  p_platform_python <- sum(oarch$platform == "Python", na.rm = TRUE) / sum(!is.na(oarch$platform))
+  p_platform_qgis <- sum(oarch$platform == "QGIS", na.rm = TRUE) / sum(!is.na(oarch$platform))
   ```
 
-`r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used platform by a large margin (@tbl-platforms).
+Some `r percent(p_platform_r)` of projects are extensions to the statistical programming language R, making it the most widely-used platform by a large margin (@tbl-platforms).
 Python, another programming language, is also relatively popular (`r percent(p_platform_python)`), as are plugins for the open source geographic information system QGIS (`r percent(p_platform_qgis)`).^[While Blender and QGIS plugins are written using the Python language, our intent while categorizing platforms was to get a sense of the developer ecosystems in which archaeological software engineers participate, rather than to simply gauge the popularity of different languages [@batist2023 2].]
@@
 
-p_license_gpl <- n_filter(oarch, str_detect(license, "GPL")) / n_filter(oarch, !is.na(license))
-p_license_mit <- n_filter(oarch, str_detect(license, "MIT")) / n_filter(oarch, !is.na(license))
-p_license_cc <- n_filter(oarch, str_detect(license, "CC")) / n_filter(oarch, !is.na(license))
+p_license_gpl <- n_filter(oarch, str_detect(license, "GPL")) / sum(!is.na(oarch$license))
+p_license_mit <- n_filter(oarch, str_detect(license, "MIT")) / sum(!is.na(oarch$license))
+p_license_cc <- n_filter(oarch, str_detect(license, "CC")) / sum(!is.na(oarch$license))
 ```
@@
 Roughly half of open-archaeo repositories are accompanied by an explicit license (@tbl-repo-licenses).
-Two common free software licenses account for the majority of these: the GNU General Public License (GPL, `r percent(p_license_gpl)`) and the MIT License (`r percent(p_license_mit)`).
+Two common free software licenses account for the majority of repositories that do contain licenses: the GNU General Public License (GPL, `r percent(p_license_gpl)`) and the MIT License (`r percent(p_license_mit)`).
 These differ primarily in the restrictions they place on reuse: the MIT License aims to be maximally permissive, while the GPL is a 'copyleft' license that specifies that all derivative works must be distributed under similar terms (in other words, it prohibits the use of open source software within non-open software, [@dusollier2007].

### 2024-05-23 analysis/_05-collaboration.qmd
Revisions based on editorial comments.

@@
 
-We measured the lifespan of a repository as the time between the first and latest commit, and its activity as the rate of commits.
+We measured the lifespan of a repository as the time between the first and latest commit, and its activity as the number of commits per day.
 Here therefore we refer to the *development* lifespan of a project, which is not necessary related to its use-life.
@@
 While git can be a great way to track changes and manage distributed contributions to a common code base, it can also be unwieldy in situations when multiple users (especially those with less experience using git for collaborative purposes) are expected to contribute within short spans of time.
-This meshes with our prior observations that projects exhibiting higher commit rates have less contributors.
+This corresponds with our prior observations that projects exhibiting higher commit rates have less contributors.
 Additionally, our analyses neglect to account for contributions that are not tracked via git or GitHub.

### 2024-05-23 analysis/_06-network.qmd
Explain regression line
Revisions based on editorial comments.

@@
 By contributing to shared repositories—whether with code (commits), issues, or comments—archaeologists using GitHub form a collaborative network which we can map using data from the GitHub API.
-Here we consider two facets of this network: repositories connected by common contributors (the repository–repository graph), and users connected by contributions to common repositories (the user–user graph).
+From a two-mode network with edges representing users’ contributions to repositories (Figure 6), we derived two one-mode networks that render co-contributions as connections in themselves: repositories connected by common contributors (the repository–repository graph, presented in Figure 7), and users connected by contributions to common repositories (the user–user graph, presented in Figure 8).
 In both cases, number of contributions constitutes a natural measure of the strength or weight of the connection, which can be further broken down by type of contribution (commit, issue/pull request, or comment).
+We identified clusters using the edge-betweenness community detection method, which operates by locating the edges that are situated along the shortest paths between all pairs of nodes, and which therefore exhibit high betweenness centrality; these edges form indirect links between otherwise completely unconnected sections of the network, and as such are the loci that distinguish groups of nodes that are more highly connected to each other than they are to others [@girvan2002].
 
@@
 
-Our data shows that there is a significant network of archaeologists collaborating on GitHub.
+Our data show that there is a significant network of archaeologists collaborating on GitHub.
 `r percent(graph_n_repos / graph_n_repos_total)` of repositories and `r percent(graph_n_users / graph_n_users_total)` of users in our dataset are connected to at least one other repository or user.
 Of these, `r percent(max(subgraphs_n_repos$n) / graph_n_repos)` of repositories and `r percent(max(subgraphs_n_users$n) / graph_n_users)` of users belong to a single connected subgraph (@fig-graph-repo-user and @fig-graph-repo-repo).
+This indicates that most repositories that have had more than one contributor, or whose contributors have worked independently on more than one repository, are not isolated, and that at least one of their contributors have, at least one time, worked with members of a broader interconnected population.
 
-We delimited `r nrow(filter(clusters_n_repos, n > 1))` distinct clusters that outline the topography of the repository–repository network.
+We delimited `r nrow(filter(clusters_n_repos, n > 1))` distinct clusters that outline the topography of the repository–repository network (@fig-graph-repo-repo).
 While many of these clusters are interconnected, some discrete components containing between 2-20 repositories appear as distinct from a primary core.
@@
 
-Clustering also reveals distinct collaborative networks within the user-user graph.
-We again see a complementary primary core connected to several more peripheral clusters, which are internally-cohesive and exhibit few connections with other peripheral clusters.
-The central core bridges all the peripheral clusters.
-The central core is not uniform, and comprises several relatively discrete clusters representing collaborative sub-communities.
-While these clusters are internally cohesive, they exhibit enough connections to other members of the central core so as to not be considered as separate or peripheral clusters.
+Clustering also reveals distinct collaborative networks within the user-user graph (@fig-graph-user-user).
+We again see a complementary primary interconnected subgraph and several unconnected subgraphs.
+The primary subgraph comprises several connected clusters, including one central cluster and several more peripheral clusters, which are internally-cohesive and exhibit few connections with other peripheral clusters.
+The central cluster bridges all the peripheral clusters.
+Moreover, the central cluster is not uniform, and comprises several relatively discrete components representing collaborative sub-communities.
+While these components are internally cohesive, they exhibit enough connections to other members of the central cluster so as to not be considered as separate or peripheral clusters.
 
@@
 ```{r fig-repo-betweenness-age-length}
-#| fig-cap: Repository centrality by age (left) and length (right). Centrality is measured by node betweenness weighted by number of contributions.
+#| fig-cap: Repository centrality by age (left) and length (right).
+Centrality is measured by node betweenness weighted by number of contributions.
+Blue line indicates a generalized linear regression, with confidence envelope shaded in grey.
 left <- oarch_graph_repo_repo |>
@@
 These are therefore relatively stable and require little additional input after release.
-On the other hand, as discussed in @sec-open-archaeology, standalone software tend to integrate multiple system components and may evolve over time to add new features or support new workflows.
+On the other hand, standalone software tend to integrate multiple system components and may evolve over time to add new features or support new workflows.
 Moreover, standalone software are generally rooted in longer-term and community-held objectives, and their development may therefore be backed by insitutions with funding and resources to support developers.
@@
 Many of the QGIS plugins add various specialized features to the extensible GIS platform, and are therefore developed by interdisciplinary teams, which explains its high rank.
-Python projects, which tend to be infrastructural or are of interest to members of other fields, are also highly ranked in terms of average centrality.
-Our findings support the notion that there are significant development patterns when working across different languages and platforms, and further analysis to qualify these observation is warranted.
+Python projects, which tend to relate to the development of information infrastructures or are of interest to members of other fields (e.g., palaeo-ecology, other fieldwork-based disciplines), are also highly ranked in terms of average centrality.
+Further analysis is warranted to qualify these observations regarding the significance of development patterns when working across different languages and platforms.
 
 ```{r fig-user-betweenness-contribs}
-#| fig-cap: User centrality by total number of contributions. Centrality is measured by node betweenness weighted by number of contributions.
+#| fig-cap: User centrality by total number of contributions.
+Centrality is measured by node betweenness weighted by number of contributions.
+Blue line indicates a generalized linear regression, with confidence envelope shaded in grey.
 oarch_graph_user_user |>

### 2024-05-23 analysis/_07-conclusion.qmd
Revisions based on editorial comments.

@@
 We sought to identify not only _what_ kinds of software archaeologists are making, but _how_ archaeologists create these tools as part of a broader community of practice. 
-Our emphasis on the collaborative experiences involved in open source software development emerged from our experience maintaining open-archaeo, through which we observed that making one's code openly available on the web does not necessarily garner the benefits often touted by open science advocates, namely that source code can be audited, forked, and appropriated for alternative use cases, which are effectively social and collaborative experiences.
+Our emphasis on the collaborative experiences involved in open source software development emerged from our experience maintaining open-archaeo.
+We observed that making one's code openly available on the web does not necessarily garner the benefits often touted by open science advocates, namely that source code can be audited, forked, and appropriated for alternative use cases, which are effectively social and collaborative experiences.
 
@@
 There is an emerging community of practice around open source research software in archaeology. 
-All but a handful of the GitHub repositories we analysed have more than commit, showing that archaeologists use it for ongoing work rather than merely to upload finished products.
+All but a handful of the GitHub repositories we analysed have more than one commit, showing that archaeologists use it for ongoing work rather than merely to upload finished products.
 They relatively frequently make use of the 'star' and 'comment' features to engage with others' repositories (@fig-collab-features) and, via these and other shared contributions, we can trace a collaborative network that includes the majority of archaeologists active on GitHub (see @sec-network).
@@
 This core exhibits complexity that corresponds with social patterns, such as the presence of various clusters representing interconnected interest or affinity groups.
-Indeed, we have found that 'real-world' social connections and institutional support structures are strong predictors of centrality, since these clusters are representative of established professional partnerships.
+Indeed, we have inferred that 'real-world' social connections and institutional support structures are strong predictors of centrality, since these clusters are representative of established professional partnerships.
 This suggests that archaeological open source is firmly embedded within existing power structures that permeate academic life, both online and offline.
@@
 Sustaining an open source project is simply not compatible with the factors that currently drive the momentum behind archaeological work.
+This is compounded by the fact that many archaeologists consider software development as a form of support work and of lesser value than traditional academic research activities, and that research software engineers are often precariously employed [@baxter2012; @cornelius-bell2021].
 
-As such, we advocate for more focused attention on specific disciplinary norms and institutional support structures that inform how knowledge is created and validated, and how varied contributions to the scholarly enterprise are mediated, credited, and valued.
+As such, we advocate for more focused attention on specific disciplinary norms and institutional support structures that inform how knowledge is created and validated, and how varied contributions to the scholarly enterprise are mediated, credited, and valued [@leonelli2023; @bennett2021; @khan2024].
 In other words, if we want to make open source effective in relation to the aforementioned goals of encouraging greater inclusivity, transparency, and productivity, we also need to foster a culture that supports active, pragmatic and humble critique, and which instills a de-territorialized attitude concerning what it means to contribute to collective knowledge [@morgan2015; @ducke2015].

### 2024-07-13 analysis/_03-data_methodology.qmd
Correct in-text stats
Also made some small wording changes, and updated the tinytable syntax for links

@@
 ```{r tbl-forges}
-#| tbl-cap: Software forges used by open archaeology projects, as per open-archaeo
+#| tbl-cap: Software forges used by open archaeology projects
 oarch_forges |>
@@
   tt() |>
-  format_tt(1, markdown = TRUE)
+  format_tt(j = 1, markdown = TRUE)
 ```
@@
 Our analysis incorporates data on `r n_repos` repositories^[We excluded three GitHub repositories from the analysis for technical reasons.
-For example, https://github.com/carpentries-incubator/R-archaeology-lesson is a repository within the scope of open-archaeo, but it was forked from a pre-existing repository that is not (https://github.com/datacarpentry/R-ecology-lesson) and thus includes in its commit history irrelevant data from the parent repository.], comprising `r n_commits` commits and `r n_issues` issues/pull requests from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
+For example, https://github.com/carpentries-incubator/R-archaeology-lesson is a repository within the scope of open-archaeo, but it was forked from a pre-existing repository that is not (https://github.com/datacarpentry/R-ecology-lesson) and thus includes in its commit history irrelevant data from the parent repository.], `r n_commits` commits, `r n_issues` issues/pull requests, and `r n_comments` from `r n_contributors` distinct users, as well as repository metadata on programming languages used, licensing, stars and forks, and so on.
 
@@
 That being said, we cannot account for practices that occur through offline or private channels, or forms of collaboration we do not know about. 
-We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves.^[JR is an extremely prolific archaeological open source software developer, who actively builds code, teaches programming methods, and conducts code reviews.
+We did not directly observe or interview archaeological software developers, though our conclusions do draw heavily from our experience as members of that community ourselves.^[JR has been active in the development of open source research software for archaeology for about ten years.
 Through his maintenance of open-archaeo over the past 5+ years, ZB has honed an extensive understanding of the people who create archaeological software and the institutional ecology that supports their work.]

### 2024-07-13 analysis/_04-open_archaeology.qmd
Correct in-text stats
Also made some small wording changes, and updated the tinytable syntax for links

@@
   group_tt(list("Software" = 1, "Documents" = 4)) |>
-  format_tt(1, markdown = TRUE)
+  format_tt(j = 1, markdown = TRUE)
 ```
@@
 
-n_packages <- nrow(filter(oarch, category == "Packages and libraries"))
-p_packages <- n_packages / nrow(oarch)
+n_platform <- sum(!is.na(oarch$platform))
+p_platform <- n_platform / nrow(oarch)
 
@@
 
-n_datasets <- nrow(filter(oarch, category == "Lists and datasets"))
+n_datasets <- nrow(filter(oarch, tag1 == "Datasets" | tag2 == "Datasets" | 
+                            tag3 == "Datasets" | tag4 == "Datasets" | 
+                            tag5 == "Datasets"))
 p_datasets <- n_datasets / nrow(oarch)
-
-# n_platform <- sum(!is.na(oarch$platform))
-# p_platform <- n_platform / nrow(oarch)
-
-# n_datasets <- nrow(filter(oarch, tag1 == "Datasets" | tag2 == "Datasets" | 
-#                             tag3 == "Datasets" | tag4 == "Datasets" | 
-#                             tag5 == "Datasets"))
-# p_datasets <- n_datasets / nrow(oarch)
 ```
@@
 From our breakdown of open-archaeo by category (see @tbl-categories), we can infer the pervasiveness of various development models, and the requisite technical capabilities that developers assume users hold.
-Most resources (`r percent(p_packages)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language (e.g.
+Most resources (`r percent(p_platform)`) included in open-archaeo are designed to be used atop an existing "platform" -- for example a package that extends a programming language (e.g.
 radiocarbon calibration is implemented in R in the package 'rcarbon' or Python in the package 'iosacal') or a plugin for an application (e.g.
@@
 ```{r data-platforms}
-  p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / sum(!is.na(oarch$platform))
-  p_platform_python <- sum(oarch$platform == "Python", na.rm = TRUE) / sum(!is.na(oarch$platform))
-  p_platform_qgis <- sum(oarch$platform == "QGIS", na.rm = TRUE) / sum(!is.na(oarch$platform))
+p_platform_r <- sum(oarch$platform == "R", na.rm = TRUE) / nrow(oarch)
+p_platform_python <- sum(oarch$platform == "Python", na.rm = TRUE) / nrow(oarch)
+p_platform_qgis <- sum(oarch$platform == "QGIS", na.rm = TRUE) / nrow(oarch)
 ```
@@
 
-p_license_gpl <- n_filter(oarch, str_detect(license, "GPL")) / sum(!is.na(oarch$license))
-p_license_mit <- n_filter(oarch, str_detect(license, "MIT")) / sum(!is.na(oarch$license))
-p_license_cc <- n_filter(oarch, str_detect(license, "CC")) / sum(!is.na(oarch$license))
+p_license_gpl <- n_filter(oarch, str_detect(license, "GPL")) / n_filter(oarch, !is.na(license))
+p_license_mit <- n_filter(oarch, str_detect(license, "MIT")) / n_filter(oarch, !is.na(license))
+p_license_cc <- n_filter(oarch, str_detect(license, "CC")) / n_filter(oarch, !is.na(license))
 ```

### 2024-07-13 analysis/_06-network.qmd
Correct in-text stats
Also made some small wording changes, and updated the tinytable syntax for links

@@
 
-By contributing to shared repositories—whether with code (commits), issues, or comments—archaeologists using GitHub form a collaborative network which we can map using data from the GitHub API.
-From a two-mode network with edges representing users’ contributions to repositories (Figure 6), we derived two one-mode networks that render co-contributions as connections in themselves: repositories connected by common contributors (the repository–repository graph, presented in Figure 7), and users connected by contributions to common repositories (the user–user graph, presented in Figure 8).
+By contributing to shared repositories—whether with code (commits), issues, or comments—archaeologists using GitHub form a collaborative network which we can map using data from the GitHub API [@fig-graph-repo-user].
+Here we consider two facets of this network: repositories connected by common contributors (the repository–repository graph, @fig-graph-repo-repo), and users connected by contributions to common repositories (the user–user graph, @fig-graph-user-user).
 In both cases, number of contributions constitutes a natural measure of the strength or weight of the connection, which can be further broken down by type of contribution (commit, issue/pull request, or comment).
@@
   tt() |>
-  format_tt(2, markdown = TRUE)
+  format_tt(j = 2, markdown = TRUE)
 ```

### 2024-07-13 analysis/_06-network.qmd
Add R2 values to plots
As suggested by JW

@@
   scale_y_log10(labels = NULL) +
-  geom_smooth(method = "glm") +
+  geom_smooth(method = "lm") +
+  stat_poly_eq(method = "lm") +
   labs(x = "Days", y = "Log centrality →") +
@@
   scale_y_log10(labels = NULL) +
-  geom_smooth(method = "glm") +
+  geom_smooth(method = "lm") +
+  stat_poly_eq(method = "lm") +
   labs(x = "Commits", y = NULL) +
@@
   scale_y_log10() +
-  geom_smooth(method = "glm") +
+  geom_smooth(method = "lm") +
+  stat_poly_eq(method = "lm") +
   labs(x = "Total contributions", y = "Centrality") +

### 2024-07-13 analysis/openarchaeo_collaboration.qmd
Add R2 values to plots
As suggested by JW

@@
 library("ggraph")
+library("ggpmisc")
 library("glue")

### 2024-07-16 analysis/openarchaeo_collaboration.qmd
Added a note referring to version of record.
Add citation to manuscript

@@
     orcid: 0000-0002-1011-1244
-date: "`r Sys.Date()`"
+date: 2024-07-18
 abstract: |
@@
   On the contrary, our results indicate that research software engineering in archaeology remains largely embedded in conventional professional norms and organizational structures of academia.
+keywords:
+  - open archaeology
+  - open source
+  - open science
+  - archaeological practice
+  - collaboration
+license: CC BY
+citation:
+  title: "Open archaeology, open source? Collaborative practices in an emerging community of archaeological software engineers"
+  container-title: Internet Archaeology
+  volume: 67
+  doi: 10.11141/ia.67.13
+funding: "Open access funding was provided by the University of Bern and swissuniversities"
 bibliography: references.bib
